---
title: "coeff of variation"
author: "Allison Davis Connelly"
date: "2024-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Initial steps 

Data collection 

```{r, echo=FALSE}
library(curl)

raw <- curl("https://raw.githubusercontent.com/allisondavis/morphology_analysis/master/adavis_morphology_raw-data_FINAL.csv")

raw <- read.csv(raw, header = TRUE, sep = ",", stringsAsFactors = TRUE)

                                                                                                           
raw1 <- raw[, -25] #removed total length, since I'm not really using it and don't want an NAs here to remove specimen
raw1 <- raw1[complete.cases(raw1[, 15:36]),] #issues with binding data later, so making a df without NAs now to see if it helps

raw1b <- raw1[-c(252), ] #this takes out the monster fish too



```

### Checking normality 

#### Checking normality {.tabset .tabset-fade .tabset-pills}

I will use Shapiro-wilke, histograms, and QQ plots to determine what traits are normal. These will only be performed on continuous variables, as discrete variables are not normal by nature.

**Conclusions:** literally all of them are NOT normal... will log transform them and double check normality. Had to square-root three traits () because still not normal with log transform. 


#### Log transformations

Since all of the continuous characters were not normal, I will log transform them, and proceed with the transformed data in all analyses.

```{r}
raw1[paste0(names(raw1), '_log')] <- log(raw1[, 26:35], 10)
#for some reason did the log of the whole data frame instead of those specific columns, so I will just remove the unnecessary columns instead of spending time on what went wrong.

raw1 <- raw1[-c(37:60)]
raw1 <- raw1[-c(48)]


```


Need to try other transformations for body depth, predorsal length, and dorsal base length. Will try squareroot or cuberoot to see what gets them normal. Cubed got them closest to normal (bd still not normal but other two are).

```{r}
#cubed root
raw1$cbR_BD ='^'(raw1$BD,1/3)
raw1$cbR_PreDL ='^'(raw1$PreDL,1/3)
raw1$cbR_DbL ='^'(raw1$DbL,1/3)


#need to cube root SL to use in regressions

raw1$cbR_SL ='^'(raw1$SL,1/3)



```

## Correcting for body size (residuals) 

### SL correction {.tabset .tabset-fade .tabset-pills}

Since amazons are in general bigger than sailfin, we don't want any results to be due to this difference in body size bias. Therefore, we will see what traits are influenced by body size (regressions) and correct for body size when necessary (absolute value of residuals). We can then use the residuals when comparing between species for traits that are influenced by body size, and raw/transformed data for traits that are not influenced by body size. I will also calculate standardized residuals to compare residuals across traits in later analyses.

**Quick results summary:** traits *not* influenced by body size are left & right pelvic, anal, scales \nabove and below lateral line (except in mexicana), scales \nbefore dorsal fin and fluctuating asymmetry; all other traits influenced by body size.

#### Regressions 

-   STEP ONE: plot trait vs body size

```{r}

library(ggplot2)
library(ggpubr)



lat <- raw1[raw1$SPP == "p.latipinna",]


form <- raw1[raw1$SPP == "p.formosa",]


mex <- raw1[raw1$SPP == "p.mexicana",]


##### LAT #####
par(mfcol=c(2,2), oma = c(0,0,2,0))

reg.lat.D <- lm(lat$D ~ lat$SL)
sd.lat.D <- rstandard(reg.lat.D)
reg.lat.D.plot <- ggplot(lat, aes(x = SL, y = D)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 11,  size=5) + theme_bw() + labs(x= "\n Standard length (mm)", y="Number of rays\n")
reg.lat.D.plot

reg.lat.P1 <- lm(lat$P1 ~ lat$SL)
sd.lat.P1 <- rstandard(reg.lat.P1)
reg.lat.P1.plot <- ggplot(lat, aes(x = SL, y = P1)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 11, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Number of rays\n")
reg.lat.P1.plot

reg.lat.A <- lm(lat$A ~ lat$SL)
sd.lat.A <- rstandard(reg.lat.A)
reg.lat.A.plot <- ggplot(lat, aes(x = SL, y = A)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 8, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Number of rays\n")
reg.lat.A.plot

reg.lat.P1.R <- lm(lat$P1.R ~ lat$SL)
sd.lat.P1.R <- rstandard(reg.lat.P1.R)
reg.lat.P1.R.plot <- ggplot(lat, aes(x = SL, y = P1.R)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 11, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Number of rays\n")
reg.lat.P1.R.plot

reg.lat.LLSC <- lm(lat$LLSC ~ lat$SL)
sd.lat.LLSC <- rstandard(reg.lat.LLSC)
reg.lat.LLSC.plot <- ggplot(lat, aes(x = SL, y = LLSC)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 20, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Number of scales\n")
reg.lat.LLSC.plot

reg.lat.SALL <- lm(lat$SALL ~ lat$SL)
sd.lat.SALL <- rstandard(reg.lat.SALL)
reg.lat.SALL.plot <- ggplot(lat, aes(x = SL, y = SALL)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 1.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Number of scales\n")
reg.lat.SALL.plot

reg.lat.SBLL <- lm(lat$SBLL ~ lat$SL)
sd.lat.SBLL <- rstandard(reg.lat.SBLL)
reg.lat.SBLL.plot <- ggplot(lat, aes(x = SL, y = SBLL)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 2, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Number of scales\n")
reg.lat.SBLL.plot

reg.lat.SBDF <- lm(lat$SBDF ~ lat$SL)
sd.lat.SBDF <- rstandard(reg.lat.SBDF)
reg.lat.SBDF.plot <- ggplot(lat, aes(x = SL, y = SBDF)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 3.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Number of scales\n")
reg.lat.SBDF.plot

reg.lat.BD <- lm(lat$cbR_BD ~ lat$cbR_SL)
sd.lat.BD <- rstandard(reg.lat.BD)
reg.lat.BD.plot <- ggplot(lat, aes(x = cbR_SL, y = cbR_BD)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 3, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, cube-root transformed)", y="Depth (mm, cube-root transformed)\n")
reg.lat.BD.plot

reg.lat.CPD <- lm(lat$CPD_log ~ lat$SL_log)
sd.lat.CPD <- rstandard(reg.lat.CPD)
reg.lat.CPD.plot <- ggplot(lat, aes(x = SL_log, y = CPD_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 1.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, log transformed)", y="Depth (mm, log transformed)\n")
reg.lat.CPD.plot

reg.lat.CPL <- lm(lat$CPL_log ~ lat$SL_log)
sd.lat.CPL <- rstandard(reg.lat.CPL)
reg.lat.CPL.plot <- ggplot(lat, aes(x = SL_log, y = CPL_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 1.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, log transformed)", y="Length (mm, log transformed)\n")
reg.lat.CPL.plot

reg.lat.PreDL <- lm(lat$cbR_PreDL ~ lat$cbR_SL)
sd.lat.PreDL <- rstandard(reg.lat.PreDL)
reg.lat.PreDL.plot <- ggplot(lat, aes(x = cbR_SL, y = cbR_PreDL)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 3.2, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, cube-root transformed)", y="Length (mm, cube-root transformed)\n")
reg.lat.PreDL.plot

reg.lat.DbL <- lm(lat$cbR_DbL ~ lat$cbR_SL)
sd.lat.DbL <- rstandard(reg.lat.DbL)
reg.lat.DbL.plot <- ggplot(lat, aes(x = cbR_SL, y = cbR_DbL)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 3, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, cube-root transformed)", y="Length (mm, cube-root transformed)\n")
reg.lat.DbL.plot

reg.lat.HL <- lm(lat$HL_log ~ lat$SL_log)
sd.lat.HL <- rstandard(reg.lat.HL)
reg.lat.HL.plot <- ggplot(lat, aes(x = SL_log, y = HL_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 1.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, log transformed)", y="Length (mm, log transformed)\n")
reg.lat.HL.plot

reg.lat.HD <- lm(lat$HD_log ~ lat$SL_log)
sd.lat.HD <- rstandard(reg.lat.HD)
reg.lat.HD.plot <- ggplot(lat, aes(x = SL_log, y = HD_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 1.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, log transformed)", y="Depth (mm, log transformed)\n")
reg.lat.HD.plot

reg.lat.HW <- lm(lat$HW_log ~ lat$SL_log)
sd.lat.HW <- rstandard(reg.lat.HW)
reg.lat.HW.plot <- ggplot(lat, aes(x = SL_log, y = HW_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 1.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, log transformed)", y="Width (mm, log transformed)\n")
reg.lat.HW.plot

reg.lat.SnL <- lm(lat$SnL_log ~ lat$SL_log)
sd.lat.SnL <- rstandard(reg.lat.SnL)
reg.lat.SnL.plot <- ggplot(lat, aes(x = SL_log, y = SnL_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 1.8, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, log transformed)", y="Length (mm, log transformed)\n")
reg.lat.SnL.plot

reg.lat.OL <- lm(lat$OL_log ~ lat$SL_log)
sd.lat.OL <- rstandard(reg.lat.OL)
reg.lat.OL.plot <- ggplot(lat, aes(x = SL_log, y = OL_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 1.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, log transformed)", y="Length (mm, log transformed)\n")
reg.lat.OL.plot

reg.lat.FLA <- lm(lat$FLA ~ lat$SL)
sd.lat.FLA <- rstandard(reg.lat.FLA)
reg.lat.FLA.plot <- ggplot(lat, aes(x = SL, y = FLA)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#78B7C5") +
  stat_cor(label.y = 3.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Difference in pectoral ray count \n (fluctuating asymmetries)\n")
reg.lat.FLA.plot



##### FORM #####

reg.form.D <- lm(form$D ~ form$SL)
sd.form.D <- rstandard(reg.form.D)
reg.form.D.plot <- ggplot(form, aes(x =SL, y = D)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 9, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Number of rays\n")
reg.form.D.plot

reg.form.P1 <- lm(form$P1 ~ form$SL)
sd.form.P1 <- rstandard(reg.form.P1)
reg.form.P1.plot <- ggplot(form, aes(x = SL, y = P1)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 11, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Number of rays\n")
reg.form.P1.plot

reg.form.A <- lm(form$A ~ form$SL)
sd.form.A <- rstandard(reg.form.A)
reg.form.A.plot <- ggplot(form, aes(x = SL, y = A)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 8, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Number of rays\n")
reg.form.A.plot

reg.form.P1.R <- lm(form$P1.R ~ form$SL)
sd.form.P1.R <- rstandard(reg.form.P1.R)
reg.form.P1.R.plot <- ggplot(form, aes(x = SL, y = P1.R)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 11, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Number of rays\n")
reg.form.P1.R.plot

reg.form.LLSC <- lm(form$LLSC ~ form$SL)
sd.form.LLSC <- rstandard(reg.form.LLSC)
reg.form.LLSC.plot <- ggplot(form, aes(x = SL, y = LLSC)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 18.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Number of scales\n")
reg.form.LLSC.plot

reg.form.SALL <- lm(form$SALL ~ form$SL)
sd.form.SALL <- rstandard(reg.form.SALL)
reg.form.SALL.plot <- ggplot(form, aes(x = SL, y = SALL)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 1.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Number of scales\n")
reg.form.SALL.plot

reg.form.SBLL <- lm(form$SBLL ~ form$SL)
sd.form.SBLL <- rstandard(reg.form.SBLL)
reg.form.SBLL.plot <- ggplot(form, aes(x = SL, y = SBLL)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 2, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Number of scales\n")
reg.form.SBLL.plot

reg.form.SBDF <- lm(form$SBDF ~ form$SL)
sd.form.SBDF <- rstandard(reg.form.SBDF)
reg.form.SBDF.plot <- ggplot(form, aes(x = SL, y = SBDF)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Number of scales\n")
reg.form.SBDF.plot

reg.form.BD <- lm(form$cbR_BD ~ form$cbR_SL)
sd.form.BD <- rstandard(reg.form.BD)
reg.form.BD.plot <- ggplot(form, aes(x = cbR_SL, y = cbR_BD)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 3, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, cube-root transformed)", y="Depth (mm, cube-root transformed)\n")
reg.form.BD.plot

reg.form.CPD <- lm(form$CPD_log ~ form$SL_log)
sd.form.CPD <- rstandard(reg.form.CPD)
reg.form.CPD.plot <- ggplot(form, aes(x = SL_log, y = CPD_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 1.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, log transformed)", y="Depth (mm, log transformed)\n")
reg.form.CPD.plot

reg.form.CPL <- lm(form$CPL_log ~ form$SL_log)
sd.form.CPL <- rstandard(reg.form.CPL)
reg.form.CPL.plot <- ggplot(form, aes(x = SL_log, y = CPL_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 1.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, log transformed)", y="Length (mm, log transformed)\n")
reg.form.CPL.plot

reg.form.PreDL <- lm(form$cbR_PreDL ~ form$cbR_SL)
sd.form.PreDL <- rstandard(reg.form.PreDL)
reg.form.PreDL.plot <- ggplot(form, aes(x = cbR_SL, y = cbR_PreDL)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 3.2, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, cube-root transformed)", y="Length (mm, cube-root transformed)\n")
reg.form.PreDL.plot

reg.form.DbL <- lm(form$cbR_DbL ~ form$cbR_SL)
sd.form.DbL <- rstandard(reg.form.DbL)
reg.form.DbL.plot <- ggplot(form, aes(x = cbR_SL, y = cbR_DbL)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 3, size=5) + theme_bw() + labs(x= "\n Standard length (mm, cube-root transformed)", y="Length (mm, cube-root transformed)\n")
reg.form.DbL.plot

reg.form.HL <- lm(form$HL_log ~ form$SL_log)
sd.form.HL <- rstandard(reg.form.HL)
reg.form.HL.plot <- ggplot(form, aes(x = SL_log, y = HL_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 1.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, log transformed)", y="Length (mm, log transformed)\n")
reg.form.HL.plot

reg.form.HD <- lm(form$HD_log ~ form$SL_log)
sd.form.HD <- rstandard(reg.form.HD)
reg.form.HD.plot <- ggplot(form, aes(x = SL_log, y = HD_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 1.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, log transformed)", y="Depth (mm, log transformed)\n")
reg.form.HD.plot

reg.form.HW <- lm(form$HW_log ~ form$SL_log)
sd.form.HW <- rstandard(reg.form.HW)
reg.form.HW.plot <- ggplot(form, aes(x = SL_log, y = HW_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 1.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, log transformed)", y="Width (mm, log transformed)\n")
reg.form.HW.plot

reg.form.SnL <- lm(form$SnL_log ~ form$SL_log)
sd.form.SnL <- rstandard(reg.form.SnL)
reg.form.SnL.plot <- ggplot(form, aes(x = SL_log, y = SnL_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 1.8, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, log transformed)", y="Length (mm, log transformed)\n")
reg.form.SnL.plot

reg.form.OL <- lm(form$OL_log ~ form$SL_log)
sd.form.OL <- rstandard(reg.form.OL)
reg.form.OL.plot <- ggplot(form, aes(x = SL_log, y = OL_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 1.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm, log transformed)", y="Length (mm, log transformed)\n")
reg.form.OL.plot

reg.form.FLA <- lm(form$FLA ~ form$SL)
sd.form.FLA <- rstandard(reg.form.FLA)
reg.form.FLA.plot <- ggplot(form, aes(x = SL, y = FLA)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "#E1AF00") +
  stat_cor(label.y = 3.5, size=5)+ theme_bw() + labs(x= "\n Standard length (mm)", y="Difference in pectoral ray count \n (fluctuating asymmetries)\n")
reg.form.FLA.plot





##### MEX #####
reg.mex.D <- lm(mex$D ~ mex$SL)
sd.mex.D <- rstandard(reg.mex.D)
reg.mex.D.plot <- ggplot(mex, aes(x = SL, y = D)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 11)
reg.mex.D.plot

reg.mex.P1 <- lm(mex$P1 ~ mex$SL)
sd.mex.P1 <- rstandard(reg.mex.P1)
reg.mex.P1.plot <- ggplot(mex, aes(x = SL, y = P1)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 11)
reg.mex.P1.plot

reg.mex.A <- lm(mex$A ~ mex$SL)
sd.mex.A <- rstandard(reg.mex.A)
reg.mex.A.plot <- ggplot(mex, aes(x = SL, y = A)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 8)
reg.mex.A.plot

reg.mex.P1.R <- lm(mex$P1.R ~ mex$SL)
sd.mex.P1.R <- rstandard(reg.mex.P1.R)
reg.mex.P1.R.plot <- ggplot(mex, aes(x = SL, y = P1.R)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 11)
reg.mex.P1.R.plot

reg.mex.LLSC <- lm(mex$LLSC ~ mex$SL)
sd.mex.LLSC <- rstandard(reg.mex.LLSC)
reg.mex.LLSC.plot <- ggplot(mex, aes(x = SL, y = LLSC)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 20)
reg.mex.LLSC.plot

reg.mex.SALL <- lm(mex$SALL ~ mex$SL)
sd.mex.SALL <- rstandard(reg.mex.SALL)
reg.mex.SALL.plot <- ggplot(mex, aes(x = SL, y = SALL)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 1.5)
reg.mex.SALL.plot

reg.mex.SBLL <- lm(mex$SBLL ~ mex$SL)
sd.mex.SBLL <- rstandard(reg.mex.SBLL)
reg.mex.SBLL.plot <- ggplot(mex, aes(x = SL, y = SBLL)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 2)
reg.mex.SBLL.plot

reg.mex.SBDF <- lm(mex$SBDF ~ mex$SL)
sd.mex.SBDF <- rstandard(reg.mex.SBDF)
reg.mex.SBDF.plot <- ggplot(mex, aes(x = SL, y = SBDF)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 5)
reg.mex.SBDF.plot

reg.mex.BD <- lm(mex$cbR_BD ~ mex$cbR_SL)
sd.mex.BD <- rstandard(reg.mex.BD)
reg.mex.BD.plot <- ggplot(mex, aes(x = cbR_SL, y = cbR_BD)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 3.5)
reg.mex.BD.plot

reg.mex.CPD <- lm(mex$CPD_log ~ mex$SL_log)
sd.mex.CPD <- rstandard(reg.mex.CPD)
reg.mex.CPD.plot <- ggplot(mex, aes(x = SL_log, y = CPD_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 3.5)
reg.mex.CPD.plot

reg.mex.CPL <- lm(mex$CPL_log ~ mex$SL_log)
sd.mex.CPL <- rstandard(reg.mex.CPL)
reg.mex.CPL.plot <- ggplot(mex, aes(x = SL_log, y = CPL_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 3.5)
reg.mex.CPL.plot

reg.mex.PreDL <- lm(mex$cbR_PreDL ~ mex$cbR_SL)
sd.mex.PreDL <- rstandard(reg.mex.PreDL)
reg.mex.PreDL.plot <- ggplot(mex, aes(x = cbR_SL, y = cbR_PreDL)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 3.5)
reg.mex.PreDL.plot

reg.mex.DbL <- lm(mex$cbR_DbL ~ mex$cbR_SL)
sd.mex.DbL <- rstandard(reg.mex.DbL)
reg.mex.DbL.plot <- ggplot(mex, aes(x = cbR_SL, y = cbR_DbL)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 3.5)
reg.mex.DbL.plot

reg.mex.HL <- lm(mex$HL_log ~ mex$SL_log)
sd.mex.HL <- rstandard(reg.mex.HL)
reg.mex.HL.plot <- ggplot(mex, aes(x = SL_log, y = HL_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 3.5)
reg.mex.HL.plot

reg.mex.HD <- lm(mex$HD_log ~ mex$SL_log)
sd.mex.HD <- rstandard(reg.mex.HD)
reg.mex.HD.plot <- ggplot(mex, aes(x = SL_log, y = HD_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 3.5)
reg.mex.HD.plot

reg.mex.HW <- lm(mex$HW_log ~ mex$SL_log)
sd.mex.HW <- rstandard(reg.mex.HW)
reg.mex.HW.plot <- ggplot(mex, aes(x = SL_log, y = HW_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 3.5)
reg.mex.HW.plot

reg.mex.SnL <- lm(mex$SnL_log ~ mex$SL_log)
sd.mex.SnL <- rstandard(reg.mex.SnL)
reg.mex.SnL.plot <- ggplot(mex, aes(x = SL_log, y = SnL_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 3.5)
reg.mex.SnL.plot

reg.mex.OL <- lm(mex$OL_log ~ mex$SL_log)
sd.mex.OL <- rstandard(reg.mex.OL)
reg.mex.OL.plot <- ggplot(mex, aes(x = SL_log, y = OL_log)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 3.5)
reg.mex.OL.plot

reg.mex.FLA <- lm(mex$FLA ~ mex$SL)
sd.mex.FLA <- rstandard(reg.mex.FLA)
reg.mex.FLA.plot <- ggplot(mex, aes(x = SL, y = FLA)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  stat_cor(label.y = 4)
reg.mex.FLA.plot


```

#### Residuals 

-   **STEP TWO:** get residuals for each individual for traits that were influenced by body size

```{r echo=FALSE}

##### LAT #####
res.lat.D <- resid(reg.lat.D)

res.lat.P1 <- resid(reg.lat.P1)

res.lat.P1.R <- resid(reg.lat.P1.R)

res.lat.LLSC <- resid(reg.lat.LLSC)

res.lat.SBLL <- resid(reg.lat.SBLL) #will only be used when comparing with p.mex; raw data used otherwise

res.lat.BD <- resid(reg.lat.BD)

res.lat.CPD <- resid(reg.lat.CPD)

res.lat.CPL <- resid(reg.lat.CPL)

res.lat.PreDL <- resid(reg.lat.PreDL)

res.lat.DbL <- resid(reg.lat.DbL)

res.lat.HL <- resid(reg.lat.HL)

res.lat.HD <- resid(reg.lat.HD)

res.lat.HW <- resid(reg.lat.HW)

res.lat.SnL <- resid(reg.lat.SnL)

res.lat.OL <- resid(reg.lat.OL)

##### FORM #####

res.form.D <- resid(reg.form.D)

res.form.P1 <- resid(reg.form.P1)

res.form.P1.R <- resid(reg.form.P1.R)

res.form.LLSC <- resid(reg.form.LLSC)

res.form.SBLL <- resid(reg.form.SBLL) #will only be used when comparing with p.mex; raw data used otherwise

res.form.BD <- resid(reg.form.BD)

res.form.CPD <- resid(reg.form.CPD)

res.form.CPL <- resid(reg.form.CPL)

res.form.PreDL <- resid(reg.form.PreDL)

res.form.DbL <- resid(reg.form.DbL)

res.form.HL <- resid(reg.form.HL)

res.form.HD <- resid(reg.form.HD)

res.form.HW <- resid(reg.form.HW)

res.form.SnL <- resid(reg.form.SnL)

res.form.OL <- resid(reg.form.OL)


##### MEX #####
res.mex.D <- resid(reg.mex.D)

res.mex.P1 <- resid(reg.mex.P1)

res.mex.P1.R <- resid(reg.mex.P1.R)

res.mex.LLSC <- resid(reg.mex.LLSC)

res.mex.SBLL <- resid(reg.mex.SBLL)

res.mex.BD <- resid(reg.mex.BD)

res.mex.CPD <- resid(reg.mex.CPD)

res.mex.CPL <- resid(reg.mex.CPL)

res.mex.PreDL <- resid(reg.mex.PreDL)

res.mex.DbL <- resid(reg.mex.DbL)

res.mex.HL <- resid(reg.mex.HL)

res.mex.HD <- resid(reg.mex.HD)

res.mex.HW <- resid(reg.mex.HW)

res.mex.SnL <- resid(reg.mex.SnL)

res.mex.OL <- resid(reg.mex.OL)


```

-   **STEP THREE:** convert residuals to absolute value

```{r}

##### LAT #####

abs.lat.D <- abs(res.lat.D)
mean(abs.lat.D)

abs.lat.P1 <- abs(res.lat.P1)
mean(abs.lat.P1)

abs.lat.P1.R <- abs(res.lat.P1.R)
mean(abs.lat.P1.R)

abs.lat.LLSC <- abs(res.lat.LLSC)
mean(abs.lat.LLSC)

abs.lat.SBLL <- abs(res.lat.SBLL)
mean(abs.lat.SBLL)

abs.lat.BD <- abs(res.lat.BD)
mean(abs.lat.BD)

abs.lat.CPD <- abs(res.lat.CPD)
mean(abs.lat.CPD)

abs.lat.CPL <- abs(res.lat.CPL)
mean(abs.lat.CPL)

abs.lat.PreDL <- abs(res.lat.PreDL)
mean(abs.lat.PreDL)

abs.lat.DbL <- abs(res.lat.DbL)
mean(abs.lat.DbL)

abs.lat.HL <- abs(res.lat.HL)
mean(abs.lat.HL)

abs.lat.HD <- abs(res.lat.HD)
mean(abs.lat.HD)

abs.lat.HW <- abs(res.lat.HW)
mean(abs.lat.HW)

abs.lat.SnL <- abs(res.lat.SnL)
mean(abs.lat.SnL)

abs.lat.OL <- abs(res.lat.OL)
mean(abs.lat.OL)


##### FORM #####

abs.form.D <- abs(res.form.D)
mean(abs.form.D)

abs.form.P1 <- abs(res.form.P1)
mean(abs.form.P1)

abs.form.P1.R <- abs(res.form.P1.R)
mean(abs.form.P1.R)

abs.form.LLSC <- abs(res.form.LLSC)
mean(abs.form.LLSC)

abs.form.SBLL <- abs(res.form.SBLL)
mean(abs.form.SBLL)

abs.form.BD <- abs(res.form.BD)
mean(abs.form.BD)

abs.form.CPD <- abs(res.form.CPD)
mean(abs.form.CPD)

abs.form.CPL <- abs(res.form.CPL)
mean(abs.form.CPL)

abs.form.PreDL <- abs(res.form.PreDL)
mean(abs.form.PreDL)

abs.form.DbL <- abs(res.form.DbL)
mean(abs.form.DbL)

abs.form.HL <- abs(res.form.HL)
mean(abs.form.HL)

abs.form.HD <- abs(res.form.HD)
mean(abs.form.HD)

abs.form.HW <- abs(res.form.HW)
mean(abs.form.HW)

abs.form.SnL <- abs(res.form.SnL)
mean(abs.form.SnL)

abs.form.OL <- abs(res.form.OL)
mean(abs.form.OL)

##### MEX #####

abs.mex.D <- abs(res.mex.D)
mean(abs.mex.D)

abs.mex.P1 <- abs(res.mex.P1)
mean(abs.mex.P1)

abs.mex.P1.R <- abs(res.mex.P1.R)
mean(abs.mex.P1.R)

abs.mex.LLSC <- abs(res.mex.LLSC)
mean(abs.mex.LLSC)

abs.mex.SBLL <- abs(res.mex.SBLL)
mean(abs.mex.SBLL)

abs.mex.BD <- abs(res.mex.BD)
mean(abs.mex.BD)

abs.mex.CPD <- abs(res.mex.CPD)
mean(abs.mex.CPD)

abs.mex.CPL <- abs(res.mex.CPL)
mean(abs.mex.CPL)

abs.mex.PreDL <- abs(res.mex.PreDL)
mean(abs.mex.PreDL)

abs.mex.DbL <- abs(res.mex.DbL)
mean(abs.mex.DbL)

abs.mex.HL <- abs(res.mex.HL)
mean(abs.mex.HL)

abs.mex.HD <- abs(res.mex.HD)
mean(abs.mex.HD)

abs.mex.HW <- abs(res.mex.HW)
mean(abs.mex.HW)

abs.mex.SnL <- abs(res.mex.SnL)
mean(abs.mex.SnL)

abs.mex.OL <- abs(res.mex.OL)
mean(abs.mex.OL)


#let's get this into the raw1 data set so that we can plot this more easily

abs.res.D <- c(abs.lat.D, abs.form.D, abs.mex.D)
abs.res.P1 <- c(abs.lat.P1, abs.form.P1, abs.mex.P1)
abs.res.P1.R <- c(abs.lat.P1.R, abs.form.P1.R, abs.mex.P1.R)
abs.res.LLSC<- c(abs.lat.LLSC, abs.form.LLSC, abs.mex.LLSC)
abs.res.SBLL<- c(abs.lat.SBLL, abs.form.SBLL, abs.mex.SBLL)
abs.res.BD<- c(abs.lat.BD, abs.form.BD, abs.mex.BD)
abs.res.CPD<- c(abs.lat.CPD, abs.form.CPD, abs.mex.CPD)
abs.res.CPL<- c(abs.lat.CPL, abs.form.CPL, abs.mex.CPL)
abs.res.PreDL <- c(abs.lat.PreDL, abs.form.PreDL, abs.mex.PreDL)
abs.res.DbL <- c(abs.lat.DbL, abs.form.DbL, abs.mex.DbL)
abs.res.HL<- c(abs.lat.HL, abs.form.HL, abs.mex.HL)
abs.res.HD<- c(abs.lat.HD, abs.form.HD, abs.mex.HD)
abs.res.HW <- c(abs.lat.HW, abs.form.HW, abs.mex.HW)
abs.res.SnL <- c(abs.lat.SnL, abs.form.SnL, abs.mex.SnL)
abs.res.OL <- c(abs.lat.OL, abs.form.OL, abs.mex.OL)


raw2 <- cbind(raw1, abs.res.D, abs.res.P1, abs.res.P1.R, abs.res.LLSC, abs.res.SBLL, abs.res.BD, abs.res.CPD, abs.res.CPL, abs.res.PreDL, abs.res.DbL, abs.res.HL, abs.res.HD, abs.res.HW, abs.res.SnL, abs.res.OL)

```

#### Residual Comparisons 

-   **STEP FOUR:** plot the mean of the |residuals| for both species, for all traits influenced by body size

```{r}

library(ggbeeswarm)
library(dplyr)


ggplot(raw2, aes(SPP, abs.res.D)) +
  geom_point(alpha=0.3) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", 
               width=0.03, colour="red", alpha=0.7) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3)


scatter_violin <- ggplot(data=raw2, aes(x=SPP, y=abs.res.D)) +
  geom_violin(trim = FALSE) + 
  stat_summary(
    fun.data = "mean_sdl",  fun.args = list(mult = 1),
    geom = "pointrange", color = "black"
    )

print(scatter_violin)

scatter_violin1 <- ggplot(data=raw2, aes(x=SPP, y=abs.res.D)) +
  geom_violin(trim = FALSE) + 
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", 
               width=0.03, colour="red", alpha=0.7) +
  stat_summary(fun=mean, geom="crossbar", fill="red", width=0.03)

print(scatter_violin1)



ggplot(raw2, aes(SPP, abs.res.P1)) +
  geom_point(alpha=0.3) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", 
               width=0.03, colour="red", alpha=0.7) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3)

ggplot(raw2, aes(SPP, abs.res.P1.R)) +
  geom_point(alpha=0.3) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", 
               width=0.03, colour="red", alpha=0.7) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3) 

ggplot(raw2, aes(SPP, abs.res.LLSC)) +
  geom_point(alpha=0.3) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", 
               width=0.03, colour="red", alpha=0.7) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3)

ggplot(raw2, aes(SPP, abs.res.SBLL)) +
  geom_point(alpha=0.3) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", 
               width=0.03, colour="red", alpha=0.7) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3)

ggplot(raw2, aes(SPP, abs.res.BD)) +
  geom_point(alpha=0.3) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", 
               width=0.03, colour="red", alpha=0.7) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3)

ggplot(raw2, aes(SPP, abs.res.CPD)) +
  geom_point(alpha=0.3) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", 
               width=0.03, colour="red", alpha=0.7) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3)

ggplot(raw2, aes(SPP, abs.res.CPL)) +
  geom_point(alpha=0.3) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", 
               width=0.03, colour="red", alpha=0.7) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3) 

ggplot(raw2, aes(SPP, abs.res.PreDL)) +
  geom_point(alpha=0.3) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", 
               width=0.03, colour="red", alpha=0.7) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3)

ggplot(raw2, aes(SPP, abs.res.DbL)) +
  geom_point(alpha=0.3) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", 
               width=0.03, colour="red", alpha=0.7) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3)

ggplot(raw2, aes(SPP, abs.res.HL)) +
  geom_point(alpha=0.3) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", 
               width=0.03, colour="red", alpha=0.7) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3)

ggplot(raw2, aes(SPP, abs.res.HD)) +
  geom_point(alpha=0.3) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", 
               width=0.03, colour="red", alpha=0.7) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3)

ggplot(raw2, aes(SPP, abs.res.HW)) +
  geom_point(alpha=0.3) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", 
               width=0.03, colour="red", alpha=0.7) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3)

ggplot(raw2, aes(SPP, abs.res.SnL)) +
  geom_point(alpha=0.3) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", 
               width=0.03, colour="red", alpha=0.7) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3)

ggplot(raw2, aes(SPP, abs.res.OL)) +
  geom_point(alpha=0.3) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", 
               width=0.03, colour="red", alpha=0.7) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3)




```

## Comparing variation 

### Variance tests {.tabset .tabset-fade .tabset-pills}

  (1) t-test for residuals, continuous 
  (2) residual discrete will go through glmer
  (3) raw discrete will go through f-test (no idea how to do this otherwise), and results will be compared to visualizations to check for discrepencies. 


F-test on residuals doesn't make much sense; the residuals themselves are representative of the variation present, as they are the distance from the mean. Therefore, F-test on residuals is like variance of the variance. Instead, I have to do a t-test on the absolute value of the residuals. In this sense, we want to see if the mean of the absolute residuals is higher or lower for asexual species--is the average amount of variation higher or lower for this trait? Based on the regressions, if the trait was influenced by body size, I will perform a t-test on the absolute value of the residuals. If the trait was not influenced by body size, I will perform an F-test of variance on the raw data.

**Quick results summary:** For the F-test on raw data, none of the traits were significantly different (P2L/R, A, SBDF, FLA). For the t-tests on residuals, the only significant traits are left pectoral (), right pectoral (lat>form), scalesabove lateral line (), scalesbelow lateral line (form>lat), and head length ().

```         
-   will do two-tailed and check out the residual means to infer direction; for traits in which we use raw data, a one-tailed f-test will be perfomed in both direction to determine which species is varying more. We will also visulize the variation using a histogram to confirm direction results.
```
#### (1) Continuous, residuals


   -    will do two-tailed and check out the means to infer direction (see beginning for trait means w/o body correction)
   -    this will be for the traits that *did* have a significant regression compared to SL; will be using the absolute value residuals. 
   -    For continuous traits, we will be using the transformed data.
    
##### T-tests

```{r, echo=FALSE}


ttest.abs.BD <- t.test(abs.lat.BD, abs.form.BD, alternative = "two.sided")
ttest.abs.BD


ttest.abs.CPD <- t.test(abs.lat.CPD, abs.form.CPD, alternative = "two.sided")
ttest.abs.CPD

ttest.abs.CPD1 <- t.test(abs.lat.CPD, abs.form.CPD, alternative = "greater")
ttest.abs.CPD1

ttest.abs.CPD2 <- t.test(abs.lat.CPD, abs.form.CPD, alternative = "less")
ttest.abs.CPD2


ttest.abs.CPL <- t.test(abs.lat.CPL, abs.form.CPL, alternative = "two.sided")
ttest.abs.CPL


ttest.abs.PreDL <- t.test(abs.lat.PreDL, abs.form.PreDL, alternative = "two.sided")
ttest.abs.PreDL


ttest.abs.DbL <- t.test(abs.lat.DbL, abs.form.DbL, alternative = "two.sided")
ttest.abs.DbL


ttest.abs.HL <- t.test(abs.lat.HL, abs.form.HL, alternative = "two.sided")
ttest.abs.HL


ttest.abs.HD <- t.test(abs.lat.HD, abs.form.HD, alternative = "two.sided")
ttest.abs.HD


ttest.abs.HW <- t.test(abs.lat.HW, abs.form.HW, alternative = "two.sided")
ttest.abs.HW


ttest.abs.SnL <- t.test(abs.lat.SnL, abs.form.SnL, alternative = "two.sided")
ttest.abs.SnL


ttest.abs.OL <- t.test(abs.lat.OL, abs.form.OL, alternative = "two.sided")
ttest.abs.OL




```

***

#### (2) Discrete, residuals

Logan suggested doing a glm with poisson distribution (because it is count data) for these relationships. 

Had to remove the poisson part, because the residuals didn't meet the non-integer requirement for poisson (it freaked out and took forever to load).

##### GLMs

```{r}

raw3 <- raw2[raw2$SPP !="p.mexicana", ]

glm_D <- glm(abs.res.D~SPP, data=raw3)
print(summary(glm_D), show.residuals=TRUE)
D.aov <- aov(glm_D)
summary(D.aov)

glm_P1 <- glm(abs.res.P1~SPP, data=raw3)
print(summary(glm_P1), show.residuals=TRUE)
P1.aov <- aov(glm_P1)
summary(P1.aov)

glm_P1.R <- glm(abs.res.P1.R~SPP, data=raw3)
print(summary(glm_P1.R), show.residuals=TRUE)
P1.R.aov <- aov(glm_P1.R)
summary(P1.R.aov)

glm_LLSC <- glm(abs.res.LLSC~SPP, data=raw3)
print(summary(glm_LLSC), show.residuals=TRUE)
LLSC.aov <- aov(glm_LLSC)
summary(LLSC.aov)

```


***

#### (3) Discrete, raw

This will only be for the traits that did **not** have a significant regression compared to SL (so P2s, A, SALL, SBDF, and FLA); none of these traits were transformed, so just doing raw for now.


##### Levene's test


Still only performing this on the traits that did **NOT** vary with SL (P2L/R, A, SALL, SBLL [between sail and amz only], SBDF, FLA).

```{r}

library(car)
library(carData)

(LT_P2L <- leveneTest(P2.L~SPP, data=raw3)) #gives nothing since it's all the same value

(LT_P2R <- leveneTest(P2.R~SPP, data=raw3)) #same as above

(LT_A <- leveneTest(A~SPP, data=raw3)) 

(LT_SALL <- leveneTest(SALL~SPP, data=raw3))

(LT_SBLL <- leveneTest(SBLL~SPP, data=raw3))

(LT_SBDF <- leveneTest(SBDF~SPP, data=raw3))

(LT_FLA <- leveneTest(FLA~SPP, data=raw3))



```

## PCA analysis 

### PCA {.tabset .tabset-fade .tabset-pills}

LOGAN: CHECK THAT EACH VARIABLES IS NEAR NORMALLY DISTRIBUTED. IF NOT, LOG TRANSFORM BEFORE PCA. ALSO CHECK THAT PCA CALCULATES Z SCORES AND PLOTS BASED ON THAT; IF NOT CONVERT TO Z SCORES THEN RUN PCA.

In this analysis, I will compare the principle components after centering and scaling the data. A PCA analysis will help us determine what aspects of morphology influence the variation in our data the most without worrying about differences in scales/measurements. Currently, data consists of **116 Sailfin and 186 Amazon**.

#### Chart

```{r}
(chart <- matrix(c("Variable chart:", "D: dorsal ray count", "P1: left pectoral ray count", "P2.R: right pelvic rays", "P2.L: left pelvic rays", "A: anal rays", "P1.R: right pectoral ray count", "LLSC: lateral line scale count", "SALL: scalesabove lateral line", "SBLL: scalesbelow lateral line", "SBDF: scalesbefore dorsal fin", "TL: total length", "SL: standard length", "BD: body depth", "CPD: caudal peduncle depth", "CPL: caudal peduncle length", "PreDL: pre-dorsal length", "DbL: dorsal base length", "HL/HW/HD: head length/width/depth", "SnL: snout length", "OL: ocular length"), ncol=1, byrow=TRUE))
```

#### Loadings

```{r}

library(stringr)

raw1a <- subset(raw1, select = -c(17:18) ) #took out P2R and L since they don't vary at all
raw1a <- subset(raw1a, select=-c(34:49))#took out transformed values
raw1a <- raw1a[raw1a$SPP !="p.mexicana", ]

#PCA <- prcomp(raw1a[, 15:32], center=TRUE, scale. = TRUE) #includes all but pelvic traits, since they are the same in all species WILL SCALE/CENTER DATA ON MY OWN, LOGAN SAID IT IS NOT ALWAYS DOING IT FOR YOU EVEN WHEN YOU SPECIFY

z.scores <- scale(raw1a[, 15:33], center = TRUE, scale = TRUE)

raw1a <- cbind(raw1a, z.scores)

colnames(raw1a)[34:52] <- str_c( "z_", colnames(raw1a)[15:33] )

PCA <- prcomp(raw1a[, 34:52])

summary(PCA)

(loadings <- PCA$rotation[, 1:5])


VM_PCA <- varimax(PCA$rotation[, 1:3]) 

VM_PCA


library(factoextra)

(eig.val <- get_eigenvalue(PCA))

ind <- get_pca_ind(PCA)
head(ind$coord[,1:4])


raw1.p <- raw1[raw1$SPP !="p.mexicana", ]
raw1.p <- cbind(raw1.p, ind$coord[,1:4])


```


#### Plots


##### PCA 1v2


```{r}

library(AMR) 
library(ggplot2)
library(ggfortify)
library(ggbiplot)

wes.yel <- "#E1AF00"
wes.blu <- "#78B7C5"

plot1<- autoplot(PCA, data = raw1a, colour="SPP", loadings=FALSE, loadings.label=FALSE, frame=TRUE, frame.type='norm')+scale_color_manual(values=c("#E1AF00","#78B7C5"))+ scale_fill_manual(values=c("#E1AF00","#78B7C5"))+ ggtitle("PCA Plot of Morphology traits") + theme_classic() 



```


##### PCA 2v3

```{r}

plot5<- autoplot(PCA, x=2, y=3, data = raw1a, colour='SPP', loadings=FALSE, loadings.label=FALSE, frame=TRUE, frame.type='norm')+scale_color_manual(values=c("#E1AF00","#78B7C5"))+ scale_fill_manual(values=c("#E1AF00","#78B7C5"))+ ggtitle("PCA Plot of Morphology traits") + theme_classic()
plot5


```


***



## Coefficient of variation

Steve mentioned that using coefficient of variation would be able to distinguish adaptability of species across the range to local environments vs genetic drift. If CVs are similar across species then this would suggest adaptability, but if they are different then it's genetic drift. 

Since some of my data is raw, some is transformed, and some is corrected for body size, I will calculate a CV for each trait in those three forms (if applicable). Could also potentially take the PC loadings and use those (would have 3 CVs for this, for each of the important PCs). I am pretty sure (judging from morph-comparisons-final), that the pca was done on the z-scores of raw values, not transformed or residuals. 

Doesn't look like there is a particular function, just typing in mean and sd for the column you want. Might make my own function??

coef_var <- function(data_frame){
  


}