---
title: "Filtered_oeciliid-morphology_2026"
author: "Allison Davis"
date: "2026-06-27"
output: 
  html_document:
    toc: TRUE
    toc_depth: 6
    toc_float:
      collapsed: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **PURPOSE**

We aim to understand the interplay of genetic variation and environmental heterogeneity in the persistence of asexual-sexual systems through comparisons of morphological variation. In this study, we compare 22 morphological traits in preserved Amazon molly (*Poecilia formosa*) and its parental species (sailfin molly (*P. latipinna*), Atlantic molly (*P. mexicana*)) from across their native and introduced sympatric range. While we expect morphological variation in all species because of local adaptations we predict the variation of the sexual sailfin and Atlantic species to be higher than the Amazon species due to the presence of genetic recombination.

In part I, we will analyze all processed samples (186 Amazon mollies, 156 female sailfin mollies, and 36 female Atlantic mollies). In part II (**Filtered_poeciliid-morphology-analysis_2026.Rmd**), we will filter for adults (standard length > 32.5mm). For the main analysis, we will focus on Amazon and sailfin mollies only. We will then extract Amazon and sailfin samples from the Tampico, Mexico region to compare to our Atlantic molly samples, as our Atlantic molly samples only originate from Tampico, MX. For our geographic analysis, we will assess variation in Amazons and sailfins only, across both latitudinal-based zones and river drainage basins, as basins are often more reflective of genetic distribution of fishes. *NOTE:* Our coefficient of variation analysis can be found in **COV_final.Rmd**.


***
***

# **PART II: MATURE FISH**

The following analysis will compare variation among adult fish (>32.5mm).After filtering for adults and removing all samples with NA values, the final numbers are: 107 Amazons, 90 sailfins, and 24 Atlantics.

For the main analysis, we will use only Amazons and sailfin; comparisons to Atlantic mollies can be found in the **Tampico MX** section.


## **Data**

All analyses were conducted using R version 4.4.2 

### Load required libraries 

Remove the '#' before `install.packages` if it is the first time installing library. 

```{r, warning=FALSE}

#install.packages("curl")             #v5.2.1
library(curl)

#install.packages("dplyr")            #v1.1.4
library(dplyr)

#install.packages("tidyverse")        #v2.0.0
library(tidyverse)

#install.packages("tidyr")            #v1.3.1
library(tidyr)

#install.packages("gt")               #v0.11.1
library(gt)

#install.packages("factoextra")       #v1.0.7
library(factoextra)

#install.packages("car")              #v3.1-2
library(car)

#install.packages("AMR")              #v2.1.1 
library(AMR)

#install.packages("ggplot2")          #v3.5.1
library(ggplot2)

#install.packages("ggfortify")        #v0.4.17
library(ggfortify)

#install.packages("ggbiplot")         #v0.6.2
library(ggbiplot)

#install.packages("emmeans")          #v2.0.1
library(emmeans)

#install.packages("glmmTMB")          #v1.1.14
library(glmmTMB)

#install.packages("DHARMa")           #v0.4.7
library(DHARMa)

#install.packages("Gifi)              #v1.0-0
library(Gifi)

#install.packages("FSA")              #v0.10.1
library(FSA)

```


### Read in data


```{r}
# OPTION 1: download 'Final_morph-data_2026.csv' into working directory 
#raw <- read.csv("Final_morph-data_2026.csv", header = TRUE, sep = ",", stringsAsFactors = TRUE)
#head(raw)


# OPTION 2: load from Github repository
raw <- curl("https://raw.githubusercontent.com/allisondavis/morphology_analysis/refs/heads/master/Final_morph-data_2026.csv")

raw <- read.csv(raw, header = TRUE, sep = ",", stringsAsFactors = TRUE)


# Removing info we don't need
## Note: Right & left pelvic (P2.L/R) identical for both species in all samples
## Note: Total length (TL) is missing data; will use standard length (SL)
## Note: Sample size too sparse in watersheds (hydro.lv8); geographic analysis will use basin (hydro.lv6; hydroBASIN level 6)
raw1 <- select(raw, -c(INSTITUTE, ID.NUM, LOCATION, LAT, LONG, SAMPLING.DATE, MEASUREMENT.DATE, PHOTO.DATE, TAG.., P2.L, P2.R, TL, HYDRO.LV8 ))


# Remove samples with missing values
raw1 <- raw1[complete.cases(raw1[, 4:23]),] 



# Make data more reader friendly
## Species
raw1$SPP <- factor(raw1$SPP,
                   levels = c("p.latipinna", "p.formosa", "p.mexicana"),
                        labels = c("Sailfin molly", "Amazon molly", "Atlantic molly"))

## Zones
raw1$ZONE <- factor(raw1$ZONE,
                    levels = c("Z1", "Z2", "Z3", "Z4"),
                    labels = c("Zone 1", "Zone 2", "Zone 3", "Zone 4"))

## Basins
raw1$HYDRO.LV6 <- factor(raw1$HYDRO.LV6, 
                         levels = c("7060048310", "7060048490", "7060048620", "7060048630", "7060048870", "7060049130", "7060049270", "7060049500", "7060049830", "7060820630", "7060824100"),
                         labels = c("Brazos-Colorado Coastal", "Lavaca-Guadalupe Coastal", "San Antonio-Guadalupe", "San Antonio-Nueces Coastal", "Upper Southwest Texas Coastal", "Lower Southwest Texas Coastal", "Lower R\u00edo Grande", "R\u00edo San Fernando", "Lagunas de San Andr\u00e9s-Morales", "R\u00edo Tames\u00ed", "R\u00edo Tamu\u00edn"))


raw1 <- raw1 %>%
  rename(Species = SPP,
         Zone = ZONE,
         Basin = HYDRO.LV6)

raw1 <- filter(raw1,SL>=32.5) 

## Trait names
trait.names <- c(
  D = "Dorsal rays",
  P1 = "Left pectoral rays",
  P1.R = "Right pectoral rays",
  A = "Anal rays", 
  LLSC = "Lateral line scale count",
  SALL = "Scales above lateral line",
  SBLL = "Scales below lateral line", 
  SBDF = "Scales before dorsal fin",
  SL = "Standard length", 
  BD = "Body depth", 
  CPD = "Caudal peduncle depth",
  CPL = "Caudal peduncle length",
  PreDL = "Predorsal length",
  DbL = "Dorsal base length",
  HL = "Head length",
  HD = "Head depth",
  HW = "Head width",
  SnL = "Snout length", 
  OL = "Orbit length",
  FLA = "Fluctuating asymmetries")


# Final data frames

## main analysis
raw2 <- raw1[raw1$Species !="Atlantic molly", ]  



## Tampico MX analysis
geo <- raw1[raw1$COUNTY.ST == "Tamaulipas/MX", ]

```


### Summary stats

```{r}

#create long df to easily summarize later
dat.long <- raw2 %>%
  pivot_longer(
    cols = c(D, P1, A, P1.R, LLSC, SALL, SBLL, SBDF, SL, BD, CPD, CPL, PreDL, DbL, HL, HD, HW, SnL, OL, FLA),
    names_to = "Trait",
    values_to = "Value")

#gather summary statistics 
summ.stats <- dat.long %>%
  group_by(Species, Trait) %>%
  summarize(
    mean = mean(Value, na.rm = TRUE),
    sd   = sd(Value, na.rm = TRUE),
    min  = min(Value, na.rm = TRUE),
    med  = median(Value, na.rm = TRUE),
    max  = max(Value, na.rm = TRUE),
    .groups = "drop")


# Species sample sizes
n.df <- raw2 %>%
  count(Species, name = "n")

# Add species label with n
summ.stats2 <- summ.stats %>%
  left_join(n.df, by = "Species") %>%
  mutate(
    Species_n = paste0(Species, " (n = ", n, ")"),
    Trait_name = trait.names[Trait],
    Trait_name = factor(
      Trait_name,
      levels = unname(trait.names))) %>%
  arrange(Species_n, Trait_name) %>%
  select(Species_n, Trait_name, mean, sd, min, med, max)


#final table
summ.stats.all <- summ.stats2 %>%
  gt(
    rowname_col = "Trait_name",
    groupname_col = "Species_n") %>%
  fmt_number(
    columns = c(mean, sd, min, med, max),
    decimals = 2) %>%
  tab_style(
      style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
      locations = cells_body(columns = c(mean, sd, min, med, max))) %>%
  opt_horizontal_padding(scale=2) %>%
  cols_align(align="center", everything())

summ.stats.all



```


### Normality

We will check the distribution for all metric traits; since meristic traits are count data, normal distributions do not apply. We will use Shapiro-Wilks tests, histograms and QQ plots to determine normality. 


```{r}
vars <- c("SL", "BD", "CPD", "CPL", "PreDL", "DbL", "HL", "HD", "HW", "SnL", "OL")

##### Shapiro-Wilk test #####
shapiro.results <- data.frame(
  Variable = character(),
  W = numeric(),
  p.value = numeric(),
  stringsAsFactors = FALSE)

for(v in vars){
  x <- raw2[[v]]
  
  
    test <- shapiro.test(x)
    shapiro.results <- rbind(shapiro.results,
                             data.frame(Variable = v,
                                        W = test$statistic,
                                        p.value = test$p.value))}


shapiro.results #all significant


##### Histograms #####

par(mfcol = c(6, 2),   
    mar = c(2, 2, 1.5, 0.5), 
    mgp = c(1, 0.4, 0))

for(v in vars){
  
  hist(raw2[[v]],
       main = paste("Histogram of", v),
       xlab = v,
       breaks = 30,
       col = "gray80",
       border = "black")}

    ## all have slight right skew

##### QQ plots #####

par(mfcol = c(6, 2),   
    mar = c(2, 2, 1.5, 0.5), 
    mgp = c(1, 0.4, 0))

for(v in vars){
  x <- raw2[[v]]
  
 qqnorm(x,
         main = paste("QQ Plot of", v),
         pch = 16, cex = 0.6)
  qqline(x, lwd = 1.5)}


    ## most have slight upward trend towards both ends


```


**CONCLUSIONS:** All metric traits fail Shapiro-Wilks test. We will log transform these values in our models, as log transformations greatly improve normality (see below).

#### *Log transformation*

```{r}

#log transform metric traits
df_trans <- raw2
df_trans[, c(13:23)] <- log(df_trans[, c(13:23)])


#recheck normality

##### Shapiro-Wilk test #####
shapiro.results2 <- data.frame(
  Variable = character(),
  W = numeric(),
  p.value = numeric(),
  stringsAsFactors = FALSE)

for(v in vars){
  x <- df_trans[[v]]
  
  
    test <- shapiro.test(x)
    shapiro.results2 <- rbind(shapiro.results2,
                             data.frame(Variable = v,
                                        W = test$statistic,
                                        p.value = test$p.value))}


shapiro.results2 #HL and OL still significant but greatly improved


##### Histograms #####

par(mfcol = c(6, 2),   
    mar = c(2, 2, 1.5, 0.5), 
    mgp = c(1, 0.4, 0))

for(v in vars){
  
  hist(df_trans[[v]],
       main = paste("Histogram of", v),
       xlab = v,
       breaks = 30,
       col = "gray80",
       border = "black")}

    ## no longer have right skew

##### QQ plots #####

par(mfcol = c(6, 2),   
    mar = c(2, 2, 1.5, 0.5), 
    mgp = c(1, 0.4, 0))

for(v in vars){
  x <- df_trans[[v]]
  
 qqnorm(x,
         main = paste("QQ Plot of", v),
         pch = 16, cex = 0.6)
  qqline(x, lwd = 1.5)}


    ## no longer have upward trend towards both ends

```

***
***




## **Analysis**

### Individual trait models

We are interested in whether there is a difference in morphological variance among the asexual Amazon molly and the sexual sailfin mollies. We will therefore need to use tests that allow for unequal variances among species, and see if this better fits the data compared to equal variances. We will also include standard length (SL) as a covariate to account for any affect of SL on the trait.

For both metric and meristic traits, we will use a ***generalized linear mixed model (GLMM) analysis*** that allows for different dispersions by species.

For metric traits, we will compare a model with equal variances (mod1 <- glmmTMB(log(trait)~log(SL) * species, family= gaussian)) to one that allows unequal variances among species (mod2 <- glmmTMB(log(trait)~log(SL) * species, dispformula=~species, family= gaussian))). We will compare models using an ANOVA test to determine if incorporating unequal variances among species better fits the data.

For meristic traits, we will first fit a model with a *poisson distribution* (mod1 <- glmmTMB(trait~SL * species, family = poisson)) and check if residuals are overdispersed. If they are, we will fit a model with a *negative binomial distribution* and allowing for different dispersion by species (mod2 <- glmmTMB(trait~SL * species, dispformula =~species, family = nbinom2)).We will compare models using delta AIC if necessary.


**CONCLUSIONS:** All metric traits showed no difference in model support, and are thus best supported by an equal-variance model (aka no significant difference in variation among the species). All meristic traits fit with a poisson model and none showed overdispersion. All effects (SL, Species, SL:Species) were non-significant for all meristic traits. No negative binomial model needed. 

```{r}

#create separate df for metric and meristic (count) data
metric <- raw2 %>%
  select(-c(5:12)) %>%
  select(-c("FLA"))

count <- raw2 %>%
  select(-c(14:23)) #still includes SL for analysis

met.vars <- c("BD", "CPD", "CPL", "PreDL", "DbL", "HL", "HD", "HW", "SnL", "OL")
cnt.vars <- c("D", "P1", "P1.R", "A", "LLSC", "SALL", "SBLL", "SBDF", "FLA")


# ───────────────────────────────────────────────────
## METRIC ANALYSIS 
# ───────────────────────────────────────────────────

# Ensure Species is a factor
metric$Species <- as.factor(metric$Species)
metric$Species <- relevel(metric$Species, ref = "Amazon molly")

results_list <- lapply(met.vars, function(trait) {
  
  # Build the formula string correctly
  form_str <- paste0("log(", trait, ") ~ log(SL) * Species")
  form <- as.formula(form_str)

  # 1. Homogeneous variance model
  mod1 <- glmmTMB(form, data = metric, family = gaussian)
  
  # 2. Heterogeneous variance model
  ## NOTE: dispformula = ~Species allows variance to differ by species
  mod2 <- glmmTMB(form, data = metric, family = gaussian, dispformula = ~Species)
  
  
  
  # Compare models
  comp <- anova(mod1, mod2)
  pval <- comp$`Pr(>Chisq)`[2]
  delta_aic <- AIC(mod2) - AIC(mod1)
  
  
  # Extract predicted log-variance per species
  rg <- ref_grid(mod2, component = "disp")
  disp_means <- as.data.frame(emmeans(rg, specs = ~ Species))
  
  # Back-transform to actual Variance (sigma^2)
  disp_means$Variance <- exp(disp_means$emmean)
  
  # Pivot to easily calculate the single global ratio
  ratio_df <- disp_means %>%
    select(Species, Variance) %>%
    pivot_wider(names_from = Species, values_from = Variance) %>%
    mutate(Var_Ratio = `Sailfin molly` / `Amazon molly`) %>%
    mutate(Trait = trait, Delta_AIC = delta_aic, LRT_p = pval)
  
  return(ratio_df)
})

results_df <- do.call(rbind, results_list)

results_df$p_adj <- p.adjust(results_df$LRT_p,
                             method = "holm")

met.results <- results_df %>%
  mutate(
    p.display = case_when(
      p_adj < 0.001 ~ paste0("*p* < 0.001*"),
      p_adj < 0.05  ~ paste0(sprintf("%.3f", p_adj), "*"),
      TRUE ~ sprintf("%.3f", p_adj))) %>%
  mutate(
    Trait_Name = factor(
      trait.names[Trait],
      levels = unname(trait.names))) %>%
  arrange(Trait_Name) %>%
  relocate(Trait_Name, .before = everything())

met.tbl <- met.results %>%
  gt() %>%
  cols_hide(c(Trait, LRT_p, p_adj, "Sailfin molly", "Amazon molly")) %>%
  cols_label(
    Trait_Name = "Trait",
    Var_Ratio = md("Variance ratio <br>(Sailfin / Amazon)"),
    Delta_AIC = md("&Delta;AIC"),
    p.display = md("*p*-value")) %>%
  fmt_number(
    columns = c(Var_Ratio, Delta_AIC),
    decimals = 2) %>%
  fmt_markdown(columns = p.display) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Trait_Name, Delta_AIC, Var_Ratio))) %>%
  cols_align("center", everything()) %>%
  opt_horizontal_padding(scale = 2)
  
  
met.tbl



# ───────────────────────────────────────────────────
## MERISTIC (COUNT) ANALYSIS 
# ───────────────────────────────────────────────────


# Ensure baseline is set
count$Species <- as.factor(count$Species)
count$Species <- relevel(count$Species, ref = "Amazon molly")

poisson_results <- lapply(cnt.vars, function(trait) {
  
  # Build formula: Linear SL is best for count link functions
  form <- as.formula(paste0(trait, " ~ SL * Species"))
  
  # 1. Fit Poisson GLM
  mod <- glmmTMB(form, data = count, family = poisson)
  
  # 2. Run Type III Wald Anova
  # tryCatch handles cases where extreme separation makes the Chidg tests unstable
  anova_res <- tryCatch({
    Anova(mod, type = "III")
  }, error = function(e) { NULL })
  
  # Extract p-values for main effects and interaction safely
  p_sl          <- if(!is.null(anova_res)) anova_res["SL", "Pr(>Chisq)"] else NA
  p_species     <- if(!is.null(anova_res)) anova_res["Species", "Pr(>Chisq)"] else NA
  p_interaction <- if(!is.null(anova_res)) anova_res["SL:Species", "Pr(>Chisq)"] else NA
  
  # 3. DHARMa Overdispersion Diagnostic
  sim_res   <- simulateResiduals(mod, plot = FALSE)
  disp_test <- testDispersion(sim_res, plot = FALSE)
  
  # Extract the raw dispersion ratio (> 1 means overdispersion) 
  # and the significance p-value
  disp_ratio  <- disp_test$statistic
  disp_pval   <- disp_test$p.value
  is_overdisp <- disp_pval < 0.05 & disp_ratio > 1

  data.frame(
    Trait = trait,
    AIC = AIC(mod),
    P_SL = p_sl,
    P_Species = p_species,
    P_Interaction = p_interaction,
    DHARMa_DispRatio = disp_ratio,
    DHARMa_p = disp_pval,
    Overdispersed = is_overdisp,
    row.names = NULL
  )
})

# Combine into a clean data frame
poisson_df <- do.call(rbind, poisson_results)

poisson_corrected <- poisson_df %>%
  mutate(
    # 1. Correct for the species main effect across all traits
    P_Species_adj = p.adjust(P_Species, method = "holm"),
    
    # 2. Correct for the size interaction across all traits
    P_Interaction_adj = p.adjust(P_Interaction, method = "holm"),
    
    # 3. Correct for the standard length effect across all traits
    P_SL_adj = p.adjust(P_SL, method = "holm")
  )

cnt.results <- poisson_corrected %>%
  mutate(
    across(
      .cols = c(P_SL_adj, P_Species_adj, P_Interaction_adj),
      .names = "{.col}_display", 
      .fns = ~ case_when(
        .x < 0.001 ~ "*p* < 0.001*",
        .x < 0.05  ~ paste0(sprintf("%.3f", .x), "*"),
        TRUE       ~ sprintf("%.3f", .x)))) %>%
  mutate(
    Trait_Name = factor(
      trait.names[Trait],
      levels = unname(trait.names))) %>%
  arrange(Trait_Name) %>%
  relocate(Trait_Name, .before = everything())

cnt.tbl <- cnt.results %>%
  gt() %>%
  cols_hide(c(Trait, P_SL, P_Species, P_Interaction, P_SL_adj, P_Species_adj, P_Interaction_adj, DHARMa_p)) %>%
  cols_label(
    Trait_Name = "Trait",
    AIC= "Model AIC",
    DHARMa_DispRatio = "Dispersion Ratio",
    Overdispersed = "Is overdispersed \n(True/False)",
    P_SL_adj_display = md("*p*-value for SL"),
    P_Species_adj_display = md("*p*-value for Species"),
    P_Interaction_adj_display = md("*p*-value for SL:Species")) %>%
  fmt_number(
    columns = c(DHARMa_DispRatio, AIC),
    decimals = 2) %>%
  fmt_markdown(columns = starts_with("P_")) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Trait_Name, AIC, DHARMa_DispRatio, Overdispersed, P_SL_adj_display, P_Species_adj_display))) %>%
  cols_align("center", everything()) %>%
  opt_horizontal_padding(scale = 2)
  


cnt.tbl

  
# ───────────────────────────────────────────────────
## PLOTS
# ───────────────────────────────────────────────────

met.long <- metric %>%
  pivot_longer(
    cols = c(SL, BD, CPD, CPL, PreDL, DbL, HL, HD, HW, SnL, OL),
    names_to = "Trait",
    values_to = "Value")


met.plots <- ggplot(met.long, aes(x=Species, y=Value, fill = Species)) + 
  geom_boxplot() +
  geom_point(color= "#8D8680", size=2, alpha=0.5) +
  facet_wrap(~ Trait, scales = "free_y", axes = "margins") +
  theme_classic(base_size = 10, base_family = "sans") + 
  theme(
    axis.title.y = element_text(size = 10, family = "sans"), 
    axis.text.y  = element_text(size = 10, family = "sans"),
    axis.text.x  = element_text(size = 10, family = "sans"),
    strip.text    = element_text(size = 10, family = "sans"),
    panel.spacing = unit(1.5, "lines")) +
  scale_fill_manual(values = c("Sailfin molly"="#78B7C5",
                               "Amazon molly"="#E1AF00")) +
  xlab("") +
  ylab("Length or Depth (mm)")
met.plots


cnt.long <- count %>%
  pivot_longer(
    cols = c(D, P1, A, P1.R, LLSC, SALL, SBLL, SBDF, SL, FLA),
    names_to = "Trait",
    values_to = "Value")

cnt.plots <- ggplot(cnt.long, aes(x=Species, y=Value, fill = Species)) + 
  geom_boxplot() +
  geom_point(color= "#8D8680", size=2, alpha=0.5) +
  facet_wrap(~ Trait, scales = "free_y", axes = "margins") +
  theme_classic(base_size = 10, base_family = "sans") + 
  theme(
    axis.title.y = element_text(size = 10, family = "sans"), 
    axis.text.y  = element_text(size = 10, family = "sans"),
    axis.text.x  = element_text(size = 10, family = "sans"),
    strip.text    = element_text(size = 10, family = "sans"),
    panel.spacing = unit(1.5, "lines")) +
  scale_fill_manual(values = c("Sailfin molly"="#78B7C5",
                               "Amazon molly"="#E1AF00")) +
  xlab("") +
  ylab("Count")
cnt.plots


```

***
***

### Overall shape variation (PCA)

While none our 10 metric traits showed a significant difference in variation among species, many of these traits are not independent of each other. Therefore, we will assess overall shape variation among species using a ***principal component analysis (PCA)*** on centered and scaled data (all metric traits will also be log transformed to better approach normality). We will then perform a mean and variance test (e.g., ANOVA and Levene's test) to determine if overall shape variation differs between the species. 

As meristic traits are count data, a normal PCA cannot be applied. Instead, we will use a ***categorical principal component analysis (CATPCA)*** that uses optimal scaling to deal with discrete integers. 

#### *Loadings*

```{r}

# ───────────────────────────────────────────────────
## METRIC ANALYSIS 
# ───────────────────────────────────────────────────


PCA.test <- prcomp(df_trans[, 13:23], center = TRUE, scale. = FALSE) 

summary(PCA.test)
(eig.val.test <- get_eigenvalue(PCA.test))


# For many morphological studies, the first component is heavily weighted by body size and is often of less interest than components two and three (Humphries et al., 1981). Therefore, we will extract the first three dimensions

ind.test <- get_pca_ind(PCA.test)
head(ind.test$coord[,1:3])


pca.df <- cbind(df_trans, ind.test$coord[,1:3])


# Will investigate the influence of each trait on the first 3 dimensions
(loadings <- PCA.test$rotation[, 1:3])

sorted.loadings.1 <- loadings[order(loadings[, 1]), 1]
myTitle <- "Loadings Plot for PC1" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.1, main=myTitle, xlab=myXlab, cex=0.5, col="blue")

sorted.loadings.2 <- loadings[order(loadings[, 2]), 2]
myTitle <- "Loadings Plot for PC2" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.2, main=myTitle, xlab=myXlab, cex=0.5, col="blue")

sorted.loadings.3 <- loadings[order(loadings[, 3]), 3]
myTitle <- "Loadings Plot for PC3" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.3, main=myTitle, xlab=myXlab, cex=0.5, col="blue")




# ───────────────────────────────────────────────────
## MERISTIC (COUNT) ANALYSIS 
# ───────────────────────────────────────────────────

# Remove SL from count df
count2 <- select(count, -c("SL"))

# Run the Categorical PCA
# 'ndim = 2' extracts the first two principal components
# 'degrees = 1' treats the scaling transformations as piecewise linear (ideal for ordinal integers)

catpca.test <- princals(count2[, 5:13], ndim = 2, degrees = 1)

# Print the summary to see how much variance your new components explain
summary(catpca.test)

(cat.loadings <- catpca.test$loadings)

sorted.loadings.4 <- cat.loadings[order(cat.loadings[, 1]), 1]
myTitle <- "Loadings Plot for CATPC1" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.4, main=myTitle, xlab=myXlab, cex=0.5, col="blue")

sorted.loadings.5 <- cat.loadings[order(cat.loadings[, 2]), 2]
myTitle <- "Loadings Plot for CATPC2" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.5, main=myTitle, xlab=myXlab, cex=0.5, col="blue")

# Extract the eigenvalues (one per dimension)
eigenvals <- catpca.test$evals

# Calculate the percentage of variation explained by each axis
# (Divided by 9 because you have 9 meristic traits)
pc1_percent <- round((eigenvals[1] / 9) * 100, 1)
pc2_percent <- round((eigenvals[2] / 9) * 100, 1)

# Quick print check to see your numbers (e.g., 24.5 and 14.2)
print(paste0("CatPC1: ", pc1_percent, "%"))
print(paste0("CatPC2: ", pc2_percent, "%"))

# 1. Extract the object scores for the first two dimensions
cat_pc_scores <- as.data.frame(catpca.test$objectscores)
colnames(cat_pc_scores) <- c("CatPC1", "CatPC2")

# 2. Combine them back with your species and size metadata
catpca.df <- cbind(count, cat_pc_scores)

```

#### *Plots*

```{r}

#### METRIC

pca.1<- autoplot(PCA.test, data = df_trans, colour="Species", shape= 'Species', loadings=FALSE, loadings.label=FALSE, frame=TRUE, frame.type='norm', label= FALSE)+
  scale_color_manual(name="Species",
                    values = c("Sailfin molly"="#78B7C5", "Amazon molly"="#E1AF00")) +
  scale_fill_manual(name="Species",
                    values = c("Sailfin molly"="#78B7C5", "Amazon molly"="#E1AF00")) +
  scale_shape_manual(name= 'Species', 
                     values = c("Sailfin molly" = 17, "Amazon molly"=16)) +
   guides(fill = guide_legend(override.aes = list(fill = NA)))+
  ggtitle("PCA Plot of Morphology traits") + theme_classic() 
pca.1


pca.2<- autoplot(PCA.test, x=2, y=3, data = df_trans, colour="Species", shape= 'Species', loadings=FALSE, loadings.label=FALSE, frame=TRUE, frame.type='norm', label= FALSE)+
  scale_color_manual(name="Species",
                    values = c("Sailfin molly"="#78B7C5", "Amazon molly"="#E1AF00")) +
  scale_fill_manual(name="Species",
                    values = c("Sailfin molly"="#78B7C5", "Amazon molly"="#E1AF00")) +
  scale_shape_manual(name= 'Species', 
                     values = c("Sailfin molly" = 17, "Amazon molly"=16)) +
   guides(fill = guide_legend(override.aes = list(fill = NA)))+
  ggtitle("PCA Plot of Morphology traits") + theme_classic() 
pca.2


#### MERISTIC

catpca <- ggplot(catpca.df, aes(x = CatPC1, y = CatPC2, color = Species, fill = Species, shape = Species)) +
  geom_point(size = 1, alpha = 0.8) +
  stat_ellipse(geom = "polygon", alpha = 0.15, type = "norm", level = 0.95) +
  scale_color_manual(name = "Species",
                     values = c("Sailfin molly" = "#78B7C5", "Amazon molly" = "#E1AF00")) +
  scale_fill_manual(name = "Species",
                    values = c("Sailfin molly" = "#78B7C5", "Amazon molly" = "#E1AF00")) +
  scale_shape_manual(name = "Species", 
                     values = c("Sailfin molly" = 17, "Amazon molly" = 16)) +
  guides(fill = guide_legend(override.aes = list(fill = NA))) +
  labs(
    title = "CATPCA Plot of Meristic Traits",
    x = "Categorical PC1",
    y = "Categorical PC2"
  ) +
  theme_classic() +
  labs(
    title = "CATPCA Plot of Meristic Traits",
    x = paste0("Categorical PC1 (", pc1_percent, "%)"),
    y = paste0("Categorical PC2 (", pc2_percent, "%)")
  )

catpca

```


#### *Post PCA tests*

**CONCLUSIONS:** For our metric data, there is no significant difference in average shape distribution for PC1, but there is a significant difference for PC2 and PC3. This is expected, as amazons and sailfins have different body shapes and therefore occupy distinct morphospace. There was no significant difference in shape variation for any PC, indicating not overall difference in morphological variation among the two species when all traits are taken into account. This corroborates our individual results.

For our meristic data, CatPC1 showed a significant difference in average meristic profiles, which makes sense as the majority of traits that load heavily on CatPC1 are species diagnostics (e.g., scales before dorsal fin, lateral line scale count, dorsal fin rays). There is also a significant difference in variation for CatPC1, with sailfins exhibiting higher variation than amazons. There was no significant difference in CatPC2 for both mean and variance. This differs from our individual trait analysis, indicating species differences in covariation among multiple meristic features. 


```{r}

#get species-only df for metric and meristic data
lat.m <- pca.df[pca.df$Species == "Sailfin molly",]
form.m <- pca.df[pca.df$Species == "Amazon molly",]

lat.c <- catpca.df[catpca.df$Species == "Sailfin molly",]
form.c <- catpca.df[catpca.df$Species == "Amazon molly",]

#quick distribution checks to see what tests to use

testmod <- lm(Dim.1 ~ Species, data = pca.df)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(Dim.2 ~ Species, data = pca.df)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(Dim.3 ~ Species, data = pca.df)
plot(simulateResiduals(testmod)) #fails outlier test; will use wilcox

testmod <- lm(CatPC1 ~ Species, data = catpca.df)
plot(simulateResiduals(testmod)) #levene's test significant; will use wilcox

testmod <- lm(CatPC2 ~ Species, data = catpca.df)
plot(simulateResiduals(testmod)) #fails KS test, outlier test, and within-group deviation; will use wilcox




# ───────────────────────────────────────────────────
## MEDIAN TESTS (Wilcox Test for all)
# ───────────────────────────────────────────────────

### Metric Data

met.mean.1 <- wilcox.test(lat.m$Dim.1, form.m$Dim.1, alternative = "two.sided", exact = FALSE, conf.int = TRUE) #not significant

met.mean.2 <- wilcox.test(lat.m$Dim.2, form.m$Dim.2, alternative = "two.sided", exact = FALSE, conf.int = TRUE) #significant (p<0.001); estimate = 0.286
    (median(lat.m$Dim.2))    #  0.152
    (median(form.m$Dim.2))   # -0.123


met.mean.3 <- wilcox.test(lat.m$Dim.3, form.m$Dim.3, alternative = "two.sided", exact = FALSE, conf.int = TRUE) #significant (p=0.008); estimate= 0.040
    (median(lat.m$Dim.3))    #  0.010
    (median(form.m$Dim.3))   # -0.025


### Meristic Data


cnt.mean.1 <- wilcox.test(lat.c$CatPC1, form.c$CatPC1, alternative = "two.sided", exact = FALSE, conf.int = TRUE) #significant (p<0.001); estimate = 1.75
    (median(lat.c$CatPC1))    #  1.0
    (median(form.c$CatPC1))   # -0.797

cnt.mean.2 <- wilcox.test(lat.c$CatPC2, form.c$CatPC2, alternative = "two.sided", exact = FALSE, conf.int = TRUE) #significant (p=0.039); estimate = 0.052
    (median(lat.c$CatPC2))    #  -0.203
    (median(form.c$CatPC2))   # -0.246

# ───────────────────────────────────────────────────
## VARIANCE TESTS (Brown-Forsythe for all)
# ───────────────────────────────────────────────────

### Metric Data
met.var.1 <- leveneTest(Dim.1~Species, data= pca.df) #not significant

met.var.2 <- leveneTest(Dim.2~Species, data= pca.df) #not significant

met.var.3 <- leveneTest(Dim.3~Species, data= pca.df) #not significant


### Meristic Data
cnt.var.1 <- leveneTest(CatPC1~Species, data= catpca.df) #significant (0.050)
    (var(lat.c$CatPC1))   # 0.289
    (var(form.c$CatPC1))  # 0.238

cnt.var.2 <- leveneTest(CatPC2~Species, data= catpca.df) #not significant


# ───────────────────────────────────────────────────
## TABLE
# ───────────────────────────────────────────────────

### Metric Data

#get mean and variance for each dimension per species
met_summary <- pca.df %>%
  group_by(Species) %>%
  summarise(
    PC1_median = median(Dim.1, na.rm = TRUE),
    PC1_mad  = mad(Dim.1, na.rm = TRUE, constant = 1),
    PC2_median = median(Dim.2, na.rm = TRUE),
    PC2_mad  = mad(Dim.2, na.rm = TRUE, constant = 1),
    PC3_median = median(Dim.3, na.rm = TRUE),
    PC3_mad  = mad(Dim.3, na.rm = TRUE, constant = 1))

#isolate results
amazon_met  <- filter(met_summary, Species == "Amazon molly")
sailfin_met <- filter(met_summary, Species == "Sailfin molly")

#extract wilcox test results, levene's results, and summary stats
postPCA_result <- tibble(
  Classifier = "Metric PCA",
  Dimension = c("PC1", "PC2", "PC3"),
  Amazon = c(
    sprintf("%.2f (%.2f)", amazon_met$PC1_median, amazon_met$PC1_mad),
    sprintf("%.2f (%.2f)", amazon_met$PC2_median, amazon_met$PC2_mad),
    sprintf("%.2f (%.2f)", amazon_met$PC3_median, amazon_met$PC3_mad)),
  Sailfin = c(
    sprintf("%.2f (%.2f)", sailfin_met$PC1_median, sailfin_met$PC1_mad),
    sprintf("%.2f (%.2f)", sailfin_met$PC2_median, sailfin_met$PC2_mad),
    sprintf("%.2f (%.2f)", sailfin_met$PC3_median, sailfin_met$PC3_mad)),
  W_value = c(met.mean.1$statistic,
              met.mean.2$statistic,
              met.mean.3$statistic),
  p_wil = c(met.mean.1$p.value,
           met.mean.2$p.value,
           met.mean.3$p.value),
  F_value = c(met.var.1$`F value`[1],
              met.var.2$`F value`[1],
              met.var.3$`F value`[1]),
  p_lv = c(met.var.1$`Pr(>F)`[1],
           met.var.2$`Pr(>F)`[1],
           met.var.3$`Pr(>F)`[1])) %>%
  mutate(
    Med_test = case_when(
      p_wil < 0.001 ~ paste0("W = ", sprintf("%.2f", W_value), 
                            "<br>*p* < 0.001*"),
      p_wil <0.05   ~ paste0("W = ", sprintf("%.2f", W_value), 
                            "<br>*p* = ", sprintf("%.3f", p_wil), "*"),
       TRUE        ~ paste0("W = ", sprintf("%.2f", W_value),
                            "<br>*p* = ", sprintf("%.3f", p_wil))),

    Variance_test = case_when(
      p_lv < 0.001 ~ paste0("F = ", sprintf("%.2f", F_value), 
                            "<br>*p* < 0.001*"),
      p_lv <0.05   ~ paste0("F = ", sprintf("%.2f", F_value), 
                            "<br>*p* = ", sprintf("%.3f", p_lv), "*"),
       TRUE        ~ paste0("F = ", sprintf("%.2f", F_value),
                            "<br>*p* = ", sprintf("%.3f", p_lv))))




### Meristic Data

#get mean and variance for each dimension per species
cnt_summary <- catpca.df %>%
  group_by(Species) %>%
  summarise(
    CPC1_median = median(CatPC1, na.rm = TRUE),
    CPC1_mad  = mad(CatPC1, na.rm = TRUE),
    CPC2_median = median(CatPC2, na.rm = TRUE),
    CPC2_mad  = mad(CatPC2, na.rm = TRUE))

#isolate results
amazon_cnt  <- filter(cnt_summary, Species == "Amazon molly")
sailfin_cnt <- filter(cnt_summary, Species == "Sailfin molly")


#extract wilcox test results, levene's results, and summary stats
postCPCA_result <- tibble(
  Classifier = "Meristic CATPCA",
  Dimension = c("CatPC1", "CatPC2"),
  Amazon = c(
    sprintf("%.2f (%.2f)", amazon_cnt$CPC1_median, amazon_cnt$CPC1_mad),
    sprintf("%.2f (%.2f)", amazon_cnt$CPC2_median, amazon_cnt$CPC2_mad)),
  Sailfin = c(
    sprintf("%.2f (%.2f)", sailfin_cnt$CPC1_median, sailfin_cnt$CPC1_mad),
    sprintf("%.2f (%.2f)", sailfin_cnt$CPC2_median, sailfin_cnt$CPC2_mad)),
  W_value = c(cnt.mean.1$statistic,
              cnt.mean.2$statistic),
  p_wil = c(cnt.mean.1$p.value,
           cnt.mean.2$p.value),
  F_value = c(cnt.var.1$`F value`[1],
              cnt.var.2$`F value`[1]),
  p_lv = c(cnt.var.1$`Pr(>F)`[1],
           cnt.var.2$`Pr(>F)`[1])) %>%
  mutate(
    Med_test = case_when(
      p_wil < 0.001 ~ paste0("W = ", sprintf("%.2f", W_value), 
                            "<br>*p* < 0.001*"),
      p_wil <0.05   ~ paste0("W = ", sprintf("%.2f", W_value), 
                            "<br>*p* = ", sprintf("%.3f", p_wil), "*"),
       TRUE        ~ paste0("W = ", sprintf("%.2f", W_value),
                            "<br>*p* = ", sprintf("%.3f", p_wil))),

    Variance_test = case_when(
      p_lv < 0.001 ~ paste0("F = ", sprintf("%.2f", F_value), 
                            "<br>*p* < 0.001*"),
      p_lv <0.05   ~ paste0("F = ", sprintf("%.2f", F_value), 
                            "<br>*p* = ", sprintf("%.3f", p_lv), "*"),
       TRUE        ~ paste0("F = ", sprintf("%.2f", F_value),
                            "<br>*p* = ", sprintf("%.3f", p_lv))))



### Combine & form table

results_combined <- bind_rows(postPCA_result, postCPCA_result) 



postPCA_table <- results_combined %>%
  select(
    Classifier,
    Dimension,
    Amazon,
    Sailfin,
    Med_test,
    Variance_test) %>%
  gt(groupname_col = "Classifier") %>%
  fmt_markdown(
    columns = c(Med_test, Variance_test)) %>%
  cols_label(
    Dimension = "         ",
    Amazon = md("**Amazon**<br>Median (MAD)"),
    Sailfin = md("**Sailfin**<br>Median (MAD)"),
    Med_test = "Median comparison",
    Variance_test = "Variance comparison") %>%
  tab_options(row_group.font.weight = "bold") %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c(Dimension, Amazon, Sailfin, Med_test))) %>%
  cols_align(align = "center", columns = everything()) %>%
  opt_horizontal_padding(scale = 2)

#final table
postPCA_table



```

***
***



## **Geographic Analysis**


### Quick summary

A quick summary, mostly of number of fish in each zone and basin.

We find that there are at least 3 fish of each species in all zones, 7 of 11 basins, and 8 of 17 basins. Classifiers marked with an asterisk only have >3 of one species.

```{r, echo=FALSE, warning=FALSE}

zone.sum <-raw2 %>%
  count(Species, Zone) %>%
  tidyr::pivot_wider(
    names_from  = Species,
    values_from = n,
    values_fill = 0) %>%
  rename(classifier = Zone) %>%
  mutate(classifier = as.character(classifier), 
         classifier_type = "Zone")

basin.sum <- raw2 %>%
  count(Species, Basin) %>%
  tidyr::pivot_wider(
    names_from  = Species,
    values_from = n,
    values_fill = 0) %>%
  rename(classifier = Basin) %>%
  mutate(classifier = as.character(classifier), 
         classifier_type = "Basin")


sum.counts <- bind_rows(zone.sum, basin.sum)

sum.counts %>%
  group_by(classifier_type) %>%
  mutate(classifier_type = paste0(classifier_type, " (n=", n(), ")")) %>%
  ungroup() %>%
  mutate(classifier_type = factor(classifier_type,
                                  levels = grep("Zone", unique(classifier_type), value = TRUE) %>%
                                           c(grep("Basin", unique(classifier_type), value = TRUE)) %>%
                                           c(grep("Watershed", unique(classifier_type), value = TRUE))))%>%
  arrange(classifier_type, classifier) %>%
  select(classifier_type, classifier, everything()) %>%
  gt(rowname_col = "classifier", groupname_col = "classifier_type") %>%
  text_transform(
    locations = cells_stub(),
    fn = function(x) {
      too.few <- sum.counts$`Amazon molly` < 3 | sum.counts$`Sailfin molly` < 3
      
      ifelse(sum.counts$classifier %in% x & too.few[match(x, sum.counts$classifier)], 
             paste0(x, "*"), x)}) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c(classifier, "Sailfin molly"))) %>%
  opt_horizontal_padding(scale=2) %>%
  cols_align(align="center", everything())



```
***
***

### Individual trait models

We will be repeating our **generalized linear mixed models (GLMM)** with the addition of zones and basins to test if variation is partitioned across a geographic range. 

#### *Zones*

**CONCLUSIONS:** Five metric traits showed greater support for a model with unequal dispersion across species and zones. Caudal peduncle depth/length, and orbital length mostly show higher variation among sailfins with highest variation in zone 4. Head width/length show amazons to be more variable in all zones except zone 2, where sailfins are more variable. No meristic traits differed by zone.

```{r}
# ───────────────────────────────────────────────────
## METRIC ANALYSIS--ZONES
# ───────────────────────────────────────────────────


results_geo <- lapply(met.vars, function(trait) {
  
  # Build the formula string correctly
  form_str <- paste0("log(", trait, ") ~ log(SL) * Species")
  form <- as.formula(form_str)

  # 1. Species variance model
  mod1 <- glmmTMB(form, data = metric, family = gaussian, dispformula = ~Species)
  
  # 2. Geographic variance model
  mod2 <- glmmTMB(form, data = metric, family = gaussian, dispformula = ~Species*Zone)
  
  
  
  # Compare models
  comp <- anova(mod1, mod2)
  pval <- comp$`Pr(>Chisq)`[2]
  delta_aic <- AIC(mod2) - AIC(mod1)

  
  # Extract predicted log-variance (dispersion) per group
  rg <- ref_grid(mod2, component = "disp")
  disp_means <- as.data.frame(emmeans(rg, specs = ~ Species * Zone))
  
  # Back-transform from log scale to get actual Variance (sigma^2)
  disp_means$Variance <- exp(disp_means$emmean)
  
  
  ratios_by_zone <- disp_means %>%
    select(Species, Zone, Variance) %>%
    pivot_wider(names_from = Species, values_from = Variance) %>%
    mutate(Var_Ratio = `Sailfin molly` / `Amazon molly`) %>% 
    mutate(Trait = trait, Delta_AIC = delta_aic, LRT_p = pval)
  
  return(ratios_by_zone)

})

geo_results_df <- do.call(rbind, results_geo)


p_adjusted_df <- geo_results_df %>%
  select(Trait, LRT_p, Delta_AIC) %>%
  distinct() %>%
  mutate(p_adj = p.adjust(LRT_p, method = "holm")) %>%
  mutate(
    p.display = case_when(
      p_adj < 0.001 ~ "*p* < 0.001*",
      p_adj < 0.05  ~ paste0(sprintf("%.3f", p_adj), "*"),
      TRUE          ~ sprintf("%.3f", p_adj)),
    Trait_Name = factor(trait.names[Trait], levels = unname(trait.names)))

ratios_wide <- geo_results_df %>%
  select(Trait, Zone, Var_Ratio) %>%
  pivot_wider(
    names_from = Zone, 
    values_from = Var_Ratio, 
    names_prefix = "Ratio_")

geo_wide_results <- p_adjusted_df %>%
  left_join(ratios_wide, by = "Trait") %>%
  arrange(Trait_Name)

geo.met.tbl <- geo_wide_results %>%
  gt() %>%
  cols_hide(c(Trait, LRT_p, p_adj)) %>%
  cols_move_to_start(columns = Trait_Name) %>%
  cols_label(
    Trait_Name = "Trait",
    Delta_AIC = md("&Delta;AIC"),
    p.display = md("*p*-value"),
    `Ratio_Zone 1` = "Zone 1",  
    `Ratio_Zone 2` = "Zone 2",  
    `Ratio_Zone 3` = "Zone 3",
    `Ratio_Zone 4` = "Zone 4") %>%
  tab_spanner(
    label = md("Variance Ratio *(Sailfin / Amazon)*"),
    columns = starts_with("Ratio_")) %>%
  fmt_number(
    columns = c(Delta_AIC, starts_with("Ratio_")),
    decimals = 2) %>%
  fmt_markdown(columns = p.display) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Trait_Name, p.display))
  ) %>%
  cols_align("center", everything()) %>%
  opt_horizontal_padding(scale = 2)

geo.met.tbl


# ───────────────────────────────────────────────────
## MERISTIC (COUNT) ANALYSIS--ZONES
# ───────────────────────────────────────────────────


poisson_geo <- lapply(cnt.vars, function(trait) {
  
  # 1. Build the formula for the count mean
  form <- as.formula(paste0(trait, " ~ SL * Species"))
  
  # To safely extract variance ratios per zone, we can look at the 
  # residual dispersion factor for each Species X Zone combination.
  # We do this by splitting the data by Zone, or mapping across them.
  
  zones <- unique(count$Zone)
  
  zone_ratios <- lapply(zones, function(z) {
    # Subset data for just this zone
    sub_data <- subset(count, Zone == z)
    
    # Fit a robust Quasipoisson GLM for this zone
    # This automatically estimates a custom dispersion parameter for the species mix
    mod_sub <- glm(form, data = sub_data, family = quasipoisson)
    
    # To get separate variances/dispersions for each species within this zone:
    # We look at the pearson residuals per species level
    sub_data$resid_sq <- (residuals(mod_sub, type = "pearson"))^2
    
    # The mean of squared Pearson residuals per group gives its dispersion factor (phi)
    phi_estimates <- aggregate(resid_sq ~ Species, data = sub_data, FUN = mean)
    
    # Extract values safely based on your level names
    phi_amazon  <- phi_estimates$resid_sq[phi_estimates$Species == "Amazon molly"]
    phi_sailfin <- phi_estimates$resid_sq[phi_estimates$Species == "Sailfin molly"]
    
    # If a group is empty or missing, safeguard it
    var_ratio <- if(length(phi_amazon) > 0 & length(phi_sailfin) > 0) {
      phi_sailfin / phi_amazon
    } else {
      NA
    }
    
    data.frame(Zone = z, Var_Ratio = var_ratio)
  })
  
  # Combine zones for this trait
  trait_zones <- do.call(rbind, zone_ratios)
  
  # 2. To get a single global significance p-value for the Zone interaction,
  # we compare a pooled model without a Zone interaction to one with it.
  mod_null_disp <- glm(form, data = count, family = poisson) # Standard Poisson
  
  # Global test for whether counts step outside poisson parameters by Zone
  # We use a standard Chi-sq comparison
  anova_global <- drop1(glm(as.formula(paste0(trait, " ~ SL * Species * Zone")), 
                            data = count, family = poisson), test = "Chisq")
  
  # Extract the interaction p-value safely
  pval <- anova_global["SL:Species:Zone", "Pr(>Chi)"]
  if(is.na(pval)) pval <- anova_global["Species:Zone", "Pr(>Chi)"] # Fallback if size interaction isn't main focus
  
  # Dummy Delta_AIC since Quasipoisson doesn't use true maximum likelihood AIC
  delta_aic <- NA 
  
  # Attach metadata to our zone matrix
  trait_zones$Trait <- trait
  trait_zones$Delta_AIC <- delta_aic
  trait_zones$LRT_p <- pval
  
  return(trait_zones)
})

# Combine results into the long data frame for your gt table
geo_poisson_df <- do.call(rbind, poisson_geo)

geo_poisson_df$p_adj <- p.adjust(geo_poisson_df$LRT_p,
                             method = "holm")

geo.cnt.results <- geo_poisson_df %>%
  mutate(
    p.display = case_when(
      p_adj < 0.001 ~ paste0("*p* < 0.001*"),
      p_adj < 0.05  ~ paste0(sprintf("%.3f", p_adj), "*"),
      TRUE ~ sprintf("%.3f", p_adj))) %>%
  mutate(
    Trait_Name = factor(
      trait.names[Trait],
      levels = unname(trait.names))) %>%
  arrange(Trait_Name, Zone) %>%
  relocate(Trait_Name, Zone, .before = everything())

geo.cnt.results

## ALL OF THAT JUST TO SHOW THAT THERE IS NO DIFFERENCE IN MERISTIC TRAITS ACROSS ZONES; ALL P-VALUES ARE 1




```



#### *Basins*

First need to filter for basins with a minimum of 3 of each species.

```{r}
# Filtering for basins with n>=3 for both species
valid_basins <- raw2 %>%
  group_by(Basin, Species) %>%
  summarise(n = n(), .groups = "drop") %>%
  tidyr::pivot_wider(names_from = Species, values_from = n, values_fill = 0) %>%
  filter(`Amazon molly` >= 3 & `Sailfin molly` >= 3) %>%
  pull(Basin)

metric2 <- metric %>% filter(Basin %in% valid_basins)
count2 <- count %>% filter(Basin %in% valid_basins)



```


**CONCLUSIONS:** All metric traits showed greater support for a model with unequal dispersion across species and basins. There is no consistent pattern across traits or across basins, indicating that local deviations are likely reflecting stochastic, demographic, or sampling-related effects rather than any differences from reproductive mode. Again, no meristic traits differed across basins. 

```{r}
# ───────────────────────────────────────────────────
## METRIC ANALYSIS--BASINS
# ───────────────────────────────────────────────────


results_geo2 <- lapply(met.vars, function(trait) {
  
  # Build the formula string correctly
  form_str <- paste0("log(", trait, ") ~ log(SL) * Species")
  form <- as.formula(form_str)

  # 1. Species variance model
  mod1 <- glmmTMB(form, data = metric2, family = gaussian, dispformula = ~Species)
  
  # 2. Geographic variance model
  mod2 <- glmmTMB(form, data = metric2, family = gaussian, dispformula = ~Species*Basin)
  
  
  
  # Compare models
  comp <- anova(mod1, mod2)
  pval <- comp$`Pr(>Chisq)`[2]
  delta_aic <- AIC(mod2) - AIC(mod1)

  
  # Extract predicted log-variance (dispersion) per group
  rg <- ref_grid(mod2, component = "disp")
  disp_means <- as.data.frame(emmeans(rg, specs = ~ Species * Basin))
  
  # Back-transform from log scale to get actual Variance (sigma^2)
  disp_means$Variance <- exp(disp_means$emmean)
  
  
  ratios_by_basin <- disp_means %>%
    select(Species, Basin, Variance) %>%
    pivot_wider(names_from = Species, values_from = Variance) %>%
    mutate(Var_Ratio = `Sailfin molly` / `Amazon molly`) %>% 
    mutate(Trait = trait, Delta_AIC = delta_aic, LRT_p = pval)
  
  return(ratios_by_basin)

})

geo_results_df2 <- do.call(rbind, results_geo2)


p_adjusted_df2 <- geo_results_df2 %>%
  select(Trait, LRT_p, Delta_AIC) %>%
  distinct() %>%
  mutate(p_adj = p.adjust(LRT_p, method = "holm")) %>%
  mutate(
    p.display = case_when(
      p_adj < 0.001 ~ "*p* < 0.001*",
      p_adj < 0.05  ~ paste0(sprintf("%.3f", p_adj), "*"),
      TRUE          ~ sprintf("%.3f", p_adj)),
    Trait_Name = factor(trait.names[Trait], levels = unname(trait.names)))

ratios_wide2 <- geo_results_df2 %>%
  select(Trait, Basin, Var_Ratio) %>%
  pivot_wider(
    names_from = Basin, 
    values_from = Var_Ratio, 
    names_prefix = "Ratio_")

geo_wide_results2 <- p_adjusted_df2 %>%
  left_join(ratios_wide2, by = "Trait") %>%
  arrange(Trait_Name)

geo.met.tbl2 <- geo_wide_results2 %>%
  gt() %>%
  cols_hide(c(Trait, LRT_p, p_adj)) %>%
  cols_move_to_start(columns = Trait_Name) %>%
  cols_label(
    Trait_Name = "Trait",
    Delta_AIC = md("&Delta;AIC"),
    p.display = md("*p*-value"),
    `Ratio_San Antonio-Guadalupe`             = "San Antonio-Guadalupe",  
    `Ratio_San Antonio-Nueces Coastal`        = "San Antonio-Nueces Coastal",  
    `Ratio_Upper Southwest Texas Coastal`     = "Upper Southwest Texas Coastal",
    `Ratio_Lower Southwest Texas Coastal`     = "Lower Southwest Texas Coastal",
    `Ratio_Lower Río Grande`                  = "Lower Río Grande",
    `Ratio_Lagunas de San Andrés-Morales`     = "Lagunas de San Andrés-Morales",
    `Ratio_Río Tamesí`                        = "Río Tamesí") %>%
  tab_spanner(
    label = md("Variance Ratio *(Sailfin / Amazon)*"),
    columns = starts_with("Ratio_")) %>%
  fmt_number(
    columns = c(Delta_AIC, starts_with("Ratio_")),
    decimals = 2) %>%
  fmt_markdown(columns = p.display) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Trait_Name, p.display))
  ) %>%
  cols_align("center", everything()) %>%
  opt_horizontal_padding(scale = 2)

geo.met.tbl2


# ───────────────────────────────────────────────────
## MERISTIC (COUNT) ANALYSIS--BASINS
# ───────────────────────────────────────────────────


poisson_geo2 <- lapply(cnt.vars, function(trait) {
  
  # 1. Build the formula for the count mean
  form <- as.formula(paste0(trait, " ~ SL * Species"))
  
  # To safely extract variance ratios per zone, we can look at the 
  # residual dispersion factor for each Species X Zone combination.
  # We do this by splitting the data by Zone, or mapping across them.
  
  basins <- unique(count2$Basin)
  
  basin_ratios <- lapply(basins, function(b) {
    # Subset data for just this basin
    sub_data <- subset(count2, Basin == b)
    
    # Fit a robust Quasipoisson GLM for this basin
    # This automatically estimates a custom dispersion parameter for the species mix
    mod_sub <- glm(form, data = sub_data, family = quasipoisson)
    
    # To get separate variances/dispersions for each species within this basin:
    # We look at the pearson residuals per species level
    sub_data$resid_sq <- (residuals(mod_sub, type = "pearson"))^2
    
    # The mean of squared Pearson residuals per group gives its dispersion factor (phi)
    phi_estimates <- aggregate(resid_sq ~ Species, data = sub_data, FUN = mean)
    
    # Extract values safely based on your level names
    phi_amazon  <- phi_estimates$resid_sq[phi_estimates$Species == "Amazon molly"]
    phi_sailfin <- phi_estimates$resid_sq[phi_estimates$Species == "Sailfin molly"]
    
    # If a group is empty or missing, safeguard it
    var_ratio <- if(length(phi_amazon) > 0 & length(phi_sailfin) > 0) {
      phi_sailfin / phi_amazon
    } else {
      NA
    }
    
    data.frame(Basin = b, Var_Ratio = var_ratio)
  })
  
  # Combine zones for this trait
  trait_basins <- do.call(rbind, basin_ratios)
  
  # 2. To get a single global significance p-value for the Basin interaction,
  # we compare a pooled model without a Zone interaction to one with it.
  mod_null_disp <- glm(form, data = count2, family = poisson) # Standard Poisson
  
  # Global test for whether counts step outside poisson parameters by Basin
  # We use a standard Chi-sq comparison
  anova_global <- drop1(glm(as.formula(paste0(trait, " ~ SL * Species * Basin")), 
                            data = count2, family = poisson), test = "Chisq")
  
  # Extract the interaction p-value safely
  pval <- anova_global["SL:Species:Basin", "Pr(>Chi)"]
  if(is.na(pval)) pval <- anova_global["Species:Basin", "Pr(>Chi)"] # Fallback if size interaction isn't main focus
  
  # Dummy Delta_AIC since Quasipoisson doesn't use true maximum likelihood AIC
  delta_aic <- NA 
  
  # Attach metadata to our zone matrix
  trait_basins$Trait <- trait
  trait_basins$Delta_AIC <- delta_aic
  trait_basins$LRT_p <- pval
  
  return(trait_basins)
})

# Combine results into the long data frame for your gt table
geo_poisson_df2 <- do.call(rbind, poisson_geo2)

geo_poisson_df2$p_adj <- p.adjust(geo_poisson_df2$LRT_p,
                             method = "holm")

geo.cnt.results2 <- geo_poisson_df2 %>%
  mutate(
    p.display = case_when(
      p_adj < 0.001 ~ paste0("*p* < 0.001*"),
      p_adj < 0.05  ~ paste0(sprintf("%.3f", p_adj), "*"),
      TRUE ~ sprintf("%.3f", p_adj))) %>%
  mutate(
    Trait_Name = factor(
      trait.names[Trait],
      levels = unname(trait.names))) %>%
  arrange(Trait_Name, Basin) %>%
  relocate(Trait_Name, Basin, .before = everything())

geo.cnt.results2

## STILL NO DIFFERENCE IN MERISTIC TRAITS ACROSS ZONES; ALL P-VALUES ARE 1




```

***
***



### Overall shape variation (PCA)

This will still use the same PCA as above, but now the post-PCA tests will include geographic classifiers (zones, basins).

#### *Zones*

**CONCLUSIONS:** Median comparisons were significant for PC1, PC2, PC3 and CatPC1, but not for CatPC2; variance was significant for PC2 and PC3. For PC1, zone 1 shows a significant differences in morphospace occupied by the species and zone 4 is a marginal difference. For PC2 all zones showed a significant difference in morphospace occupied by the species, but only zone 1 shows a difference in variation with Amazons exhibiting more variation. For PC3, only zones 3 and 4 show a significant difference in morphospace occupation, and again only zone 1 shows a difference in variaiton with Amazons exhibiting more variation. For CatPC1, all zones show a significant difference in morphospace occupied.

```{r}
#quick distribution check
testmod <- lm(Dim.1~Species*Zone, data= pca.df)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(Dim.2~Species*Zone, data= pca.df)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(Dim.3~Species*Zone, data= pca.df)
plot(simulateResiduals(testmod)) #fails outlier test and levene's test significant; will use glmm

testmod <- lm(CatPC1~Species*Zone, data= catpca.df)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(CatPC2~Species*Zone, data= catpca.df)
plot(simulateResiduals(testmod)) #fails KS test, outlier test, and within-group deviation; will use glmm


# ───────────────────────────────────────────────────
## MEDIAN TESTS (GLMM for all)
# ───────────────────────────────────────────────────

### Metric Data
geo.met.mean.1 <- glmmTMB(Dim.1 ~Species*Zone, dispformula = ~Species * Zone, data = pca.df)
(gmm1 <- Anova(geo.met.mean.1, type = "III")) #all effects significant (all p<0.001)
    post_hoc1 <- emmeans(geo.met.mean.1,  ~ Species | Zone)
    summary(post_hoc1)
    pw1 <- pairs(post_hoc1)
    (pair1 <- as.data.frame(rbind(pw1, adjust = "holm"))) #significant species difference in zone 1 (p<0.001) and marginal for zone 4 (p=0.056)

geo.met.mean.2 <- glmmTMB(Dim.2 ~Species*Zone, dispformula = ~Species*Zone, data = pca.df) 
(gmm2 <- Anova(geo.met.mean.2, type = "III")) #all effects significant: species (p<0.001), zone (p<0.001), species:zone (p=0.003)
    post_hoc2 <- emmeans(geo.met.mean.2,  ~ Species | Zone)
    summary(post_hoc2)
    pw2 <- pairs(post_hoc2)
    (pair2 <- as.data.frame(rbind(pw2, adjust = "holm"))) #all zones significant (p<0.001)


geo.met.mean.3 <- glmmTMB(Dim.3 ~Species*Zone, dispformula = ~Species*Zone, data = pca.df) 
(gmm3<- Anova(geo.met.mean.3, type = "III")) #zone and species:zone significant (p<0.001) but not species
    post_hoc3 <- emmeans(geo.met.mean.3,  ~ Species | Zone)
    summary(post_hoc3)
    pw3 <- pairs(post_hoc3)
    (pair3 <- as.data.frame(rbind(pw3, adjust = "holm"))) #only zone 3 (p=0.010) and zone 4 (p<0.001) significant

### Meristic Data
geo.cnt.mean.1 <- glmmTMB(CatPC1 ~Species*Zone, dispformula = ~Species*Zone, data = catpca.df)
(gcm1<- Anova(geo.cnt.mean.1, type = "III")) # species (p<0.001) and species:zone (p=0.002) significant, but not zones
    post_hoc4 <- emmeans(geo.cnt.mean.1,  ~ Species | Zone)
    summary(post_hoc4)
    pw4 <- pairs(post_hoc4)
    (pair4 <- as.data.frame(rbind(pw4, adjust = "holm"))) #all zones significant (p<0.001)

geo.cnt.mean.2 <- glmmTMB(CatPC2 ~Species*Zone, dispformula = ~Species*Zone, data = catpca.df)
(gcm2<- Anova(geo.cnt.mean.2, type = "III")) #not significant

# ───────────────────────────────────────────────────
## VARIANCE TESTS (Brown-Forsythe for all)
# ───────────────────────────────────────────────────

### Metric Data
geo.met.var.1 <- leveneTest(Dim.1~Species*Zone, data= pca.df) #not significant

geo.met.var.2 <- leveneTest(Dim.2~Species*Zone, data= pca.df) #significant (p=0.03)
    post_hoc5 <- emmeans(geo.met.mean.2,  ~ Species | Zone, component = "disp")
    summary(post_hoc5)
    pw5 <- pairs(post_hoc5)
    (pair5 <- as.data.frame(rbind(pw5, adjust = "holm"))) #only zone 1 significant (p=0.03)
    
geo.met.var.3 <- leveneTest(Dim.3~Species*Zone, data= pca.df) #significant (p=0.003)
    post_hoc6 <- emmeans(geo.met.mean.3,  ~ Species | Zone, component = "disp")
    summary(post_hoc6)
    pw6 <- pairs(post_hoc6)
    (pair6 <- as.data.frame(rbind(pw6, adjust = "holm"))) #only zone 1 significant (p<0.001)
    

### Meristic Data
geo.cnt.var.1 <- leveneTest(CatPC1~Species*Zone, data= catpca.df) #not significant

geo.cnt.var.2 <- leveneTest(CatPC2~Species*Zone, data= catpca.df) #not significant


# ───────────────────────────────────────────────────
## TABLE
# ───────────────────────────────────────────────────

### Metric Data

#get medians and median absolute deviations for each dimension per species per zone
geo_met_summary <- pca.df %>%
  group_by(Species, Zone) %>%
  summarise(
    PC1_median = median(Dim.1, na.rm = TRUE),
    PC1_mad  = mad(Dim.1, na.rm = TRUE, constant = 1),
    PC2_median = median(Dim.2, na.rm = TRUE),
    PC2_mad  = mad(Dim.2, na.rm = TRUE, constant = 1),
    PC3_median = median(Dim.3, na.rm = TRUE),
    PC3_mad  = mad(Dim.3, na.rm = TRUE, constant = 1))

#format summary stats
geo_met_long <- geo_met_summary %>%
  pivot_longer(
    cols = starts_with("PC"),
    names_to = c("Dimension", ".value"),
    names_pattern = "(PC\\d)_(.*)") %>%
  mutate(stat_string = sprintf("%.2f (%.2f)", median, mad)) %>%
  select(Species, Zone, Dimension, stat_string) %>%
  pivot_wider(names_from = Species, values_from = stat_string)


#extract Anova(model) results and levene's results
stats_results <- tibble(
  Dimension = c("PC1", "PC2", "PC3"),
  Chi_value = c(
    gmm1["Species:Zone", "Chisq"], 
    gmm2["Species:Zone", "Chisq"], 
    gmm3["Species:Zone", "Chisq"]),
  p_int     = c(
    gmm1["Species:Zone", "Pr(>Chisq)"], 
    gmm2["Species:Zone", "Pr(>Chisq)"], 
    gmm3["Species:Zone", "Pr(>Chisq)"]),
  F_value   = c(
    geo.met.var.1$`F value`[1], 
    geo.met.var.2$`F value`[1], 
    geo.met.var.3$`F value`[1]),
  p_lv      = c(
    geo.met.var.1$`Pr(>F)`[1], 
    geo.met.var.2$`Pr(>F)`[1], 
    geo.met.var.3$`Pr(>F)`[1]))


#format results for final table
postPCA_geo_result <- geo_met_long %>%
  left_join(stats_results, by = "Dimension") %>%
  mutate(
    Classifier = "Metric PCA",
    Med_test = case_when(
      p_int < 0.001 ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chi_value), "<br>*p* < 0.001*"),
      p_int < 0.05  ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chi_value), "<br>*p* = ", sprintf("%.3f", p_int), "*"),
      TRUE          ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chi_value), "<br>*p* = ", sprintf("%.3f", p_int))),
    Variance_test = case_when(
      p_lv < 0.001 ~ paste0("F = ", sprintf("%.2f", F_value), "<br>*p* < 0.001*"),
      p_lv < 0.05  ~ paste0("F = ", sprintf("%.2f", F_value), "<br>*p* = ", sprintf("%.3f", p_lv), "*"),
      TRUE          ~ paste0("F = ", sprintf("%.2f", F_value), "<br>*p* = ", sprintf("%.3f", p_lv))))




### Meristic Data


#get medians and median absolute deviations for each dimension per species per zone
geo_cnt_summary <- catpca.df %>%
  group_by(Species, Zone) %>%
  summarise(
    CatPC1_median = median(CatPC1, na.rm = TRUE),
    CatPC1_mad  = mad(CatPC1, na.rm = TRUE, constant = 1),
    CatPC2_median = median(CatPC2, na.rm = TRUE),
    CatPC2_mad  = mad(CatPC2, na.rm = TRUE, constant = 1))

#format summary stats
geo_cnt_long <- geo_cnt_summary %>%
  pivot_longer(
    cols = starts_with("CatPC"),
    names_to = c("Dimension", ".value"),
    names_pattern = "(CatPC\\d)_(.*)") %>%
  mutate(stat_string = sprintf("%.2f (%.2f)", median, mad)) %>%
  select(Species, Zone, Dimension, stat_string) %>%
  pivot_wider(names_from = Species, values_from = stat_string)


#extract Anova(model) results and levene's results
stats_results2 <- tibble(
  Dimension = c("CatPC1", "CatPC2"),
  Chi_value = c(
    gcm1["Species:Zone", "Chisq"], 
    gcm2["Species:Zone", "Chisq"]),
  p_int     = c(
    gcm1["Species:Zone", "Pr(>Chisq)"], 
    gcm2["Species:Zone", "Pr(>Chisq)"]),
  F_value   = c(
    geo.cnt.var.1$`F value`[1], 
    geo.cnt.var.2$`F value`[1]),
  p_lv      = c(
    geo.cnt.var.1$`Pr(>F)`[1], 
    geo.cnt.var.2$`Pr(>F)`[1]))


#format results for final table
postCPCA_geo_result <- geo_cnt_long %>%
  left_join(stats_results2, by = "Dimension") %>%
  mutate(
    Classifier = "Meristic CatPCA",
    Med_test = case_when(
      p_int < 0.001 ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chi_value), "<br>*p* < 0.001*"),
      p_int < 0.05  ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chi_value), "<br>*p* = ", sprintf("%.3f", p_int), "*"),
      TRUE          ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chi_value), "<br>*p* = ", sprintf("%.3f", p_int))),
    Variance_test = case_when(
      p_lv < 0.001 ~ paste0("F = ", sprintf("%.2f", F_value), "<br>*p* < 0.001*"),
      p_lv < 0.05  ~ paste0("F = ", sprintf("%.2f", F_value), "<br>*p* = ", sprintf("%.3f", p_lv), "*"),
      TRUE          ~ paste0("F = ", sprintf("%.2f", F_value), "<br>*p* = ", sprintf("%.3f", p_lv))))


### Add posthoc test results
extract_pairs <- function(pair_obj, dim_name, type = "Med") {
  if (is.null(pair_obj)) return(NULL)
  
  pair_obj %>%
    mutate(
      Dimension = dim_name,
      Zone = as.character(Zone),
      p_fmt = case_when(
        p.value < 0.001 ~ "*p* < 0.001*",
        p.value < 0.05  ~ sprintf("*p* = %.3f*", p.value),
        TRUE            ~ sprintf("*p* = %.3f", p.value)),
      PostHoc_Sum = sprintf("<b>%s:</b> %.2f (&plusmn;%.2f); %s", type, estimate, SE, p_fmt)) %>%
    select(Dimension, Zone, PostHoc_Sum)
}

# Bind all pairwise comparisons together
posthoc_ref <- bind_rows(
  extract_pairs(pair1, "PC1", "M"),
  extract_pairs(pair2, "PC2", "M"),
  extract_pairs(pair3, "PC3", "M"),
  extract_pairs(pair4, "CatPC1", "M"),
  extract_pairs(pair5, "PC2", "V"),
  extract_pairs(pair6, "PC3", "V")) %>%
  group_by(Dimension, Zone) %>%
  summarise(
    PostHoc_Summary = paste(PostHoc_Sum, collapse = "<br>"),
    .groups = "drop")


### Combine & form table

geo_results_combined <- bind_rows(postPCA_geo_result, postCPCA_geo_result) %>%
  mutate(
    Med_Part = paste0("Median: &chi;<sup>2</sup> = ", sprintf("%.2f", Chi_value), 
                      " (*p* ", ifelse(p_int < 0.001, "< 0.001*", paste0("= ", sprintf("%.3f", p_int), ifelse(p_int < 0.05, "*", ""))), ")"),
    Var_Part = paste0("Variance: F = ", sprintf("%.2f", F_value), 
                      " (*p* ", ifelse(p_lv < 0.001, "< 0.001*", paste0("= ", sprintf("%.3f", p_lv), ifelse(p_lv < 0.05, "*", ""))), ")"),
    Test_Summary = paste0(Med_Part, " &nbsp;&bull;&nbsp; ", Var_Part),
    Group_Label = paste0("**", Classifier, "**<br>",
                         "***", Dimension, "***<br>",
                         "<span style='font-size: 85%'>", Test_Summary, "</span>"),
    Zone = as.character(Zone)) %>%
  left_join(posthoc_ref, by = c("Dimension", "Zone")) %>%
  mutate(PostHoc_Summary = coalesce(PostHoc_Summary, "&mdash;"))
  


postPCA_geo_table <- geo_results_combined %>%
  select(
    Group_Label,
    Zone,
    `Amazon molly`,
    `Sailfin molly`,
    PostHoc_Summary) %>%
  gt(groupname_col = "Group_Label", process_md = TRUE) %>%
  fmt_markdown(columns = everything()) %>%
  cols_label(
    Zone = " ",
    `Amazon molly` = md("**Amazon**<br>Median (MAD)"),
    `Sailfin molly` = md("**Sailfin**<br>Median (MAD)"),
    PostHoc_Summary = md("**Post-hoc Contrasts**<br>(M)edian / (V)ariance<br><span style='font-size:90%; font-weight:normal;'>est. (SE); adj. p</span>")) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c(Zone, `Amazon molly`, `Sailfin molly`))) %>%
  cols_align(align = "center", columns = everything()) %>%
  opt_horizontal_padding(scale = 2)

#final table
postPCA_geo_table





```

***
***

#### *Basins*

**CONCLUSIONS:** Median comparisons were significant for all PCs and CatPCs; variation comparisons were significant for PC1 and PC3 (marginally). For PC1 species occupied different morphospace in: San Antonio-Nueces Coastal, Upper and Lower Southwest Texas Coastal, and Rio Tamesi; variation was significantly different in Upper SWTX coastal and Lower Rio Grande (marginally), where Amazons had higher variation. For PC2 species occupied different morphospace in all basins. For PC3 species occupied different morphospace in San Antonio-Guadalupe and Lower SWTX coastal; variation was significantly different in Lower SWTX coastal, where Amazons exhibited more variation. For CatPC1 species occupied different morphospace in all basins. For CatPC2, the medians were not equivalent across basins (leading to a significant result for median comparison) but posthoc analysis did not show any significant basin (indicating that they were not significantly different from a 0 estimate).

```{r}
#filter for basins with >3 for each species
pca.df2 <- pca.df %>% filter(Basin %in% valid_basins)
catpca.df2 <- catpca.df %>% filter(Basin %in% valid_basins)

#quick distribution check
testmod <- lm(Dim.1~Species*Basin, data= pca.df2)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(Dim.2~Species*Basin, data= pca.df2)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(Dim.3~Species*Basin, data= pca.df2)
plot(simulateResiduals(testmod)) #fails outlier test; will use glmm

testmod <- lm(CatPC1~Species*Basin, data= catpca.df2)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(CatPC2~Species*Basin, data= catpca.df2)
plot(simulateResiduals(testmod)) #fails KS test, outlier test, and combined adjusted quantile test; will use glmm


# ───────────────────────────────────────────────────
## MEDIAN TESTS (GLMM for all)
# ───────────────────────────────────────────────────

### Metric Data
geo2.met.mean.1 <- glmmTMB(Dim.1 ~Species*Basin, dispformula = ~Species * Basin, data = pca.df2)
(g2mm1 <- Anova(geo2.met.mean.1, type = "III")) #basin and species:basin significant (p<0.001) but not species
    post_hoc2.1 <- emmeans(geo2.met.mean.1,  ~ Species | Basin)
    summary(post_hoc2.1)
    pw2.1 <- pairs(post_hoc2.1)
    (pair2.1 <- as.data.frame(rbind(pw2.1, adjust = "holm"))) #significant species difference in San Antonio-Nueces Coastal (p<0.001), Upper and Lower Southwest Texas coastal (p=0.003 and p<0.001) and Rio Tamesi (p=0.014)

geo2.met.mean.2 <- glmmTMB(Dim.2 ~Species*Basin, dispformula = ~Species*Basin, data = pca.df2) 
(g2mm2 <- Anova(geo2.met.mean.2, type = "III")) #all effects significant: species (all p<0.001)
    post_hoc2.2 <- emmeans(geo2.met.mean.2,  ~ Species | Basin)
    summary(post_hoc2.2)
    pw2.2 <- pairs(post_hoc2.2)
    (pair2.2 <- as.data.frame(rbind(pw2.2, adjust = "holm"))) #all Basins significant (p<0.001)


geo2.met.mean.3 <- glmmTMB(Dim.3 ~Species*Basin, dispformula = ~Species*Basin, data = pca.df2) 
(g2mm3<- Anova(geo2.met.mean.3, type = "III")) #all effects significant (p<0.001)
    post_hoc2.3 <- emmeans(geo2.met.mean.3,  ~ Species | Basin)
    summary(post_hoc2.3)
    pw2.3 <- pairs(post_hoc2.3)
    (pair2.3 <- as.data.frame(rbind(pw2.3, adjust = "holm"))) #only San Antonio-Guadalupe (p<0.001) and Lower Southwest Texas coastal (p=0.003) significant

### Meristic Data
geo2.cnt.mean.1 <- glmmTMB(CatPC1 ~Species*Basin, dispformula = ~Species*Basin, data = catpca.df2)
(g2cm1<- Anova(geo2.cnt.mean.1, type = "III")) # species (p<0.001) and species:Basin (p=0.013) significant, but not Basins
    post_hoc2.4 <- emmeans(geo2.cnt.mean.1,  ~ Species | Basin)
    summary(post_hoc2.4)
    pw2.4 <- pairs(post_hoc2.4)
    (pair2.4 <- as.data.frame(rbind(pw2.4, adjust = "holm"))) #all Basins significant (p<0.001)

geo2.cnt.mean.2 <- glmmTMB(CatPC2 ~Species*Basin, dispformula = ~Species*Basin, data = catpca.df2)
(g2cm2<- Anova(geo2.cnt.mean.2, type = "III")) #basin (p<0.001) and species:basin (p=0.037) significant, but not species
    post_hoc2.5 <- emmeans(geo2.cnt.mean.2,  ~ Species | Basin)
    summary(post_hoc2.5)
    pw2.5 <- pairs(post_hoc2.5)
    (pair2.5 <- as.data.frame(rbind(pw2.5, adjust = "holm"))) #no Basins significant

# ───────────────────────────────────────────────────
## VARIANCE TESTS (Brown-Forsythe for all)
# ───────────────────────────────────────────────────

### Metric Data
geo2.met.var.1 <- leveneTest(Dim.1~Species*Basin, data= pca.df2) #significant (p<0.001)
    post_hoc2.6 <- emmeans(geo2.met.mean.1,  ~ Species | Basin, component = "disp")
    summary(post_hoc2.6)
    pw2.6 <- pairs(post_hoc2.6)
    (pair2.6 <- as.data.frame(rbind(pw2.6, adjust = "holm"))) #only Upper SW Texas Coastal (p<0.001) and Lower Rio Grande (p=0.049) significant


geo2.met.var.2 <- leveneTest(Dim.2~Species*Basin, data= pca.df2) #not significant
   
    
geo2.met.var.3 <- leveneTest(Dim.3~Species*Basin, data= pca.df2) #marginal (p=0.050)
    post_hoc2.7 <- emmeans(geo2.met.mean.3,  ~ Species | Basin, component = "disp")
    summary(post_hoc2.7)
    pw2.7 <- pairs(post_hoc2.7)
    (pair2.7 <- as.data.frame(rbind(pw2.7, adjust = "holm"))) #only Lower SW Texas Coastal significant (p<0.001)
    

### Meristic Data
geo2.cnt.var.1 <- leveneTest(CatPC1~Species*Basin, data= catpca.df2) #not significant

geo2.cnt.var.2 <- leveneTest(CatPC2~Species*Basin, data= catpca.df2) #not significant


# ───────────────────────────────────────────────────
## TABLE
# ───────────────────────────────────────────────────

### Metric Data

#get medians and median absolute deviations for each dimension per species per Basin
geo2_met_summary <- pca.df2 %>%
  group_by(Species, Basin) %>%
  summarise(
    PC1_median = median(Dim.1, na.rm = TRUE),
    PC1_mad  = mad(Dim.1, na.rm = TRUE, constant = 1),
    PC2_median = median(Dim.2, na.rm = TRUE),
    PC2_mad  = mad(Dim.2, na.rm = TRUE, constant = 1),
    PC3_median = median(Dim.3, na.rm = TRUE),
    PC3_mad  = mad(Dim.3, na.rm = TRUE, constant = 1))

#format summary stats
geo2_met_long <- geo2_met_summary %>%
  pivot_longer(
    cols = starts_with("PC"),
    names_to = c("Dimension", ".value"),
    names_pattern = "(PC\\d)_(.*)") %>%
  mutate(stat_string = sprintf("%.2f (%.2f)", median, mad)) %>%
  select(Species, Basin, Dimension, stat_string) %>%
  pivot_wider(names_from = Species, values_from = stat_string)


#extract Anova(model) results and levene's results
stats_results2.1 <- tibble(
  Dimension = c("PC1", "PC2", "PC3"),
  Chi_value = c(
    g2mm1["Species:Basin", "Chisq"], 
    g2mm2["Species:Basin", "Chisq"], 
    g2mm3["Species:Basin", "Chisq"]),
  p_int     = c(
    g2mm1["Species:Basin", "Pr(>Chisq)"], 
    g2mm2["Species:Basin", "Pr(>Chisq)"], 
    g2mm3["Species:Basin", "Pr(>Chisq)"]),
  F_value   = c(
    geo2.met.var.1$`F value`[1], 
    geo2.met.var.2$`F value`[1], 
    geo2.met.var.3$`F value`[1]),
  p_lv      = c(
    geo2.met.var.1$`Pr(>F)`[1], 
    geo2.met.var.2$`Pr(>F)`[1], 
    geo2.met.var.3$`Pr(>F)`[1]))


#format results for final table
postPCA_geo2_result <- geo2_met_long %>%
  left_join(stats_results2.1, by = "Dimension") %>%
  mutate(
    Classifier = "Metric PCA",
    Med_test = case_when(
      p_int < 0.001 ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chi_value), "<br>*p* < 0.001*"),
      p_int < 0.05  ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chi_value), "<br>*p* = ", sprintf("%.3f", p_int), "*"),
      TRUE          ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chi_value), "<br>*p* = ", sprintf("%.3f", p_int))),
    Variance_test = case_when(
      p_lv < 0.001 ~ paste0("F = ", sprintf("%.2f", F_value), "<br>*p* < 0.001*"),
      p_lv < 0.05  ~ paste0("F = ", sprintf("%.2f", F_value), "<br>*p* = ", sprintf("%.3f", p_lv), "*"),
      TRUE          ~ paste0("F = ", sprintf("%.2f", F_value), "<br>*p* = ", sprintf("%.3f", p_lv))))




### Meristic Data


#get medians and median absolute deviations for each dimension per species per Basin
geo2_cnt_summary <- catpca.df2 %>%
  group_by(Species, Basin) %>%
  summarise(
    CatPC1_median = median(CatPC1, na.rm = TRUE),
    CatPC1_mad  = mad(CatPC1, na.rm = TRUE, constant = 1),
    CatPC2_median = median(CatPC2, na.rm = TRUE),
    CatPC2_mad  = mad(CatPC2, na.rm = TRUE, constant = 1))

#format summary stats
geo2_cnt_long <- geo2_cnt_summary %>%
  pivot_longer(
    cols = starts_with("CatPC"),
    names_to = c("Dimension", ".value"),
    names_pattern = "(CatPC\\d)_(.*)") %>%
  mutate(stat_string = sprintf("%.2f (%.2f)", median, mad)) %>%
  select(Species, Basin, Dimension, stat_string) %>%
  pivot_wider(names_from = Species, values_from = stat_string)


#extract Anova(model) results and levene's results
stats_results2.2 <- tibble(
  Dimension = c("CatPC1", "CatPC2"),
  Chi_value = c(
    g2cm1["Species:Basin", "Chisq"], 
    g2cm2["Species:Basin", "Chisq"]),
  p_int     = c(
    g2cm1["Species:Basin", "Pr(>Chisq)"], 
    g2cm2["Species:Basin", "Pr(>Chisq)"]),
  F_value   = c(
    geo2.cnt.var.1$`F value`[1], 
    geo2.cnt.var.2$`F value`[1]),
  p_lv      = c(
    geo2.cnt.var.1$`Pr(>F)`[1], 
    geo2.cnt.var.2$`Pr(>F)`[1]))


#format results for final table
postCPCA_geo2_result <- geo2_cnt_long %>%
  left_join(stats_results2.2, by = "Dimension") %>%
  mutate(
    Classifier = "Meristic CatPCA",
    Med_test = case_when(
      p_int < 0.001 ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chi_value), "<br>*p* < 0.001*"),
      p_int < 0.05  ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chi_value), "<br>*p* = ", sprintf("%.3f", p_int), "*"),
      TRUE          ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chi_value), "<br>*p* = ", sprintf("%.3f", p_int))),
    Variance_test = case_when(
      p_lv < 0.001 ~ paste0("F = ", sprintf("%.2f", F_value), "<br>*p* < 0.001*"),
      p_lv < 0.05  ~ paste0("F = ", sprintf("%.2f", F_value), "<br>*p* = ", sprintf("%.3f", p_lv), "*"),
      TRUE          ~ paste0("F = ", sprintf("%.2f", F_value), "<br>*p* = ", sprintf("%.3f", p_lv))))



### Add posthoc test results
extract_pairs2 <- function(pair_obj, dim_name, type = "Med") {
  if (is.null(pair_obj)) return(NULL)
  
  pair_obj %>%
    mutate(
      Dimension = dim_name,
      Basin = as.character(Basin),
      p_fmt = case_when(
        p.value < 0.001 ~ "*p* < 0.001*",
        p.value < 0.05  ~ sprintf("*p* = %.3f*", p.value),
        TRUE            ~ sprintf("*p* = %.3f", p.value)),
      PostHoc_Sum = sprintf("<b>%s:</b> %.2f (&plusmn;%.2f); %s", type, estimate, SE, p_fmt)) %>%
    select(Dimension, Basin, PostHoc_Sum)
}

# Bind all pairwise comparisons together
posthoc_ref2 <- bind_rows(
  extract_pairs2(pair2.1, "PC1", "M"),
  extract_pairs2(pair2.2, "PC2", "M"),
  extract_pairs2(pair2.3, "PC3", "M"),
  extract_pairs2(pair2.4, "CatPC1", "M"),
  extract_pairs2(pair2.5, "CatPC2", "M"),
  extract_pairs2(pair2.6, "PC1", "V"),
  extract_pairs2(pair2.7, "PC3", "V")) %>%
  group_by(Dimension, Basin) %>%
  summarise(
    PostHoc_Summary = paste(PostHoc_Sum, collapse = "<br>"),
    .groups = "drop")


### Combine & form table

geo2_results_combined <- bind_rows(postPCA_geo2_result, postCPCA_geo2_result) %>%
  mutate(
    Med_Part = paste0("Median: &chi;<sup>2</sup> = ", sprintf("%.2f", Chi_value), 
                      " (*p* ", ifelse(p_int < 0.001, "< 0.001*", paste0("= ", sprintf("%.3f", p_int), ifelse(p_int < 0.05, "*", ""))), ")"),
    Var_Part = paste0("Variance: F = ", sprintf("%.2f", F_value), 
                      " (*p* ", ifelse(p_lv < 0.001, "< 0.001*", paste0("= ", sprintf("%.3f", p_lv), ifelse(p_lv < 0.05, "*", ""))), ")"),
    Test_Summary = paste0(Med_Part, " &nbsp;&bull;&nbsp; ", Var_Part),
    Group_Label = paste0("**", Classifier, "**<br>",
                         "***", Dimension, "***<br>",
                         "<span style='font-size: 85%'>", Test_Summary, "</span>"),
    Basin = as.character(Basin)) %>%
  left_join(posthoc_ref2, by = c("Dimension", "Basin")) %>%
  mutate(PostHoc_Summary = coalesce(PostHoc_Summary, "&mdash;"))
  


postPCA_geo2_table <- geo2_results_combined %>%
  select(
    Group_Label,
    Basin,
    `Amazon molly`,
    `Sailfin molly`,
    PostHoc_Summary) %>%
  gt(groupname_col = "Group_Label", process_md = TRUE) %>%
  fmt_markdown(columns = everything()) %>%
  cols_label(
    Basin = " ",
    `Amazon molly` = md("**Amazon**<br>Median (MAD)"),
    `Sailfin molly` = md("**Sailfin**<br>Median (MAD)"),
    PostHoc_Summary = md("**Post-hoc Contrasts**<br>(M)edian / (V)ariance<br><span style='font-size:90%; font-weight:normal;'>est. (SE); adj. p</span>")) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c(Basin, `Amazon molly`, `Sailfin molly`))) %>%
  cols_align(align = "center", columns = everything()) %>%
  opt_horizontal_padding(scale = 2)

#final table
postPCA_geo2_table





```
***
***

## **Tampico MX Analysis**

For this analysis, we will compare sailfin, amazon and atlantic mollies from the Tampico, Mexico region. This is the proposed origin of Amazon mollies, and the only region in which all three species coexist. 

After filtering for individuals where the county/state information was recorded as Tamaulipas/MX, we have 32 sailfin molly, 28 Amazon molly, and 32 Atlantic molly.

### Individual trait models

**CONCLUSIONS:** Body depth, predorsal length, and dorsal base length were best supported with an unequal variance model; in all cases, the sexual species (sailfin & atlantic) were more variable than the amazon. All remaining metric traits and all meristic traits were best supported with an equal variance model (aka no significant difference in variation among the species). 


```{r}
#create separate df for metric and meristic (count) data
metric2 <- geo %>%
  select(-c(5:12)) %>%
  select(-c("FLA"))

count2 <- geo %>%
  select(-c(14:23)) #still includes SL for analysis



# ───────────────────────────────────────────────────
## METRIC ANALYSIS 
# ───────────────────────────────────────────────────

# Ensure Species is a factor
metric2$Species <- as.factor(metric2$Species)
metric2$Species <- relevel(metric2$Species, ref = "Amazon molly")

tam_results_list <- lapply(met.vars, function(trait) {
  
  # Build the formula string correctly
  form_str <- paste0("log(", trait, ") ~ log(SL) * Species")
  form <- as.formula(form_str)

  # 1. Homogeneous variance model
  mod1 <- glmmTMB(form, data = metric2, family = gaussian)
  
  # 2. Heterogeneous variance model
  ## NOTE: dispformula = ~Species allows variance to differ by species
  mod2 <- glmmTMB(form, data = metric2, family = gaussian, dispformula = ~Species)
  
  
  
  # Compare models
  comp <- anova(mod1, mod2)
  pval <- comp$`Pr(>Chisq)`[2]
  delta_aic <- AIC(mod2) - AIC(mod1)
  
  
  # Extract predicted log-variance per species
  rg <- ref_grid(mod2, component = "disp")
  disp_means <- as.data.frame(emmeans(rg, specs = ~ Species))
  
  # Back-transform to actual Variance (sigma^2)
  disp_means$Variance <- exp(disp_means$emmean)
  
  # Pivot to easily calculate the single global ratio
  ratio_df <- disp_means %>%
    select(Species, Variance) %>%
    pivot_wider(names_from = Species, values_from = Variance) %>%
    mutate(
      SA_Ratio = `Sailfin molly` / `Amazon molly`,
      AA_Ratio = `Atlantic molly` / `Amazon molly`) %>%
    mutate(Trait = trait, Delta_AIC = delta_aic, LRT_p = pval)
  
  return(ratio_df)
})

tam_results_df <- do.call(rbind, tam_results_list)

tam_results_df$p_adj <- p.adjust(tam_results_df$LRT_p,
                             method = "holm")

tam.met.results <- tam_results_df %>%
  mutate(
    p.display = case_when(
      p_adj < 0.001 ~ paste0("*p* < 0.001*"),
      p_adj < 0.05  ~ paste0(sprintf("%.3f", p_adj), "*"),
      TRUE ~ sprintf("%.3f", p_adj))) %>%
  mutate(
    Trait_Name = factor(
      trait.names[Trait],
      levels = unname(trait.names))) %>%
  arrange(Trait_Name) %>%
  relocate(Trait_Name, .before = everything())

tam.met.tbl <- tam.met.results %>%
  gt() %>%
  cols_hide(c(Trait, LRT_p, p_adj, "Sailfin molly", "Amazon molly", "Atlantic molly")) %>%
  cols_label(
    Trait_Name = "Trait",
    SA_Ratio = md("Variance ratio <br>(Sailfin / Amazon)"),
    AA_Ratio = md("Variance ratio <br>(Atlantic / Amazon)"),
    Delta_AIC = md("&Delta;AIC"),
    p.display = md("*p*-value")) %>%
  fmt_number(
    columns = c(SA_Ratio, AA_Ratio, Delta_AIC),
    decimals = 2) %>%
  fmt_markdown(columns = p.display) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Trait_Name, Delta_AIC, SA_Ratio, AA_Ratio))) %>%
  cols_align("center", everything()) %>%
  opt_horizontal_padding(scale = 2)
  
  
tam.met.tbl



# ───────────────────────────────────────────────────
## MERISTIC (COUNT) ANALYSIS 
# ───────────────────────────────────────────────────


# Ensure baseline is set
count2$Species <- as.factor(count2$Species)
count2$Species <- relevel(count2$Species, ref = "Amazon molly")

tam_poisson_results <- lapply(cnt.vars, function(trait) {
  
  # Build formula: Linear SL is best for count link functions
  form <- as.formula(paste0(trait, " ~ SL * Species"))
  
  # 1. Fit Poisson GLM
  mod <- glmmTMB(form, data = count2, family = poisson)
  
  # 2. Run Type III Wald Anova
  # tryCatch handles cases where extreme separation makes the Chidg tests unstable
  anova_res <- tryCatch({
    Anova(mod, type = "III")
  }, error = function(e) { NULL })
  
  # Extract p-values for main effects and interaction safely
  p_sl          <- if(!is.null(anova_res)) anova_res["SL", "Pr(>Chisq)"] else NA
  p_species     <- if(!is.null(anova_res)) anova_res["Species", "Pr(>Chisq)"] else NA
  p_interaction <- if(!is.null(anova_res)) anova_res["SL:Species", "Pr(>Chisq)"] else NA
  
  # 3. DHARMa Overdispersion Diagnostic
  sim_res   <- simulateResiduals(mod, plot = FALSE)
  disp_test <- testDispersion(sim_res, plot = FALSE)
  
  # Extract the raw dispersion ratio (> 1 means overdispersion) 
  # and the significance p-value
  disp_ratio  <- disp_test$statistic
  disp_pval   <- disp_test$p.value
  is_overdisp <- disp_pval < 0.05 & disp_ratio > 1

  data.frame(
    Trait = trait,
    AIC = AIC(mod),
    P_SL = p_sl,
    P_Species = p_species,
    P_Interaction = p_interaction,
    DHARMa_DispRatio = disp_ratio,
    DHARMa_p = disp_pval,
    Overdispersed = is_overdisp,
    row.names = NULL
  )
})

# Combine into a clean data frame
tam_poisson_df <- do.call(rbind, tam_poisson_results)

tam_poisson_corrected <- tam_poisson_df %>%
  mutate(
    # 1. Correct for the species main effect across all traits
    P_Species_adj = p.adjust(P_Species, method = "holm"),
    
    # 2. Correct for the size interaction across all traits
    P_Interaction_adj = p.adjust(P_Interaction, method = "holm"),
    
    # 3. Correct for the standard length effect across all traits
    P_SL_adj = p.adjust(P_SL, method = "holm")
  )

tam.cnt.results <- tam_poisson_corrected %>%
  mutate(
    across(
      .cols = c(P_SL_adj, P_Species_adj, P_Interaction_adj),
      .names = "{.col}_display", 
      .fns = ~ case_when(
        .x < 0.001 ~ "*p* < 0.001*",
        .x < 0.05  ~ paste0(sprintf("%.3f", .x), "*"),
        TRUE       ~ sprintf("%.3f", .x)))) %>%
  mutate(
    Trait_Name = factor(
      trait.names[Trait],
      levels = unname(trait.names))) %>%
  arrange(Trait_Name) %>%
  relocate(Trait_Name, .before = everything())

tam.cnt.tbl <- tam.cnt.results %>%
  gt() %>%
  cols_hide(c(Trait, P_SL, P_Species, P_Interaction, P_SL_adj, P_Species_adj, P_Interaction_adj, DHARMa_p)) %>%
  cols_label(
    Trait_Name = "Trait",
    AIC= "Model AIC",
    DHARMa_DispRatio = "Dispersion Ratio",
    Overdispersed = "Is overdispersed <br>(True/False)",
    P_SL_adj_display = md("*p*-value for SL"),
    P_Species_adj_display = md("*p*-value for Species"),
    P_Interaction_adj_display = md("*p*-value for SL:Species")) %>%
  fmt_number(
    columns = c(DHARMa_DispRatio, AIC),
    decimals = 2) %>%
  fmt_markdown(columns = starts_with("P_")) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Trait_Name, AIC, DHARMa_DispRatio, Overdispersed, P_SL_adj_display, P_Species_adj_display))) %>%
  cols_align("center", everything()) %>%
  opt_horizontal_padding(scale = 2)
  


tam.cnt.tbl

  
# ───────────────────────────────────────────────────
## PLOTS
# ───────────────────────────────────────────────────

met.long2 <- metric2 %>%
  pivot_longer(
    cols = c(SL, BD, CPD, CPL, PreDL, DbL, HL, HD, HW, SnL, OL),
    names_to = "Trait",
    values_to = "Value")


tam.met.plots <- ggplot(met.long2, aes(x=Species, y=Value, fill = Species)) + 
  geom_boxplot() +
  geom_point(color= "#8D8680", size=2, alpha=0.5) +
  facet_wrap(~ Trait, scales = "free_y", axes = "margins") +
  theme_classic(base_size = 10, base_family = "sans") + 
  theme(
    axis.title.y = element_text(size = 10, family = "sans"), 
    axis.text.y  = element_text(size = 10, family = "sans"),
    axis.text.x  = element_text(size = 10, family = "sans"),
    strip.text    = element_text(size = 10, family = "sans"),
    panel.spacing = unit(1.5, "lines")) +
  scale_fill_manual(values = c("Sailfin molly"="#78B7C5",
                               "Amazon molly"="#E1AF00",
                               "Atlantic molly"="#9986A5")) +
  scale_x_discrete(labels = c("Sailfin molly" = "Sailfin\nmolly", 
                              "Amazon molly" = "Amazon\nmolly", 
                              "Atlantic molly" = "Atlantic\nmolly")) +
  xlab("") +
  ylab("Length or Depth (mm)")
tam.met.plots


cnt.long2 <- count2 %>%
  pivot_longer(
    cols = c(D, P1, A, P1.R, LLSC, SALL, SBLL, SBDF, FLA),
    names_to = "Trait",
    values_to = "Value")

tam.cnt.plots <- ggplot(cnt.long2, aes(x=Species, y=Value, fill = Species)) + 
  geom_boxplot() +
  geom_point(color= "#8D8680", size=2, alpha=0.5) +
  facet_wrap(~ Trait, scales = "free_y", axes = "margins") +
  theme_classic(base_size = 10, base_family = "sans") + 
  theme(
    axis.title.y = element_text(size = 10, family = "sans"), 
    axis.text.y  = element_text(size = 10, family = "sans"),
    axis.text.x  = element_text(size = 10, family = "sans"),
    strip.text    = element_text(size = 10, family = "sans"),
    panel.spacing = unit(1.5, "lines")) +
  scale_fill_manual(values = c("Sailfin molly"="#78B7C5",
                               "Amazon molly"="#E1AF00",
                               "Atlantic molly"="#9986A5")) +
  scale_x_discrete(labels = c("Sailfin molly" = "Sailfin\nmolly", 
                              "Amazon molly" = "Amazon\nmolly", 
                              "Atlantic molly" = "Atlantic\nmolly")) +
  xlab("") +
  ylab("Count")
tam.cnt.plots


```

***
***

### Overall shape variation (PCA)

As before, we will assess overall shape variation among species using a ***principal component analysis (PCA)*** on centered and scaled data (all metric traits will also be log transformed to better approach normality). We will then perform a mean and variance test (e.g., ANOVA and Levene's test) to determine if overall shape variation differs between the species. 

As meristic traits are count data, a normal PCA cannot be applied. Instead, we will use a ***categorical principal component analysis (CATPCA)*** that uses optimal scaling to deal with discrete integers. 

#### *Loadings*

```{r}
#log transform metric traits
df_trans2 <- geo
df_trans2[, c(13:23)] <- log(df_trans2[, c(13:23)])

# ───────────────────────────────────────────────────
## METRIC ANALYSIS 
# ───────────────────────────────────────────────────


PCA.test2 <- prcomp(df_trans2[, 13:23], center = TRUE, scale. = FALSE) 

summary(PCA.test2)
(eig.val.test <- get_eigenvalue(PCA.test2))


# For many morphological studies, the first component is heavily weighted by body size and is often of less interest than components two and three (Humphries et al., 1981). Therefore, we will extract the first three dimensions

ind.test <- get_pca_ind(PCA.test2)
head(ind.test$coord[,1:3])


pca.df2 <- cbind(df_trans2, ind.test$coord[,1:3])


# Will investigate the influence of each trait on the first 3 dimensions
(loadings <- PCA.test2$rotation[, 1:3])

sorted.loadings.1 <- loadings[order(loadings[, 1]), 1]
myTitle <- "Loadings Plot for PC1" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.1, main=myTitle, xlab=myXlab, cex=0.5, col="blue")

sorted.loadings.2 <- loadings[order(loadings[, 2]), 2]
myTitle <- "Loadings Plot for PC2" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.2, main=myTitle, xlab=myXlab, cex=0.5, col="blue")

sorted.loadings.3 <- loadings[order(loadings[, 3]), 3]
myTitle <- "Loadings Plot for PC3" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.3, main=myTitle, xlab=myXlab, cex=0.5, col="blue")




# ───────────────────────────────────────────────────
## MERISTIC (COUNT) ANALYSIS 
# ───────────────────────────────────────────────────

# Remove SL from count df
count3 <- select(count2, -c("SL"))

# Run the Categorical PCA
# 'ndim = 2' extracts the first two principal components
# 'degrees = 1' treats the scaling transformations as piecewise linear (ideal for ordinal integers)

catpca.test2 <- princals(count3[, 5:13], ndim = 2, degrees = 1)

# Print the summary to see how much variance your new components explain
summary(catpca.test2)

(cat.loadings <- catpca.test2$loadings)

sorted.loadings.4 <- cat.loadings[order(cat.loadings[, 1]), 1]
myTitle <- "Loadings Plot for CATPC1" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.4, main=myTitle, xlab=myXlab, cex=0.5, col="blue")

sorted.loadings.5 <- cat.loadings[order(cat.loadings[, 2]), 2]
myTitle <- "Loadings Plot for CATPC2" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.5, main=myTitle, xlab=myXlab, cex=0.5, col="blue")

# Extract the eigenvalues (one per dimension)
eigenvals <- catpca.test2$evals

# Calculate the percentage of variation explained by each axis
# (Divided by 9 because you have 9 meristic traits)
pc1_percent <- round((eigenvals[1] / 9) * 100, 1)
pc2_percent <- round((eigenvals[2] / 9) * 100, 1)

# Quick print check to see your numbers (e.g., 24.5 and 14.2)
print(paste0("CatPC1: ", pc1_percent, "%"))
print(paste0("CatPC2: ", pc2_percent, "%"))

# 1. Extract the object scores for the first two dimensions
cat_pc_scores <- as.data.frame(catpca.test2$objectscores)
colnames(cat_pc_scores) <- c("CatPC1", "CatPC2")

# 2. Combine them back with your species and size metadata
catpca.df2 <- cbind(count2, cat_pc_scores)

```

#### *Plots*

```{r}

#### METRIC

tam.pca.1<- autoplot(PCA.test2, data = df_trans2, colour="Species", shape= 'Species', loadings=FALSE, loadings.label=FALSE, frame=TRUE, frame.type='norm', label= FALSE)+
  scale_color_manual(name="Species",
                    values = c("Sailfin molly"="#78B7C5", "Amazon molly"="#E1AF00", "Atlantic molly"="#9986A5")) +
  scale_fill_manual(name="Species",
                    values = c("Sailfin molly"="#78B7C5", "Amazon molly"="#E1AF00", "Atlantic molly"="#9986A5")) +
  scale_shape_manual(name= 'Species', 
                     values = c("Sailfin molly" = 17, "Amazon molly"=16, "Atlantic molly"= 3)) +
   guides(fill = guide_legend(override.aes = list(fill = NA)))+
  ggtitle("PCA Plot of Morphology traits") + theme_classic() 
tam.pca.1


tam.pca.2<- autoplot(PCA.test2, x=2, y=3, data = df_trans2, colour="Species", shape= 'Species', loadings=FALSE, loadings.label=FALSE, frame=TRUE, frame.type='norm', label= FALSE)+
  scale_color_manual(name="Species",
                    values = c("Sailfin molly"="#78B7C5", "Amazon molly"="#E1AF00", "Atlantic molly"="#9986A5")) +
  scale_fill_manual(name="Species",
                    values = c("Sailfin molly"="#78B7C5", "Amazon molly"="#E1AF00", "Atlantic molly"="#9986A5")) +
  scale_shape_manual(name= 'Species', 
                     values = c("Sailfin molly" = 17, "Amazon molly"=16, "Atlantic molly"= 3)) +
   guides(fill = guide_legend(override.aes = list(fill = NA)))+
  ggtitle("PCA Plot of Morphology traits") + theme_classic() 
tam.pca.2


#### MERISTIC

tam.catpca <- ggplot(catpca.df2, aes(x = CatPC1, y = CatPC2, color = Species, fill = Species, shape = Species)) +
  geom_point(size = 1, alpha = 0.8) +
  stat_ellipse(geom = "polygon", alpha = 0.15, type = "norm", level = 0.95) +
  scale_color_manual(name = "Species",
                     values = c("Sailfin molly" = "#78B7C5", "Amazon molly" = "#E1AF00", "Atlantic molly"="#9986A5")) +
  scale_fill_manual(name = "Species",
                    values = c("Sailfin molly" = "#78B7C5", "Amazon molly" = "#E1AF00", "Atlantic molly"="#9986A5")) +
  scale_shape_manual(name = "Species", 
                     values = c("Sailfin molly" = 17, "Amazon molly" = 16, "Atlantic molly"= 3)) +
  guides(fill = guide_legend(override.aes = list(fill = NA))) +
  labs(
    x = "Categorical PC1",
    y = "Categorical PC2"
  ) +
  theme_classic() +
  labs(
    x = paste0("Categorical PC1 (", pc1_percent, "%)"),
    y = paste0("Categorical PC2 (", pc2_percent, "%)")
  )

tam.catpca

```


#### *Post PCA tests*

**CONCLUSIONS:** For our metric data, there is no significant difference in average shape distribution for PC1 or PC3, but there is a significant difference for PC2. This is expected, as all three species have different body shapes and therefore occupy distinct morphospace. PC1 is influenced by body size, so PC2 and PC3 are of more interest. The species occupy a different morphospace for PC2 which explains 27.1% of the shape variation, but occupy similar morphospace for PC3 which only explains 4.48%. 

PC1 and PC2 both show a significant difference in variation, with pochoc tests revealing the sailfin-Atlantic comparisons to be driving this. Some of the traits that load heavily on PC1 and PC2 (e.g. body depth, dorsal base length) are also the traits that were best supported by unequal variances in our individual models, with the sexual species again having more variation than the asexual, so the PCA results seem to corroborate this. This indicates that for both individual traits and when traits are taken all together, the sexual species in Tampico, Mexico are more variable than the asexual species.


For our meristic data, both CatPC1 and CatPC2 showed a significant difference in average meristic profiles, which again makes sense as many of the meristic traits are species specific. There is also a significant difference in variation for CatPC1, with the pairwise comparisons showing a significant difference for sailfin and atlantic only; there was no significant difference in CatPC2. This differs from our individual trait analysis, indicating species differences in covariation among multiple meristic features. 


```{r}

#quick distribution checks to see what tests to use

testmod <- lm(Dim.1 ~ Species, data = pca.df2)
plot(simulateResiduals(testmod)) #fails levene's test; will use kruskal-wallis

testmod <- lm(Dim.2 ~ Species, data = pca.df2)
plot(simulateResiduals(testmod)) #fails levene's test; will use kruskal-wallis

testmod <- lm(Dim.3 ~ Species, data = pca.df2)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(CatPC1 ~ Species, data = catpca.df2)
plot(simulateResiduals(testmod)) #fails outlier and levene's test; will use kruskal-wallis

testmod <- lm(CatPC2 ~ Species, data = catpca.df2)
plot(simulateResiduals(testmod)) #fails KS test and outlier test; will use kruskal-wallis




# ───────────────────────────────────────────────────
## MEDIAN TESTS (Kruskal-Wallis test for all)
# ───────────────────────────────────────────────────

### Metric Data

tam.met.mean.1 <- kruskal.test(Dim.1~Species, data = pca.df2) #not significant

tam.met.mean.2 <- kruskal.test(Dim.2~Species, data = pca.df2) #significant (p<0.001)
    tam_post_hoc1 <- dunnTest(Dim.2 ~ Species, data = pca.df2, method = "holm")
    tam_post_hoc1 #all comparisons significant (p<0.001)


tam.met.mean.3 <- kruskal.test(Dim.3~Species, data = pca.df2) #not significant


### Meristic Data


tam.cnt.mean.1 <- kruskal.test(CatPC1~Species, data = catpca.df2) #significant (p<0.001)
    tam_post_hoc2 <- dunnTest(CatPC1 ~ Species, data = catpca.df2, method = "holm")
    tam_post_hoc2 #all comparisons significant (p<0.001)

tam.cnt.mean.2 <- kruskal.test(CatPC2~Species, data = catpca.df2) #significant (p=0.009)
    tam_post_hoc3 <- dunnTest(CatPC2 ~ Species, data = catpca.df2, method = "holm")
    tam_post_hoc3 #amazon-atlantic (p=0.0.019) and sailfin-atlatnic (p=0.021) significant but not amazon-sailfin

# ───────────────────────────────────────────────────
## VARIANCE TESTS (Brown-Forsythe for all)
# ───────────────────────────────────────────────────

### Metric Data
tam.met.var.1 <- leveneTest(Dim.1~Species, data= pca.df2) #significant (p=0.016)

    tam_post_hoc4 <- with(pca.df2, {
      medians <- tapply(Dim.1, Species, median, na.rm = TRUE)
      abs_devs <- abs(Dim.1 - medians[Species])
      pairwise.t.test(abs_devs, Species, p.adjust.method = "holm", pool.sd = FALSE)})
    tam_post_hoc4 #amazon-sailfin marginal (p=0.056), sailfin-atlantic significant (p=0.009) and amazon-atlatnic not significant


tam.met.var.2 <- leveneTest(Dim.2~Species, data= pca.df2) #significant (p=0.005)
    tam_post_hoc5 <- with(pca.df2, {
      medians <- tapply(Dim.2, Species, median, na.rm = TRUE)
      abs_devs <- abs(Dim.2 - medians[Species])
      pairwise.t.test(abs_devs, Species, p.adjust.method = "holm", pool.sd = FALSE)})
    tam_post_hoc5 #amazon-atlantic marginal (p=0.049), sailfin-atlantic significant (p=0.012) and amazon-sailfin not significant

tam.met.var.3 <- leveneTest(Dim.3~Species, data= pca.df2) #not significant


### Meristic Data
tam.cnt.var.1 <- leveneTest(CatPC1~Species, data= catpca.df2) #significant (p=0.002)
    tam_post_hoc6 <- with(catpca.df2, {
      medians <- tapply(CatPC1, Species, median, na.rm = TRUE)
      abs_devs <- abs(CatPC1 - medians[Species])
      pairwise.t.test(abs_devs, Species, p.adjust.method = "holm", pool.sd = FALSE)})
    tam_post_hoc6 #sailfin-atlantic significant (p=0.004) and amazon-sailfin and amazon-atlantic not significant

tam.cnt.var.2 <- leveneTest(CatPC2~Species, data= catpca.df2) #not significant


# ───────────────────────────────────────────────────
## TABLE
# ───────────────────────────────────────────────────

### Metric Data

#get mean and variance for each dimension per species
met_summary2 <- pca.df2 %>%
  group_by(Species) %>%
  summarise(
    PC1_median = median(Dim.1, na.rm = TRUE),
    PC1_mad  = mad(Dim.1, na.rm = TRUE, constant = 1),
    PC2_median = median(Dim.2, na.rm = TRUE),
    PC2_mad  = mad(Dim.2, na.rm = TRUE, constant = 1),
    PC3_median = median(Dim.3, na.rm = TRUE),
    PC3_mad  = mad(Dim.3, na.rm = TRUE, constant = 1))

#isolate results
amazon_met  <- filter(met_summary2, Species == "Amazon molly")
sailfin_met <- filter(met_summary2, Species == "Sailfin molly")
atlantic_met <- filter(met_summary2, Species == "Atlantic molly")

#extract wilcox test results, levene's results, and summary stats
tam_postPCA_result <- tibble(
  Classifier = "Metric PCA",
  Dimension = c("PC1", "PC2", "PC3"),
  Amazon = c(
    sprintf("%.2f (%.2f)", amazon_met$PC1_median, amazon_met$PC1_mad),
    sprintf("%.2f (%.2f)", amazon_met$PC2_median, amazon_met$PC2_mad),
    sprintf("%.2f (%.2f)", amazon_met$PC3_median, amazon_met$PC3_mad)),
  Sailfin = c(
    sprintf("%.2f (%.2f)", sailfin_met$PC1_median, sailfin_met$PC1_mad),
    sprintf("%.2f (%.2f)", sailfin_met$PC2_median, sailfin_met$PC2_mad),
    sprintf("%.2f (%.2f)", sailfin_met$PC3_median, sailfin_met$PC3_mad)),
  Atlantic = c(
    sprintf("%.2f (%.2f)", atlantic_met$PC1_median, atlantic_met$PC1_mad),
    sprintf("%.2f (%.2f)", atlantic_met$PC2_median, atlantic_met$PC2_mad),
    sprintf("%.2f (%.2f)", atlantic_met$PC3_median, atlantic_met$PC3_mad)),
  H_value = c(tam.met.mean.1$statistic,
              tam.met.mean.2$statistic,
              tam.met.mean.3$statistic),
  p_kw = c(tam.met.mean.1$p.value,
           tam.met.mean.2$p.value,
           tam.met.mean.3$p.value),
  F_value = c(tam.met.var.1$`F value`[1],
              tam.met.var.2$`F value`[1],
              tam.met.var.3$`F value`[1]),
  p_lv = c(tam.met.var.1$`Pr(>F)`[1],
           tam.met.var.2$`Pr(>F)`[1],
           tam.met.var.3$`Pr(>F)`[1])) %>%
  mutate(
    Med_test = case_when(
      p_kw < 0.001 ~ paste0("H = ", sprintf("%.2f", H_value), 
                            "<br>*p* < 0.001*"),
      p_kw <0.05   ~ paste0("H = ", sprintf("%.2f", H_value), 
                            "<br>*p* = ", sprintf("%.3f", p_kw), "*"),
       TRUE        ~ paste0("H = ", sprintf("%.2f", H_value),
                            "<br>*p* = ", sprintf("%.3f", p_kw))),

    Variance_test = case_when(
      p_lv < 0.001 ~ paste0("F = ", sprintf("%.2f", F_value), 
                            "<br>*p* < 0.001*"),
      p_lv <0.05   ~ paste0("F = ", sprintf("%.2f", F_value), 
                            "<br>*p* = ", sprintf("%.3f", p_lv), "*"),
       TRUE        ~ paste0("F = ", sprintf("%.2f", F_value),
                            "<br>*p* = ", sprintf("%.3f", p_lv))))




### Meristic Data

#get mean and variance for each dimension per species
cnt_summary2 <- catpca.df2 %>%
  group_by(Species) %>%
  summarise(
    CPC1_median = median(CatPC1, na.rm = TRUE),
    CPC1_mad  = mad(CatPC1, na.rm = TRUE),
    CPC2_median = median(CatPC2, na.rm = TRUE),
    CPC2_mad  = mad(CatPC2, na.rm = TRUE))

#isolate results
amazon_cnt  <- filter(cnt_summary2, Species == "Amazon molly")
sailfin_cnt <- filter(cnt_summary2, Species == "Sailfin molly")
atlantic_cnt <- filter(cnt_summary2, Species == "Atlantic molly")


#extract wilcox test results, levene's results, and summary stats
tam_postCPCA_result <- tibble(
  Classifier = "Meristic CATPCA",
  Dimension = c("CatPC1", "CatPC2"),
  Amazon = c(
    sprintf("%.2f (%.2f)", amazon_cnt$CPC1_median, amazon_cnt$CPC1_mad),
    sprintf("%.2f (%.2f)", amazon_cnt$CPC2_median, amazon_cnt$CPC2_mad)),
  Sailfin = c(
    sprintf("%.2f (%.2f)", sailfin_cnt$CPC1_median, sailfin_cnt$CPC1_mad),
    sprintf("%.2f (%.2f)", sailfin_cnt$CPC2_median, sailfin_cnt$CPC2_mad)),
  Atlantic = c(
    sprintf("%.2f (%.2f)", atlantic_cnt$CPC1_median, atlantic_cnt$CPC1_mad),
    sprintf("%.2f (%.2f)", atlantic_cnt$CPC2_median, atlantic_cnt$CPC2_mad)),
  H_value = c(tam.cnt.mean.1$statistic,
              tam.cnt.mean.2$statistic),
  p_kw = c(tam.cnt.mean.1$p.value,
           tam.cnt.mean.2$p.value),
  F_value = c(tam.cnt.var.1$`F value`[1],
              tam.cnt.var.2$`F value`[1]),
  p_lv = c(tam.cnt.var.1$`Pr(>F)`[1],
           tam.cnt.var.2$`Pr(>F)`[1])) %>%
  mutate(
    Med_test = case_when(
      p_kw < 0.001 ~ paste0("H = ", sprintf("%.2f", H_value), 
                            "<br>*p* < 0.001*"),
      p_kw <0.05   ~ paste0("H = ", sprintf("%.2f", H_value), 
                            "<br>*p* = ", sprintf("%.3f", p_kw), "*"),
       TRUE        ~ paste0("H = ", sprintf("%.2f", H_value),
                            "<br>*p* = ", sprintf("%.3f", p_kw))),

    Variance_test = case_when(
      p_lv < 0.001 ~ paste0("F = ", sprintf("%.2f", F_value), 
                            "<br>*p* < 0.001*"),
      p_lv <0.05   ~ paste0("F = ", sprintf("%.2f", F_value), 
                            "<br>*p* = ", sprintf("%.3f", p_lv), "*"),
       TRUE        ~ paste0("F = ", sprintf("%.2f", F_value),
                            "<br>*p* = ", sprintf("%.3f", p_lv))))




### Add posthoc test results
tam_extract_pairs <- function(pair_obj, dim_name, test.type = "M") {
  if (is.null(pair_obj)) return(NULL)
  
  # CASE A: If it's a Dunn Test object (FSA package for Medians)
  if (inherits(pair_obj, "dunnTest") || "res" %in% names(pair_obj)) {
    df <- pair_obj$res %>%
      rename(Contrast = Comparison, p.value = P.adj)
  } 
  # CASE B: If it's a pairwise.t.test object (Base R matrix for Variances)
  else if (inherits(pair_obj, "pairwise.htest")) {
    p_mat <- pair_obj$p.value
    df <- as.data.frame(as.table(p_mat)) %>%
      filter(!is.na(Freq)) %>%
      mutate(Contrast = paste0(Var2, " - ", Var1)) %>%
      rename(p.value = Freq)
  } else {
    return(NULL)
  }
  
  df %>%
    mutate(
      Dimension = dim_name,
      type = test.type,
      p_fmt = case_when(
        p.value < 0.001 ~ "*p* < 0.001*",
        p.value < 0.05  ~ sprintf("*p* = %.3f*", p.value),
        TRUE            ~ sprintf("*p* = %.3f", p.value)),
      Contrast = gsub(" molly", "", Contrast), 
      PostHoc_Sum = sprintf("%s: %s", Contrast, p_fmt)) %>%
    select(Dimension, type, PostHoc_Sum)
}


posthoc_ref2 <- bind_rows(
  tam_extract_pairs(tam_post_hoc1, "PC2", "M"),
  tam_extract_pairs(tam_post_hoc2, "CatPC1", "M"),
  tam_extract_pairs(tam_post_hoc3, "CatPC2", "M"),
  tam_extract_pairs(tam_post_hoc4, "PC1", "V"),
  tam_extract_pairs(tam_post_hoc5, "PC2", "V"),
  tam_extract_pairs(tam_post_hoc6, "CatPC1", "V")) %>%
  group_by(Dimension, type) %>%
  summarise(
    Combined_Pairs = paste(PostHoc_Sum, collapse = "<br>"), .groups = "drop") %>%
  pivot_wider(
    names_from = type,
    values_from = Combined_Pairs,
    names_prefix = "PostHoc_")

### Combine & form table

tam_results_combined <- bind_rows(tam_postPCA_result, tam_postCPCA_result) %>%
  mutate(
    Med_Part = paste0("Median: H = ", sprintf("%.2f", H_value), 
                      " (*p* ", ifelse(p_kw < 0.001, "< 0.001*", paste0("= ", sprintf("%.3f", p_kw), ifelse(p_kw < 0.05, "*", ""))), ")"),
    Var_Part = paste0("Variance: F = ", sprintf("%.2f", F_value), 
                      " (*p* ", ifelse(p_lv < 0.001, "< 0.001*", paste0("= ", sprintf("%.3f", p_lv), ifelse(p_lv < 0.05, "*", ""))), ")"),
    Test_Summary = paste0(Med_Part, "<br> ", Var_Part),
    Group_Label = paste0("**", Classifier, "**<br>",
                         "***", Dimension, "***<br>",
                         "<span style='font-size: 85%'>", Test_Summary, "</span>")) %>%
  left_join(posthoc_ref2, by = "Dimension") %>%
  mutate(PostHoc_M = coalesce(PostHoc_M, "&mdash;"),
         PostHoc_V = coalesce(PostHoc_V, "&mdash;"))
  



tam_postPCA_table <- tam_results_combined %>%
  select(
    Group_Label,
    Amazon,
    Sailfin,
    Atlantic,
    PostHoc_M,
    PostHoc_V) %>%
  gt() %>%
  fmt_markdown(columns = everything()) %>%
  cols_label(
    Group_Label ="",
    Amazon = md("**Amazon**<br>Median (MAD)"),
    Sailfin = md("**Sailfin**<br>Median (MAD)"),
    Atlantic = md("**Atlantic**<br>Median (MAD)"),
    PostHoc_M = md("**Pairwise Medians**<br><span style='font-size:85%; font-weight:normal;'>Holm-adj. p</span>"),
    PostHoc_V = md("**Pairwise Variances**<br><span style='font-size:85%; font-weight:normal;'>Holm-adj. p</span>")) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c(Group_Label, Amazon, Sailfin, Atlantic, PostHoc_M))) %>%
  cols_align(align = "center", columns = everything()) %>%
  opt_horizontal_padding(scale = 2)

#final table
tam_postPCA_table




```

***
***