---
title: "Filtered_poeciliid-morphology_2026"
author: "Allison Davis"
date: "2026-06-27"
output: 
  html_document:
    toc: TRUE
    toc_depth: 6
    toc_float:
      collapsed: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **PURPOSE**

We aim to understand the interplay of genetic variation and environmental heterogeneity in the persistence of asexual-sexual systems through comparisons of morphological variation. In this study, we compare 22 morphological traits in preserved Amazon molly (*Poecilia formosa*) and its parental species (sailfin molly (*P. latipinna*), Atlantic molly (*P. mexicana*)) from across their native and introduced sympatric range. While we expect morphological variation in all species because of local adaptations we predict the variation of the sexual sailfin and Atlantic species to be higher than the Amazon species due to the presence of genetic recombination.

In part I, we analyzed all processed samples (186 Amazon mollies, 156 female sailfin mollies, and 36 female Atlantic mollies)[***See Poeciliid-morphology-analysis_2026.Rmd***]. 

In part II, we will filter for adults (standard length > 32.5mm), leaving us with 107 Amazon mollies, 90 female sailfin mollies, and 24 female Atlantic mollies. For the main analysis, we will focus on Amazon and sailfin mollies only. We will then extract Amazon and sailfin samples from the Tampico, Mexico region to compare to our Atlantic molly samples, as our Atlantic molly samples only originate from Tampico, MX. For our geographic analysis, we will assess variation in Amazons and sailfins only, across both latitudinal-based zones and river drainage basins, as basins are often more reflective of genetic distribution of fishes. *NOTE:* Our coefficient of variation analysis can be found in **COV_final.Rmd**.


***
***

# **PART II: MATURE FISH**

The following analysis will compare variation among adult fish (>32.5mm).After filtering for adults and removing all samples with NA values, the final numbers are: 107 Amazons, 90 sailfins, and 24 Atlantics.

For the main analysis, we will use only Amazons and sailfin; comparisons to Atlantic mollies can be found in the **Tampico MX** section.


## **Data**

All analyses were conducted using R version 4.6.1

### Load required libraries 

Remove the '#' before `install.packages` if it is the first time installing library. 

```{r, warning=FALSE}

#install.packages("curl")             #v7.1.0
library(curl)

#install.packages("dplyr")            #v2.6.0
library(dplyr)

#install.packages("tidyverse")        #v2.0.0
library(tidyverse)

#install.packages("tidyr")            #v1.3.2
library(tidyr)

#install.packages("gt")               #v1.3.0
library(gt)

#install.packages("factoextra")       #v2.1.0
library(factoextra)

#install.packages("car")              #v3.1-5
library(car)

#install.packages("AMR")              #v3.0.1
library(AMR)

#install.packages("ggplot2")          #v4.0.3
library(ggplot2)

#install.packages("ggfortify")        #v0.4.19
library(ggfortify)

#install.packages("ggbiplot")         #v0.6.2
library(ggbiplot)

#install.packages("emmeans")          #v2.0.3
library(emmeans)

#install.packages("glmmTMB")          #v1.1.14
library(glmmTMB)

#install.packages("DHARMa")           #v0.5.0
library(DHARMa)

#install.packages("Gifi)              #v1.0-0
library(Gifi)

#install.packages("FSA")              #v0.10.1
library(FSA)

```




### Read in data


```{r}
# OPTION 1: download 'Final_morph-data_2026.csv' into working directory 
#raw <- read.csv("Final_morph-data_2026.csv", header = TRUE, sep = ",", stringsAsFactors = TRUE)
#head(raw)


# OPTION 2: load from Github repository
raw <- curl("https://raw.githubusercontent.com/allisondavis/morphology_analysis/refs/heads/master/Final_morph-data_2026.csv")

raw <- read.csv(raw, header = TRUE, sep = ",", stringsAsFactors = TRUE)


# Removing info we don't need
## Note: Right & left pelvic (P2.L/R) identical for both species in all samples
## Note: Total length (TL) is missing data; will use standard length (SL)
## Note: Sample size too sparse in watersheds (hydro.lv8); geographic analysis will use basin (hydro.lv6; hydroBASIN level 6)
raw1 <- select(raw, -c(INSTITUTE, ID.NUM, LOCATION, LAT, LONG, SAMPLING.DATE, MEASUREMENT.DATE, PHOTO.DATE, TAG.., P2.L, P2.R, TL, HYDRO.LV8 ))


# Remove samples with missing values
raw1 <- raw1[complete.cases(raw1[, 4:23]),] 



# Make data more reader friendly
## Species
raw1$SPP <- factor(raw1$SPP,
                   levels = c("p.latipinna", "p.formosa", "p.mexicana"),
                        labels = c("Sailfin molly", "Amazon molly", "Atlantic molly"))

## Zones
raw1$ZONE <- factor(raw1$ZONE,
                    levels = c("Z1", "Z2", "Z3", "Z4"),
                    labels = c("Zone 1", "Zone 2", "Zone 3", "Zone 4"))

## Basins
raw1$HYDRO.LV6 <- factor(raw1$HYDRO.LV6, 
                         levels = c("7060048310", "7060048490", "7060048620", "7060048630", "7060048870", "7060049130", "7060049270", "7060049500", "7060049830", "7060820630", "7060824100"),
                         labels = c("Brazos-Colorado Coastal", "Lavaca-Guadalupe Coastal", "San Antonio-Guadalupe", "San Antonio-Nueces Coastal", "Upper Southwest Texas Coastal", "Lower Southwest Texas Coastal", "Lower R\u00edo Grande", "R\u00edo San Fernando", "Lagunas de San Andr\u00e9s-Morales", "R\u00edo Tames\u00ed", "R\u00edo Tamu\u00edn"))


raw1 <- raw1 %>%
  rename(Species = SPP,
         Zone = ZONE,
         Basin = HYDRO.LV6)

raw1 <- filter(raw1,SL>=32.5) ## FILTER: mature adults only

## Trait names
trait.names <- c(
  D = "Dorsal rays",
  P1 = "Left pectoral rays",
  P1.R = "Right pectoral rays",
  A = "Anal rays", 
  LLSC = "Lateral line scale count",
  SALL = "Scales above lateral line",
  SBLL = "Scales below lateral line", 
  SBDF = "Scales before dorsal fin",
  SL = "Standard length", 
  BD = "Body depth", 
  CPD = "Caudal peduncle depth",
  CPL = "Caudal peduncle length",
  PreDL = "Predorsal length",
  DbL = "Dorsal base length",
  HL = "Head length",
  HD = "Head depth",
  HW = "Head width",
  SnL = "Snout length", 
  OL = "Orbit length",
  FLA = "Fluctuating asymmetries")


# Final data frames

## main analysis
raw2 <- raw1 %>% 
  filter(Species != "Atlantic molly") %>% 
  droplevels()

raw2$Species <- factor(raw2$Species, levels = sort(unique(as.character(raw2$Species))))



## Tampico MX analysis
geo <- raw1 %>%
  filter(COUNTY.ST == "Tamaulipas/MX") %>%
  droplevels()

geo$Species <- factor(geo$Species, levels = sort(unique(as.character(geo$Species))))
```



### Summary stats

```{r}

#create long df to easily summarize later
dat.long <- raw2 %>%
  pivot_longer(
    cols = c(D, P1, A, P1.R, LLSC, SALL, SBLL, SBDF, SL, BD, CPD, CPL, PreDL, DbL, HL, HD, HW, SnL, OL, FLA),
    names_to = "Trait",
    values_to = "Value")

#gather summary statistics 
summ.stats <- dat.long %>%
  group_by(Species, Trait) %>%
  summarize(
    mean = mean(Value, na.rm = TRUE),
    sd   = sd(Value, na.rm = TRUE),
    min  = min(Value, na.rm = TRUE),
    med  = median(Value, na.rm = TRUE),
    max  = max(Value, na.rm = TRUE),
    .groups = "drop")


# Species sample sizes
n.df <- raw2 %>%
  count(Species, name = "n")

# Add species label with n
summ.stats2 <- summ.stats %>%
  left_join(n.df, by = "Species") %>%
  mutate(
    Species_n = paste0(Species, " (n = ", n, ")"),
    Trait_name = trait.names[Trait],
    Trait_name = factor(
      Trait_name,
      levels = unname(trait.names))) %>%
  arrange(Species_n, Trait_name) %>%
  select(Species_n, Trait_name, mean, sd, min, med, max)


#final table
summ.stats.all <- summ.stats2 %>%
  gt(
    rowname_col = "Trait_name",
    groupname_col = "Species_n") %>%
  fmt_number(
    columns = c(mean, sd, min, med, max),
    decimals = 2) %>%
  tab_style(
      style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
      locations = cells_body(columns = c(mean, sd, min, med, max))) %>%
  opt_horizontal_padding(scale=2) %>%
  cols_align(align="center", everything())

summ.stats.all



```


### Normality

Since the data is now artificially truncated for SL >= 32.5mm, normality checks on raw data are not as useful. Instead, I will check model residuals and alter family if needed to adjust for the truncated data.

***
***




## **Analysis**

### Individual trait models

We are interested in whether there is a difference in morphological variance among the asexual Amazon molly and the sexual sailfin mollies. We will therefore need to use tests that allow for unequal variances among species, and see if this better fits the data compared to equal variances. We will also include standard length (SL) as a covariate to account for any affect of SL on the trait.

For both metric and meristic traits, we will use a ***generalized linear mixed model (GLMM) analysis*** that allows for different dispersions by species.

For metric traits, we will compare a model with equal variances (mod1 <- glmmTMB(log(trait)~log(SL) * species, family= gaussian)) to one that allows unequal variances among species (mod2 <- glmmTMB(log(trait)~log(SL) * species, dispformula=~species, family= gaussian))). We will compare models using an ANOVA test to determine if incorporating unequal variances among species better fits the data.

For meristic traits, we will compare a model with equal variances (mod1 <- glmmTMB(trait~SL * species, family= compois)) to one that allows unequal variances among species (mod2 <- glmmTMB(trait~SL * species, dispformula=~species, family= compois))). We will compare models using an ANOVA test to determine if incorporating unequal variances among species better fits the data. A *Conway-Maxwell-poisson distribution* will be used to control for underdispersion (see note below).

**NOTE:** we previously performed the meristic model analysis with a *poisson distribution* but all our dispersion ratios were <0.05 indicating severe underdispersion. This is not unexpected, given that most traits had very few values that all fish fell into (e.g. dorsal rays of 12, 13, or 14). So the analysis below is now using a *Conway-Maxwell-poisson distribution* to adjust for underdispersion. See **Poisson-meristic-analysis.R** for previous results. 

**CONCLUSIONS:** All metric and meristic traits showed no difference in model support, and are thus best supported by an equal-variance model (aka no significant difference in variation among the species). *Same as main analysis*

```{r}

#create separate df for metric and meristic (count) data
metric <- raw2 %>%
  select(-c(5:12)) %>%
  select(-c("FLA"))

count <- raw2 %>%
  select(-c(14:23)) #still includes SL for analysis

met.vars <- c("BD", "CPD", "CPL", "PreDL", "DbL", "HL", "HD", "HW", "SnL", "OL")
cnt.vars <- c("D", "P1", "P1.R", "A", "LLSC", "SALL", "SBLL", "SBDF", "FLA")


# ───────────────────────────────────────────────────
## METRIC ANALYSIS 
# ───────────────────────────────────────────────────


# Ensure Species is a factor
metric$Species <- as.factor(metric$Species)
metric$Species <- relevel(metric$Species, ref = "Amazon molly")

results_list <- lapply(met.vars, function(trait) {
  
  # Build the formula string correctly
  form_str <- paste0("log(", trait, ") ~ log(SL) * Species")
  form <- as.formula(form_str)

  # 1. Homogeneous variance model (with safety catch)
  mod1 <- tryCatch({
    glmmTMB(form, data = metric, family = gaussian)
  }, error = function(e) { NULL }, warning = function(w) { NULL })
  
  # 2. Heterogeneous variance model (with strict optimization catching)
  mod2 <- tryCatch({
    glmmTMB(form, data = metric, family = gaussian, dispformula = ~Species)
  }, 
  error = function(e) { NULL }, 
  warning = function(w) { 
    if(grepl("NA/NaN|calc_loglambda|iteration|failure", w$message, ignore.case = TRUE)) {
      return(NULL) 
    }
    invokeRestart("muffleWarning") 
  })
  
  # Verify clean convergence of both models
  models_valid <- !is.null(mod1) && !is.null(mod2) && 
                  !is.na(AIC(mod1)) && !is.na(AIC(mod2)) && 
                  !any(is.nan(summary(mod2)$coefficients$disp[,2]))

  unique_species <- levels(as.factor(metric$Species))
  
  # --- PATH A: MODEL CONVERGED CLEANLY ---
  if (models_valid) {
    # Compare models
    comp <- anova(mod1, mod2)
    pval <- comp$`Pr(>Chisq)`[2]
    delta_aic <- AIC(mod2) - AIC(mod1)
    
    # Run DHARMa residual check
    best_mod <- if (AIC(mod1) < AIC(mod2)) mod1 else mod2
    best_mod_name <- if (AIC(mod1) < AIC(mod2)) "Homogeneous (mod1)" else "Heterogeneous (mod2)"
    
    message(paste0("--- DHARMa Diagnostics for Trait: ", trait, " (Best Model: ", best_mod_name, ") ---"))
    sim_res <- simulateResiduals(best_mod, plot = FALSE)
    
    # Plot the residuals automatically
    plot(sim_res) 
    title(main = paste(trait, "-", best_mod_name), line = 2.5)
    
    # Extract predicted log-variance per species
    rg <- ref_grid(mod2, component = "disp")
    disp_means <- as.data.frame(emmeans(rg, specs = ~ Species))
    disp_means$Variance <- exp(disp_means$emmean)
    
    # Pivot to easily calculate the single global ratio
    ratio_df <- disp_means %>%
      select(Species, Variance) %>%
      pivot_wider(names_from = Species, values_from = Variance) %>%
      mutate(
        Var_Ratio  = `Sailfin molly` / `Amazon molly`,
        Delta_AIC  = delta_aic, 
        LRT_p      = pval,
        Best_Model = best_mod_name)
      
  } else {
    # --- PATH B: FALLBACK TO SIZE-CORRECTED LEVENE'S TEST ---
    best_mod_name <- "Model Failed - Levene's Test Fallback"
    message(paste0("!!! glmmTMB Failed for Trait: ", trait, ". Running Levene's Fallback. !!!"))
    
    # Step 1: Strip out body size dependencies using linear model on log traits
    lm_check <- lm(as.formula(paste0("log(", trait, ") ~ log(SL)")), data = metric)
    metric$Abs_Resid <- abs(residuals(lm_check))
    
    # Step 2: Run a formal Levene's-style test via anova on absolute deviations
    lev_test <- tryCatch({
      leveneTest(Abs_Resid ~ Species, data = metric)
    }, error = function(e) { NULL })
    fallback_p <- if(!is.null(lev_test)) lev_test$`Pr(>F)`[1] else NA
    
    # Step 3: Calculate empirical size-corrected variances per group
    empirical_vars <- metric %>%
      group_by(Species) %>%
      summarise(Variance = var(log(!!sym(trait)), na.rm = TRUE), .groups = 'drop') %>%
      complete(Species = unique_species, fill = list(Variance = NA))
    
    ratio_df <- empirical_vars %>%
      pivot_wider(names_from = Species, values_from = Variance) %>%
      mutate(
        Var_Ratio  = `Sailfin molly` / `Amazon molly`,
        Delta_AIC  = NA, # Marked NA so it shows as blank/dash in your table
        LRT_p      = fallback_p,
        Best_Model = best_mod_name)
  }
  
  # Append baseline trait tracker string
  ratio_df$Trait <- trait
  return(ratio_df)
})

results_df <- do.call(rbind, results_list)

results_df$p_adj <- p.adjust(results_df$LRT_p,
                             method = "holm")

met.results <- results_df %>%
  mutate(
    p.display = case_when(
      p_adj < 0.001 ~ paste0("*p* < 0.001*"),
      p_adj < 0.05  ~ paste0(sprintf("%.3f", p_adj), "*"),
      TRUE ~ sprintf("%.3f", p_adj))) %>%
  mutate(
    Trait_Name = factor(
      trait.names[Trait],
      levels = unname(trait.names))) %>%
  arrange(Trait_Name) %>%
  relocate(Trait_Name, .before = everything())

met.tbl <- met.results %>%
  gt() %>%
  cols_hide(c(Trait, LRT_p, p_adj, Best_Model, "Sailfin molly", "Amazon molly")) %>%
  cols_label(
    Trait_Name = "Trait",
    Var_Ratio = md("Variance ratio <br>(Sailfin / Amazon)"),
    Delta_AIC = md("&Delta;AIC"),
    p.display = md("adj. *p*-value")) %>%
  cols_move(
    columns = c(Delta_AIC, p.display),
    after = Trait_Name) %>%
  fmt_number(
    columns = c(Var_Ratio, Delta_AIC),
    decimals = 2) %>%
  fmt_markdown(columns = p.display) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Trait_Name, Delta_AIC, Var_Ratio))) %>%
  cols_align("center", everything()) %>%
  opt_horizontal_padding(scale = 2)
  
  
met.tbl



# ───────────────────────────────────────────────────
## MERISTIC (COUNT) ANALYSIS 
# ───────────────────────────────────────────────────

# Ensure Species is a factor matching the metric settings
count$Species <- as.factor(count$Species)
count$Species <- relevel(count$Species, ref = "Amazon molly")

results_list_cnt <- lapply(cnt.vars, function(trait) {
  
  form_str <- paste0(trait, " ~ SL * Species")
  form <- as.formula(form_str)
  
  # 1. Fit Baseline Homogeneous Model
  mod1 <- tryCatch({
    glmmTMB(form, data = count, family = compois)
  }, error = function(e) { NULL }, warning = function(w) { NULL })
  
  # 2. Fit Heterogeneous Model (with proper warning muffling)
  mod2 <- tryCatch({
    withCallingHandlers({
      glmmTMB(form, data = count, family = compois, dispformula = ~Species)
    }, warning = function(w) {
      # If it's a fatal mathematical breakdown, stop the model so tryCatch catches it
      if(grepl("NA/NaN|calc_loglambda|iteration|failure", w$message, ignore.case = TRUE)) {
        stop(w$message) 
      }
      # Otherwise, safely muffle the minor convergence warning and keep going
      invokeRestart("muffleWarning") 
    })
  }, error = function(e) { NULL })

  # Verify if both models successfully converged
  models_valid <- !is.null(mod1) && !is.null(mod2) && 
                  !is.na(AIC(mod1)) && !is.na(AIC(mod2)) && 
                  !any(is.nan(summary(mod2)$coefficients$disp[,2]))

  # Keep original factor levels intact
  species_levels <- levels(count$Species)
  
  # --- PATH A: GLMMTMB CONVERGED CLEANLY ---
  if (models_valid) {
    comp <- anova(mod1, mod2)
    pval <- comp$`Pr(>Chisq)`[2]
    delta_aic <- AIC(mod2) - AIC(mod1)
    
    best_mod <- if (AIC(mod1) < AIC(mod2)) mod1 else mod2
    best_mod_name <- if (AIC(mod1) < AIC(mod2)) "Homogeneous (mod1)" else "Heterogeneous (mod2)"
    
    message(paste0("--- DHARMa Diagnostics for Trait: ", trait, " (Best Model: ", best_mod_name, ") ---"))
    sim_res <- simulateResiduals(best_mod, plot = FALSE)
    plot(sim_res) 
    title(main = paste(trait, "-", best_mod_name), line = 2.5)
    
 # --- FIXED PATH A: BYPASS PREDICT() COMPLETELY TO AVOID MATRIX CRASH ---
    mean_sl <- mean(count$SL, na.rm = TRUE)
    
    # Manually compute Mean (mu) and Variance for each species at mean_sl 
    # to completely circumvent glmmTMB's predict.glmmTMB internal bugs
    variances <- sapply(species_levels, function(sp) {
      
      # 1. Calculate Expected Mean (mu) from conditional fixed effects
      # Formula scale is log-link for compois family
      beta <- fixef(mod2)$cond
      log_mu <- beta["(Intercept)"] + 
                beta["SL"] * mean_sl + 
                (if(sp == "Amazon molly") 0 else beta[paste0("Species", sp)]) +
                (if(sp == "Amazon molly") 0 else beta[paste0("SL:Species", sp)]) * mean_sl
      mu <- exp(as.numeric(log_mu))
      
      # 2. Calculate Dispersion (phi) from dispersion fixed effects
      # Formula scale is log-link for dispersion component
      alpha <- fixef(mod2)$disp
      log_phi <- alpha["(Intercept)"] + 
                 (if(sp == "Amazon molly") 0 else alpha[paste0("Species", sp)])
      phi <- exp(as.numeric(log_phi))
      
      # COM-Poisson approximate variance product
      return(mu * phi)
    })
    
    # Shape into the required output data frame structure
    output_df <- tibble(
      Species = species_levels,
      Variance = variances
    ) %>%
      pivot_wider(names_from = Species, values_from = Variance) %>%
      mutate(
        Var_Ratio   = `Sailfin molly` / `Amazon molly`,
        Delta_AIC  = delta_aic,
        LRT_p      = pval,
        Best_Model = best_mod_name)
  } else {
    # --- PATH B: CONVERGENCE FAILED -> FALLBACK TO LEVENE'S TEST ---
    best_mod_name <- "Model Failed Optimization - Used Levene's Test Fallback"
    message(paste0("!!! glmmTMB Failed for Trait: ", trait, ". Running Levene's Test Fallback. !!!"))
    
    levene_form <- as.formula(paste0(trait, " ~ Species"))
    lev_test <- tryCatch({
      leveneTest(levene_form, data = count)
    }, error = function(e) { NULL })
    
    raw_vars <- count %>%
      group_by(Species) %>%
      summarise(Variance = var(!!sym(trait), na.rm = TRUE), .groups = 'drop') %>%
      complete(Species = species_levels, fill = list(Variance = NA))
    
    wide_vars <- raw_vars %>%
      pivot_wider(names_from = Species, values_from = Variance)
    
    l_p <- if(!is.null(lev_test)) lev_test$`Pr(>F)`[1] else NA
    
    output_df <- wide_vars %>%
      mutate(
        Var_Ratio   = `Sailfin molly` / `Amazon molly`,
        Delta_AIC  = NA,   
        LRT_p      = l_p,  
        Best_Model = best_mod_name)
  }
  
  output_df$Trait <- trait
  return(output_df)
})

final_cnt_df <- do.call(rbind, results_list_cnt)

final_cnt_df$p_adj <- p.adjust(final_cnt_df$LRT_p, method = "holm")

cnt.results <- final_cnt_df %>%
  mutate(
    p.display = case_when(
      p_adj < 0.001 ~ paste0("*p* < 0.001*"),
      p_adj < 0.05  ~ paste0(sprintf("%.3f", p_adj), "*"),
      TRUE ~ sprintf("%.3f", p_adj))) %>%
  mutate(
    Trait_Name = factor(
      trait.names[Trait],
      levels = unname(trait.names))) %>%
  arrange(Trait_Name) %>%
  relocate(Trait_Name, .before = everything())

cnt.tbl <- cnt.results %>%
  gt() %>%
  cols_hide(c(Trait, LRT_p, p_adj, Best_Model, "Sailfin molly", "Amazon molly")) %>%
  cols_label(
    Trait_Name = "Trait",
    Var_Ratio = md("Variance ratio <br>(Sailfin / Amazon)"),
    Delta_AIC = md("&Delta;AIC"),
    p.display = md("adj. *p*-value")) %>%
  cols_move(
    columns = c(Delta_AIC, p.display),
    after = Trait_Name) %>%
  fmt_number(
    columns = c(Var_Ratio, Delta_AIC),
    decimals = 2) %>%
  fmt_markdown(columns = p.display) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Trait_Name, Delta_AIC, Var_Ratio))) %>%
  cols_align("center", everything()) %>%
  opt_horizontal_padding(scale = 2)

cnt.tbl

  
# ───────────────────────────────────────────────────
## PLOTS
# ───────────────────────────────────────────────────

met.long <- metric %>%
  pivot_longer(
    cols = c(SL, BD, CPD, CPL, PreDL, DbL, HL, HD, HW, SnL, OL),
    names_to = "Trait",
    values_to = "Value")


met.plots <- ggplot(met.long, aes(x=Species, y=Value, fill = Species)) + 
  geom_boxplot() +
  geom_point(color= "#8D8680", size=2, alpha=0.5) +
  facet_wrap(~ Trait, scales = "free_y", axes = "margins") +
  theme_classic(base_size = 10, base_family = "sans") + 
  theme(
    axis.title.y = element_text(size = 10, family = "sans"), 
    axis.text.y  = element_text(size = 10, family = "sans"),
    axis.text.x  = element_text(size = 10, family = "sans"),
    strip.text    = element_text(size = 10, family = "sans"),
    panel.spacing = unit(1.5, "lines")) +
  scale_fill_manual(values = c("Sailfin molly"="#78B7C5",
                               "Amazon molly"="#E1AF00")) +
  xlab("") +
  ylab("Length or Depth (mm)")
met.plots


cnt.long <- count %>%
  pivot_longer(
    cols = c(D, P1, A, P1.R, LLSC, SALL, SBLL, SBDF, SL, FLA),
    names_to = "Trait",
    values_to = "Value")

cnt.plots <- ggplot(cnt.long, aes(x=Species, y=Value, fill = Species)) + 
  geom_boxplot() +
  geom_point(color= "#8D8680", size=2, alpha=0.5) +
  facet_wrap(~ Trait, scales = "free_y", axes = "margins") +
  theme_classic(base_size = 10, base_family = "sans") + 
  theme(
    axis.title.y = element_text(size = 10, family = "sans"), 
    axis.text.y  = element_text(size = 10, family = "sans"),
    axis.text.x  = element_text(size = 10, family = "sans"),
    strip.text    = element_text(size = 10, family = "sans"),
    panel.spacing = unit(1.5, "lines")) +
  scale_fill_manual(values = c("Sailfin molly"="#78B7C5",
                               "Amazon molly"="#E1AF00")) +
  xlab("") +
  ylab("Count")
cnt.plots


```

***
***

### Overall shape variation (PCA)

While none our 10 metric traits showed a significant difference in variation among species, many of these traits are not independent of each other. Therefore, we will assess overall shape variation among species using a ***principal component analysis (PCA)*** on centered and scaled data. Previously, we log transformed our data to better approach normality, but this is no longer the case with the truncated data. Instead, we will run a PCA on size-corrected residuals. We will then perform a mean and variance test to determine if overall shape variation differs between the species. 

As meristic traits are count data, a normal PCA cannot be applied. Instead, we will use a ***categorical principal component analysis (CATPCA)*** that uses optimal scaling to deal with discrete integers. 

#### *Loadings*

```{r}

# ───────────────────────────────────────────────────
## METRIC ANALYSIS 
# ───────────────────────────────────────────────────

## Get residuals of traits ~ standard length
shape_residuals <- lapply(met.vars, function(trait) {
  mod <- lm(metric[[trait]] ~ metric$SL)
  return(residuals(mod))
})

residual_df <- as.data.frame(shape_residuals)
colnames(residual_df) <- met.vars

## run PCA
PCA.test <- prcomp(residual_df, center = TRUE, scale. = TRUE)


summary(PCA.test)
(eig.val.test <- get_eigenvalue(PCA.test))


# Since we already corrected for body size, we will focus on the first two dimensions

ind.test <- get_pca_ind(PCA.test)
head(ind.test$coord[,1:2])


pca.df <- cbind(metric, ind.test$coord[,1:2])


# Will investigate the influence of each trait on the first 2 dimensions
(loadings <- PCA.test$rotation[, 1:2])

sorted.loadings.1 <- loadings[order(loadings[, 1]), 1]
myTitle <- "Loadings Plot for PC1" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.1, main=myTitle, xlab=myXlab, cex=0.5, col="blue")

sorted.loadings.2 <- loadings[order(loadings[, 2]), 2]
myTitle <- "Loadings Plot for PC2" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.2, main=myTitle, xlab=myXlab, cex=0.5, col="blue")



# ───────────────────────────────────────────────────
## MERISTIC (COUNT) ANALYSIS 
# ───────────────────────────────────────────────────

# Remove SL from count df
count2 <- select(count, -c("SL"))

# Run the Categorical PCA
# 'ndim = 2' extracts the first two principal components
# 'degrees = 1' treats the scaling transformations as piecewise linear (ideal for ordinal integers)

catpca.test <- princals(count2[, 5:13], ndim = 2, degrees = 1)

# Print the summary to see how much variance your new components explain
summary(catpca.test)

(cat.loadings <- catpca.test$loadings)

sorted.loadings.4 <- cat.loadings[order(cat.loadings[, 1]), 1]
myTitle <- "Loadings Plot for CATPC1" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.4, main=myTitle, xlab=myXlab, cex=0.5, col="blue")

sorted.loadings.5 <- cat.loadings[order(cat.loadings[, 2]), 2]
myTitle <- "Loadings Plot for CATPC2" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.5, main=myTitle, xlab=myXlab, cex=0.5, col="blue")

# Extract the eigenvalues (one per dimension)
eigenvals <- catpca.test$evals

# Calculate the percentage of variation explained by each axis
# (Divided by 9 because you have 9 meristic traits)
pc1_percent <- round((eigenvals[1] / 9) * 100, 1)
pc2_percent <- round((eigenvals[2] / 9) * 100, 1)

# Quick print check to see your numbers (e.g., 24.5 and 14.2)
print(paste0("CatPC1: ", pc1_percent, "%"))
print(paste0("CatPC2: ", pc2_percent, "%"))

# 1. Extract the object scores for the first two dimensions
cat_pc_scores <- as.data.frame(catpca.test$objectscores)
colnames(cat_pc_scores) <- c("CatPC1", "CatPC2")

# 2. Combine them back with your species and size metadata
catpca.df <- cbind(count, cat_pc_scores)

```

#### *Plots*

```{r}

#### METRIC

pca.1<- autoplot(PCA.test, data = metric, colour="Species", shape= 'Species', loadings=FALSE, loadings.label=FALSE, frame=TRUE, frame.type='norm', label= FALSE)+
  scale_color_manual(name="Species",
                    values = c("Sailfin molly"="#78B7C5", "Amazon molly"="#E1AF00")) +
  scale_fill_manual(name="Species",
                    values = c("Sailfin molly"="#78B7C5", "Amazon molly"="#E1AF00")) +
  scale_shape_manual(name= 'Species', 
                     values = c("Sailfin molly" = 17, "Amazon molly"=16)) +
   guides(fill = guide_legend(override.aes = list(fill = NA)))+
  ggtitle("PCA Plot of Morphology traits") + theme_classic() 
pca.1



#### MERISTIC

catpca <- ggplot(catpca.df, aes(x = CatPC1, y = CatPC2, color = Species, fill = Species, shape = Species)) +
  geom_point(size = 1, alpha = 0.8) +
  stat_ellipse(geom = "polygon", alpha = 0.15, type = "norm", level = 0.95) +
  scale_color_manual(name = "Species",
                     values = c("Sailfin molly" = "#78B7C5", "Amazon molly" = "#E1AF00")) +
  scale_fill_manual(name = "Species",
                    values = c("Sailfin molly" = "#78B7C5", "Amazon molly" = "#E1AF00")) +
  scale_shape_manual(name = "Species", 
                     values = c("Sailfin molly" = 17, "Amazon molly" = 16)) +
  guides(fill = guide_legend(override.aes = list(fill = NA))) +
  labs(
    title = "CATPCA Plot of Meristic Traits",
    x = "Categorical PC1",
    y = "Categorical PC2"
  ) +
  theme_classic() +
  labs(
    title = "CATPCA Plot of Meristic Traits",
    x = paste0("Categorical PC1 (", pc1_percent, "%)"),
    y = paste0("Categorical PC2 (", pc2_percent, "%)")
  )

catpca

```


#### *Post PCA tests*

**CONCLUSIONS:** For our metric data, there is a significant difference in average shape distribution for both PCs. This is expected, as amazons and sailfins have different body shapes and therefore occupy distinct morphospace. Both PCs also differ significantly in variation, with sailfin being more variable for PC1 and amazons more variable for PC2. This is different from our individual results, which suggests morphological variation exists among the species in a multi-trait manner. *Different to main analysis*

For our meristic data, both CatPC1 and CatPC2 showed a significant difference in average meristic profiles, which makes sense as the majority of the meristic traits are species diagnostics (e.g., scales before dorsal fin, lateral line scale count, dorsal fin rays). CatPC2 shows a difference in variation, with amazons exhibiting more variation, but the overall estimate variance is quite low (amazon 0.12, sailfin 0.07). This again could suggest a multi-trait covariation among species. *Different to main analysis*


```{r}
#quick distribution checks to see what tests to use

testmod <- lm(Dim.1 ~ Species, data = pca.df)
plot(simulateResiduals(testmod)) #levene's test significant

testmod <- lm(Dim.2 ~ Species, data = pca.df)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(CatPC1 ~ Species, data = catpca.df)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(CatPC2 ~ Species, data = catpca.df)
plot(simulateResiduals(testmod)) #fails KS test, outlier test, and within-group deviation

#since some models fail residual checks, I will use a GLMM with a by-species dispersion factor and test mean and variance differences from that; default family=gaussian unless otherwise specified


# ───────────────────────────────────────────────────
## MEAN TESTS (GLMM w/ ~Species dispersion for all)
# ───────────────────────────────────────────────────

### Metric Data

met.mean.1 <- glmmTMB(Dim.1 ~ Species, dispformula = ~Species, data = pca.df) 
(mm1<- Anova(met.mean.1, type = "III")) #not significant
plot(simulateResiduals(met.mean.1)) #all good
    ph1 <- as.data.frame(emmeans(met.mean.1, ~Species)) #to use for table later
    
met.mean.2 <- glmmTMB(Dim.2 ~ Species, dispformula = ~Species, data = pca.df) 
(mm2<- Anova(met.mean.2, type = "III")) #significant (p<0.001)
plot(simulateResiduals(met.mean.2)) #all good
    ph2 <- as.data.frame(emmeans(met.mean.2,  ~ Species)) #to use for table later


### Meristic Data


cnt.mean.1 <- glmmTMB(CatPC1 ~ Species, dispformula = ~Species, data = catpca.df) 
(cm1<- Anova(cnt.mean.1, type = "III")) #significant (p<0.001)
plot(simulateResiduals(cnt.mean.1)) #all good
    ph4 <- as.data.frame(emmeans(cnt.mean.1,  ~ Species))
    summary(ph4)


#default gaussian family failed residual checks; changing to Student t-distribution due to extreme tails
cnt.mean.2 <- glmmTMB(CatPC2 ~ Species, dispformula = ~Species, family = t_family, data = catpca.df) 
(cm2<- Anova(cnt.mean.2, type = "III")) #significant (p<0.001)
plot(simulateResiduals(cnt.mean.2)) #failed KS test but qqplot w/ good alignment
    ph5 <- as.data.frame(emmeans(cnt.mean.2,  ~ Species))
    summary(ph5)

# ───────────────────────────────────────────────────
## VARIANCE TESTS (Extract dispersion info for all)
# ───────────────────────────────────────────────────

### Metric Data
met.var.1 <- emmeans(met.mean.1, ~ Species, component = "disp", type = "response")
(mv1 <- as.data.frame(pairs(met.var.1))) #significant (p=0.045)

met.var.2 <- emmeans(met.mean.2, ~ Species, component = "disp", type = "response")
(mv2 <- as.data.frame(pairs(met.var.2))) #not significant


### Meristic Data
cnt.var.1 <- emmeans(cnt.mean.1, ~ Species, component = "disp", type = "response")
(cv1 <- as.data.frame(pairs(cnt.var.1))) #not significant

cnt.var.2 <- emmeans(cnt.mean.2, ~ Species, component = "disp", type = "response")
(cv2 <- as.data.frame(pairs(cnt.var.2))) #significant (p=0.010)


# ───────────────────────────────────────────────────
## TABLE
# ───────────────────────────────────────────────────

### Metric data 

## ensure all results are as df
met.var.1 <- as.data.frame(met.var.1)
met.var.2 <- as.data.frame(met.var.2)


## pull species-specific model results
# Means
amazon_m1  <- sprintf("%.2f (%.2f)", ph1$emmean[1], ph1$SE[1])
sailfin_m1 <- sprintf("%.2f (%.2f)", ph1$emmean[2], ph1$SE[2])

amazon_m2  <- sprintf("%.2f (%.2f)", ph2$emmean[1], ph2$SE[1])
sailfin_m2 <- sprintf("%.2f (%.2f)", ph2$emmean[2], ph2$SE[2])


Amazon_mean <- c(amazon_m1, amazon_m2)

Sailfin_mean <- c(sailfin_m1, sailfin_m2)

# variance
Amazon_var <- sprintf("%.2f", c(met.var.1$response[1], met.var.2$response[1]))

Sailfin_var <- sprintf("%.2f", c(met.var.1$response[2], met.var.2$response[2]))


## pull model stats
Chisq_mean <- c(mm1["Species", "Chisq"], mm2["Species", "Chisq"])

p_mean <- c(mm1["Species", "Pr(>Chisq)"], mm2["Species", "Pr(>Chisq)"])

Chisq_var <- c(Anova(met.mean.1, component = "disp", type = "III")["Species", "Chisq"], Anova(met.mean.2, component = "disp", type = "III")["Species", "Chisq"])

p_var <- c(Anova(met.mean.1, component = "disp", type = "III")["Species", "Pr(>Chisq)"], Anova(met.mean.2, component = "disp", type = "III")["Species", "Pr(>Chisq)"])
  

## create tibble
postPCA_result <- tibble(
  Classifier   = "Metric PCA",
  Dimension    = c("PC1", "PC2"),
  Amazon_mean  = Amazon_mean,
  Sailfin_mean = Sailfin_mean,
  Chisq_mean   = Chisq_mean,
  p_mean       = p_mean,
  Amazon_var   = Amazon_var,
  Sailfin_var  = Sailfin_var,
  Chisq_var    = Chisq_var,
  p_var        = p_var) 


### Meristic Data

## ensure all results are as df
cnt.var.1 <- as.data.frame(cnt.var.1)
cnt.var.2 <- as.data.frame(cnt.var.2)

## pull species-specific model results
# Means
amazon_c1  <- sprintf("%.2f (%.2f)", ph4$emmean[1], ph4$SE[1])
sailfin_c1 <- sprintf("%.2f (%.2f)", ph4$emmean[2], ph4$SE[2])

amazon_c2  <- sprintf("%.2f (%.2f)", ph5$emmean[1], ph5$SE[1])
sailfin_c2 <- sprintf("%.2f (%.2f)", ph5$emmean[2], ph5$SE[2])

Amazon_mean <- c(amazon_c1, amazon_c2)

Sailfin_mean <- c(sailfin_c1, sailfin_c2)

# variance
Amazon_var <- sprintf("%.2f", c(cnt.var.1$response[1], cnt.var.2$response[1]))

Sailfin_var <- sprintf("%.2f", c(cnt.var.1$response[2], cnt.var.2$response[2]))


## pull model stats
Chisq_mean <- c(cm1["Species", "Chisq"], cm2["Species", "Chisq"])

p_mean <- c(cm1["Species", "Pr(>Chisq)"], cm2["Species", "Pr(>Chisq)"])

Chisq_var <- c(Anova(cnt.mean.1, component = "disp", type = "III")["Species", "Chisq"], Anova(cnt.mean.2, component = "disp", type = "III")["Species", "Chisq"])

p_var <- c(Anova(cnt.mean.1, component = "disp", type = "III")["Species", "Pr(>Chisq)"], Anova(cnt.mean.2, component = "disp", type = "III")["Species", "Pr(>Chisq)"])

## create tibble
postCPCA_result <- tibble(
  Classifier   = "Meristic CatPCA",
  Dimension    = c("CatPC1", "CatPC2"),
  Amazon_mean  = Amazon_mean,
  Sailfin_mean = Sailfin_mean,
  Chisq_mean   = Chisq_mean,
  p_mean       = p_mean,
  Amazon_var   = Amazon_var,
  Sailfin_var  = Sailfin_var,
  Chisq_var    = Chisq_var,
  p_var        = p_var) 


  
### Combine & form table

results_combined <- bind_rows(postPCA_result, postCPCA_result) %>%
  group_by(Classifier) %>%
  mutate(
  p_mean_adj = p.adjust(p_mean, method = "holm"),
  p_var_adj = p.adjust(p_var, method = "holm")) %>%
  mutate(
    Mean_test = case_when(
      p_mean_adj < 0.001 ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chisq_mean), "<br>*p* < 0.001*"),
      p_mean_adj < 0.05  ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chisq_mean), "<br>*p* = ", sprintf("%.3f", p_mean_adj), "*"),
      TRUE           ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chisq_mean), "<br>*p* = ", sprintf("%.3f", p_mean_adj))),
    
    Variance_test = case_when(
      p_var_adj < 0.001 ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chisq_var), "<br>*p* < 0.001*"),
      p_var_adj < 0.05  ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chisq_var), "<br>*p* = ", sprintf("%.3f", p_var_adj), "*"),
      TRUE          ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chisq_var), "<br>*p* = ", sprintf("%.3f", p_var_adj)))) %>%
  ungroup()




postPCA_table <- results_combined %>%
  select(
    Classifier,
    Dimension,
    Amazon_mean,
    Sailfin_mean,
    Mean_test,
    Amazon_var,
    Sailfin_var,
    Variance_test) %>%
  gt(groupname_col = "Classifier") %>%
  fmt_markdown(
    columns = everything()) %>%
  cols_label(
    Dimension = " ",
    Amazon_mean = md("**Amazon**"),
    Sailfin_mean = md("**Sailfin**"),
    Mean_test = html("Mean comparison<br>(Wald &chi;<sup>2</sup>)"),
    Amazon_var = md("**Amazon**"),
    Sailfin_var = md("**Sailfin**"),
    Variance_test = html("Variance comparison<br>(Wald &chi;<sup>2</sup>)")) %>%
  tab_spanner(
    label = html("Est. Marginal Mean (&plusmn;SE)"),
    columns = c(Amazon_mean, Sailfin_mean)) %>%
  tab_spanner(
    label = "Est. Variance",
    columns = c(Amazon_var, Sailfin_var)) %>%
  tab_options(row_group.font.weight = "bold") %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c(Dimension, Sailfin_mean, Mean_test, Sailfin_var))) %>%
  cols_align(align = "center", columns = everything()) %>%
  opt_horizontal_padding(scale = 1.5)

### Final table
postPCA_table

```

***
***



## **Geographic Analysis**


### Quick summary

A quick summary, mostly of number of fish in each zone and basin.

We find that there are at least 3 fish of each species in all zones and 5 of 9 basins Classifiers marked with an asterisk only have >3 of *one* species.

```{r, echo=FALSE, warning=FALSE}

zone.sum <-raw2 %>%
  count(Species, Zone) %>%
  tidyr::pivot_wider(
    names_from  = Species,
    values_from = n,
    values_fill = 0) %>%
  rename(classifier = Zone) %>%
  mutate(classifier = as.character(classifier), 
         classifier_type = "Zone")

basin.sum <- raw2 %>%
  count(Species, Basin) %>%
  tidyr::pivot_wider(
    names_from  = Species,
    values_from = n,
    values_fill = 0) %>%
  rename(classifier = Basin) %>%
  mutate(classifier = as.character(classifier), 
         classifier_type = "Basin")


sum.counts <- bind_rows(zone.sum, basin.sum)

sum.counts %>%
  group_by(classifier_type) %>%
  mutate(classifier_type = paste0(classifier_type, " (n=", n(), ")")) %>%
  ungroup() %>%
  mutate(classifier_type = factor(classifier_type,
                                  levels = grep("Zone", unique(classifier_type), value = TRUE) %>%
                                           c(grep("Basin", unique(classifier_type), value = TRUE)) %>%
                                           c(grep("Watershed", unique(classifier_type), value = TRUE))))%>%
  arrange(classifier_type, classifier) %>%
  select(classifier_type, classifier, everything()) %>%
  gt(rowname_col = "classifier", groupname_col = "classifier_type") %>%
  text_transform(
    locations = cells_stub(),
    fn = function(x) {
      too.few <- sum.counts$`Amazon molly` < 3 | sum.counts$`Sailfin molly` < 3
      
      ifelse(sum.counts$classifier %in% x & too.few[match(x, sum.counts$classifier)], 
             paste0(x, "*"), x)}) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c(classifier, "Sailfin molly"))) %>%
  opt_horizontal_padding(scale=2) %>%
  cols_align(align="center", everything())



```
***
***

### Individual trait models

We will be repeating our **generalized linear mixed models (GLMM)** with the addition of zones and basins to test if variation is partitioned across a geographic range. 

#### *Zones*

**CONCLUSIONS:** Three metric traits showed greater support for a model with unequal dispersion across species and zones. Caudal peduncle depth, and orbital length mostly show higher variation among sailfins with highest variation in zone 4. Head length show amazons to be more variable in all zones except zone 2, where sailfins are more variable. *Different from main analysis*

No meristic traits showed greater support for a model with unequal dispersion across species and zones. Just about all models failed convergence and used the fallback residual ANOVA. The fallback residual ANOVA for anal rays showed a significant difference in variation with sailfins containing more variation in all but zone 4, where amazons contained more variation. *Different from main analysis* 

```{r}
# ───────────────────────────────────────────────────
## METRIC ANALYSIS--ZONES
# ───────────────────────────────────────────────────


results_geo <- lapply(met.vars, function(trait) {
  
  form_str <- paste0("log(", trait, ") ~ log(SL) * Species")
  form = as.formula(form_str)

  # 1. Species variance model
  mod1 <- tryCatch({
    glmmTMB(form, data = metric, family = gaussian, dispformula = ~Species)
  }, error = function(e) { NULL }, warning = function(w) { NULL })
  
  # 2. Geographic variance model (with strict optimization catching)
  mod2 <- tryCatch({
    glmmTMB(form, data = metric, family = gaussian, dispformula = ~Species*Zone)
  }, 
  error = function(e) { NULL },
  warning = function(w) {
    if(grepl("NA/NaN|calc_loglambda|iteration|failure", w$message, ignore.case = TRUE)) {
      return(NULL)
    }
    invokeRestart("muffleWarning")
  })
  
  # Verify clean convergence of both models
  models_valid <- !is.null(mod1) && !is.null(mod2) && 
                  !is.na(AIC(mod1)) && !is.na(AIC(mod2)) &&
                  !any(is.nan(summary(mod2)$coefficients$disp[,2]))
  
  unique_zones <- levels(as.factor(metric$Zone))
  
  # --- PATH A: MODELS CONVERGED CLEANLY ---
  if (models_valid) {
    comp <- anova(mod1, mod2)
    pval <- comp$`Pr(>Chisq)`[2]
    delta_aic <- AIC(mod2) - AIC(mod1)
    
    best_mod <- if (AIC(mod1) < AIC(mod2)) mod1 else mod2
    best_mod_name <- if (AIC(mod1) < AIC(mod2)) "Species-Only (mod1)" else "Geographic (mod2)"
    
    message(paste0("--- DHARMa Diagnostics for Trait: ", trait, " (Best Model: ", best_mod_name, ") ---"))
    sim_res <- simulateResiduals(best_mod, plot = FALSE)
    plot(sim_res) 
    title(main = paste(trait, "-", best_mod_name), line = 2.5)
    
    rg <- ref_grid(mod2, component = "disp")
    disp_means <- as.data.frame(emmeans(rg, specs = ~ Species * Zone))
    disp_means$Variance <- exp(disp_means$emmean)
    
    ratios_by_zone <- disp_means %>%
      select(Species, Zone, Variance) %>%
      pivot_wider(names_from = Species, values_from = Variance) %>%
      mutate(
        Var_Ratio = `Sailfin molly` / `Amazon molly`,
        Trait = trait, 
        Delta_AIC = delta_aic, 
        LRT_p = pval,
        Best_Model = best_mod_name)
      
  } else {
    # --- PATH B: FALLBACK SYSTEM FOR METRIC ---
    best_mod_name <- "Model Failed - Residual ANOVA Fallback"
    message(paste0("!!! glmmTMB Failed for Trait: ", trait, ". Running Fallback. !!!"))
    
    # Non-parametric fallback: test if absolute size-corrected residuals differ by Species * Zone
    lm_check <- lm(as.formula(paste0("log(", trait, ") ~ log(SL)")), data = metric)
    metric$Abs_Resid <- abs(residuals(lm_check))
    aov_fallback <- anova(lm(Abs_Resid ~ Species * Zone, data = metric))
    fallback_p <- aov_fallback$`Pr(>F)`[3] # p-value for the interaction term
    
    # Calculate direct empirical variance ratios per zone
    empirical_ratios <- metric %>%
      group_by(Zone, Species) %>%
      summarise(Variance = var(!!sym(trait), na.rm = TRUE), .groups = 'drop') %>%
      complete(Zone = unique_zones, Species = c("Amazon molly", "Sailfin molly"), fill = list(Variance = NA)) %>%
      pivot_wider(names_from = Species, values_from = Variance) %>%
      mutate(
        Var_Ratio = `Sailfin molly` / `Amazon molly`,
        Trait = trait,
        Delta_AIC = NA,
        LRT_p = fallback_p,
        Best_Model = best_mod_name)
    
    ratios_by_zone <- empirical_ratios
  }
  
  return(ratios_by_zone)
})


geo_results_df <- do.call(rbind, results_geo)


p_adjusted_df <- geo_results_df %>%
  select(Trait, LRT_p, Delta_AIC) %>%
  distinct() %>%
  mutate(p_adj = p.adjust(LRT_p, method = "holm")) %>%
  mutate(
    p.display = case_when(
      p_adj < 0.001 ~ "*p* < 0.001*",
      p_adj < 0.05  ~ paste0(sprintf("%.3f", p_adj), "*"),
      TRUE          ~ sprintf("%.3f", p_adj)),
    Trait_Name = factor(trait.names[Trait], levels = unname(trait.names)))

ratios_wide <- geo_results_df %>%
  select(Trait, Zone, Var_Ratio) %>%
  pivot_wider(
    names_from = Zone, 
    values_from = Var_Ratio, 
    names_prefix = "Ratio_")

geo_wide_results <- p_adjusted_df %>%
  left_join(ratios_wide, by = "Trait") %>%
  arrange(Trait_Name)

geo.met.tbl <- geo_wide_results %>%
  gt() %>%
  cols_hide(c(Trait, LRT_p, p_adj)) %>%
  cols_move_to_start(columns = Trait_Name) %>%
  cols_label(
    Trait_Name = "Trait",
    Delta_AIC = md("&Delta;AIC"),
    p.display = md("adj. *p*-value"),
    `Ratio_Zone 1` = "Zone 1",  
    `Ratio_Zone 2` = "Zone 2",  
    `Ratio_Zone 3` = "Zone 3",
    `Ratio_Zone 4` = "Zone 4") %>%
  tab_spanner(
    label = md("Variance Ratio *(Sailfin / Amazon)*"),
    columns = starts_with("Ratio_")) %>%
  fmt_number(
    columns = c(Delta_AIC, starts_with("Ratio_")),
    decimals = 2) %>%
  fmt_markdown(columns = p.display) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Trait_Name, p.display))
  ) %>%
  cols_align("center", everything()) %>%
  opt_horizontal_padding(scale = 2)

geo.met.tbl


# ───────────────────────────────────────────────────
## MERISTIC (COUNT) ANALYSIS--ZONES
# ───────────────────────────────────────────────────


results_geo_cnt <- lapply(cnt.vars, function(trait) {
  
  form_str <- paste0(trait, " ~ SL * Species")
  form <- as.formula(form_str)

  # 1. Species variance model 
  mod1 <- tryCatch({
    glmmTMB(form, data = count, family = compois, dispformula = ~Species)
  }, error = function(e) { NULL }, warning = function(w) { NULL })
  
  # 2. Geographic variance model (compois with strict warning catching)
  mod2 <- tryCatch({
    withCallingHandlers(
      glmmTMB(form, data = count, family = compois, dispformula = ~Species*Zone),
      warning = function(w) {
        if (grepl("NA/NaN|calc_loglambda|iteration|failure", w$message, ignore.case = TRUE)) {
          stop(w$message) }
        invokeRestart("muffleWarning")})}, error = function(e) { NULL })
  
  models_valid <- !is.null(mod1) && !is.null(mod2) && 
                  !is.na(AIC(mod1)) && !is.na(AIC(mod2)) && 
                  !any(is.nan(summary(mod2)$coefficients$disp[,2]))

  unique_zones <- levels(as.factor(count$Zone))
  
  # --- PATH A: MODELS CONVERGED CLEANLY ---
  if (models_valid) {
    comp <- anova(mod1, mod2)
    pval <- comp$`Pr(>Chisq)`[2]
    delta_aic <- AIC(mod2) - AIC(mod1)
    
    best_mod <- if (AIC(mod1) < AIC(mod2)) mod1 else mod2
    best_mod_name <- if (AIC(mod1) < AIC(mod2)) "Species-Only (mod1)" else "Geographic (mod2)"
    
    message(paste0("--- DHARMa Diagnostics for Trait: ", trait, " (Best Model: ", best_mod_name, ") ---"))
    sim_res <- simulateResiduals(best_mod, plot = FALSE)
    plot(sim_res) 
    title(main = paste(trait, "-", best_mod_name), line = 2.5)
    
    mean_sl <- mean(count$SL, na.rm = TRUE)
    pred_data <- expand.grid(Species = c("Amazon molly", "Sailfin molly"), Zone = unique_zones)
    pred_data$SL <- mean_sl
    
    pred_data$Mean <- predict(mod2, newdata = pred_data, type = "response")
    pred_data$Nu   <- exp(predict(mod2, newdata = pred_data, type = "disp"))
    pred_data$Variance <- pred_data$Mean / pred_data$Nu # COM-Poisson variance block
    
    zone_output <- pred_data %>%
      select(Zone, Species, Variance) %>%
      pivot_wider(names_from = Species, values_from = Variance) %>%
      mutate(
        Var_Ratio = `Sailfin molly` / `Amazon molly`,
        Trait     = trait,
        Delta_AIC = delta_aic,
        LRT_p     = pval,
        Best_Model = best_mod_name)
  } else {
    # --- PATH B: FALLBACK SYSTEM FOR MERISTIC ---
    best_mod_name <- "Model Failed - Residual ANOVA Fallback"
    message(paste0("!!! glmmTMB Failed for Trait: ", trait, ". Running Fallback. !!!"))
    
    # Fit basic poisson to get residuals safely without optimization breaking
    glance_mod <- glm(as.formula(paste0(trait, " ~ SL")), data = count, family = poisson)
    count$Abs_Resid <- abs(residuals(glance_mod, type = "response"))
    aov_fallback <- anova(lm(Abs_Resid ~ Species * Zone, data = count))
    fallback_p <- aov_fallback$`Pr(>F)`[3]
    
    empirical_ratios <- count %>%
      group_by(Zone, Species) %>%
      summarise(Variance = var(!!sym(trait), na.rm = TRUE), .groups = 'drop') %>%
      complete(Zone = unique_zones, Species = c("Amazon molly", "Sailfin molly"), fill = list(Variance = NA)) %>%
      pivot_wider(names_from = Species, values_from = Variance) %>%
      mutate(
        Var_Ratio = `Sailfin molly` / `Amazon molly`,
        Trait     = trait,
        Delta_AIC = NA,
        LRT_p     = fallback_p,
        Best_Model = best_mod_name)
    
    zone_output <- empirical_ratios
  }
  
  return(zone_output)
})

geo_cnt_df <- do.call(rbind, results_geo_cnt)


p_adj_cnt_df <- geo_cnt_df %>%
  select(Trait, LRT_p, Delta_AIC) %>%
  distinct() %>%
  mutate(p_adj = p.adjust(LRT_p, method = "holm")) %>%
  mutate(
    p.display = case_when(
      p_adj < 0.001 ~ "*p* < 0.001*",
      p_adj < 0.05  ~ paste0(sprintf("%.3f", p_adj), "*"),
      TRUE          ~ sprintf("%.3f", p_adj)),
    Trait_Name = factor(trait.names[Trait], levels = unname(trait.names)))

ratios_wide2 <- geo_cnt_df %>%
  select(Trait, Zone, Var_Ratio) %>%
  pivot_wider(
    names_from = Zone, 
    values_from = Var_Ratio, 
    names_prefix = "Ratio_")

geo_wide_cnt <- p_adj_cnt_df %>%
  left_join(ratios_wide2, by = "Trait") %>%
  arrange(Trait_Name)

geo.cnt.tbl <- geo_wide_cnt %>%
  gt() %>%
  cols_hide(c(Trait, LRT_p, p_adj)) %>%
  cols_move_to_start(columns = Trait_Name) %>%
  cols_label(
    Trait_Name = "Trait",
    Delta_AIC = md("&Delta;AIC"),
    p.display = md("adj. *p*-value"),
    `Ratio_Zone 1` = "Zone 1",  
    `Ratio_Zone 2` = "Zone 2",  
    `Ratio_Zone 3` = "Zone 3",
    `Ratio_Zone 4` = "Zone 4") %>%
  tab_spanner(
    label = md("Variance Ratio *(Sailfin / Amazon)*"),
    columns = starts_with("Ratio_")) %>%
  fmt_number(
    columns = c(Delta_AIC, starts_with("Ratio_")),
    decimals = 2) %>%
  fmt_markdown(columns = p.display) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Trait_Name, p.display))
  ) %>%
  cols_align("center", everything()) %>%
  opt_horizontal_padding(scale = 2)

geo.cnt.tbl

```



#### *Basins*

First need to filter for basins with a minimum of 3 of each species.

```{r}
# Filtering for basins with n>=3 for both species
valid_basins <- raw2 %>%
  group_by(Basin, Species) %>%
  summarise(n = n(), .groups = "drop") %>%
  tidyr::pivot_wider(names_from = Species, values_from = n, values_fill = 0) %>%
  filter(`Amazon molly` >= 3 & `Sailfin molly` >= 3) %>%
  pull(Basin)

metric2 <- metric %>% filter(Basin %in% valid_basins) %>% droplevels()
count2 <- count %>% filter(Basin %in% valid_basins) %>% droplevels()



```


**CONCLUSIONS:** Only one metric trait (orbit length) showed greater support for a model with unequal dispersion across species and basins. There is no consistent pattern across across basins, indicating that local deviations are likely reflecting stochastic, demographic, or sampling-related effects rather than any differences from reproductive mode. The rest of the metric traits show no significant difference in variation. *Different from main analysis* 

Again, nearly all models failed convergence and relied on fallback residual ANOVA. Only anal rays showed a significant difference, with amazons more variable than sailfin in all basins except Lower SWTX coastal. However, it is important to note that while we filtered for basins with >=3 of each species, some sample sizes were still very small or variation very sparse and may be affecting overall results. *Different from main analysis*

```{r}
# ───────────────────────────────────────────────────
## METRIC ANALYSIS--BASINS
# ───────────────────────────────────────────────────


results_geo2 <- lapply(met.vars, function(trait) {
  
  form_str <- paste0("log(", trait, ") ~ log(SL) * Species")
  form = as.formula(form_str)

  # 1. Species variance model
  mod1 <- tryCatch({
    glmmTMB(form, data = metric2, family = gaussian, dispformula = ~Species)
  }, error = function(e) { NULL }, warning = function(w) { NULL })
  
  # 2. Geographic variance model (with strict optimization catching)
  mod2 <- tryCatch({
    glmmTMB(form, data = metric2, family = gaussian, dispformula = ~Species*Basin)
  }, 
  error = function(e) { NULL },
  warning = function(w) {
    if(grepl("NA/NaN|calc_loglambda|iteration|failure", w$message, ignore.case = TRUE)) {
      return(NULL)
    }
    invokeRestart("muffleWarning")
  })
  
  # Verify clean convergence of both models
  models_valid <- !is.null(mod1) && !is.null(mod2) && 
                  !is.na(AIC(mod1)) && !is.na(AIC(mod2)) &&
                  !any(is.nan(summary(mod2)$coefficients$disp[,2]))
  
  unique_basins <- levels(as.factor(metric2$Basin))
  
  # --- PATH A: MODELS CONVERGED CLEANLY ---
  if (models_valid) {
    comp <- anova(mod1, mod2)
    pval <- comp$`Pr(>Chisq)`[2]
    delta_aic <- AIC(mod2) - AIC(mod1)
    
    best_mod <- if (AIC(mod1) < AIC(mod2)) mod1 else mod2
    best_mod_name <- if (AIC(mod1) < AIC(mod2)) "Species-Only (mod1)" else "Geographic (mod2)"
    
    message(paste0("--- DHARMa Diagnostics for Trait: ", trait, " (Best Model: ", best_mod_name, ") ---"))
    sim_res <- simulateResiduals(best_mod, plot = FALSE)
    plot(sim_res) 
    title(main = paste(trait, "-", best_mod_name), line = 2.5)
    
    rg <- ref_grid(mod2, component = "disp")
    disp_means <- as.data.frame(emmeans(rg, specs = ~ Species * Basin))
    disp_means$Variance <- exp(disp_means$emmean)
    
    ratios_by_basin <- disp_means %>%
      select(Species, Basin, Variance) %>%
      pivot_wider(names_from = Species, values_from = Variance) %>%
      mutate(
        Var_Ratio = `Sailfin molly` / `Amazon molly`,
        Trait = trait, 
        Delta_AIC = delta_aic, 
        LRT_p = pval,
        Best_Model = best_mod_name)
      
  } else {
    # --- PATH B: FALLBACK SYSTEM FOR METRIC ---
    best_mod_name <- "Model Failed - Residual ANOVA Fallback"
    message(paste0("!!! glmmTMB Failed for Trait: ", trait, ". Running Fallback. !!!"))
    
    # Non-parametric fallback: test if absolute size-corrected residuals differ by Species * Zone
    lm_check <- lm(as.formula(paste0("log(", trait, ") ~ log(SL)")), data = metric2)
    metric2$Abs_Resid <- abs(residuals(lm_check))
    aov_fallback <- anova(lm(Abs_Resid ~ Species * Basin, data = metric2))
    fallback_p <- aov_fallback$`Pr(>F)`[3] # p-value for the interaction term
    
    # Calculate direct empirical variance ratios per zone
    empirical_ratios <- metric2 %>%
      group_by(Basin, Species) %>%
      summarise(Variance = var(!!sym(trait), na.rm = TRUE), .groups = 'drop') %>%
      complete(Basin = unique_basins, Species = c("Amazon molly", "Sailfin molly"), fill = list(Variance = NA)) %>%
      pivot_wider(names_from = Species, values_from = Variance) %>%
      mutate(
        Var_Ratio = `Sailfin molly` / `Amazon molly`,
        Trait = trait,
        Delta_AIC = NA,
        LRT_p = fallback_p,
        Best_Model = best_mod_name)
    
    ratios_by_basin <- empirical_ratios
  }
  
  return(ratios_by_basin)
})


geo_results_df2 <- do.call(rbind, results_geo2)


p_adjusted_df2 <- geo_results_df2 %>%
  select(Trait, LRT_p, Delta_AIC) %>%
  distinct() %>%
  mutate(p_adj = p.adjust(LRT_p, method = "holm")) %>%
  mutate(
    p.display = case_when(
      p_adj < 0.001 ~ "*p* < 0.001*",
      p_adj < 0.05  ~ paste0(sprintf("%.3f", p_adj), "*"),
      TRUE          ~ sprintf("%.3f", p_adj)),
    Trait_Name = factor(trait.names[Trait], levels = unname(trait.names)))

ratios_wide_bas <- geo_results_df2 %>%
  select(Trait, Basin, Var_Ratio) %>%
  pivot_wider(
    names_from = Basin, 
    values_from = Var_Ratio, 
    names_prefix = "Ratio_")

geo_wide_results2 <- p_adjusted_df2 %>%
  left_join(ratios_wide_bas, by = "Trait") %>%
  arrange(Trait_Name)

geo.met.tbl2 <- geo_wide_results2 %>%
  gt() %>%
  cols_hide(c(Trait, LRT_p, p_adj)) %>%
  cols_move_to_start(columns = Trait_Name) %>%
  cols_label(
    Trait_Name = "Trait",
    Delta_AIC = md("&Delta;AIC"),
    p.display = md("adj. *p*-value"),
    `Ratio_San Antonio-Guadalupe`             = "San Antonio-Guadalupe",  
    `Ratio_Upper Southwest Texas Coastal`     = "Upper Southwest Texas Coastal",
    `Ratio_Lower Southwest Texas Coastal`     = "Lower Southwest Texas Coastal",
    `Ratio_Lagunas de San Andrés-Morales`     = "Lagunas de San Andrés-Morales",
    `Ratio_Río Tamesí`                        = "Río Tamesí") %>%
  tab_spanner(
    label = md("Variance Ratio *(Sailfin / Amazon)*"),
    columns = starts_with("Ratio_")) %>%
  fmt_number(
    columns = c(Delta_AIC, starts_with("Ratio_")),
    decimals = 2) %>%
  fmt_markdown(columns = p.display) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Trait_Name, p.display))) %>%
  cols_align("center", everything()) %>%
  opt_horizontal_padding(scale = 2)

geo.met.tbl2


# ───────────────────────────────────────────────────
## MERISTIC (COUNT) ANALYSIS--BASINS
# ───────────────────────────────────────────────────


results_geo_cnt2 <- lapply(cnt.vars, function(trait) {
  
  form_str <- paste0(trait, " ~ SL * Species")
  form <- as.formula(form_str)

  # 1. Species variance model 
  mod1 <- tryCatch({
    glmmTMB(form, data = count2, family = compois, dispformula = ~Species)
  }, error = function(e) { NULL }, warning = function(w) { NULL })
  
  # 2. Geographic variance model (compois with strict warning catching)
  mod2 <- tryCatch({
    withCallingHandlers(
      glmmTMB(form, data = count2, family = compois, dispformula = ~Species*Basin),
      warning = function(w) {
        if (grepl("NA/NaN|calc_loglambda|iteration|failure", w$message, ignore.case = TRUE)) {
          stop(w$message) }
        invokeRestart("muffleWarning")})}, error = function(e) { NULL })
  
  
  models_valid <- !is.null(mod1) && !is.null(mod2) && 
                  !is.na(AIC(mod1)) && !is.na(AIC(mod2)) && 
                  !any(is.nan(summary(mod2)$coefficients$disp[,2]))

  unique_basins <- levels(as.factor(count2$Basin))
  
  # --- PATH A: MODELS CONVERGED CLEANLY ---
  if (models_valid) {
    comp <- anova(mod1, mod2)
    pval <- comp$`Pr(>Chisq)`[2]
    delta_aic <- AIC(mod2) - AIC(mod1)
    
    best_mod <- if (AIC(mod1) < AIC(mod2)) mod1 else mod2
    best_mod_name <- if (AIC(mod1) < AIC(mod2)) "Species-Only (mod1)" else "Geographic (mod2)"
    
    message(paste0("--- DHARMa Diagnostics for Trait: ", trait, " (Best Model: ", best_mod_name, ") ---"))
    sim_res <- simulateResiduals(best_mod, plot = FALSE)
    plot(sim_res) 
    title(main = paste(trait, "-", best_mod_name), line = 2.5)
    
    mean_sl <- mean(count2$SL, na.rm = TRUE)
    pred_data <- expand.grid(Species = c("Amazon molly", "Sailfin molly"), Basin = unique_basins)
    pred_data$SL <- mean_sl
    
    pred_data$Mean <- predict(mod2, newdata = pred_data, type = "response")
    pred_data$Nu   <- exp(predict(mod2, newdata = pred_data, type = "disp"))
    pred_data$Variance <- pred_data$Mean / pred_data$Nu # COM-Poisson variance block
    
    basin_output <- pred_data %>%
      select(Basin, Species, Variance) %>%
      pivot_wider(names_from = Species, values_from = Variance) %>%
      mutate(
        Var_Ratio = `Sailfin molly` / `Amazon molly`,
        Trait     = trait,
        Delta_AIC = delta_aic,
        LRT_p     = pval,
        Best_Model = best_mod_name)
  } else {
    # --- PATH B: FALLBACK SYSTEM FOR MERISTIC ---
    best_mod_name <- "Model Failed - Residual ANOVA Fallback"
    message(paste0("!!! glmmTMB Failed for Trait: ", trait, ". Running Fallback. !!!"))
    
    # Fit basic poisson to get residuals safely without optimization breaking
    glance_mod <- glm(as.formula(paste0(trait, " ~ SL")), data = count2, family = poisson)
    count2$Abs_Resid <- abs(residuals(glance_mod, type = "response"))
    aov_fallback <- anova(lm(Abs_Resid ~ Species * Basin, data = count2))
    fallback_p <- aov_fallback$`Pr(>F)`[3]
    
    empirical_ratios <- count2 %>%
      group_by(Basin, Species) %>%
      summarise(Variance = var(!!sym(trait), na.rm = TRUE), .groups = 'drop') %>%
      complete(Basin = unique_basins, Species = c("Amazon molly", "Sailfin molly"), fill = list(Variance = NA)) %>%
      pivot_wider(names_from = Species, values_from = Variance) %>%
      mutate(
        Var_Ratio = `Sailfin molly` / `Amazon molly`,
        Trait     = trait,
        Delta_AIC = NA,
        LRT_p     = fallback_p,
        Best_Model = best_mod_name)
    
    basin_output <- empirical_ratios
  }
  
  return(basin_output)
})

geo_cnt_df2 <- do.call(rbind, results_geo_cnt2)


p_adj_cnt_df2 <- geo_cnt_df2 %>%
  select(Trait, LRT_p, Delta_AIC) %>%
  distinct() %>%
  mutate(p_adj = p.adjust(LRT_p, method = "holm")) %>%
  mutate(
    p.display = case_when(
      p_adj < 0.001 ~ "*p* < 0.001*",
      p_adj < 0.05  ~ paste0(sprintf("%.3f", p_adj), "*"),
      TRUE          ~ sprintf("%.3f", p_adj)),
    Trait_Name = factor(trait.names[Trait], levels = unname(trait.names)))

ratios_wide_bas2 <- geo_cnt_df2 %>%
  select(Trait, Basin, Var_Ratio) %>%
  pivot_wider(
    names_from = Basin, 
    values_from = Var_Ratio, 
    names_prefix = "Ratio_")

geo_wide_cnt2 <- p_adj_cnt_df %>%
  left_join(ratios_wide_bas2, by = "Trait") %>%
  arrange(Trait_Name)

geo.cnt.tbl2 <- geo_wide_cnt2 %>%
  gt() %>%
  cols_hide(c(Trait, LRT_p, p_adj)) %>%
  cols_move_to_start(columns = Trait_Name) %>%
  cols_label(
    Trait_Name = "Trait",
    Delta_AIC = md("&Delta;AIC"),
    p.display = md("adj. *p*-value"),
    `Ratio_San Antonio-Guadalupe`             = "San Antonio-Guadalupe",  
    `Ratio_Upper Southwest Texas Coastal`     = "Upper Southwest Texas Coastal",
    `Ratio_Lower Southwest Texas Coastal`     = "Lower Southwest Texas Coastal",
    `Ratio_Lagunas de San Andrés-Morales`     = "Lagunas de San Andrés-Morales",
    `Ratio_Río Tamesí`                        = "Río Tamesí") %>%
  tab_spanner(
    label = md("Variance Ratio *(Sailfin / Amazon)*"),
    columns = starts_with("Ratio_")) %>%
  fmt_number(
    columns = c(Delta_AIC, starts_with("Ratio_")),
    decimals = 2) %>%
  fmt_markdown(columns = p.display) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Trait_Name, p.display))) %>%
  cols_align("center", everything()) %>%
  opt_horizontal_padding(scale = 2)

geo.cnt.tbl2

```


***
***



### Overall shape variation (PCA)

This will still use the same PCA as above, but now the post-PCA tests will include geographic classifiers (zones, basins).

#### *Zones*

**CONCLUSIONS:** Species:zone interactions for PC2 were significant, but not PC1 or either CatPC; variance was significant for PC2 with zone 1 being the only significant zone. PC2: zones 1, 2, and 4 showed differences in means. *Different from main analysis*

NOTE: Many pairwise results show a significant difference among zones, but this is an artifact from how emmeans splits data. The global test showed species:zone interaction for PC2 only.

```{r}
#quick distribution check
testmod <- lm(Dim.1~Species*Zone, data= pca.df)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(Dim.2~Species*Zone, data= pca.df)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(CatPC1~Species*Zone, data= catpca.df)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(CatPC2~Species*Zone, data= catpca.df)
plot(simulateResiduals(testmod)) #fails KS test and within-group deviation


# ───────────────────────────────────────────────────
## MEAN TESTS (GLMM for all)
# ───────────────────────────────────────────────────

### Metric Data
geo.met.mean.1 <- glmmTMB(Dim.1 ~Species*Zone, dispformula = ~Species * Zone, data = pca.df)
(gmm1 <- Anova(geo.met.mean.1, type = "III")) #only zone significant (p=0.008)
plot(simulateResiduals(geo.met.mean.1)) #all good

    post_hoc1 <- emmeans(geo.met.mean.1,  ~ Species | Zone)
    summary(post_hoc1)
    pw1 <- pairs(post_hoc1)
    (pair1 <- as.data.frame(rbind(pw1, adjust = "holm"))) #no significant zone



geo.met.mean.2 <- glmmTMB(Dim.2 ~Species*Zone, dispformula = ~Species*Zone, data = pca.df) 
(gmm2 <- Anova(geo.met.mean.2, type = "III")) #species and zone significant (p<0.001) but not species:zone
plot(simulateResiduals(geo.met.mean.2)) #all good

    post_hoc2 <- emmeans(geo.met.mean.2,  ~ Species | Zone)
    summary(post_hoc2)
    pw2 <- pairs(post_hoc2)
    (pair2 <- as.data.frame(rbind(pw2, adjust = "holm"))) #all zones sig (p<0.001)



### Meristic Data
geo.cnt.mean.1 <- glmmTMB(CatPC1 ~Species*Zone, dispformula = ~Species*Zone, data = catpca.df)
(gcm1<- Anova(geo.cnt.mean.1, type = "III")) # species (p<0.001) sig
plot(simulateResiduals(geo.cnt.mean.1)) #all good

    post_hoc4 <- emmeans(geo.cnt.mean.1,  ~ Species | Zone)
    summary(post_hoc4)
    pw4 <- pairs(post_hoc4)
    (pair4 <- as.data.frame(rbind(pw4, adjust = "holm"))) #all zones sig (p<0.001)

#default gaussian family still failed residual tests; switching to Student-t distribution due to extreme tail ends
geo.cnt.mean.2 <- glmmTMB(CatPC2 ~Species*Zone, dispformula = ~Species*Zone, family = t_family(), data = catpca.df)
(gcm2<- Anova(geo.cnt.mean.2, type = "III")) #species (p<0.001) and zone sig (p=0.046) 
plot(simulateResiduals(geo.cnt.mean.2)) #all good
    
    post_hoc5 <- emmeans(geo.cnt.mean.2,  ~ Species | Zone)
    summary(post_hoc5)
    pw5 <- pairs(post_hoc5)
    (pair5 <- as.data.frame(rbind(pw5, adjust = "holm"))) #zone 1 and zone 4 sig (p<0.001)
# ───────────────────────────────────────────────────
## VARIANCE TESTS (Extract dispersion)
# ───────────────────────────────────────────────────

### Metric Data
geo.met.var.1 <- emmeans(geo.met.mean.1, ~ Species | Zone, component = "disp", type = "response")
gmv1 <- pairs(geo.met.var.1) 
(pair6 <- as.data.frame(rbind(gmv1, adjust = "holm"))) #zone 4 sig (p<0.001)

geo.met.var.2 <- emmeans(geo.met.mean.2, ~ Species | Zone, component = "disp", type = "response")
gmv2 <- pairs(geo.met.var.2) 
(pair7 <- as.data.frame(rbind(gmv2, adjust = "holm"))) #none sig
    

### Meristic Data
geo.cnt.var.1 <- emmeans(geo.cnt.mean.1, ~ Species | Zone, component = "disp", type = "response")
gcv1 <- pairs(geo.cnt.var.1) 
(pair9 <- as.data.frame(rbind(gcv1, adjust = "holm"))) #none significant

geo.cnt.var.2 <- emmeans(geo.cnt.mean.2, ~ Species | Zone, component = "disp", type = "response")
gcv2 <- pairs(geo.cnt.var.2) 
(pair10 <- as.data.frame(rbind(gcv2, adjust = "holm"))) #none significant


# ───────────────────────────────────────────────────
## TABLE
# ───────────────────────────────────────────────────

### Metric data 

## ensure all results are as df
post_hoc1 <- as.data.frame(post_hoc1)
post_hoc2 <- as.data.frame(post_hoc2)
geo.met.var.1 <- as.data.frame(geo.met.var.1)
geo.met.var.2 <- as.data.frame(geo.met.var.2)


# Combine means and variances per dimension 
m1_all <- left_join(post_hoc1, geo.met.var.1, by = c("Species", "Zone")) %>% 
  mutate(Dimension = "PC1", 
         Mean_Str = sprintf("%.2f (%.2f)", emmean, SE.x), 
         Var_Str = sprintf("%.2f", response))

m2_all <- left_join(post_hoc2, geo.met.var.2, by = c("Species", "Zone")) %>% 
  mutate(Dimension = "PC2", 
         Mean_Str = sprintf("%.2f (%.2f)", emmean, SE.x), 
         Var_Str = sprintf("%.2f", response))

# Bind all metric dimensions together and pivot wide to separate species columns
metric_combined <- bind_rows(m1_all, m2_all) %>%
  select(Dimension, Zone, Species, Mean_Str, Var_Str) %>%
  pivot_wider(
    names_from = Species,
    values_from = c(Mean_Str, Var_Str))

# Extract Pairwise Post-hoc P-values from 'pair' data frames
p_means_met <- bind_rows(
  mutate(pair1, Dimension = "PC1"),
  mutate(pair2, Dimension = "PC2")) %>% 
  select(Dimension, Zone, p.value_mean = p.value)

p_vars_met <- bind_rows(
  mutate(pair6, Dimension = "PC1"),
  mutate(pair7, Dimension = "PC2")) %>% 
  select(Dimension, Zone, p.value_var = p.value)

# build tibble
postPCA_geo_result <- metric_combined %>%
  left_join(p_means_met, by = c("Dimension", "Zone")) %>%
  left_join(p_vars_met, by = c("Dimension", "Zone")) %>%
  mutate(
    Classifier = "Metric PCA",
    Mean_test = case_when(
      p.value_mean < 0.001 ~ "*p* < 0.001*",
      p.value_mean < 0.05  ~ paste0("*p* = ", sprintf("%.3f", p.value_mean), "*"),
      TRUE                 ~ paste0("*p* = ", sprintf("%.3f", p.value_mean))),
    Variance_test = case_when(
      p.value_var < 0.001 ~ "*p* < 0.001*",
      p.value_var < 0.05  ~ paste0("*p* = ", sprintf("%.3f", p.value_var), "*"),
      TRUE                ~ paste0("*p* = ", sprintf("%.3f", p.value_var)))) %>%
  rename(Amazon_mean = `Mean_Str_Amazon molly`, Sailfin_mean = `Mean_Str_Sailfin molly`,
         Amazon_var = `Var_Str_Amazon molly`, Sailfin_var = `Var_Str_Sailfin molly`)

### Meristic Data

## ensure all results are as df
post_hoc4 <- as.data.frame(post_hoc4)
post_hoc5 <- as.data.frame(post_hoc5)
geo.cnt.var.1 <- as.data.frame(geo.cnt.var.1)
geo.cnt.var.2 <- as.data.frame(geo.cnt.var.2)

# Combine means and variances per dimension 
c1_all <- left_join(post_hoc4, geo.cnt.var.1, by = c("Species", "Zone")) %>% 
  mutate(Dimension = "CatPC1", 
         Mean_Str = sprintf("%.2f (%.2f)", emmean, SE.x), 
         Var_Str = sprintf("%.2f", response))

c2_all <- left_join(post_hoc5, geo.cnt.var.2, by = c("Species", "Zone")) %>% 
  mutate(Dimension = "CatPC2", 
         Mean_Str = sprintf("%.2f (%.2f)", emmean, SE.x), 
         Var_Str = sprintf("%.2f", response))


# Bind all meristic dimensions together and pivot wide to separate species columns
meristic_combined <- bind_rows(c1_all, c2_all) %>%
  select(Dimension, Zone, Species, Mean_Str, Var_Str) %>%
  pivot_wider(
    names_from = Species,
    values_from = c(Mean_Str, Var_Str))

# Extract Pairwise Post-hoc P-values from 'pair' data frames
p_means_met <- bind_rows(
  mutate(pair4, Dimension = "CatPC1"),
  mutate(pair5, Dimension = "CatPC2")) %>% 
  select(Dimension, Zone, p.value_mean = p.value)

p_vars_met <- bind_rows(
  mutate(pair9, Dimension = "CatPC1"),
  mutate(pair10, Dimension = "CatPC2")) %>% 
  select(Dimension, Zone, p.value_var = p.value)

# build tibble
postCPCA_geo_result <- meristic_combined %>%
  left_join(p_means_met, by = c("Dimension", "Zone")) %>%
  left_join(p_vars_met, by = c("Dimension", "Zone")) %>%
  mutate(
    Classifier = "Meristic CatPCA",
    Mean_test = case_when(
      p.value_mean < 0.001 ~ "*p* < 0.001*",
      p.value_mean < 0.05  ~ paste0("*p* = ", sprintf("%.3f", p.value_mean), "*"),
      TRUE                 ~ paste0("*p* = ", sprintf("%.3f", p.value_mean))),
    Variance_test = case_when(
      p.value_var < 0.001 ~ "*p* < 0.001*",
      p.value_var < 0.05  ~ paste0("*p* = ", sprintf("%.3f", p.value_var), "*"),
      TRUE                ~ paste0("*p* = ", sprintf("%.3f", p.value_var)))) %>%
  rename(Amazon_mean = `Mean_Str_Amazon molly`, Sailfin_mean = `Mean_Str_Sailfin molly`,
         Amazon_var = `Var_Str_Amazon molly`, Sailfin_var = `Var_Str_Sailfin molly`)

## Grab global stats

global_mean_chisq_met <- c(gmm1["Species:Zone", "Chisq"], gmm2["Species:Zone", "Chisq"])

global_mean_p_met     <- c(gmm1["Species:Zone", "Pr(>Chisq)"], gmm2["Species:Zone", "Pr(>Chisq)"])

global_var_chisq_met <- c(Anova(geo.met.mean.1, component = "disp", type = "III")["Species:Zone", "Chisq"], Anova(geo.met.mean.2, component = "disp", type = "III")["Species:Zone", "Chisq"])

global_var_p_met <- c(Anova(geo.met.mean.1, component = "disp", type = "III")["Species:Zone", "Pr(>Chisq)"], Anova(geo.met.mean.2, component = "disp", type = "III")["Species:Zone", "Pr(>Chisq)"])


global_mean_chisq_cnt <- c(gcm1["Species:Zone", "Chisq"], gcm2["Species:Zone", "Chisq"])

global_mean_p_cnt     <- c(gcm1["Species:Zone", "Pr(>Chisq)"], gcm2["Species:Zone", "Pr(>Chisq)"])

global_var_chisq_cnt <- c(Anova(geo.cnt.mean.1, component = "disp", type = "III")["Species:Zone", "Chisq"], Anova(geo.cnt.mean.2, component = "disp", type = "III")["Species:Zone", "Chisq"])

global_var_p_cnt <- c(Anova(geo.cnt.mean.1, component = "disp", type = "III")["Species:Zone", "Pr(>Chisq)"], Anova(geo.cnt.mean.2, component = "disp", type = "III")["Species:Zone", "Pr(>Chisq)"])

# Combine global stats and map to dimensions
global_lookup <- tibble(
  Dimension = c("PC1", "PC2", "CatPC1", "CatPC2"),
  g_m_chi   = c(global_mean_chisq_met, global_mean_chisq_cnt),
  g_m_p     = c(global_mean_p_met, global_mean_p_cnt),
  g_v_chi   = c(global_var_chisq_met, global_var_chisq_cnt),
  g_v_p     = c(global_var_p_met, global_var_p_cnt))

##Combine tables
geo_results_combined <- bind_rows(postPCA_geo_result, postCPCA_geo_result) %>%
  left_join(global_lookup, by = "Dimension") %>%
  group_by(Classifier, Dimension) %>%
  mutate(
  gmp_adj = p.adjust(g_m_p, method = "holm"),
  gvp_adj = p.adjust(g_v_p, method = "holm")) %>%
  mutate(
    Mean_test = if_else(gmp_adj >= 0.05, "&mdash;", Mean_test),
    Variance_test = if_else(gvp_adj >= 0.05, "&mdash;", Variance_test),
    
    mean_str = case_when(
      gmp_adj < 0.001 ~ paste0("Mean: &chi;<sup>2</sup>=", sprintf("%.2f", g_m_chi), ", *p*<0.001*"),
      gmp_adj < 0.05  ~ paste0("Mean: &chi;<sup>2</sup>=", sprintf("%.2f", g_m_chi), ", *p*=", sprintf("%.3f", gmp_adj), "*"),
      TRUE          ~ paste0("Mean: &chi;<sup>2</sup>=", sprintf("%.2f", g_m_chi), ", *p*=", sprintf("%.3f", gmp_adj))),
    
    var_str = case_when(
      gvp_adj < 0.001 ~ paste0("Var: &chi;<sup>2</sup>=", sprintf("%.2f", g_v_chi), ", *p*<0.001*"),
      gvp_adj < 0.05  ~ paste0("Var: &chi;<sup>2</sup>=", sprintf("%.2f", g_v_chi), ", *p*=", sprintf("%.3f", gvp_adj), "*"),
      TRUE          ~ paste0("Var: &chi;<sup>2</sup>=", sprintf("%.2f", g_v_chi), ", *p*=", sprintf("%.3f", gvp_adj))),
    
    Dimension_Display = case_when(
      row_number() == 1 ~ paste0("**", Dimension, "**"),
      row_number() == 2 ~ mean_str,
      row_number() == 3 ~ var_str,
      TRUE              ~ "")) %>%
  ungroup()


#build table 
postPCA_geo_table <- geo_results_combined %>%
  select(
    Classifier,
    Dimension_Display,
    Zone,
    Amazon_mean,
    Sailfin_mean,
    Mean_test,
    Amazon_var,
    Sailfin_var,
    Variance_test) %>%
  gt(groupname_col = "Classifier") %>%
  fmt_markdown(columns = everything()) %>%
  cols_label(
    Dimension_Display = " ",
    Zone             = " ",
    Amazon_mean       = md("**Amazon**"),
    Sailfin_mean      = md("**Sailfin**"),
    Mean_test         = md("Pairwise Mean Contrast<br>(adj. *p*)"), 
    Amazon_var        = md("**Amazon**"),
    Sailfin_var       = md("**Sailfin**"),
    Variance_test     = md("Pairwise Var Contrast<br>(adj. *p*)")) %>%
  tab_spanner(
    label = html("Est. Marginal Mean (&plusmn;SE)"),
    columns = c(Amazon_mean, Sailfin_mean)) %>%
  tab_spanner(
    label = "Est. Variance",
    columns = c(Amazon_var, Sailfin_var)) %>%
  tab_options(row_group.font.weight = "bold") %>%
  tab_style(
    style = cell_borders(sides = c("top", "bottom"), color = "transparent"),
    locations = cells_body(
      columns = Dimension_Display)) %>%
  tab_style(
    style = cell_text(align = "left", size = "small"),
    locations = cells_body(
      columns = Dimension_Display,
      rows = grepl("Mean:|Var:", Dimension_Display))) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c(Dimension_Display, Zone, Sailfin_mean, Mean_test, Sailfin_var))) %>%
  tab_style(
    style = cell_borders(sides = "bottom", color = "lightgray", weight = px(2.5)),
    locations = cells_body(rows = Zone == "Zone 4")) %>%
  cols_align(align = "center", columns = -Dimension_Display) %>%
  opt_horizontal_padding(scale = 1.5)

### Final table
postPCA_geo_table

```

***
***

#### *Basins*

**CONCLUSIONS:** Species:basin interaction effects significant for both PCs and CatPC2. No consistent pattern across basins (e.g. not a select few basins with higher mean differences than the rest) suggesting these are random local differences or sampling effects. Pairwise results show all basins significant for PC1, only 2 basins significat for PC2, and 4 basins significant for CatPC2. *Different from main analysis*

Global variance is not significant for any PC or CatPc. *Same from main analysis*

```{r}
#filter for basins with >3 for each species
pca.df2 <- pca.df %>% filter(Basin %in% valid_basins) %>% droplevels()
catpca.df2 <- catpca.df %>% filter(Basin %in% valid_basins) %>% droplevels()

#quick distribution check
testmod <- lm(Dim.1~Species*Basin, data= pca.df2)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(Dim.2~Species*Basin, data= pca.df2)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(CatPC1~Species*Basin, data= catpca.df2)
plot(simulateResiduals(testmod)) #all good

testmod <- lm(CatPC2~Species*Basin, data= catpca.df2)
plot(simulateResiduals(testmod)) #fails KS test, outlier test & combined adjusted quantile test


# ───────────────────────────────────────────────────
## MEAN TESTS (GLMM for all)
# ───────────────────────────────────────────────────

### Metric Data
geo2.met.mean.1 <- glmmTMB(Dim.1 ~Species*Basin, dispformula = ~Species * Basin, data = pca.df2)
(g2mm1 <- Anova(geo2.met.mean.1, type = "III")) #basin and species:basin significant (p<0.001) 
plot(simulateResiduals(geo2.met.mean.1)) #all good

    post_hoc2.1 <- emmeans(geo2.met.mean.1,  ~ Species | Basin)
    summary(post_hoc2.1)
    pw2.1 <- pairs(post_hoc2.1)
    (pair2.1 <- as.data.frame(rbind(pw2.1, adjust = "holm"))) #significant species difference in  Upper and Lower Southwest Texas coastal (p=0.001 and p<0.001) 

geo2.met.mean.2 <- glmmTMB(Dim.2 ~Species*Basin, dispformula = ~Species*Basin, data = pca.df2) 
(g2mm2 <- Anova(geo2.met.mean.2, type = "III")) #all effects significant (all p<0.001)
plot(simulateResiduals(geo2.met.mean.2)) #all good

    post_hoc2.2 <- emmeans(geo2.met.mean.2,  ~ Species | Basin)
    summary(post_hoc2.2)
    pw2.2 <- pairs(post_hoc2.2)
    (pair2.2 <- as.data.frame(rbind(pw2.2, adjust = "holm"))) #all Basins significant (p<0.001)


### Meristic Data
geo2.cnt.mean.1 <- glmmTMB(CatPC1 ~Species*Basin, dispformula = ~Species*Basin, data = catpca.df2)
(g2cm1<- Anova(geo2.cnt.mean.1, type = "III")) # species (p<0.001) sig
plot(simulateResiduals(geo2.cnt.mean.1)) #all good

    post_hoc2.4 <- emmeans(geo2.cnt.mean.1,  ~ Species | Basin)
    summary(post_hoc2.4)
    pw2.4 <- pairs(post_hoc2.4)
    (pair2.4 <- as.data.frame(rbind(pw2.4, adjust = "holm"))) #all Basins significant (p<0.001)

    
#default gaussian family still failed residual tests; switching to Student-t distribution due to extreme tail ends
geo2.cnt.mean.2 <- glmmTMB(CatPC2 ~Species*Basin, dispformula = ~Species*Basin, family= t_family, data = catpca.df2)
(g2cm2<- Anova(geo2.cnt.mean.2, type = "III")) #all effect sig: species (p<0.001), basin (p=0.022) and species:basin (p<0.001)
plot(simulateResiduals(geo2.cnt.mean.2)) #all good

    post_hoc2.5 <- emmeans(geo2.cnt.mean.2,  ~ Species | Basin)
    summary(post_hoc2.5)
    pw2.5 <- pairs(post_hoc2.5)
    (pair2.5 <- as.data.frame(rbind(pw2.5, adjust = "holm"))) #San Antonio-Guadalupe (p<0.001), Upper SWTX Coastal (p<0.001), Lower SWTX Coastal (p=0.002), Rio Tamesi (p=0.031)

# ───────────────────────────────────────────────────
## VARIANCE TESTS (Extract dispersion)
# ───────────────────────────────────────────────────

### Metric Data
geo2.met.var.1 <- emmeans(geo2.met.mean.1, ~ Species | Basin, component = "disp", type = "response")
gmv1 <- pairs(geo2.met.var.1) 
(pair2.6 <- as.data.frame(rbind(gmv1, adjust = "holm"))) #San Antonio-Guadalupe (p<0.001), Upper SWTX Coastal (p<0.001) and Rio Tamesi (p=0.033) significant


geo2.met.var.2 <- emmeans(geo2.met.mean.2, ~ Species | Basin, component = "disp", type = "response")
gmv2 <- pairs(geo2.met.var.2) 
(pair2.7 <- as.data.frame(rbind(gmv2, adjust = "holm"))) #none significant
   
    

### Meristic Data
geo2.cnt.var.1 <- emmeans(geo2.cnt.mean.1, ~ Species | Basin, component = "disp", type = "response")
gcv1 <- pairs(geo2.cnt.var.1) 
(pair2.9 <- as.data.frame(rbind(gcv1, adjust = "holm"))) #none significant

geo2.cnt.var.2 <- emmeans(geo2.cnt.mean.2, ~ Species | Basin, component = "disp", type = "response")
gcv2 <- pairs(geo2.cnt.var.2) 
(pair2.10 <- as.data.frame(rbind(gcv2, adjust = "holm"))) #none significant


# ───────────────────────────────────────────────────
## TABLE
# ───────────────────────────────────────────────────

### Metric data 

## ensure all results are as df
post_hoc2.1 <- as.data.frame(post_hoc2.1)
post_hoc2.2 <- as.data.frame(post_hoc2.2)
geo2.met.var.1 <- as.data.frame(geo2.met.var.1)
geo2.met.var.2 <- as.data.frame(geo2.met.var.2)


# Combine means and variances per dimension 
m1_all <- left_join(post_hoc2.1, geo2.met.var.1, by = c("Species", "Basin")) %>% 
  mutate(Dimension = "PC1", 
         Mean_Str = sprintf("%.2f (%.2f)", emmean, SE.x), 
         Var_Str = sprintf("%.2f", response))

m2_all <- left_join(post_hoc2.2, geo2.met.var.2, by = c("Species", "Basin")) %>% 
  mutate(Dimension = "PC2", 
         Mean_Str = sprintf("%.2f (%.2f)", emmean, SE.x), 
         Var_Str = sprintf("%.2f", response))


# Bind all metric dimensions together and pivot wide to separate species columns
metric_combined <- bind_rows(m1_all, m2_all) %>%
  select(Dimension, Basin, Species, Mean_Str, Var_Str) %>%
  pivot_wider(
    names_from = Species,
    values_from = c(Mean_Str, Var_Str))

# Extract Pairwise Post-hoc P-values from 'pair' data frames
p_means_met <- bind_rows(
  mutate(pair2.1, Dimension = "PC1"),
  mutate(pair2.2, Dimension = "PC2")) %>% 
  select(Dimension, Basin, p.value_mean = p.value)

p_vars_met <- bind_rows(
  mutate(pair2.6, Dimension = "PC1"),
  mutate(pair2.7, Dimension = "PC2")) %>% 
  select(Dimension, Basin, p.value_var = p.value)

# build tibble
postPCA_geo2_result <- metric_combined %>%
  left_join(p_means_met, by = c("Dimension", "Basin")) %>%
  left_join(p_vars_met, by = c("Dimension", "Basin")) %>%
  mutate(
    Classifier = "Metric PCA",
    Mean_test = case_when(
      p.value_mean < 0.001 ~ "*p* < 0.001*",
      p.value_mean < 0.05  ~ paste0("*p* = ", sprintf("%.3f", p.value_mean), "*"),
      TRUE                 ~ paste0("*p* = ", sprintf("%.3f", p.value_mean))),
    Variance_test = case_when(
      p.value_var < 0.001 ~ "*p* < 0.001*",
      p.value_var < 0.05  ~ paste0("*p* = ", sprintf("%.3f", p.value_var), "*"),
      TRUE                ~ paste0("*p* = ", sprintf("%.3f", p.value_var)))) %>%
  rename(Amazon_mean = `Mean_Str_Amazon molly`, Sailfin_mean = `Mean_Str_Sailfin molly`,
         Amazon_var = `Var_Str_Amazon molly`, Sailfin_var = `Var_Str_Sailfin molly`)

### Meristic Data

## ensure all results are as df
post_hoc2.4 <- as.data.frame(post_hoc2.4)
post_hoc2.5 <- as.data.frame(post_hoc2.5)
geo2.cnt.var.1 <- as.data.frame(geo2.cnt.var.1)
geo2.cnt.var.2 <- as.data.frame(geo2.cnt.var.2)

# Combine means and variances per dimension 
c1_all <- left_join(post_hoc2.4, geo2.cnt.var.1, by = c("Species", "Basin")) %>% 
  mutate(Dimension = "CatPC1", 
         Mean_Str = sprintf("%.2f (%.2f)", emmean, SE.x), 
         Var_Str = sprintf("%.2f", response))

c2_all <- left_join(post_hoc2.5, geo2.cnt.var.2, by = c("Species", "Basin")) %>% 
  mutate(Dimension = "CatPC2", 
         Mean_Str = sprintf("%.2f (%.2f)", emmean, SE.x), 
         Var_Str = sprintf("%.2f", response))


# Bind all meristic dimensions together and pivot wide to separate species columns
meristic_combined <- bind_rows(c1_all, c2_all) %>%
  select(Dimension, Basin, Species, Mean_Str, Var_Str) %>%
  pivot_wider(
    names_from = Species,
    values_from = c(Mean_Str, Var_Str))

# Extract Pairwise Post-hoc P-values from 'pair' data frames
p_means_met <- bind_rows(
  mutate(pair2.4, Dimension = "CatPC1"),
  mutate(pair2.5, Dimension = "CatPC2")) %>% 
  select(Dimension, Basin, p.value_mean = p.value)

p_vars_met <- bind_rows(
  mutate(pair2.9, Dimension = "CatPC1"),
  mutate(pair2.10, Dimension = "CatPC2")) %>% 
  select(Dimension, Basin, p.value_var = p.value)

# build tibble
postCPCA_geo2_result <- meristic_combined %>%
  left_join(p_means_met, by = c("Dimension", "Basin")) %>%
  left_join(p_vars_met, by = c("Dimension", "Basin")) %>%
  mutate(
    Classifier = "Meristic CatPCA",
    Mean_test = case_when(
      p.value_mean < 0.001 ~ "*p* < 0.001*",
      p.value_mean < 0.05  ~ paste0("*p* = ", sprintf("%.3f", p.value_mean), "*"),
      TRUE                 ~ paste0("*p* = ", sprintf("%.3f", p.value_mean))),
    Variance_test = case_when(
      p.value_var < 0.001 ~ "*p* < 0.001*",
      p.value_var < 0.05  ~ paste0("*p* = ", sprintf("%.3f", p.value_var), "*"),
      TRUE                ~ paste0("*p* = ", sprintf("%.3f", p.value_var)))) %>%
  rename(Amazon_mean = `Mean_Str_Amazon molly`, Sailfin_mean = `Mean_Str_Sailfin molly`,
         Amazon_var = `Var_Str_Amazon molly`, Sailfin_var = `Var_Str_Sailfin molly`)

  
## Grab global stats

global_mean_chisq_met <- c(g2mm1["Species:Basin", "Chisq"], g2mm2["Species:Basin", "Chisq"])

global_mean_p_met     <- c(g2mm1["Species:Basin", "Pr(>Chisq)"], g2mm2["Species:Basin", "Pr(>Chisq)"])

global_var_chisq_met <- c(Anova(geo2.met.mean.1, component = "disp", type = "III")["Species:Basin", "Chisq"], Anova(geo2.met.mean.2, component = "disp", type = "III")["Species:Basin", "Chisq"])

global_var_p_met <- c(Anova(geo2.met.mean.1, component = "disp", type = "III")["Species:Basin", "Pr(>Chisq)"], Anova(geo2.met.mean.2, component = "disp", type = "III")["Species:Basin", "Pr(>Chisq)"])

global_mean_chisq_cnt <- c(g2cm1["Species:Basin", "Chisq"], g2cm2["Species:Basin", "Chisq"])

global_mean_p_cnt     <- c(g2cm1["Species:Basin", "Pr(>Chisq)"], g2cm2["Species:Basin", "Pr(>Chisq)"])

global_var_chisq_cnt <- c(Anova(geo2.cnt.mean.1, component = "disp", type = "III")["Species:Basin", "Chisq"], Anova(geo2.cnt.mean.2, component = "disp", type = "III")["Species:Basin", "Chisq"])

global_var_p_cnt <- c(Anova(geo2.cnt.mean.1, component = "disp", type = "III")["Species:Basin", "Pr(>Chisq)"], Anova(geo2.cnt.mean.2, component = "disp", type = "III")["Species:Basin", "Pr(>Chisq)"])

# Combine global stats and map to dimensions
global_lookup <- tibble(
  Dimension = c("PC1", "PC2", "CatPC1", "CatPC2"),
  g_m_chi   = c(global_mean_chisq_met, global_mean_chisq_cnt),
  g_m_p     = c(global_mean_p_met, global_mean_p_cnt),
  g_v_chi   = c(global_var_chisq_met, global_var_chisq_cnt),
  g_v_p     = c(global_var_p_met, global_var_p_cnt))


##Combine tables
geo2_results_combined <- bind_rows(postPCA_geo2_result, postCPCA_geo2_result) %>%
  left_join(global_lookup, by = "Dimension") %>%
  group_by(Classifier, Dimension) %>%
   mutate(
  gmp_adj = p.adjust(g_m_p, method = "holm"),
  gvp_adj = p.adjust(g_v_p, method = "holm")) %>%
  mutate(
    Mean_test = if_else(gmp_adj >= 0.05, "&mdash;", Mean_test),
    Variance_test = if_else(gvp_adj >= 0.05, "&mdash;", Variance_test),
    
    mean_str = case_when(
      gmp_adj < 0.001 ~ paste0("Mean: &chi;<sup>2</sup>=", sprintf("%.2f", g_m_chi), ", *p*<0.001*"),
      gmp_adj < 0.05  ~ paste0("Mean: &chi;<sup>2</sup>=", sprintf("%.2f", g_m_chi), ", *p*=", sprintf("%.3f", gmp_adj), "*"),
      TRUE          ~ paste0("Mean: &chi;<sup>2</sup>=", sprintf("%.2f", g_m_chi), ", *p*=", sprintf("%.3f", gmp_adj))),
    
    var_str = case_when(
      gvp_adj < 0.001 ~ paste0("Var: &chi;<sup>2</sup>=", sprintf("%.2f", g_v_chi), ", *p*<0.001*"),
      gvp_adj < 0.05  ~ paste0("Var: &chi;<sup>2</sup>=", sprintf("%.2f", g_v_chi), ", *p*=", sprintf("%.3f", gvp_adj), "*"),
      TRUE          ~ paste0("Var: &chi;<sup>2</sup>=", sprintf("%.2f", g_v_chi), ", *p*=", sprintf("%.3f", gvp_adj))),
    
    Dimension_Display = case_when(
      row_number() == 1 ~ paste0("**", Dimension, "**"),
      row_number() == 2 ~ mean_str,
      row_number() == 3 ~ var_str,
      TRUE              ~ "")) %>%
  ungroup()

#build table 
postPCA_geo2_table <- geo2_results_combined %>%
  select(
    Classifier,
    Dimension_Display,
    Basin,
    Amazon_mean,
    Sailfin_mean,
    Mean_test,
    Amazon_var,
    Sailfin_var,
    Variance_test) %>%
  gt(groupname_col = "Classifier") %>%
  fmt_markdown(columns = everything()) %>%
  cols_label(
    Dimension_Display = " ",
    Basin             = " ",
    Amazon_mean       = md("**Amazon**"),
    Sailfin_mean      = md("**Sailfin**"),
    Mean_test         = md("Pairwise Mean Contrast<br>(adj. *p*)"), 
    Amazon_var        = md("**Amazon**"),
    Sailfin_var       = md("**Sailfin**"),
    Variance_test     = md("Pairwise Var Contrast<br>(adj. *p*)")) %>%
  tab_spanner(
    label = html("Est. Marginal Mean (&plusmn;SE)"),
    columns = c(Amazon_mean, Sailfin_mean)) %>%
  tab_spanner(
    label = "Est. Variance",
    columns = c(Amazon_var, Sailfin_var)) %>%
  tab_options(row_group.font.weight = "bold") %>%
  tab_style(
    style = cell_borders(sides = c("top", "bottom"), color = "transparent"),
    locations = cells_body(
      columns = Dimension_Display)) %>%
  tab_style(
    style = cell_text(align = "left", size = "small"),
    locations = cells_body(
      columns = Dimension_Display,
      rows = grepl("Mean:|Var:", Dimension_Display))) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c(Dimension_Display, Basin, Sailfin_mean, Mean_test, Sailfin_var))) %>%
  tab_style(
    style = cell_borders(sides = "bottom", color = "lightgray", weight = px(2.5)),
    locations = cells_body(rows = Basin == "Río Tamesí")) %>%
  cols_align(align = "center", columns = -Dimension_Display) %>%
  opt_horizontal_padding(scale = 1.5)

### Final table
postPCA_geo2_table





```
***
***

## **Tampico MX Analysis**

For this analysis, we will compare sailfin, amazon and atlantic mollies from the Tampico, Mexico region. This is the proposed origin of Amazon mollies, and the only region in which all three species coexist. 

After filtering for individuals where the county/state information was recorded as Tamaulipas/MX, we have 32 sailfin molly, 28 Amazon molly, and 32 Atlantic molly.

### Individual trait models

**CONCLUSIONS:** Body depth, predorsal length, and dorsal base length were best supported with an unequal variance model; in all cases, the sexual species (sailfin & atlantic) were more variable than the amazon except for predorsal length, in which amazons were more variable than sailfin (atlantic still more variable than amazon). All remaining metric traits were best supported with an equal variance model (aka no significant difference in variation among the species). About half of the meristic traits failed model comparisons due to lack of variation; fallback Levene's tests and models that did converge show no significant variation differences for any traits. *Slightly different from main analysis*


```{r}
#create separate df for metric and meristic (count) data
metric3 <- geo %>%
  select(-c(5:12)) %>%
  select(-c("FLA"))

count3 <- geo %>%
  select(-c(14:23)) #still includes SL for analysis



# ───────────────────────────────────────────────────
## METRIC ANALYSIS 
# ───────────────────────────────────────────────────

# Ensure Species is a factor
metric3$Species <- as.factor(metric3$Species)
metric3$Species <- relevel(metric3$Species, ref = "Amazon molly")

tam_results_list <- lapply(met.vars, function(trait) {
  
  # Build the formula string correctly
  form_str <- paste0("log(", trait, ") ~ log(SL) * Species")
  form <- as.formula(form_str)

  # 1. Homogeneous variance model (with safety catch)
  mod1 <- tryCatch({
    glmmTMB(form, data = metric3, family = gaussian)
  }, error = function(e) { NULL }, warning = function(w) { NULL })
  
  mod2 <- tryCatch({
    withCallingHandlers({
      glmmTMB(form, data = metric3, family = gaussian, dispformula = ~Species)
    }, warning = function(w) {
      # If it's a bad warning, escalate to an error to abort the model gracefully
      if(grepl("NA/NaN|calc_loglambda|iteration|failure", w$message, ignore.case = TRUE)) {
        stop(w$message) 
      }
      # Otherwise, safely muffle minor warnings while the stack is still active
      invokeRestart("muffleWarning") 
    })
  }, error = function(e) { NULL })
  
  # Verify clean convergence of both models
  models_valid <- !is.null(mod1) && !is.null(mod2) && 
                  !is.na(AIC(mod1)) && !is.na(AIC(mod2)) && 
                  !any(is.nan(summary(mod2)$coefficients$disp[,2]))

  unique_species <- levels(as.factor(metric3$Species))
  
  # --- PATH A: MODEL CONVERGED CLEANLY ---
  if (models_valid) {
    # Compare models
    comp <- anova(mod1, mod2)
    pval <- comp$`Pr(>Chisq)`[2]
    delta_aic <- AIC(mod2) - AIC(mod1)
    
    # Run DHARMa residual check
    best_mod <- if (AIC(mod1) < AIC(mod2)) mod1 else mod2
    best_mod_name <- if (AIC(mod1) < AIC(mod2)) "Homogeneous (mod1)" else "Heterogeneous (mod2)"
    
    message(paste0("--- DHARMa Diagnostics for Trait: ", trait, " (Best Model: ", best_mod_name, ") ---"))
    sim_res <- simulateResiduals(best_mod, plot = FALSE)
    
    # Plot the residuals automatically
    plot(sim_res) 
    title(main = paste(trait, "-", best_mod_name), line = 2.5)
    
    # Extract predicted log-variance per species
    rg <- ref_grid(mod2, component = "disp")
    disp_means <- as.data.frame(emmeans(rg, specs = ~ Species))
    disp_means$Variance <- exp(disp_means$emmean)
    
    # Pivot to easily calculate the single global ratio
    ratio_df <- disp_means %>%
      select(Species, Variance) %>%
      pivot_wider(names_from = Species, values_from = Variance) %>%
      mutate(
        SA_Ratio  = `Sailfin molly` / `Amazon molly`,
        AA_Ratio  = `Atlantic molly` / `Amazon molly`,
        Delta_AIC  = delta_aic, 
        LRT_p      = pval,
        Best_Model = best_mod_name)
      
  } else {
    # --- PATH B: FALLBACK TO SIZE-CORRECTED LEVENE'S TEST ---
    best_mod_name <- "Model Failed - Levene's Test Fallback"
    message(paste0("!!! glmmTMB Failed for Trait: ", trait, ". Running Levene's Fallback. !!!"))
    
    # Step 1: Strip out body size dependencies using linear model on log traits
    lm_check <- lm(as.formula(paste0("log(", trait, ") ~ log(SL)")), data = metric3)
    metric$Abs_Resid <- abs(residuals(lm_check))
    
    # Step 2: Run a formal Levene's-style test via anova on absolute deviations
    lev_test <- tryCatch({
      leveneTest(Abs_Resid ~ Species, data = metric3)
    }, error = function(e) { NULL })
    fallback_p <- if(!is.null(lev_test)) lev_test$`Pr(>F)`[1] else NA
    
    # Step 3: Calculate empirical size-corrected variances per group
    empirical_vars <- metric3 %>%
      group_by(Species) %>%
      summarise(Variance = var(log(!!sym(trait)), na.rm = TRUE), .groups = 'drop') %>%
      complete(Species = unique_species, fill = list(Variance = NA))
    
    ratio_df <- empirical_vars %>%
      pivot_wider(names_from = Species, values_from = Variance) %>%
      mutate(
        SA_Ratio  = `Sailfin molly` / `Amazon molly`,
        AA_Ratio  = `Atlantic molly` / `Amazon molly`,
        Delta_AIC  = NA, # Marked NA so it shows as blank/dash in your table
        LRT_p      = fallback_p,
        Best_Model = best_mod_name)
  }
  
  # Append baseline trait tracker string
  ratio_df$Trait <- trait
  return(ratio_df)
})

tam_results_df <- do.call(rbind, tam_results_list)

tam_results_df$p_adj <- p.adjust(tam_results_df$LRT_p,
                             method = "holm")

tam.met.results <- tam_results_df %>%
  mutate(
    p.display = case_when(
      p_adj < 0.001 ~ paste0("*p* < 0.001*"),
      p_adj < 0.05  ~ paste0(sprintf("%.3f", p_adj), "*"),
      TRUE ~ sprintf("%.3f", p_adj))) %>%
  mutate(
    Trait_Name = factor(
      trait.names[Trait],
      levels = unname(trait.names))) %>%
  arrange(Trait_Name) %>%
  relocate(Trait_Name, .before = everything())

tam.met.tbl <- tam.met.results %>%
  gt() %>%
  cols_hide(c(Trait, LRT_p, p_adj, "Sailfin molly", "Amazon molly", "Atlantic molly", Best_Model)) %>%
  cols_label(
    Trait_Name = "Trait",
    SA_Ratio = md("Sailfin / Amazon"),
    AA_Ratio = md("Atlantic / Amazon"),
    Delta_AIC = md("&Delta;AIC"),
    p.display = md("adj. *p*-value")) %>%
  tab_spanner(
    label = "Variance Ratio",
    columns = c(SA_Ratio, AA_Ratio))%>%
  cols_move(
    columns = c(Delta_AIC, p.display),
    after = Trait_Name) %>%
  fmt_number(
    columns = c(SA_Ratio, AA_Ratio, Delta_AIC),
    decimals = 2) %>%
  fmt_markdown(columns = p.display) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Trait_Name, Delta_AIC, SA_Ratio, AA_Ratio))) %>%
  cols_align("center", everything()) %>%
  opt_horizontal_padding(scale = 2)
  
  
tam.met.tbl



# ───────────────────────────────────────────────────
## MERISTIC (COUNT) ANALYSIS 
# ───────────────────────────────────────────────────
### NOTE: issues with convergence since variation is so small and data is truncated

tam_results_list_count <- lapply(cnt.vars, function(trait) {
  
  form_str <- paste0(trait, " ~ SL * Species")
  form <- as.formula(form_str)
  
  # Initialize status tracking flags
  mod2_failed <- FALSE
  levene_failed <- FALSE
  
  # 1. Fit Baseline Homogeneous Model
  mod1 <- tryCatch({
    glmmTMB(form, data = count3, family = compois)
  }, error = function(e) { NULL }, warning = function(w) { NULL }) # Rejects if optimizer warns
  
  mod2 <- tryCatch({
    withCallingHandlers({
      glmmTMB(form, data = count3, family = compois, dispformula = ~Species)
    }, warning = function(w) {
      # Check internal optimization/evaluation warnings and escalate to fail cleanly
      if(grepl("NA/NaN|calc_loglambda|iteration|failure", w$message, ignore.case = TRUE)) {
        stop(w$message) 
      }
      # Muffle minor, harmless warnings safely
      invokeRestart("muffleWarning") 
    })
  }, error = function(e) { NULL })

  # Verify if both models successfully and cleanly converged
  models_valid <- !is.null(mod1) && !is.null(mod2) && 
                  !is.na(AIC(mod1)) && !is.na(AIC(mod2)) && 
                  !any(is.nan(summary(mod2)$coefficients$disp[,2]))

  unique_species <- levels(as.factor(count3$Species))
  
  # --- PATH A: GLMMTMB CONVERGED CLEANLY ---
  if (models_valid) {
    comp <- anova(mod1, mod2)
    pval <- comp$`Pr(>Chisq)`[2]
    delta_aic <- AIC(mod2) - AIC(mod1)
    
    best_mod <- if (AIC(mod1) < AIC(mod2)) mod1 else mod2
    best_mod_name <- if (AIC(mod1) < AIC(mod2)) "Homogeneous (mod1)" else "Heterogeneous (mod2)"
    
    # Run DHARMa diagnostics safely
    message(paste0("--- DHARMa Diagnostics for Trait: ", trait, " (Best Model: ", best_mod_name, ") ---"))
    sim_res <- simulateResiduals(best_mod, plot = FALSE)
    plot(sim_res) 
    title(main = paste(trait, "-", best_mod_name), line = 2.5)
    
    # Extract Model Variances
    mean_sl <- mean(count3$SL, na.rm = TRUE)
    pred_data <- data.frame(Species = unique_species, SL = mean_sl)
    pred_data$Mean <- predict(mod2, newdata = pred_data, type = "response")
    pred_data$Phi  <- exp(predict(mod2, newdata = pred_data, type = "disp"))
    pred_data$Variance <- pred_data$Mean * pred_data$Phi
    
    output_df <- pred_data %>%
      select(Species, Variance) %>%
      pivot_wider(names_from = Species, values_from = Variance) %>%
      mutate(
        SA_Ratio   = `Sailfin molly` / `Amazon molly`,
        AA_Ratio   = `Atlantic molly` / `Amazon molly`,
        Delta_AIC  = delta_aic,
        LRT_p      = pval,
        Best_Model = best_mod_name)
      
  } else {
    # --- PATH B: CONVERGENCE FAILED -> FALLBACK TO LEVENE'S TEST ---
    best_mod_name <- "Model Failed Optimization - Used Levene's Test Fallback"
    message(paste0("!!! glmmTMB Failed for Trait: ", trait, ". Running Levene's Test Fallback. !!!"))
    
    # Run Levene's Test on the raw response variable across species
    levene_form <- as.formula(paste0(trait, " ~ Species"))
    lev_test <- tryCatch({
      leveneTest(levene_form, data = count3)
    }, error = function(e) { NULL })
    
    # Extract empirical variances directly from the raw data
    raw_vars <- count3 %>%
      group_by(Species) %>%
      summarise(Variance = var(!!sym(trait), na.rm = TRUE), .groups = 'drop') %>%
      complete(Species = unique_species, fill = list(Variance = NA)) # handle missing classes
    
    # Calculate ratios using the empirical data
    wide_vars <- raw_vars %>%
      pivot_wider(names_from = Species, values_from = Variance)
    
    # Pull p-value from Levene's test table
    l_p <- if(!is.null(lev_test)) lev_test$`Pr(>F)`[1] else NA
    
    output_df <- wide_vars %>%
      mutate(
        SA_Ratio   = `Sailfin molly` / `Amazon molly`,
        AA_Ratio   = `Atlantic molly` / `Amazon molly`,
        Delta_AIC  = NA,   # AIC is invalid because the likelihood engine failed
        LRT_p      = l_p,  # Storing Levene's p-value here so it seamlessly passes into p_adj
        Best_Model = best_mod_name)
  }
  
  # Append the raw trait identifier string
  output_df$Trait <- trait
  return(output_df)
})


# Combine into a clean data frame
tam_cnt_df <- do.call(rbind, tam_results_list_count)

tam_cnt_df$p_adj <- p.adjust(tam_cnt_df$LRT_p,
                             method = "holm")

tam.cnt.results <- tam_cnt_df %>%
  mutate(
    p.display = case_when(
      p_adj < 0.001 ~ paste0("*p* < 0.001*"),
      p_adj < 0.05  ~ paste0(sprintf("%.3f", p_adj), "*"),
      TRUE ~ sprintf("%.3f", p_adj))) %>%
  mutate(
    Trait_Name = factor(
      trait.names[Trait],
      levels = unname(trait.names))) %>%
  arrange(Trait_Name) %>%
  relocate(Trait_Name, .before = everything())

tam.cnt.tbl <- tam.cnt.results %>%
  gt() %>%
  cols_hide(c(Trait, LRT_p, p_adj, "Sailfin molly", "Amazon molly", "Atlantic molly", Best_Model)) %>%
  cols_label(
    Trait_Name = "Trait",
    SA_Ratio = md("Sailfin / Amazon"),
    AA_Ratio = md("Atlantic / Amazon"),
    Delta_AIC = md("&Delta;AIC"),
    p.display = md("adj. *p*-value")) %>%
  tab_spanner(
    label = "Variance Ratio",
    columns = c(SA_Ratio, AA_Ratio)) %>%
  cols_move(
    columns = c(Delta_AIC, p.display),
    after = Trait_Name) %>%
  fmt_number(
    columns = c(SA_Ratio, AA_Ratio, Delta_AIC),
    decimals = 2) %>%
  fmt_markdown(columns = p.display) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Trait_Name, Delta_AIC, SA_Ratio, AA_Ratio))) %>%
  cols_align("center", everything()) %>%
  opt_horizontal_padding(scale = 2)
  
  
tam.cnt.tbl

# ───────────────────────────────────────────────────
## PLOTS
# ───────────────────────────────────────────────────

met.long2 <- metric3 %>%
  pivot_longer(
    cols = c(SL, BD, CPD, CPL, PreDL, DbL, HL, HD, HW, SnL, OL),
    names_to = "Trait",
    values_to = "Value")


tam.met.plots <- ggplot(met.long2, aes(x=Species, y=Value, fill = Species)) + 
  geom_boxplot() +
  geom_point(color= "#8D8680", size=2, alpha=0.5) +
  facet_wrap(~ Trait, scales = "free_y", axes = "margins") +
  theme_classic(base_size = 10, base_family = "sans") + 
  theme(
    axis.title.y = element_text(size = 10, family = "sans"), 
    axis.text.y  = element_text(size = 10, family = "sans"),
    axis.text.x  = element_text(size = 10, family = "sans"),
    strip.text    = element_text(size = 10, family = "sans"),
    panel.spacing = unit(1.5, "lines")) +
  scale_fill_manual(values = c("Sailfin molly"="#78B7C5",
                               "Amazon molly"="#E1AF00",
                               "Atlantic molly"="#9986A5")) +
  scale_x_discrete(labels = c("Sailfin molly" = "Sailfin\nmolly", 
                              "Amazon molly" = "Amazon\nmolly", 
                              "Atlantic molly" = "Atlantic\nmolly")) +
  xlab("") +
  ylab("Length or Depth (mm)")
tam.met.plots


cnt.long2 <- count3 %>%
  pivot_longer(
    cols = c(D, P1, A, P1.R, LLSC, SALL, SBLL, SBDF, FLA),
    names_to = "Trait",
    values_to = "Value")

tam.cnt.plots <- ggplot(cnt.long2, aes(x=Species, y=Value, fill = Species)) + 
  geom_boxplot() +
  geom_point(color= "#8D8680", size=2, alpha=0.5) +
  facet_wrap(~ Trait, scales = "free_y", axes = "margins") +
  theme_classic(base_size = 10, base_family = "sans") + 
  theme(
    axis.title.y = element_text(size = 10, family = "sans"), 
    axis.text.y  = element_text(size = 10, family = "sans"),
    axis.text.x  = element_text(size = 10, family = "sans"),
    strip.text    = element_text(size = 10, family = "sans"),
    panel.spacing = unit(1.5, "lines")) +
  scale_fill_manual(values = c("Sailfin molly"="#78B7C5",
                               "Amazon molly"="#E1AF00",
                               "Atlantic molly"="#9986A5")) +
  scale_x_discrete(labels = c("Sailfin molly" = "Sailfin\nmolly", 
                              "Amazon molly" = "Amazon\nmolly", 
                              "Atlantic molly" = "Atlantic\nmolly")) +
  xlab("") +
  ylab("Count")
tam.cnt.plots


```

***
***

### Overall shape variation (PCA)

As before, we will assess overall shape variation among species using a ***principal component analysis (PCA)*** on centered and scaled data (all metric traits will also be log transformed to better approach normality). We will then perform a mean and variance test (e.g., ANOVA and Levene's test) to determine if overall shape variation differs between the species. 

As meristic traits are count data, a normal PCA cannot be applied. Instead, we will use a ***categorical principal component analysis (CATPCA)*** that uses optimal scaling to deal with discrete integers. 

#### *Loadings*

```{r}

# ───────────────────────────────────────────────────
## METRIC ANALYSIS 
# ───────────────────────────────────────────────────

#get size-corrected residuals
shape_residuals <- lapply(met.vars, function(trait) {
  # Fit a simple linear model of the raw trait against SL
  mod <- lm(metric3[[trait]] ~ metric3$SL)
  return(residuals(mod))
})

residual_df2 <- as.data.frame(shape_residuals)
colnames(residual_df2) <- met.vars

# run PCA
PCA.test2 <- prcomp(residual_df2, center = TRUE, scale. = TRUE)

summary(PCA.test2)
(eig.val.test <- get_eigenvalue(PCA.test2))


# Since we corrected for size, we will focus on only the first 2 dimensions

ind.test <- get_pca_ind(PCA.test2)
head(ind.test$coord[,1:2])


pca.df2 <- cbind(metric3, ind.test$coord[,1:2])


# Will investigate the influence of each trait on the first 2 dimensions
(loadings <- PCA.test2$rotation[, 1:2])

sorted.loadings.1 <- loadings[order(loadings[, 1]), 1]
myTitle <- "Loadings Plot for PC1" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.1, main=myTitle, xlab=myXlab, cex=0.5, col="blue")

sorted.loadings.2 <- loadings[order(loadings[, 2]), 2]
myTitle <- "Loadings Plot for PC2" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.2, main=myTitle, xlab=myXlab, cex=0.5, col="blue")


# ───────────────────────────────────────────────────
## MERISTIC (COUNT) ANALYSIS 
# ───────────────────────────────────────────────────

# Remove SL from count df
count4 <- select(count3, -c("SL"))

# Run the Categorical PCA
# 'ndim = 2' extracts the first two principal components
# 'degrees = 1' treats the scaling transformations as piecewise linear (ideal for ordinal integers)

catpca.test2 <- princals(count4[, 5:13], ndim = 2, degrees = 1)

# Print the summary to see how much variance your new components explain
summary(catpca.test2)

(cat.loadings <- catpca.test2$loadings)

sorted.loadings.4 <- cat.loadings[order(cat.loadings[, 1]), 1]
myTitle <- "Loadings Plot for CATPC1" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.4, main=myTitle, xlab=myXlab, cex=0.5, col="blue")

sorted.loadings.5 <- cat.loadings[order(cat.loadings[, 2]), 2]
myTitle <- "Loadings Plot for CATPC2" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.5, main=myTitle, xlab=myXlab, cex=0.5, col="blue")

# Extract the eigenvalues (one per dimension)
eigenvals <- catpca.test2$evals

# Calculate the percentage of variation explained by each axis
# (Divided by 9 because you have 9 meristic traits)
pc1_percent <- round((eigenvals[1] / 9) * 100, 1)
pc2_percent <- round((eigenvals[2] / 9) * 100, 1)

# Quick print check to see your numbers (e.g., 24.5 and 14.2)
print(paste0("CatPC1: ", pc1_percent, "%"))
print(paste0("CatPC2: ", pc2_percent, "%"))

# 1. Extract the object scores for the first two dimensions
cat_pc_scores <- as.data.frame(catpca.test2$objectscores)
colnames(cat_pc_scores) <- c("CatPC1", "CatPC2")

# 2. Combine them back with your species and size metadata
catpca.df2 <- cbind(count4, cat_pc_scores)

```

#### *Plots*

```{r}

#### METRIC

tam.pca.1<- autoplot(PCA.test2, data = metric3, colour="Species", shape= 'Species', loadings=FALSE, loadings.label=FALSE, frame=TRUE, frame.type='norm', label= FALSE)+
  scale_color_manual(name="Species",
                    values = c("Sailfin molly"="#78B7C5", "Amazon molly"="#E1AF00", "Atlantic molly"="#9986A5")) +
  scale_fill_manual(name="Species",
                    values = c("Sailfin molly"="#78B7C5", "Amazon molly"="#E1AF00", "Atlantic molly"="#9986A5")) +
  scale_shape_manual(name= 'Species', 
                     values = c("Sailfin molly" = 17, "Amazon molly"=16, "Atlantic molly"= 3)) +
   guides(fill = guide_legend(override.aes = list(fill = NA)))+
  ggtitle("PCA Plot of Morphology traits") + theme_classic() 
tam.pca.1



#### MERISTIC

tam.catpca <- ggplot(catpca.df2, aes(x = CatPC1, y = CatPC2, color = Species, fill = Species, shape = Species)) +
  geom_point(size = 1, alpha = 0.8) +
  stat_ellipse(geom = "polygon", alpha = 0.15, type = "norm", level = 0.95) +
  scale_color_manual(name = "Species",
                     values = c("Sailfin molly" = "#78B7C5", "Amazon molly" = "#E1AF00", "Atlantic molly"="#9986A5")) +
  scale_fill_manual(name = "Species",
                    values = c("Sailfin molly" = "#78B7C5", "Amazon molly" = "#E1AF00", "Atlantic molly"="#9986A5")) +
  scale_shape_manual(name = "Species", 
                     values = c("Sailfin molly" = 17, "Amazon molly" = 16, "Atlantic molly"= 3)) +
  guides(fill = guide_legend(override.aes = list(fill = NA))) +
  labs(
    x = "Categorical PC1",
    y = "Categorical PC2"
  ) +
  theme_classic() +
  labs(
    x = paste0("Categorical PC1 (", pc1_percent, "%)"),
    y = paste0("Categorical PC2 (", pc2_percent, "%)")
  )

tam.catpca


```


#### *Post PCA tests*

**CONCLUSIONS:** PC1 and CatPC1 have significant mean differences but not PC2 or CatPC2. This is expected, as all three species have different body shapes and therefore occupy distinct morphospace. *Different from main analysis*

CatPC1 shows a significant difference in variation among the species, with sailfins exhibiting the highest amount of variation followed by atlantic; Amazons contain the lowest amount of variation. *Slightly different from main analysis*

Only sailfin and atlantic differ for CatPC1. *Slightly different from main analysis* This differs slightly from our individual trait results: in individual trait results, only a few metric traits showed support for a unequal-dispersion model, and none of our meristic traits showed support for unequal-disperson models. This indicates species differences in covariation among multiple  meristic features. 

```{r}
testmod <- lm(Dim.1 ~ Species, data = pca.df2)
plot(simulateResiduals(testmod)) #levene's test sig

testmod <- lm(Dim.2 ~ Species, data = pca.df2)
plot(simulateResiduals(testmod)) #levene's test sig

testmod <- lm(CatPC1 ~ Species, data = catpca.df2)
plot(simulateResiduals(testmod)) #levene's test sig

testmod <- lm(CatPC2 ~ Species, data = catpca.df2)
plot(simulateResiduals(testmod)) #fails KS test

#since some models fail residual checks, I will use a GLMM with a by-species dispersion factor and test mean and variance differences from that; default family=gaussian unless otherwise specified


# ───────────────────────────────────────────────────
## MEAN TESTS (GLMM w/ ~Species dispersion for all)
# ───────────────────────────────────────────────────

### Metric Data

tam.met.mean.1 <- glmmTMB(Dim.1 ~ Species, dispformula = ~Species, data = pca.df2) 
(tmm1<- Anova(tam.met.mean.1, type = "III")) #significant (p=0.011)
plot(simulateResiduals(tam.met.mean.1)) #all good
    tph1 <- emmeans(tam.met.mean.1, ~Species)
    summary(tph1)
    tpw1 <- pairs(tph1)
    (tpair1 <- as.data.frame(rbind(tpw1, adjust = "holm"))) #sail/atl(p=0.022), amz/atl(p=0.022)
    
tam.met.mean.2 <- glmmTMB(Dim.2 ~ Species, dispformula = ~Species, data = pca.df2) 
(tmm2<- Anova(tam.met.mean.2, type = "III")) #significant (p<0.001)
plot(simulateResiduals(tam.met.mean.2)) #all good
    tph2 <- emmeans(tam.met.mean.2,  ~ Species)
    summary(tph2)
    tpw2 <- pairs(tph2)
    (tpair2 <- as.data.frame(rbind(tpw2, adjust = "holm"))) #all sig (p<0.001)
    
    

### Meristic Data


tam.cnt.mean.1 <- glmmTMB(CatPC1 ~ Species, dispformula = ~Species, data = catpca.df2) 
(tcm1<- Anova(tam.cnt.mean.1, type = "III")) #significant (p<0.001)
plot(simulateResiduals(tam.cnt.mean.1)) #all good
    tph4 <- emmeans(tam.cnt.mean.1,  ~ Species)
    summary(tph4)
    tpw4 <- pairs(tph4)
    (tpair4 <- as.data.frame(rbind(tpw4, adjust = "holm"))) #all sig (p<0.001)
    


#default gaussian family failed residual checks; changing to Student t-distribution due to extreme tails
tam.cnt.mean.2 <- glmmTMB(CatPC2 ~ Species, dispformula = ~Species, family = t_family, data = catpca.df2) 
(tcm2<- Anova(tam.cnt.mean.2, type = "III")) #not significant
plot(simulateResiduals(tam.cnt.mean.2)) #failed KS test but qqplot w/ good alignment
    tph5 <- emmeans(tam.cnt.mean.2,  ~ Species)
    summary(tph5)
    tpw5 <- pairs(tph5)
    (tpair5 <- as.data.frame(rbind(tpw5, adjust = "holm"))) #for table later
    

# ───────────────────────────────────────────────────
## VARIANCE TESTS (Extract dispersion info for all)
# ───────────────────────────────────────────────────

### Metric Data
tam.met.var.1 <- emmeans(tam.met.mean.1, ~ Species, component = "disp", type = "response")
tmv1 <- pairs(tam.met.var.1)
(tpair6 <- as.data.frame(rbind(tmv1, adjust = "holm"))) #sail/amz (p<0.001, sail/atl (p=0.007)

tam.met.var.2 <- emmeans(tam.met.mean.2, ~ Species, component = "disp", type = "response")
tmv2 <- pairs(tam.met.var.2) 
(tpair7 <- as.data.frame(rbind(tmv2, adjust = "holm"))) #sail/amz (p=0.18), sail/atl (p=0.008)



### Meristic Data
tam.cnt.var.1 <- emmeans(tam.cnt.mean.1, ~ Species, component = "disp", type = "response")
tcv1 <- pairs(tam.cnt.var.1) 
(tpair9 <- as.data.frame(rbind(tcv1, adjust = "holm"))) #sail/atl (p=0.007)


tam.cnt.var.2 <- emmeans(tam.cnt.mean.2, ~ Species, component = "disp", type = "response")
tcv2 <- pairs(tam.cnt.var.2) #not significant
(tpair10 <- as.data.frame(rbind(tcv2, adjust = "holm"))) #for later table


# ───────────────────────────────────────────────────
## TABLE
# ───────────────────────────────────────────────────

### Metric data 

## ensure all results are as df
tph1 <- as.data.frame(tph1)
tph2 <- as.data.frame(tph2)
tam.met.var.1 <- as.data.frame(tam.met.var.1)
tam.met.var.2 <- as.data.frame(tam.met.var.2)


## pull species-specific model results
# Means
amazon_m1  <- sprintf("%.2f (%.2f)", tph1$emmean[1], tph1$SE[1])
atlantic_m1 <- sprintf("%.2f (%.2f)", tph1$emmean[2], tph1$SE[2])
sailfin_m1 <- sprintf("%.2f (%.2f)", tph1$emmean[3], tph1$SE[3])

amazon_m2  <- sprintf("%.2f (%.2f)", tph2$emmean[1], tph2$SE[1])
atlantic_m2 <- sprintf("%.2f (%.2f)", tph2$emmean[2], tph2$SE[2])
sailfin_m2 <- sprintf("%.2f (%.2f)", tph2$emmean[3], tph2$SE[3])


Amazon_mean <- c(amazon_m1, amazon_m2)

Atlantic_mean <- c(atlantic_m1, atlantic_m2)

Sailfin_mean <- c(sailfin_m1, sailfin_m2)

# variance
Amazon_var <- sprintf("%.2f", c(tam.met.var.1$response[1], tam.met.var.2$response[1]))

Atlantic_var <- sprintf("%.2f", c(tam.met.var.1$response[2], tam.met.var.2$response[2]))

Sailfin_var <- sprintf("%.2f", c(tam.met.var.1$response[3], tam.met.var.2$response[3]))


## pull model stats
Chisq_mean <- c(tmm1["Species", "Chisq"], tmm2["Species", "Chisq"])

p_mean <- c(tmm1["Species", "Pr(>Chisq)"], tmm2["Species", "Pr(>Chisq)"])

#need to run an anova on dispersion for 3 species
Chisq_var <- c(Anova(tam.met.mean.1, component = "disp", type = "III")["Species", "Chisq"],
               Anova(tam.met.mean.2, component = "disp", type = "III")["Species", "Chisq"])

p_var <- c(Anova(tam.met.mean.1, component = "disp", type = "III")["Species", "Pr(>Chisq)"],
           Anova(tam.met.mean.2, component = "disp", type = "III")["Species", "Pr(>Chisq)"])
  

## create tibble
tam_postPCA_result <- tibble(
  Classifier   = "Metric PCA",
  Dimension    = c("PC1", "PC2"),
  Amazon_mean  = Amazon_mean,
  Atlantic_mean  = Atlantic_mean,
  Sailfin_mean = Sailfin_mean,
  Chisq_mean   = Chisq_mean,
  p_mean       = p_mean,
  Amazon_var   = Amazon_var,
  Atlantic_var   = Atlantic_var,
  Sailfin_var  = Sailfin_var,
  Chisq_var    = Chisq_var,
  p_var        = p_var) 
  

### Meristic Data

## ensure all results are as df
tph4 <- as.data.frame(tph4)
tph5 <- as.data.frame(tph5)
tam.cnt.var.1 <- as.data.frame(tam.cnt.var.1)
tam.cnt.var.2 <- as.data.frame(tam.cnt.var.2)


## pull species-specific model results
# Means
amazon_c1  <- sprintf("%.2f (%.2f)", tph4$emmean[1], tph4$SE[1])
atlantic_c1 <- sprintf("%.2f (%.2f)", tph4$emmean[2], tph4$SE[2])
sailfin_c1 <- sprintf("%.2f (%.2f)", tph4$emmean[3], tph4$SE[3])

amazon_c2  <- sprintf("%.2f (%.2f)", tph5$emmean[1], tph5$SE[1])
atlantic_c2 <- sprintf("%.2f (%.2f)", tph5$emmean[2], tph5$SE[2])
sailfin_c2 <- sprintf("%.2f (%.2f)", tph5$emmean[3], tph5$SE[3])


Amazon_mean <- c(amazon_c1, amazon_c2)

Atlantic_mean <- c(atlantic_c1, atlantic_c2)

Sailfin_mean <- c(sailfin_c1, sailfin_c2)

# variance
Amazon_var <- sprintf("%.2f", c(tam.cnt.var.1$response[1], tam.cnt.var.2$response[1]))

Atlantic_var <- sprintf("%.2f", c(tam.cnt.var.1$response[2], tam.cnt.var.2$response[2]))

Sailfin_var <- sprintf("%.2f", c(tam.cnt.var.1$response[3], tam.cnt.var.2$response[3]))


## pull model stats
Chisq_mean <- c(tcm1["Species", "Chisq"], tcm2["Species", "Chisq"])

p_mean <- c(tcm1["Species", "Pr(>Chisq)"], tcm2["Species", "Pr(>Chisq)"])

#need to run an anova on dispersion for 3 species
Chisq_var <- c(Anova(tam.cnt.mean.1, component = "disp", type = "III")["Species", "Chisq"],
               Anova(tam.cnt.mean.2, component = "disp", type = "III")["Species", "Chisq"])

p_var <- c(Anova(tam.cnt.mean.1, component = "disp", type = "III")["Species", "Pr(>Chisq)"],
           Anova(tam.cnt.mean.2, component = "disp", type = "III")["Species", "Pr(>Chisq)"])
  

## create tibble
tam_postCPCA_result <- tibble(
  Classifier   = "Meristic CatPCA",
  Dimension    = c("CatPC1", "CatPC2"),
  Amazon_mean  = Amazon_mean,
  Atlantic_mean  = Atlantic_mean,
  Sailfin_mean = Sailfin_mean,
  Chisq_mean   = Chisq_mean,
  p_mean       = p_mean,
  Amazon_var   = Amazon_var,
  Atlantic_var   = Atlantic_var,
  Sailfin_var  = Sailfin_var,
  Chisq_var    = Chisq_var,
  p_var        = p_var) 
  


  
### Combine & form table

tam_results_combined <- bind_rows(tam_postPCA_result, tam_postCPCA_result) %>%
  group_by(Classifier) %>%
  mutate(
  p_mean_adj = p.adjust(p_mean, method = "holm"),
  p_var_adj = p.adjust(p_var, method = "holm")) %>%
  mutate(
    Mean_test = case_when(
      p_mean_adj < 0.001 ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chisq_mean), "<br>*p* < 0.001*"),
      p_mean_adj < 0.05  ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chisq_mean), "<br>*p* = ", sprintf("%.3f", p_mean_adj), "*"),
      TRUE           ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chisq_mean), "<br>*p* = ", sprintf("%.3f", p_mean_adj))),
    
    Variance_test = case_when(
      p_var_adj < 0.001 ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chisq_var), "<br>*p* < 0.001*"),
      p_var_adj < 0.05  ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chisq_var), "<br>*p* = ", sprintf("%.3f", p_var_adj), "*"),
      TRUE          ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chisq_var), "<br>*p* = ", sprintf("%.3f", p_var_adj)))) %>%
  ungroup()




tam_postPCA_table <- tam_results_combined %>%
  select(
    Classifier,
    Dimension,
    Amazon_mean,
    Atlantic_mean,
    Sailfin_mean,
    Mean_test,
    Amazon_var,
    Atlantic_var,
    Sailfin_var,
    Variance_test) %>%
  gt(groupname_col = "Classifier") %>%
  fmt_markdown(
    columns = everything()) %>%
  cols_label(
    Dimension = " ",
    Amazon_mean = md("**Amazon**"),
    Atlantic_mean = md("**Atlantic**"),
    Sailfin_mean = md("**Sailfin**"),
    Mean_test = html("Mean comparison<br>(Wald &chi;<sup>2</sup>)"),
    Amazon_var = md("**Amazon**"),
    Atlantic_var = md("**Atlantic**"),
    Sailfin_var = md("**Sailfin**"),
    Variance_test = html("Variance comparison<br>(Wald &chi;<sup>2</sup>)")) %>%
  tab_spanner(
    label = html("Est. Marginal Mean (&plusmn;SE)"),
    columns = c(Amazon_mean, Atlantic_mean, Sailfin_mean)) %>%
  tab_spanner(
    label = "Est. Variance",
    columns = c(Amazon_var, Atlantic_var, Sailfin_var)) %>%
  tab_options(row_group.font.weight = "bold") %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c(Dimension, Sailfin_mean, Mean_test, Sailfin_var))) %>%
  cols_align(align = "center", columns = everything()) %>%
  opt_horizontal_padding(scale = 1.5)

### Final table
tam_postPCA_table

```

```{r}

#additional table of just the variances

### Metric data

#grab the anova variance test again
Chisq_var <- c(Anova(tam.met.mean.1, component = "disp", type = "III")["Species", "Chisq"],
               Anova(tam.met.mean.2, component = "disp", type = "III")["Species", "Chisq"])

p_var <- c(Anova(tam.met.mean.1, component = "disp", type = "III")["Species", "Pr(>Chisq)"],
           Anova(tam.met.mean.2, component = "disp", type = "III")["Species", "Pr(>Chisq)"])

#extract the ratios and p-values after adjusting
AA_ratio <- sprintf("%.2f", c(tpair6$ratio[1], tpair7$ratio[1]))
SAm_ratio <- sprintf("%.2f", c(tpair6$ratio[2], tpair7$ratio[2]))
SAt_ratio <- sprintf("%.2f", c(tpair6$ratio[3], tpair7$ratio[3]))

AA_p_raw <- c(tpair6$p.value[1], tpair7$p.value[1])
SAm_p_raw <- c(tpair6$p.value[2], tpair7$p.value[2])
SAt_p_raw <- c(tpair6$p.value[3], tpair7$p.value[3])

tam.pca.var.table <- tibble(
  Classifier   = "Metric PCA",
  Dimension    = c("PC1", "PC2"),
  Chisq_var    = Chisq_var,
  p_var        = p_var,
  SAm_ratio    = SAm_ratio,
  SAt_ratio    = SAt_ratio,
  AA_ratio     = AA_ratio,
  SAm_p        = SAm_p_raw,
  SAt_p        = SAt_p_raw,
  AA_p         = AA_p_raw) 


### Meristic data

#grab the anova variance test again
Chisq_var <- c(Anova(tam.cnt.mean.1, component = "disp", type = "III")["Species", "Chisq"],
               Anova(tam.cnt.mean.2, component = "disp", type = "III")["Species", "Chisq"])

p_var <- c(Anova(tam.cnt.mean.1, component = "disp", type = "III")["Species", "Pr(>Chisq)"],
           Anova(tam.cnt.mean.2, component = "disp", type = "III")["Species", "Pr(>Chisq)"])

#extract the ratios and p-values after adjusting
AA_ratio <- sprintf("%.2f", c(tpair9$ratio[1], tpair10$ratio[1]))
SAm_ratio <- sprintf("%.2f", c(tpair9$ratio[2], tpair10$ratio[2]))
SAt_ratio <- sprintf("%.2f", c(tpair9$ratio[3], tpair10$ratio[3]))

AA_p_raw <- c(tpair9$p.value[1], tpair10$p.value[1])
SAm_p_raw <- c(tpair9$p.value[2], tpair10$p.value[2])
SAt_p_raw <- c(tpair9$p.value[3], tpair10$p.value[3])

tam.cpca.var.table <- tibble(
  Classifier   = "Meristic CatPCA",
  Dimension    = c("CatPC1", "CatPC2"),
  Chisq_var    = Chisq_var,
  p_var        = p_var,
  SAm_ratio    = SAm_ratio,
  SAt_ratio    = SAt_ratio,
  AA_ratio     = AA_ratio,
  SAm_p        = SAm_p_raw,
  SAt_p        = SAt_p_raw,
  AA_p         = AA_p_raw) 


### Combine & form table

var_results_combined_tam <- bind_rows(tam.pca.var.table, tam.cpca.var.table) %>%
group_by(Classifier) %>%
  mutate(
  SAm_p_adj = p.adjust(SAm_p, method = "holm"),
  SAt_p_adj = p.adjust(SAt_p, method = "holm"),
  AA_p_adj = p.adjust(AA_p, method = "holm"),
  p_var_adj = p.adjust(p_var, method = "holm")) %>%
  mutate(
    Variance_test = case_when(
      p_var_adj < 0.001 ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chisq_var), "<br>*p* < 0.001*"),
      p_var_adj < 0.05  ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chisq_var), "<br>*p* = ", sprintf("%.3f", p_var_adj), "*"),
      TRUE          ~ paste0("&chi;<sup>2</sup> = ", sprintf("%.2f", Chisq_var), "<br>*p* = ", sprintf("%.3f", p_var_adj))),
    
    Pairwise_tests = if_else(
      p_var_adj >= 0.05, 
      "&mdash;", 
      paste0(
        "Amazon/Sailfin: ", case_when(SAm_p_adj < 0.001 ~ "*p* < 0.001*", SAm_p_adj < 0.05 ~ paste0("*p* = ", sprintf("%.3f", SAm_p_adj), "*"), TRUE ~ paste0("*p* = ", sprintf("%.3f", SAm_p_adj))),
        "<br>",
        "Atlantic/Sailfin: ", case_when(SAt_p_adj < 0.001 ~ "*p* < 0.001*", SAt_p_adj < 0.05 ~ paste0("*p* = ", sprintf("%.3f", SAt_p_adj), "*"), TRUE ~ paste0("*p* = ", sprintf("%.3f", SAt_p_adj))),
        "<br>",
        "Amazon/Atlantic: ", case_when(AA_p_adj < 0.001  ~ "*p* < 0.001*", AA_p_adj < 0.05  ~ paste0("*p* = ", sprintf("%.3f", AA_p_adj), "*"),  TRUE ~ paste0("*p* = ", sprintf("%.3f", AA_p_adj)))))) %>%
  ungroup()




var_postPCA_table_tam <- var_results_combined_tam %>%
  select(
    Classifier,
    Dimension,
    SAm_ratio,
    SAt_ratio,
    AA_ratio,
    SAm_p, 
    SAt_p,
    AA_p,
    Variance_test,
    Pairwise_tests) %>%
  gt(groupname_col = "Classifier") %>%
  cols_hide(c(SAm_p, SAt_p, AA_p)) %>%
  fmt_markdown(
    columns = c(SAm_ratio, SAt_ratio, AA_ratio, Variance_test, Pairwise_tests)) %>%
  cols_label(
    Dimension = " ",
    SAm_ratio = "Amazon/Sailfin",
    SAt_ratio = "Atlantic/Sailfin",
    AA_ratio  = "Amazon/Atlantic",
    Pairwise_tests = md("Pairwise tests (adj. *p*)"),
    Variance_test = html("Variance comparison<br>(Wald &chi;<sup>2</sup>)")) %>%
  tab_spanner(
    label = "Variance Ratio",
    columns = c(SAm_ratio, SAt_ratio, AA_ratio)) %>%
  cols_move(
    columns = Variance_test,
    after = Dimension) %>%
  tab_options(row_group.font.weight = "bold") %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c(Dimension, Variance_test, AA_ratio))) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1), style = "dashed"), 
    locations = cells_body(columns = c(SAm_ratio, SAt_ratio))) %>%
  cols_align(align = "center", columns = everything()) %>%
  opt_horizontal_padding(scale = 1.5)

### Final table
var_postPCA_table_tam



```

***
***