---
title: "geomorph_final"
author: "Allison Davis Connelly & Jackson Horvath"
date: "2025-11-12"
output: 
  html_document:
    toc: TRUE
    toc_depth: 5
    toc_float:
      collapsed: FALSE
  
      
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **PURPOSE**

There are two main goals of this analysis. First, we want to compare morphological variation among the asexual Amazon molly and their two parental species (sailfin and Atlantic molly) across their native and introduced geographic range [PART 1]. Second, we want to compare the precision between the two data collection methods used (manual measurements versus digital geomorphometrics) [PART 2]. 

# **PART I: VARIATION**

The following analyses will compare variation among the species. 

## **Data**

Original TPS files were imported into MorphoJ where we performed a procrustes analysis and a multivariate regression on shape (procrustes coordinates) and size (centroid size) to correct for overall body size differences between the species. We then imported the residuals of this regression below to perform our analyses.

All analyses were conducted using R version 4.4.2


### Load required libraries

Remove the '#' in the code below if it is the first time installing library. 

```{r}

#install.packages("curl")             #v5.2.1
library(curl)

#install.packages("dplyr")            #v1.1.4
library(dplyr)

#install.packages("tidyverse")        #v2.0.0
library(tidyverse)

#install.packages("tidyr")            #v1.3.1
library(tidyr)

#install.packages("gt")               #v0.11.1
library(gt)

#install.packages("factoextra")       #v1.0.7
library(factoextra)

#install.packages("car")              #v3.1-2
library(car)

#install.packages("AMR")              #v2.1.1 DOUBLE CHECK I NEED ALL OF THE FOLLOWING 4
library(AMR)

#install.packages("ggplot2")          #v3.5.1
library(ggplot2)

#install.packages("ggfortify")        #v0.4.17
library(ggfortify)

#install.packages("ggbiplot")         #v0.6.2
library(ggbiplot)

#install.packages("gridExtra")        #v2.3
library(gridExtra)

#install.packages("blandr")           #v0.6.0
library(blandr)

#install.packages("stringr")          #v1.5.1
library(stringr)

```


### Read in data

```{r}

test.df <- curl("https://raw.githubusercontent.com/allisondavis/morphology_analysis/refs/heads/master/MV%20regression%2C%20results.txt")

test.df <- read.table(test.df, header = TRUE, sep = "\t")

# Replace HydroBASIN basin codes with basin names
test.df$Basin <- factor(test.df$Basin,
                        levels = c("7040048420", "7040049270", "7040048310", "7040049990", "7040049280"),
                        labels = c("Nueces-Rio Grande Coastal", "Rio Grande", "Brazos-Colorado Coastal", "Panuco", "San Fernando-Soto la Marina"))

# Replace default residual column names with the corresponding landmark (x,y). This will make it easier to interpret which size-corrected shape coordinates influence our results later on.

n_landmarks <- 16

coord_names <- as.vector(rbind(
  paste0("LM", 1:n_landmarks, ".X"),
  paste0("LM", 1:n_landmarks, ".Y")
))

colnames(test.df)[7:(6 + length(coord_names))] <- coord_names



```


## **Analysis**

### ***PCA***

We'll create a PCA to visualize overall shape differences. We'll then perform Levene's tests on the PC scores to assess differences in variation among the species. 

#### Loadings

```{r}

PCA.test <- prcomp(test.df[, 7:38], center = TRUE, scale. = FALSE) 
#NOTE: scale=false uses covariance matrix like in MorphoJ (what we imported); scale=true only useful if variables are measured on different scales

summary(PCA.test)
(eig.val.test <- get_eigenvalue(PCA.test))


# the first 2 dimensions account for nearly half of the variation (45.2%). We will extract the individual scores for the first 2 dimensions and attach it to the original data frame.

ind.test <- get_pca_ind(PCA.test)
head(ind.test$coord[,1:2])


test.df <- cbind(test.df, ind.test$coord[,1:2])


# Will investigate the influence of each landmark coordinates on the first 2 dimensions
(loadings <- PCA.test$rotation[, 1:2])

sorted.loadings.1 <- loadings[order(loadings[, 1]), 1]
myTitle <- "Loadings Plot for PC1" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.1, main=myTitle, xlab=myXlab, cex=0.5, col="blue")

sorted.loadings.2 <- loadings[order(loadings[, 2]), 2]
myTitle <- "Loadings Plot for PC2" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.2, main=myTitle, xlab=myXlab, cex=0.5, col="blue")


```

#### Post PCA tests

We perform a Levene's Test for each dimension to assess variation among the species. This will inform us if the spread of shape variation is different among the species. 

```{r}

pc1V <- leveneTest(Dim.1~Species, data=test.df) # not significant (p=0.708)

pc2V <- leveneTest(Dim.2~Species, data=test.df) # significant (p<0.001)


# Will display results in a table
levene_result <- tibble(
  Dimension = c("PC1", "PC2"),
  F_value = c(pc1V$`F value`[1], pc2V$`F value`[1]),
  p_value = c(pc1V$`Pr(>F)`[1], pc2V$`Pr(>F)`[1])) %>%
  mutate(
    p_display = if_else(
      p_value < 0.001,
      "*p* < 0.001^*^",
      sprintf("%.3f", p_value))
  )

levenes_PCA_results <- levene_result %>%
  gt() %>%
  cols_hide(columns = c("p_value")) %>%
  cols_label(
    Dimension = "Dimension",
    F_value = "F-value",
    p_display = "p-value") %>%
  fmt_markdown(columns = p_display) %>%
  fmt_number(columns = F_value, decimals = 3) %>%
  opt_horizontal_padding(scale=2) %>%
  cols_align(align="center", everything()) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = "p_display",
      rows = p_value < 0.05))
levenes_PCA_results




# Summary of variation for each dimension in each species
  ## Sailfin mollies have highest amount of variation in PC1 and PC2 (followed by Atlantic and then Amazon)

var.PCA.spp <- test.df %>%
  group_by(Species) %>%
  summarise(var_dim1=var(Dim.1, na.rm = TRUE),
            var_dim2=var(Dim.2, na.rm = TRUE)) %>%
  mutate(Species = case_when(
    Species == "p.latipinna" ~ "Sailfin molly",
    Species == "p.formosa" ~ "Amazon molly",
    Species == "p.mexicana" ~ "Atlantic molly")) %>%
  gt() %>% tab_header(title = "Variation of Principal Components 1 & 2") %>%
  cols_label(Species = "Species",
             var_dim1 = "PC1",
             var_dim2 = "PC2") %>%
  fmt_scientific(columns = c("var_dim1", "var_dim2"), decimals = 2) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c("Species", "var_dim1"))) %>%
  opt_horizontal_padding(scale=2) %>%
  cols_align(align="center", everything())

var.PCA.spp

gtsave(var.PCA.spp, "TB5_var-pca-spp.png")
gtsave(var.PCA.spp, "TB5_var-pca-spp.rtf")

# pairwise comparisons for significant dimensions
  ## Only significant pairwise comparison (even after Holms correction) is between Amazon's and Sailfins for both PC2 and PC3


# will create a function to group all pairwise comparisons into one table

pairwise_variance_test <- function(df, pcs = c("PC1", "PC2"), group = "Species", adjust = "holm", digits = 5) {
  
  if (!group %in% names(df)) stop(paste("Grouping variable", group, "not found in data frame."))
  
  groups <- unique(df[[group]])
  
  all_results <- list()
  
  for (pc in pcs) {
    if (!pc %in% names(df)) {
      warning(paste("Column", pc, "not found in data-skipping."))
    }
  
    pairs <- combn(groups, 2, simplify = FALSE)
    
    res_list <- lapply(pairs, function(pair) {
      sp1 <- df[[pc]][df[[group]] == pair[1]]
      sp2 <- df[[pc]][df[[group]] == pair[2]]
      
      # Look for possible errors (e.g., too few values)
      test <- tryCatch(var.test(sp1, sp2), error = function(e) NULL)
      if (is.null(test)) return(NULL)
      
      data.frame(
        PC = pc,
        group1 = pair[1],
        group2 = pair[2],
        F.value = as.numeric(test$statistic),
        p.value = test$p.value
      )
    }) 
    
    pc_results <- do.call(rbind, res_list)
    all_results[[pc]] <- pc_results
    
    results <- do.call(rbind, all_results)
  }
  
  # P-value adjustment and formatting
  results$p.adjusted <- p.adjust(results$p.value, method = adjust)
  results$p.value <- round(results$p.value, digits)
  results$p.adjusted <- round(results$p.adjusted, digits)
  results$F.value <- round(results$F.value, digits)
  
  attr(results, "adjustment_method") <- adjust
  attr(results, "grouping_variable") <- group
  
  return(results)
} # END OF FUNCTION




# run test
(PW_results <- pairwise_variance_test(test.df, pcs = c("Dim.2")))



# Results in a prettier table 
pairwise.results <- PW_results %>%
  filter(PC == "Dim.2") %>%
  arrange(PC, group1, group2) %>%
  mutate(
    group1 = dplyr::recode(group1, "p.latipinna" = "Sailfin", "p.formosa" = "Amazon", "p.mexicana" = "Atlantic"),
    group2 = dplyr::recode(group2, "p.latipinna" = "Sailfin", "p.formosa" = "Amazon", "p.mexicana" = "Atlantic"),
    PC = dplyr::recode(PC, "Dim.2" = "PC2"),
    PC = if_else(row_number() == 2, PC, ""),
    p_display = if_else(
      p.value < 0.001,
      "*p* < 0.001^*^",
      sprintf("%.3f", p.value)), 
    p_display.adj = if_else(
      p.adjusted < 0.001,
      "*p* < 0.001^*^",
      sprintf("%.3f", p.adjusted))) %>%
  ungroup() %>%
  gt() %>% 
  tab_header(title = "Pairwise Comparisons of Variation") %>%
  cols_hide(columns = c("p.value", "p.adjusted")) %>%
  cols_label(PC = "PC",
             group1 = "Species A",
             group2 = "Species B",
             F.value = "F Statistic",
             p_display = "p-value",
             p_display.adj = "Adjusted p-value") %>%
  fmt_number(columns = c(F.value), decimals = 3) %>%
  fmt_markdown(columns = c("p_display", "p_display.adj")) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c("PC", "group2", "F.value", "p_display"))) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1), style = "dashed"),
    locations = cells_body(columns = c("group1"))) %>%
  tab_style(
  style = cell_borders(sides = c("top"), color = "white", weight = px(1)),
  locations = cells_body(columns = "PC")) %>%
  opt_horizontal_padding(scale=2) %>%
  cols_align(align="center", everything()) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = c(p_display, p_display.adj),
      rows = p.value < 0.05 | p.adjusted < 0.05))
pairwise.results

gtsave(pairwise.results, "TB4_pairwise-results.png")
gtsave(pairwise.results, "TB4_pairwise-results.rtf")

```

#### *Plots*

```{r}



plot1<- autoplot(PCA.test, data = test.df, colour="Species", shape= 'Species', loadings=FALSE, loadings.label=FALSE, frame=TRUE, frame.type='norm', label= FALSE)+
  scale_color_manual(name="Species",
                    values = c("p.latipinna"="#78B7C5", "p.formosa"="#E1AF00", "p.mexicana"="#9986A5"),
                    labels= c("p.latipinna"="Sailfin molly", "p.formosa"= "Amazon molly", "p.mexicana"= "Atlantic molly")) +
  scale_fill_manual(name="Species",
                    values = c("p.latipinna"="#78B7C5", "p.formosa"="#E1AF00", "p.mexicana"="#9986A5"),
                    labels= c("p.latipinna"="Sailfin molly", "p.formosa"= "Amazon molly", "p.mexicana"= "Atlantic molly")) +
  scale_shape_manual(name= 'Species', 
                     values = c("p.latipinna" = 17, "p.formosa"=16, "p.mexicana"=3),
                     labels= c("p.latipinna"="Sailfin molly", "p.formosa"= "Amazon molly", "p.mexicana"= "Atlantic molly")) +
   guides(fill = guide_legend(override.aes = list(fill = NA)))+
  ggtitle("PCA Plot of Morphology traits") + theme_classic() 
plot1

#ggsave(plot1, filename = "Figure_4b_PCA-spp.pdf")
ggsave(plot1, filename = "Fig4b_PCA-spp.png")






```

### ***Geographic analysis***

We are also interested in how the present variation is partitioned across the sampled geographic range. We have two geographic factors to categorize samples: zones (as determined by latitudes) and river basins. While the zone categories allowed us avoid sampling bias, basins may be a better indicator of gene flow. Therfore, we will repeat the above analyses for both zone and basins.  

As our Atlantic molly (*P.mexicana*) samples are from one geographic location, we excluded them from our geographic analyses.

#### Species across Zones

```{r}

geo.df <- test.df[test.df$Species !="p.mexicana", ]

pc1V <- leveneTest(Dim.1~Species*Zone, data=geo.df) # not significant (p=0.223)

pc2V <- leveneTest(Dim.2~Species*Zone, data=geo.df) # significant (p<0.001)


# Will display results in a table

zone_result <- tibble(
  Classifier = "Zone", #used when combining tables later
  Dimension = c("PC1", "PC2"),
  F_value = c(pc1V$`F value`[1], pc2V$`F value`[1]),
  p_value = c(pc1V$`Pr(>F)`[1], pc2V$`Pr(>F)`[1])
) %>%
  mutate(p_display = if_else(
      p_value < 0.001,
      "*p* < 0.001^*^",
      sprintf("%.3f", p_value)))

zone_result %>%
  gt() %>%
  cols_hide(columns = c(Classifier, p_value)) %>%
  cols_label(
    Dimension = "Dimension",
    F_value = "F-value",
    p_display = "p-value") %>%
  fmt_markdown(columns = p_display) %>%
  fmt_number(columns = c("F_value"), decimals = 3)  %>%
  opt_horizontal_padding(scale=2) %>%
  cols_align(align="center", everything()) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = "p_display",
      rows = p_value < 0.05)
  )




# Summary of variation for each dimension in each species
  ## Sailfin mollies contain more variation than Amazon mollies for zones 2-4 in PC1, zones 1-4 in PC2, and zone 4 in PC3. 

table_zone <- geo.df %>%
  group_by(Species, Zone) %>%
  summarise(var_dim1 = var(Dim.1, na.rm = TRUE),
            var_dim2 = var(Dim.2, na.rm = TRUE),
            .groups = "drop") %>%
  pivot_wider(names_from = Species, values_from = c(var_dim1, var_dim2))

var.geo.zone <- table_zone %>%
  gt()  %>%
  cols_label(var_dim1_p.latipinna = "Sailfin", var_dim1_p.formosa = "Amazon",
             var_dim2_p.latipinna = "Sailfin", var_dim2_p.formosa = "Amazon") %>%
  tab_spanner(label = "PC1", columns = c("var_dim1_p.latipinna", "var_dim1_p.formosa")) %>%
  tab_spanner(label = "PC2", columns = c("var_dim2_p.latipinna", "var_dim2_p.formosa")) %>%
  fmt_scientific(columns = c("var_dim1_p.latipinna", "var_dim1_p.formosa", "var_dim2_p.latipinna", "var_dim2_p.formosa"), decimals = 2) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c("Zone", "var_dim1_p.formosa"))) %>%
  tab_style(style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
            locations = cells_column_labels(columns = c("var_dim1_p.formosa"))) %>% 
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_column_spanners(spanners = c("PC1"))) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1), style = "dashed"),
    locations = cells_body(columns = c("var_dim1_p.latipinna", "var_dim2_p.latipinna"))) %>%
  opt_horizontal_padding(scale=2) %>%
  cols_align(align="center", everything())

var.geo.zone

gtsave(var.geo.zone, "TB7_var-geo-zone.png")

```

##### *Plots*

```{r}

geo_long <- geo.df %>%
  pivot_longer(cols = c(Dim.1, Dim.2),
               names_to = "PC",
               values_to = "Value") %>%
  mutate(PC = factor(PC, levels = c("Dim.1", "Dim.2"),
                     labels = c("PC1", "PC2")),
         Zone = factor(Zone, levels = c("Z1","Z2","Z3","Z4"),
                      labels = c("Zone 1","Zone 2","Zone 3","Zone 4")))


zone_backgrounds <- geo_long %>%
  filter(Zone %in% c("Z2","Z4")) %>%
  distinct(PC, Zone)

zones.plot <- ggplot(geo_long, aes(x = Species, y = Value, fill = Species)) +
  geom_rect(data = zone_backgrounds,
            aes(xmin = 0.5, xmax = 2.5, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE,
            fill = "lightgray", alpha = 0.3) +
  geom_boxplot(show.legend = FALSE) +
  facet_grid(PC ~ Zone, scales = "free_y", switch = "y") +
  scale_x_discrete(labels = c("p.latipinna" = "Sailfin \nmolly",
                              "p.formosa" = "Amazon \nmolly")) +
  scale_fill_manual(values = c("p.latipinna"="#78B7C5",
                               "p.formosa"="#E1AF00")) +
  theme_classic() +
  theme(
    strip.placement = "outside",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    axis.ticks.x = element_blank(),
    panel.spacing.y = unit(1.5, "lines")) +
  xlab("") +
  ylab("")

zones.plot

ggsave(zones.plot, filename = "Fig5a_zones.png")

```


#### Species across basins

Note that Brazos-Colorado Coastal is only represented by sailfin mollies and not Amazon mollies.

```{r}
pc1V <- leveneTest(Dim.1~Species*Basin, data=geo.df) # not significant (p=0.182)

pc2V <- leveneTest(Dim.2~Species*Basin, data=geo.df) # significant (p<0.001)


# Will display results in a table

basin_result <- tibble(
  Classifier = "Basin", #used when combining tables later
  Dimension = c("PC1", "PC2"),
  F_value = c(pc1V$`F value`[1], pc2V$`F value`[1]),
  p_value = c(pc1V$`Pr(>F)`[1], pc2V$`Pr(>F)`[1])
) %>%
  mutate(p_display = if_else(
      p_value < 0.001,
      "*p* < 0.001^*^",
      sprintf("%.3f", p_value)))

basin_result %>%
  gt() %>%
  cols_hide(columns = c(Classifier, p_value)) %>%
  cols_label(
    Dimension = "Principal Component",
    F_value = "F-value",
    p_display = "p-value") %>%
  fmt_number(columns = F_value, decimals = 3) %>%
  fmt_markdown(columns = p_display)  %>%
  opt_horizontal_padding(scale=2) %>%
  cols_align(align="center", everything()) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = "p_display",
      rows = p_value < 0.05))

# Summary of variation for each dimension in each species
  ## 

table_basin <- geo.df %>%
  group_by(Species, Basin) %>%
  summarise(var_dim1 = var(Dim.1, na.rm = TRUE),
            var_dim2 = var(Dim.2, na.rm = TRUE),
            .groups = "drop") %>%
  pivot_wider(names_from = Species, values_from = c(var_dim1, var_dim2))

var.geo.basin <- table_basin %>%
  gt()  %>%
  cols_label(var_dim1_p.latipinna = "Sailfin", var_dim1_p.formosa = "Amazon",
             var_dim2_p.latipinna = "Sailfin", var_dim2_p.formosa = "Amazon") %>%
  tab_spanner(label = "PC1", columns = c(var_dim1_p.latipinna, var_dim1_p.formosa)) %>%
  tab_spanner(label = "PC2", columns = c(var_dim2_p.latipinna, var_dim2_p.formosa)) %>%
  fmt_scientific(columns = c("var_dim1_p.latipinna", "var_dim1_p.formosa","var_dim2_p.latipinna", "var_dim2_p.formosa"), decimals = 2) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c(Basin, var_dim1_p.formosa))) %>%
  tab_style(style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
            locations = cells_column_labels(columns = c("var_dim1_p.formosa"))) %>% 
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_column_spanners(spanners = c("PC1"))) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1), style = "dashed"),
    locations = cells_body(columns = c("var_dim1_p.latipinna", "var_dim2_p.latipinna"))) %>%
  opt_horizontal_padding(scale=2) %>%
  cols_align(align="center", everything())

var.geo.basin

gtsave(var.geo.basin, "TB8_var-geo-basin.png")


```

##### *Plots*

```{r}
geo_long2 <- geo_long %>%
  filter(Basin != "Brazos-Colorado Coastal") %>%
  mutate(Basin = factor(Basin,
                        levels = unique(Basin),
                        labels = c("Nueces-Rio Grande \nCoastal", "Rio Grande", "Panuco", "San Fernando-Soto \nla Marina")))

basin_backgrounds <- geo_long2 %>%
  filter(Basin %in% c("Rio Grande","San Fernando-Soto \nla Marina")) %>%
  distinct(PC, Basin)

basins.plot <- ggplot(geo_long2, aes(x = Species, y = Value, fill = Species)) +
  geom_rect(data = basin_backgrounds,
            aes(xmin = 0.5, xmax = 2.5, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE,
            fill = "lightgray", alpha = 0.3) +
  geom_boxplot(show.legend = FALSE) +
  facet_grid(PC ~ Basin, scales = "free_y", switch = "y") +
  scale_x_discrete(labels = c("p.latipinna" = "Sailfin \nmolly",
                              "p.formosa" = "Amazon \nmolly")) +
  scale_fill_manual(values = c("p.latipinna"="#78B7C5",
                               "p.formosa"="#E1AF00")) +
  theme_classic() +
  theme(
    strip.placement = "outside",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    axis.ticks.x = element_blank(),
    panel.spacing.y = unit(1.5, "lines")) +
  xlab("") +
  ylab("")

basins.plot

ggsave(basins.plot, filename = "Fig5b_basins.png")

```



## **Summary: Part I**

As expected, we see that Amazons are intermediate to both parents in overall shape. This is most evident in PC1, but in PC 2, there is a lot of overlap between the species. This is likely due to PC 2 explaining much less variation, so there is less that distinguishes the groups. 

When we perform our Levene's test, we see that there is no significant difference in variation among the species for PC1, but there is a significant difference for PC2 (p=0.0006). Landmarks 3 (x-coor), 9 (x-coor), and 10 (y-coor) all load heavily on PC2, which correspond to dorsal position and body depth. We see that sailfins contain significantly more variation than Amazons and more variation (NS) than Atlantics for PC2 (note: PC1 included in table but only PC2 had significant difference). 

It is good to remember though, that while PC 2 shows significant differences in PC shape variation,this PC explains much less variation than PC1, which shows no significant difference. 

***

Previous results are echoed in the geographic analysis, regardless of whether 'zones' or 'basins' was selected as the geographic characteristic. PC1 was not different between species across zones (p=0.223) or basins (p=0.182). PC2 was significantly different between species across both zones (p<0.001) and basins (p<0.001). For zones, sailfin mollies contained more variation than Amazons in all zones for PC2. For basins, sailfins contained more variation than Amazons in all river basins for PC2 (note: one river basin only contained sailfins). 


***
***
***

# **PART II: METHODOLOGY**

The following analyses will compare the precision of digital versus manual measurements. 

## **Data**

Manual measurement data were collected for a separate study (Connelly & Ryan, in prep). Data include 12 merestic and 10 metric traits. In order to compare to the digital measurements, only 10 metric traits from the manual measured data were included. 

Digital morpohometric data consisted of 16 landmarks and 2 curves. To obtain comparable distance measurements, we calculated the euclidean distance between the landmark coordinates that best matched the features used in manual measurement; see table below for associated landmarks and manual measurements. Digital distance is originally formatted in pixels so using the scale factor for each image in the original TPS files, we converted the distance into millimeters. 

For more details on data inclusion and distance calculation, please see Methods in the associated paper. 

```{r}
dig.meas <- data.frame(
  type= c("Standard length (SL)", "Body depth 1 (BD1)", "Body depth 2 (BD2)", "Caudal peduncle depth (CPD)", "Caudal peduncle length (CPL)", "Predorsal length (PreDL)", "Dorsal base length (DbL)", "Head length (HL)", "Head depth 1 (HD1)", "Head depth 2 (HD2)", "Snout length (SnL)", "Orbital length (OL)"),
  First.landmark = c("LM1", "LM3", "LM3", "LM5", "LM6", "LM1", "LM3", "LM1", "LM2", "LM2", "LM1", "LM15"),
  Second.landmark= c("LM6", "LM10", "LM11", "LM7", "LM8", "LM3", "LM4", "LM2", "LM11", "LM12", "LM15", "LM16"))

landmark.tbl <-dig.meas %>%
  gt() %>%
  cols_label(
    type = "Trait", 
    First.landmark = "First Landmark", 
    Second.landmark = "Second Landmark") %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_column_labels(columns = c(type, First.landmark))) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = type)) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1), style = "dashed"),
    locations = cells_body(columns = c(First.landmark))) %>%
  cols_align(
    align = "center",
    columns = everything())
landmark.tbl

gtsave(landmark.tbl, "TB3_landmark-table.png")

```


### Read in Data

```{r}


raw <- curl("https://raw.githubusercontent.com/allisondavis/morphology_analysis/refs/heads/master/comparison_data2.csv")

raw <- read.csv(raw, header = TRUE, sep = ",", stringsAsFactors = TRUE)

#AD20-084 is missing all manual values, so I will remove to avoid issues with NAs
raw<- raw[raw$ID !="AD20-082", ]

```

### Normality check

We used Shapiro-Wilk tests, histograms, and QQ plots to determine what traits were normal, and performed transformations as needed. The original raw data set was separated into the two factors of interest: digital and manual measurements.


```{r}


digital <- raw %>%
  select(ID, SPP, ends_with(".dig"))

manual <- raw %>%
  select(ID, SPP, ends_with(".man"))

```

#### *Digital* 

All but OL fail the Shapiro-Wilk test with a slight skew/deviation to the right.

```{r, fig.width=10, fig.height=25}
vars.dig <- c("SL.dig","BD1.dig", "BD2.dig","CPD.dig","CPL.dig","PreDL.dig",
              "DbL.dig","HL.dig","HD1.dig", "HD2.dig","SnL.dig","OL.dig")

##### Shapiro-Wilk test #####
shapiro.results.dig <- data.frame(
  Variable = character(),
  W = numeric(),
  p.value = numeric(),
  stringsAsFactors = FALSE)

for(v in vars.dig){
  x <- digital[[v]]
  
  
    test <- shapiro.test(x)
    shapiro.results.dig <- rbind(shapiro.results.dig,
                             data.frame(Variable = v,
                                        W = test$statistic,
                                        p.value = test$p.value))}


shapiro.results.dig


##### Histograms #####

par(mfcol = c(6, 2),   
    mar = c(4,4,2,1))

for(v in vars.dig){
  
  hist(digital[[v]],
       main = paste("Histogram of", v),
       xlab = v,
       breaks = 30,
       col = "gray80",
       border = "black")}


##### QQ plots #####

par(mfcol = c(6, 2),   
    mar = c(4,4,2,1))

for(v in vars.dig){
  x <- digital[[v]]
  
 qqnorm(x,
         main = paste("QQ Plot of", v),
         pch = 16, cex = 0.6)
  qqline(x, lwd = 1.5)}

```


#### *Manual*

All fail the SW test, with a slight skew/deviation to the right. 

```{r, fig.width=10, fig.height=25}

vars.man <- c("SL.man","BD1.man","CPD.man","CPL.man","PreDL.man",
              "DbL.man","HL.man","HD1.man","SnL.man","OL.man")

##### Shapiro-Wilk test #####
shapiro.results.man <- data.frame(
  Variable = character(),
  W = numeric(),
  p.value = numeric(),
  stringsAsFactors = FALSE)

for(v in vars.man){
  x <- manual[[v]]
  
  
    test <- shapiro.test(x)
    shapiro.results.man <- rbind(shapiro.results.man,
                             data.frame(Variable = v,
                                        W = test$statistic,
                                        p.value = test$p.value))}


shapiro.results.man


##### Histograms #####

par(mfcol = c(5, 2),   
    mar = c(4,4,2,1))

for(v in vars.man){
  
  hist(manual[[v]],
       main = paste("Histogram of", v),
       xlab = v,
       breaks = 30,
       col = "gray80",
       border = "black")}


##### QQ plots #####

par(mfcol = c(5, 2),   
    mar = c(4,4,2,1))

for(v in vars.man){
  x <- manual[[v]]
  
 qqnorm(x,
         main = paste("QQ Plot of", v),
         pch = 16, cex = 0.6)
  qqline(x, lwd = 1.5)}



```

#### Log transform & recheck

While some values are still significant, log transformations vastly improve normality (eg p=2e-16 to p=0.002). The histograms and QQ plots look close to normal, so I will stick with log transformed values. 

Orbital length (OL) was normal in the raw check of the digital measurements, but since it was not normal in manual measurements, and I want to compare the two data sets, I will transform the digital OL data.

Log transformed data sets

```{r}
dig_trans <- digital
dig_trans[, c(3:14)] <- log(dig_trans[, c(3:14)])

man_trans <- manual
man_trans[, c(3:12)] <- log(man_trans[, c(3:12)])


```

##### *Digital Recheck*

```{r, fig.width=10, fig.height=25}
##### Shapiro-Wilk test #####
shapiro.results.dig2 <- data.frame(
  Variable = character(),
  W = numeric(),
  p.value = numeric(),
  stringsAsFactors = FALSE)

for(v in vars.dig){
  x <- dig_trans[[v]]
  
  
    test <- shapiro.test(x)
    shapiro.results.dig2 <- rbind(shapiro.results.dig2,
                             data.frame(Variable = v,
                                        W = test$statistic,
                                        p.value = test$p.value))}


shapiro.results.dig2


##### Histograms #####

par(mfcol = c(6, 2),   
    mar = c(4,4,2,1))

for(v in vars.dig){
  
  hist(dig_trans[[v]],
       main = paste("Histogram of", v),
       xlab = v,
       breaks = 30,
       col = "gray80",
       border = "black")}


##### QQ plots #####

par(mfcol = c(6, 2),   
    mar = c(4,4,2,1))

for(v in vars.dig){
  x <- dig_trans[[v]]
  
 qqnorm(x,
         main = paste("QQ Plot of", v),
         pch = 16, cex = 0.6)
  qqline(x, lwd = 1.5)}


```


##### *Manual Recheck*

```{r, fig.width=10, fig.height=25}
##### Shapiro-Wilk test #####
shapiro.results.man2 <- data.frame(
  Variable = character(),
  W = numeric(),
  p.value = numeric(),
  stringsAsFactors = FALSE)

for(v in vars.man){
  x <- man_trans[[v]]
  
  
    test <- shapiro.test(x)
    shapiro.results.man2 <- rbind(shapiro.results.man2,
                             data.frame(Variable = v,
                                        W = test$statistic,
                                        p.value = test$p.value))}


shapiro.results.man2


##### Histograms #####

par(mfcol = c(5, 2),   
    mar = c(4,4,2,1))

for(v in vars.man){
  
  hist(man_trans[[v]],
       main = paste("Histogram of", v),
       xlab = v,
       breaks = 30,
       col = "gray80",
       border = "black")}


##### QQ plots #####

par(mfcol = c(5, 2),   
    mar = c(4,4,2,1))

for(v in vars.man){
  x <- man_trans[[v]]
  
 qqnorm(x,
         main = paste("QQ Plot of", v),
         pch = 16, cex = 0.6)
  qqline(x, lwd = 1.5)}



```

### Formatting 

I will perform 3 different analyses: Bland-Altman (BA), MANOVA and PCA. These vary in whether they need long or wide formats. Also, the raw data contains columns that are not needed for these analyses (eg location or photo date) so I will remove those to leave just the species, tag ID, and measurements.

Additionally, since we have two measurements for BD and HD, we will create two datasets: one with BD1 and HD1, and another with BD2 and HD2 (this applies to digital only, as manual only has one measurement for each). Once we run BA plots, we can determine which of these measurements is best to continue analysis with.


**Long format**

```{r}


#log transform entire data set

raw2 <- raw
raw2[, c(3:24)] <- log(raw2[, c(3:24)])

# this will create a data frame with species, ID, characteristic, method and value
df_long <- raw2 %>%
  pivot_longer(
    cols = ends_with(".dig") | ends_with(".man"),
    names_to = c("characteristic", "method"),
    names_sep = "\\.",
    values_to = "value"
  )

# method and characteristic are initially a character; need to be a factor
df_long$method <- as.factor(df_long$method)
df_long$characteristic <- as.factor(df_long$characteristic)


#create data frames that only have one measurement for BD and HD

## BD1 and HD1 for both digital and manual measurements 
df_long2 <- df_long[-grep("BD2", df_long$characteristic),]
df_long2 <- df_long2[-grep("HD2", df_long2$characteristic),]#only contains BD1 & HD1

## BD1/HD1 for manual (since they only have one) and BD2/HD2 for digital
df_long3 <- df_long %>%
    filter(
    # Keep everything for Hand measurements
    method == "man" |
    # Keep all digital measurements except BD1 and HD1
    !(method == "dig" & characteristic %in% c("BD1", "HD1"))
  )

#to ensure comparisons between BD2 and BD1 in this new data frame, will combine characteristic names to just BD or HD

df_long3<- df_long3 %>%
  mutate(
    characteristic = case_when(
      characteristic %in% c("BD1", "BD2") ~ "BD",
      characteristic %in% c("HD1", "HD2") ~ "HD",
      TRUE ~ characteristic  # Leave other characteristics unchanged
    )
  )


```


**Wide format**

```{r}
df_wide <- df_long2 %>%
  pivot_wider(
    names_from = method,
    values_from = value
  )


df_wide2 <- df_long3 %>%
  pivot_wider(
    names_from = method,
    values_from = value
  )

```



## **Bland-Altman Plot**

Used to assess agreement between two measurement methods.

### BD/HD plots

We're first interested in how BD1, HD1, BD2, and HD2 compare, so we isolated those.


```{r}

## Isolate data for each characteristic (BD1, BD2, HD1, HD2)
data_BD1 <- df_wide %>%
  filter(characteristic == "BD1")
data_BD1 <- data_BD1 %>%
  mutate(
    average = (dig + man) / 2,
    difference = dig - man)


data_HD1 <- df_wide %>%
  filter(characteristic == "HD1")
data_HD1 <- data_HD1 %>%
  mutate(
    average = (dig + man) / 2,
    difference = dig - man)


data_BD2 <- df_wide2 %>%
  filter(characteristic == "BD")
data_BD2 <- data_BD2 %>%
  mutate(
    average = (dig + man) / 2,
    difference = dig - man)


data_HD2 <- df_wide2 %>%
  filter(characteristic == "HD")
data_HD2 <- data_HD2 %>%
  mutate(
    average = (dig + man) / 2,
    difference = dig - man)


## Create Bland-Altman plots for each characteristic 

ba.BD1 <- blandr.draw(data_BD1$dig, data_BD1$man, ciDisplay = FALSE) +
  geom_hline(yintercept = mean(data_BD1$difference, na.rm = TRUE), color = "#b22d11", linetype = "dashed") +  # Mean line
  geom_hline(yintercept = mean(data_BD1$difference, na.rm = TRUE) + c(-1.96, 1.96) * sd(data_BD1$difference, na.rm = TRUE),
             color = "dodgerblue", linetype = "dashed") +  # Limits of agreement
  labs(title = "Body depth 1",
       x = "",
       y = "Difference (Digital - Hand)") +
  theme_minimal()
#ba.BD1


ba.BD2 <- blandr.draw(data_BD2$dig, data_BD2$man, ciDisplay = FALSE) +
  geom_hline(yintercept = mean(data_BD2$difference, na.rm = TRUE), color = "#b22d11", linetype = "dashed") +  # Mean line
  geom_hline(yintercept = mean(data_BD2$difference, na.rm = TRUE) + c(-1.96, 1.96) * sd(data_BD2$difference, na.rm = TRUE),
             color = "dodgerblue", linetype = "dashed") +  # Limits of agreement
  labs(title = "Body depth 2",
       x = "",
       y = "") +
  theme_minimal()
#ba.BD2


ba.HD1 <- blandr.draw(data_HD1$dig, data_HD1$man, ciDisplay = FALSE) +
  geom_hline(yintercept = mean(data_HD1$difference, na.rm = TRUE), color = "#b22d11", linetype = "dashed") +  # Mean line
  geom_hline(yintercept = mean(data_HD1$difference, na.rm = TRUE) + c(-1.96, 1.96) * sd(data_HD1$difference, na.rm = TRUE),
             color = "dodgerblue", linetype = "dashed") +  # Limits of agreement
  labs(title = "Head depth 1",
       x = "Means",
       y = "Differences (digital-manual)") +
  theme_minimal()
#ba.HD1


ba.HD2 <- blandr.draw(data_HD2$dig, data_HD2$man, ciDisplay = FALSE) +
  geom_hline(yintercept = mean(data_HD2$difference, na.rm = TRUE), color = "#b22d11", linetype = "dashed") +  # Mean line
  geom_hline(yintercept = mean(data_HD2$difference, na.rm = TRUE) + c(-1.96, 1.96) * sd(data_HD2$difference, na.rm = TRUE),
             color = "dodgerblue", linetype = "dashed") +  # Limits of agreement
  labs(title = "Head depth 2",
       x = "Means",
       y = "") +
  theme_minimal()
#ba.HD2


plots.BDHD <- grid.arrange(ba.BD1, ba.BD2, ba.HD1, ba.HD2)
plots.BDHD
ggsave(plots.BDHD, filename = "SFig3_BDHD-BA-plots.png")
```

We see that BD1 has less scatter, less outside the limits of agreement (blue dashed lines; LOA) and a smaller bias compared to BD2 (red dashed line; 0.12 vs 0.36 respectively). In addition, BD2's LOA don't include 0, so BD1 is more appropriate to use. In contrast, HD1, while it has less scatter with fewer points outside the LOA, it has a bias further away from 0 compared to HD2 (red dashed line; 0.06 vs -0.008 respectively), so HD2 is more appropriate to use. 

I will therefore create a final data set with BD1 and HD2 for all subsequent analyses. 

### BA plots

```{r, echo=FALSE}
###create final dataset
df_long_fin <- df_long[-grep("BD2", df_long$characteristic),] #removes BD2

df_long_fin <- df_long_fin %>%
    filter(
    # Keep everything for Hand measurements
    method == "man" |
    # Keep all digital measurements except HD1
    !(method == "dig" & characteristic %in% c("HD1"))
  )

df_long_fin<- df_long_fin %>%
  mutate(
    characteristic = case_when(
      characteristic %in% c("BD1", "BD2") ~ "BD",
      characteristic %in% c("HD1", "HD2") ~ "HD",
      TRUE ~ characteristic  # Leave other characteristics unchanged
    )
  )

df_wide_fin <- df_long_fin %>%
  pivot_wider(
    names_from = method,
    values_from = value
  )


```

#### *All traits*

We see major clusters on the BA plot for all the data. When colored by species, we see that this does not differentiate the clusters, so it is not a species difference we're seeing. When colored by characteristic, we see this nicely corresponds to the clustering. This is due to differences in range of the measurements we used (eg standard length can range from 14-40mm but snout length ranges from 1-3mm). 

```{r}

df_ba_fin <- df_wide_fin %>%
  mutate(
    average = (dig + man) / 2,
    difference = dig - man
  )

ba_full_fin <- blandr.draw(df_ba_fin$dig, df_ba_fin$man, ciDisplay = FALSE) +
  geom_hline(yintercept = mean(df_ba_fin$difference, na.rm = TRUE), color = "#b22d11", linetype = "dashed") +  # Mean line
  geom_hline(yintercept = mean(df_ba_fin$difference, na.rm = TRUE) + c(-1.96, 1.96) * sd(df_ba_fin$difference, na.rm = TRUE),
             color = "dodgerblue", linetype = "dashed") +  # Limits of agreement
  labs(title = "Bland-Altman plot for all characteristics",
       x = "Means",
       y = "Difference (Digital - Hand)") +
  theme_minimal()

ba_full_fin


##### By species

ba_full_SPP_fin <- blandr.draw(df_ba_fin$dig, df_ba_fin$man, ciDisplay = FALSE) +
  geom_point(data = df_ba_fin, aes(color=SPP)) +
  scale_color_manual(
    name = "Species",
    values=c("#E1AF00","#78B7C5", "#564670"),
    labels = c("Amazon", "Sailfin", "Atlantic")) +
  geom_hline(yintercept = mean(df_ba_fin$difference, na.rm = TRUE), color = "#b22d11", linetype = "dashed") +  # Mean line
  geom_hline(yintercept = mean(df_ba_fin$difference, na.rm = TRUE) + c(-1.96, 1.96) * sd(df_ba_fin$difference, na.rm = TRUE),
             color = "dodgerblue", linetype = "dashed") +  # Limits of agreement
  labs(title = "Bland-Altman plot for all characteristics",
       x = "Average of Digital and Hand Measurements",
       y = "Difference (Digital - Hand)") +
  theme_minimal()

ba_full_SPP_fin


##### By characteristics

ba_full_CR_fin <- blandr.draw(df_ba_fin$dig, df_ba_fin$man, ciDisplay = FALSE) +
  geom_point(data = df_ba_fin, aes(color=characteristic)) +
  scale_color_manual(
    name = "Traits",
    values= hcl.colors(10, palette = "Roma")) +
  geom_hline(yintercept = mean(df_ba_fin$difference, na.rm = TRUE), color = "#b22d11", linetype = "dashed") +  # Mean line
  geom_hline(yintercept = mean(df_ba_fin$difference, na.rm = TRUE) + c(-1.96, 1.96) * sd(df_ba_fin$difference, na.rm = TRUE),
             color = "dodgerblue", linetype = "dashed") +  # Limits of agreement
  labs(title = "Bland-Altman plot for all characteristics",
       x = "Means",
       y = "Difference (Digital - Hand)") +
  theme_minimal()

ba_full_CR_fin
ggsave(ba_full_CR_fin, filename= "Fig6a_BA-full.png")

```


#### *By trait*

```{r}

# NOTE: could not figure out how to facet wrap with blandr.draw, so created a Bland-Altman plot through ggplot instead

ba_by_char_fin <- ggplot(df_ba_fin, aes(x = average, y = difference)) +
  geom_point(color= "slategrey", alpha = 0.4) +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  geom_hline(yintercept = mean(df_ba_fin$difference, na.rm = TRUE), color = "#b22d11", linetype = "dashed") +  # Mean line
  geom_hline(yintercept = mean(df_ba_fin$difference, na.rm = TRUE) + c(-1.96, 1.96) * sd(df_ba_fin$difference, na.rm = TRUE),
             color = "dodgerblue", linetype = "dashed") +  # Limits of agreement
  labs(title = "Bland-Altman plots for each characteristic",
       x = "Means",
       y = "Difference (Digital - Hand)") +
  facet_wrap(~ characteristic, scales = "free") +
  theme_minimal()


ba_by_char_fin
ggsave(ba_by_char_fin, filename = "Fig6b_BA-by-trait.png")

```


### BA Stats

We can extract the important information (bias, LOAs, confidence intervals, and p-value after a one-sided t-test to see if bias is different from 0) using the blandr package. I will perform this on the whole data set and then by characteristic. 

```{r}



### Whole data set

(bar_full <- blandr.statistics(df_wide_fin$dig, df_wide_fin$man))

### By characteristic

ba_stats_all <- df_wide_fin %>%
  group_by(characteristic) %>%
  summarise(
    stats = list(blandr.statistics(dig, man)),
    .groups = "drop") %>%
  rowwise() %>% 
  mutate(
    t_value = stats$bias / stats$biasSEM,
    p_value = 2 * pt(-abs(stats$bias / stats$biasSEM), df = stats$no.of.observations - 1),
    bias_CI = paste0(round(stats$bias,3),
                     " (", round(stats$biasLowerCI,3),
                     ", ", round(stats$biasUpperCI,3), ")"),
    LOA_range = paste0(round(stats$lowerLOA,3), " to ", round(stats$upperLOA,3))) %>%
    mutate(p_display = case_when(
    p_value < 0.001 ~ "*p* < 0.001^*^",
    p_value < 0.05  ~ paste0(sprintf("%.3f", p_value), "^*^"),
    TRUE            ~ sprintf("%.3f", p_value))) %>%
  ungroup() 

ba.stats.char <- ba_stats_all %>%
  select(characteristic, t_value, p_value, p_display, bias_CI, LOA_range) %>%
  gt() %>%
  cols_hide(columns = p_value) %>%
  tab_header(title = "Bland–Altman Statistics by Trait") %>%
  cols_label(
    characteristic = "Trait",
    t_value = "t value",
    p_display = "p-value",
    bias_CI = "Bias (CI)",
    LOA_range = "Limit of Agreement Range") %>%
  fmt_markdown(columns = p_display) %>%
  fmt_number(columns = c(t_value), decimals = 3) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c(characteristic, t_value, p_display, bias_CI))) %>%
  cols_align(
    align = "center",
    columns = everything()) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = "p_display",
      rows = p_value < 0.05))
ba.stats.char

gtsave(ba.stats.char, "Table7_ba-stats-char.png")

```


While p-values for the whole data set and just about all of the characteristics are significant, we see that the bias itself is incredibly close to zero. The significant t-test may be a result of very little deviation overall, so any small deviation that does exist is flagged as significant. However, due to the biases being so close to zero, and the LOAs including zero, we feel that any significant difference is not methodologically relevant (ie there is not actual difference in methods).


## **MANOVA**

We will run a repeated measures MANOVA on our data to corroborate our Bland-Altman results. We will follow up our MANOVA with an effect size calculation to see if any statistical significance is methodologically relevant. 

### Data frame

First, need a data frame with two rows per individual (one for digital and one for manual), plus 10 columns corresponding to each characteristic. Will create new long/wide formats since the ones above don't have separate columns for the characteristics. 

```{r}


raw3 <- raw2 %>%
  rename(HD.dig = HD2.dig, HD.man = HD1.man)

raw3<- raw3[-5] #removes BD2
raw3<- raw3[-10] #removes HD1 for digital

df_long_ex <- raw3 %>%
  pivot_longer(cols = -c(ID, SPP), 
               names_to = "Measurement", 
               values_to = "Value") %>%
  mutate(Method = ifelse(str_detect(Measurement, ".dig"), "Digital", "Manual"),
         Character = str_remove_all(Measurement, ".dig|.man")) %>%
  select(ID, Method, Character, Value)

df_wide_ex <- df_long_ex %>%
  pivot_wider(names_from = Character, values_from = Value)


```

### Results

Similar to our BA stats, we see a significant but not practical difference between methods. The MANOVA is statistically significant (p<0.001) but have a small effect size (0.0267). This means that method only explains 2.67% of the variance in our data. 

```{r}

manova_mod <- manova(cbind(SL, BD1, CPD, CPL, PreDL, DbL, HL, HD, SnL, OL) ~ Method + Error(ID/Method), data=df_wide_ex)

summary(manova_mod, test= "Pillai")

# resutls from summary
pillai_method <- 0.90343
num_df_method <- 10
den_df_method <- 329

# Calculate partial eta-squared for Method
eta_squared_method <- pillai_method / (pillai_method + (den_df_method / num_df_method))

# Print the result
cat("Partial Eta-Squared (Method):", eta_squared_method, "\n")


```


## **PCA**

In this analysis, I will compare the principle components after centering and scaling the data. A PCA analysis will help us determine which methodology influences the variation in our data the most without worrying about differences in scales/measurements.


### Data sets

In order to create ellipses around each method, we need to create a data frame where we have a method column. 

```{r}

z.scores <- raw3

z.scores[, c(3:22)] <- scale(z.scores[, 3:22], center = TRUE, scale = TRUE)

z.long <- z.scores %>%
  pivot_longer(cols = -c(ID, SPP), 
               names_to = "Measurement", 
               values_to = "Value") %>%
  mutate(Method = ifelse(str_detect(Measurement, ".dig"), "Digital", "Manual"),
         Character = str_remove_all(Measurement, ".dig|.man")) %>%
  select(ID, Method, Character, Value)

z.wide <- z.long %>%
  pivot_wider(names_from = Character, values_from = Value)


```

### Loadings


```{r}


PCA_z <- prcomp(z.wide[, 3:12])

summary(PCA_z)

(loadings_z <- PCA_z$rotation[, 1:5])

(z.eig.val <- get_eigenvalue(PCA_z)) # We will keep the first two PCs (explains ~88% of variation)

z.ind <- get_pca_ind(PCA_z)

z.wide <- cbind(z.wide, z.ind$coord[,1:2])


z.sorted.loadings.1 <- loadings_z[order(loadings_z[, 1]), 1]
z.myTitle <- "Loadings Plot for PC1" 
z.myXlab  <- "Variable Loadings"
dotchart(z.sorted.loadings.1, main=z.myTitle, xlab=z.myXlab, cex=0.5, col="blue")

z.sorted.loadings.2 <- loadings_z[order(loadings_z[, 2]), 2]
z.myTitle <- "Loadings Plot for PC2" 
z.myXlab  <- "Variable Loadings"
dotchart(z.sorted.loadings.2, main=z.myTitle, xlab=z.myXlab, cex=0.5, col="blue")

```



### Post PCA Tests

We will now make comparisons between the two methods using both t-tests and Levene's test for variance. The t-test will tell us if there is a difference in mean shape distribution whereas the the Levene's test will tell us if there is a difference in the spread of the data.

#### *T-tests*

No significant differences.

```{r}
zw.dig <- z.wide %>%
  filter(Method == "Digital")

zw.man <- z.wide %>%
  filter(Method == "Manual")


meth.pc1 <- t.test(zw.dig$Dim.1, zw.man$Dim.1, paired= TRUE, alternative = "two.sided") # Not signficant (p=1)

meth.pc2 <- t.test(zw.dig$Dim.2, zw.man$Dim.2, paired= TRUE, alternative = "two.sided") # Not significant (p=1)


# Will display results in a table

ttest_result <- tibble(
  Classifier = "T-test", #for combining tables later
  Dimension = c("PC1", "PC2"),
  t_value = c(meth.pc1$statistic[1], meth.pc2$statistic[1]),
  p_value = c(meth.pc1$p.value[1], meth.pc2$p.value[1])
) %>%
  mutate(p_display = if_else(
      p_value < 0.001,
      "*p* < 0.001^*^",
      sprintf("%.3f", p_value)))

ttest_result %>%
  gt() %>%
  cols_hide(columns = c(Classifier, p_value)) %>%
  cols_label(
    Dimension = "Dimension",
    t_value = "t-value",
    p_display = "p-value") %>%
  fmt_number(columns = c("t_value"), decimals = 3) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Dimension, t_value)))





```

#### *Levene's tests*

No significant differences. 

```{r}

meth.pc1V <- leveneTest(Dim.1~Method, data=z.wide) # Not significant (p=0.96)
meth.pc2V <- leveneTest(Dim.2~Method, data=z.wide) # Not significant (p=0.18)

# Will display results in a table

levenes_result_meth <- tibble(
  Classifier = "Levene's Test",
  Dimension = c("PC1", "PC2"),
  F_value = c(meth.pc1V$`F value`[1], meth.pc2V$`F value`[1]),
  p_value = c(meth.pc1V$`Pr(>F)`[1], meth.pc2V$`Pr(>F)`[1])
) %>%
  mutate(p_display = if_else(
      p_value < 0.001,
      "*p* < 0.001^*^",
      sprintf("%.3f", p_value)))

levenes_result_meth %>%
  gt() %>%
  cols_hide(columns = c(Classifier, p_value)) %>%
  cols_label(
    Dimension = "Dimension",
    F_value = "F-value",
    p_display = "p-value") %>%
  fmt_number(columns = c("F_value"), decimals = 3) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Dimension, F_value)))


```

### Plots


```{r}

digVman <- autoplot(PCA_z, data = z.wide, colour="Method", shape='Method', loadings=FALSE, loadings.label=FALSE, frame=TRUE, frame.type='norm')+scale_color_manual(values=c("#E1AF00","#78B7C5"))+ scale_fill_manual(values=c("#E1AF00","#78B7C5"))+ scale_shape_manual(values = c(16,17)) + ggtitle("PCA Plot of Methods") + theme_classic() 

digVman

ggsave(digVman, filename = "Fig7_PCA-methods.png")

```


## **Summary: Part II**

Our main goal with the Part II analysis was to compare the precision of manual and digital measurements. 

We first use Bland-Altman plots which compares how closely two methods are to each other. We see that for each characteristic and for our combined data, the bias is close to zero (range: -0.1 to 0.1) with LOAs that include zero, indicating little to no difference between the two methods. Note that for biases greater than zero, this indicates digital measurements were on average larger, whereas biases less than zero indicate larger manual measurements. While the statistical tests for the Bland-Altman plots (one-sided t-test to see if bias is different than zero) indicate a significant difference between the methods, we feel that the biases and LOAs themselves indicate no practical difference. The significant results are likely due to small deviations in an otherwise consistent data set that will statistically indicate a difference, but these differences in reality are minuscule. 

Since these characteristics are not independent of each other, a more robust statistical test (than the t-tests on individual characteristics) is a repeated measures MANOVA. This allowed us to compare the manual and digital methods while treating all the characteristics as related variables. Similar to our Bland-Altman results, we see a significant difference (p<0.001). However, when we calculate the effect size, we see a small effect size (0.0267). This again illustrates that while there is a statistical difference between the two methods, this difference is not practical. Only 2.67% of the variance in our data is explained by the difference in measurement method. 

Lastly, to characterize the overall shape difference and compare variation among the methods, we performed a principal component analysis. Both paired t-tests and Levene's tests on the PC scores showed no significant difference between the two methods, corroborating our interpretation of the practical results from our previous analyses.


Overall, we conclude there is no practical difference between manual and digital measurements. 

***
***
***


# Additional Figures & Tables

Below are additional figures and tables used in the publication.

## Part 1

```{r}
# COMBINED ZONE AND BASIN LEVENE'S RESULTS

levene_combined <- bind_rows(zone_result, basin_result)

levene.geo.results <- levene_combined %>%
  gt() %>%
  cols_hide(columns = "p_value") %>%
  cols_label(
    Classifier = "Classifier",
    Dimension = "Dimension",
    F_value = "F-value",
    p_display = "p-value"
  ) %>%
  fmt_markdown(columns = p_display) %>%
  fmt_number(columns = c("F_value"), decimals = 3)  %>%
  opt_horizontal_padding(scale = 2) %>%
  cols_align(align = "center", everything()) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = "p_display",
      rows = p_value < 0.05)) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Classifier, Dimension, F_value)))

levene.geo.results

gtsave(levene.geo.results, "Table5_levene-geo-results.png")

```


```{r}
# COMBINED ZONE AND BASIN VARIATION LEVELS

table_zone <- geo.df %>%
  group_by(Species, Zone) %>%
  summarise(var_dim1 = var(Dim.1, na.rm = TRUE),
            var_dim2 = var(Dim.2, na.rm = TRUE),
            .groups = "drop") %>%
  pivot_wider(names_from = Species, values_from = c(var_dim1, var_dim2)) %>%
  rename(Classifier = Zone)

table_basin <- geo.df %>%
  group_by(Species, Basin) %>%
  summarise(var_dim1 = var(Dim.1, na.rm = TRUE),
            var_dim2 = var(Dim.2, na.rm = TRUE),
            .groups = "drop") %>%
  pivot_wider(names_from = Species, values_from = c(var_dim1, var_dim2)) %>%
  rename(Classifier = Basin)

var_combined <- bind_rows(table_zone, table_basin) %>%
  mutate(
    Classifier = case_when(
      grepl("^Z", Classifier) ~ sub("^Z", "Zone ", Classifier),
      TRUE ~ Classifier))

geo.variation <- var_combined %>%
  gt() %>%
  cols_label(
    Classifier = "Classifier",
    var_dim1_p.formosa = "Amazon",
    var_dim1_p.latipinna = "Sailfin",
    var_dim2_p.formosa = "Amazon",
    var_dim2_p.latipinna = "Sailfin"
  ) %>%
  tab_spanner(label = "PC1", columns = c("var_dim1_p.latipinna", "var_dim1_p.formosa")) %>%
  tab_spanner(label = "PC2", columns = c("var_dim2_p.latipinna", "var_dim2_p.formosa")) %>%
  fmt_scientific(columns = c("var_dim1_p.latipinna", "var_dim1_p.formosa", "var_dim2_p.latipinna", "var_dim2_p.formosa"), decimals = 2) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)), 
    locations = cells_body(columns = c("Classifier", "var_dim1_p.formosa"))) %>%
  tab_style(style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
            locations = cells_column_labels(columns = c("var_dim1_p.formosa"))) %>% 
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_column_spanners(spanners = c("PC1"))) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1), style = "dashed"),
    locations = cells_body(columns = c("var_dim1_p.latipinna", "var_dim2_p.latipinna"))) %>%
  tab_style(
  style = cell_borders(
    sides = "bottom",
    color = "darkgray",
    weight = px(2.5)),
  locations = cells_body(rows = 4)) %>%
  opt_horizontal_padding(scale=2) %>%
  cols_align(align="center", everything())


geo.variation

gtsave(geo.variation, "Table6_geo-variation.png")

```

## Part 2

```{r}
#QQPLOTS BEFORE AND AFTER TRANSFORMATION

vars.man <- c("SL.man","BD1.man","CPD.man","CPL.man","PreDL.man",
          "DbL.man","HL.man","HD1.man","SnL.man","OL.man")

tiff("QQplots-man.tiff",
     width = 2400,
     height = 6000,
     res = 600,
     compression = "lzw")

par(mfcol = c(length(vars.man), 2),
    mar = c(2,2,1,1),      # small inner margins
    oma = c(2,12,4,1))      # outer margins (left + top space)

for(i in seq_along(vars.man)){
  
  v <- vars.man[i]
  
  x1 <- manual[[v]]
  x2 <- man_trans[[v]]
  
  p1 <- shapiro.test(x1)$p.value
  p2 <- shapiro.test(x2)$p.value
  
  ## Original
  qqnorm(x1, main="", pch=16, cex=0.6)
  qqline(x1, lwd=1.5)
  
  usr <- par("usr")

x_pos <- usr[2] - 0.05 * (usr[2] - usr[1])  # 5% inward from right
y_pos <- usr[3] + 0.05 * (usr[4] - usr[3])  # 5% upward from bottom

text(x_pos, y_pos,
     labels = paste0("p = ", format.pval(p1, digits=3)),
     adj = c(1,0),
     cex = 0.8)
  
  ## Transformed
  qqnorm(x2, main="", pch=16, cex=0.6)
  qqline(x2, lwd=1.5)

  usr <- par("usr")

x_pos <- usr[2] - 0.05 * (usr[2] - usr[1])  # 5% inward from right
y_pos <- usr[3] + 0.05 * (usr[4] - usr[3])  # 5% upward from bottom

text(x_pos, y_pos,
     labels = paste0("p = ", format.pval(p1, digits=3)),
     adj = c(1,0),
     cex = 0.8)

  mtext(v,
        side=2,
        line=6,
        at = 1 - (i - 0.5)/length(vars.man),
        outer=TRUE,
        las=2,
        cex=0.9)
}
# Column headers
mtext("Original", side = 3, line = 1.5, at = 0.27, outer = TRUE, cex = 1.2)
mtext("Transformed", side = 3, line = 1.5, at = 0.73, outer = TRUE, cex = 1.2)

dev.off()

##############

vars.dig <- c("SL.dig","BD1.dig", "BD2.dig","CPD.dig","CPL.dig","PreDL.dig",
          "DbL.dig","HL.dig","HD1.dig", "HD2.dig","SnL.dig","OL.dig")

tiff("QQplots-dig.tiff",
     width = 2400,
     height = 6000,
     res = 600,
     compression = "lzw")

par(mfcol = c(length(vars.dig), 2),
    mar = c(2,2,1,1),      # small inner margins
    oma = c(2,12,4,1))      # outer margins (left + top space)

for(i in seq_along(vars.dig)){
  
  v <- vars.dig[i]
  
  x1 <- digital[[v]]
  x2 <- dig_trans[[v]]
  
  p1 <- shapiro.test(x1)$p.value
  p2 <- shapiro.test(x2)$p.value
  
  ## Original
  qqnorm(x1, main="", pch=16, cex=0.6)
  qqline(x1, lwd=1.5)
  
  usr <- par("usr")

x_pos <- usr[2] - 0.05 * (usr[2] - usr[1])  # 5% inward from right
y_pos <- usr[3] + 0.05 * (usr[4] - usr[3])  # 5% upward from bottom

text(x_pos, y_pos,
     labels = paste0("p = ", format.pval(p1, digits=3)),
     adj = c(1,0),
     cex = 0.8)
  
  ## Transformed
  qqnorm(x2, main="", pch=16, cex=0.6)
  qqline(x2, lwd=1.5)

  usr <- par("usr")

x_pos <- usr[2] - 0.05 * (usr[2] - usr[1])  # 5% inward from right
y_pos <- usr[3] + 0.05 * (usr[4] - usr[3])  # 5% upward from bottom

text(x_pos, y_pos,
     labels = paste0("p = ", format.pval(p1, digits=3)),
     adj = c(1,0),
     cex = 0.8)

  mtext(v,
        side=2,
        line=6,
        at = 1 - (i - 0.5)/length(vars.dig),
        outer=TRUE,
        las=2,
        cex=0.9)
}
# Column headers
mtext("Original", side = 3, line = 1.5, at = 0.27, outer = TRUE, cex = 1.2)
mtext("Transformed", side = 3, line = 1.5, at = 0.73, outer = TRUE, cex = 1.2)

dev.off()

```


```{r}
#POST PCA TESTS IN ONE TABLE
post.pca.combo <- bind_rows(ttest_result, levenes_result_meth) %>%
 mutate(Test_Statistic = case_when(
      Classifier == "T-test"       ~ paste0("t = ", sprintf("%.3f", t_value)),   
      Classifier == "Levene's Test" ~ paste0("F = ", sprintf("%.3f", F_value))))

post.pca.tbl <- post.pca.combo %>%
  gt() %>%
  cols_hide(columns = c("F_value", "t_value", "p_value")) %>%
  cols_move_to_start(columns = c(Classifier, Dimension, Test_Statistic)) %>%
  cols_label(
    Classifier = "Test type",
    Dimension = "Dimension",
    Test_Statistic = "Test statistic",
    p_display = "p-value") %>%
  fmt_markdown(columns = p_display) %>%
  opt_horizontal_padding(scale = 2) %>%
  cols_align(align = "center", everything()) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = "p_display",
      rows = p_value < 0.05)) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "lightgray", weight = px(1.5)),
    locations = cells_body(columns = c(Classifier, Dimension, Test_Statistic)))

post.pca.tbl

gtsave(post.pca.tbl, "Table8_post-pca-methods.png")


```
