---
title: "Student template"
author: "Allison Davis Connelly"
date: "2024-05-23"
output: 
  html_document:
    toc: TRUE
    toc_depth: 5
    toc_float:
      collapsed: FALSE
  
      
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Experiment overview

*Purpose:* Briefly outline the purpose of your experiment.

*Hypothesis:* Outline the hypotheses if applicable.

*Predictions:*
  - Prediction 1 (if applicable).
  
  - How it relates to hypothesis: describe how this prediction relates to the hypothesis

*Data:* Briefly describe the type of data collected.


***





# Data 

Since the data is in tps format, I will read it directly from my current working directory. 

For this step: geomorph() package, readland.tps()

Output should indicate that there are fixed landmarks and curve points for each specimen.
  
**Note**: initial read of tps file using readland.tps returned an error, stating that there were curves detected but uneven number for each specimen. According to <>, this indicated that there were different number of landmarks for each specimen. Upon inspection, there indeed were 3 fish with 16+ landmarks. I re-landmarked these photos and updated the tps with the new coordinates. Once this happened, it returned the correct response. 

```{r}
#load in the geomorph package

library(geomorph)

#read TPS file and assign name
##tried specifying directory but got errors saying it didn't exist and couldn't connect-->since my working directory was already the morph analysis, I just needed to specify the document from that folder, didn't need to re-specify the morph analysis folder
### use getwd() to see what directory you're working in and put the file there

gmData <- readland.tps ("ALL-tps2.TPS", specID = "ID")



```

## Adding classifiers from excel 

Online tutorials say there is a way to import an excel file for this, but the easiest way was to transpose classifiers in excel, make it a list, and then copy them here to a named list. ***This did not work for me!*** For some reason, it returned a + indicating it was looking for more instructions, but everything was closed with parentheses and quotations appropriately. Could not for the life of me figure out what the issue was (tried concatenating in excel first, but no luck).

SO I will just load in the csv of the classifiers and see if that works for further analysis...

For this step, you will need: curl package, curl(), read.csv()
  
  
  
```{r}
#individual lists not working for me (returning + but everything is closed). will import classifiers csv and see if that works for further analysis.

library(curl)
classifiers <- curl("https://raw.githubusercontent.com/allisondavis/morphology_analysis/master/classifiers2.csv")

classifiers <- read.csv(classifiers, header = TRUE, sep = ",", stringsAsFactors = TRUE)


#still have to make lat, long, basin and sub basin factors

classifiers$Latitude <- as.factor(classifiers$Latitude)
classifiers$Longitude <- as.factor(classifiers$Longitude)
classifiers$Basin <- as.factor(classifiers$Basin)
classifiers$Sub.basin <- as.factor(classifiers$Sub.basin)

```



```{r}

#name and copy classifiers from excel for each individual classifier
##guess the way I did it in excel wouldn't let me just copy the first cell and get all the values, had to copy the whole row

#ID <- factor(c("AD19-001","AD19-002","AD19-003","AD19-004","AD19-005","AD19-006","AD19-007","AD19-008","AD19-009","AD19-010","AD19-011","AD19-012","AD19-013","AD19-014","AD19-015","AD19-016","AD19-017","AD19-018","AD19-019","AD19-020","AD19-021","AD19-022","AD19-023","AD19-024","AD19-025","AD19-051","AD19-052","AD19-053","AD20-086","AD20-087","AD20-088","AD22-020","AD22-021","AD22-022","AD22-038","AD22-039","AD22-040","AD22-041","AD22-042","AD22-043","AD22-044","AD22-045","AD22-046","AD22-047","AD22-048","AD22-049","AD22-050","AD22-051","AD22-052","AD22-053","AD22-054","AD22-055","AD22-056","AD22-057","AD22-058","AD22-059","AD22-060","AD22-061","AD22-062","AD22-063","AD22-064","AD22-065","AD22-066","AD22-067","AD22-068","AD22-069","AD22-070","AD22-071","AD22-072","AD22-073","AD22-074","AD22-089","AD22-090","AD22-091","AD22-092","AD22-093","AD22-094","AD22-095","AD22-096","AD22-097","AD22-098","AD22-099","AD22-100","AD22-101","AD23-001","AD23-002","AD23-003","AD23-004","AD23-005","AD23-006","AD23-007","AD23-008","AD23-013","AD23-014","AD23-015","AD23-016","AD23-017","AD23-018","AD23-019","AD23-020","AD23-021","AD23-022","AD23-023","AD23-024","AD23-025","AD23-026","AD23-027","AD23-028","AD23-029","AD23-030","AD23-031","AD23-032","AD23-033","AD23-034","AD23-035","AD23-036","AD23-120","AD23-121","AD23-122","AD23-123","AD23-124","AD23-125","AD23-126","AD23-127","AD23-128","AD23-129","AD23-130","AD23-131","AD23-132","AD23-133","AD23-134","AD23-135","AD23-136","AD23-137","AD23-138","AD23-139","AD23-140","AD23-141","AD23-142","AD23-143","AD23-144","AD23-145","AD23-146","AD23-147","AD23-148","AD23-149","AD23-150","AD23-151","AD23-152","AD23-153","AD23-154","AD23-155","AD23-156","AD23-157","AD23-158","AD23-159","AD19-026","AD19-027","AD19-028","AD19-029","AD19-030","AD19-031","AD19-032","AD19-033","AD19-034","AD19-035","AD19-036","AD19-037","AD19-038","AD19-039","AD19-040","AD19-041","AD19-042","AD19-043","AD19-044","AD19-045","AD19-046","AD19-047","AD19-048","AD19-049","AD19-050","AD20-054","AD20-055","AD20-056","AD20-057","AD20-058","AD20-059","AD20-060","AD20-061","AD20-062","AD20-063","AD20-064","AD20-065","AD20-066","AD20-067","AD20-068","AD20-069","AD20-070","AD20-071","AD20-072","AD20-073","AD20-074","AD20-075","AD20-076","AD20-077","AD20-078","AD20-079","AD20-080","AD20-081","AD20-082","AD20-083","AD20-084","AD20-085","AD20-089","AD20-090","AD22-001","AD22-002","AD22-003","AD22-004","AD22-005","AD22-006","AD22-007","AD22-008","AD22-009","AD22-010","AD22-011","AD22-012","AD22-013","AD22-014","AD22-015","AD22-016","AD22-017","AD22-018","AD22-019","AD22-027","AD22-028","AD22-029","AD22-030","AD22-031","AD22-032","AD22-033","AD22-034","AD22-035","AD22-036","AD22-037","AD22-075","AD22-076","AD22-077","AD22-078","AD22-079","AD22-080","AD22-081","AD22-082","AD22-083","AD22-084","AD22-085","AD22-086","AD22-087","AD22-088","AD23-037","AD23-038","AD23-039","AD23-040","AD23-041","AD23-042","AD23-043","AD23-044","AD23-045","AD23-046","AD23-047","AD23-048","AD23-049","AD23-050","AD23-051","AD23-052","AD23-053","AD23-054","AD23-055","AD23-056","AD23-057","AD23-058","AD23-059","AD23-060","AD23-061","AD23-062","AD23-063","AD23-064","AD23-065","AD23-066","AD23-067","AD23-068","AD23-069","AD23-070","AD23-071","AD23-072","AD23-073","AD23-074","AD23-075","AD23-076","AD23-077","AD23-078","AD23-079","AD23-080","AD23-081","AD23-082","AD23-083","AD23-084","AD23-085","AD23-086","AD23-087","AD23-088","AD23-089","AD23-090","AD23-091","AD23-092","AD23-093","AD23-094","AD23-095","AD23-096","AD23-097","AD23-098","AD23-099","AD23-100","AD23-101","AD23-102","AD23-103","AD23-104","AD23-105","AD23-106","AD23-107","AD23-108","AD23-109","AD23-110","AD23-111","AD23-112","AD23-113","AD23-114","AD23-115","AD23-116","AD23-117","AD23-118","AD23-119","AD23-160","AD23-161","AD23-162","AD23-163","AD23-164","AD23-165","AD23-166","AD23-167","AD23-168","AD23-169","AD23-170","AD23-171","AD23-172","AD23-173","AD23-174","AD23-175","AD23-176","AD23-177","AD23-178","AD23-179","AD23-180","AD23-181","AD23-182","AD23-183","AD23-184","AD23-185","AD23-186","AD23-187","AD23-188","AD23-189","AD23-190","AD23-191","AD23-192","AD23-193","AD23-194","AD23-195" + ))


#spp <- factor(c("p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.latipinna",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.formosa",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana",	"p.mexicana"))

#latitude <- factor(c("27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.8188",	"27.6825",	"27.6825",	"27.6825",	"26.07",	"26.07",	"26.07",	"28.4966",	"28.4966",	"28.4966",	"28.88297",	"28.88297",	"28.88297",	"28.88297",	"28.88297",	"28.88297",	"28.88297",	"28.88297",	"28.88297",	"28.88297",	"28.95043",	"28.95043",	"28.95043",	"29.2182",	"29.2182",	"29.8617",	"29.8733",	"29.8733",	"29.8733",	"29.8733",	"29.8733",	"29.8733",	"29.8733",	"29.8733",	"29.8733",	"29.8733",	"29.8733",	"29.8733",	"29.8733",	"29.8733",	"29.8571",	"29.8571",	"29.8571",	"29.8571",	"29.8571",	"29.8571",	"29.8571",	"28.6566",	"28.6566",	"28.6566",	"28.6566",	"28.6566",	"28.6566",	"28.6566",	"28.6566",	"28.6566",	"28.6566",	"28.6566",	"28.6566",	"28.5023",	"28.5023",	"28.5023",	"28.5023",	"28.5023",	"28.5023",	"28.5023",	"28.5023",	"28.5023",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"27.6825",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.0223",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"26.07",	"28.4966",	"28.4966",	"28.4966",	"28.4966",	"28.4966",	"28.4966",	"28.4966",	"28.2924",	"28.2924",	"28.2924",	"28.2924",	"28.5789",	"29.8639",	"21.5",	"21.5",	"21.5",	"21.5",	"21.5",	"21.5",	"21.5",	"21.5",	"21.5",	"21.5",	"21.5",	"21.5",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"24.9666667",	"24.9666667",	"24.9666667",	"24.9666667",	"24.9666667",	"24.9666667",	"24.9666667",	"24.9666667",	"24.9666667",	"24.9666667",	"24.9666667",	"24.9666667",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"29.5904",	"29.5904",	"29.5904",	"29.5904",	"29.5904",	"29.5904",	"29.5904",	"29.5904",	"29.5904",	"29.5904",	"29.5904",	"29.5904",	"29.5904",	"29.5904",	"29.5904",	"29.5904",	"29.5904",	"29.5904",	"29.8617",	"29.8617",	"29.8617",	"29.8617",	"29.8617",	"29.8617",	"29.8639",	"29.8639",	"29.8639",	"29.8639",	"29.8639",	"29.8639",	"29.8809",	"29.8809",	"29.8809",	"29.8809",	"29.8809",	"29.8809",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3",	"22.3"))

#longitude <- factor(c("-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.0734",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.9612",	"-97.9612",	"-97.9612",	"-96.8436",	"-96.8436",	"-96.8436",	"-95.58666",	"-95.58666",	"-95.58666",	"-95.58666",	"-95.58666",	"-95.58666",	"-95.58666",	"-95.58666",	"-95.58666",	"-95.58666",	"-95.55517",	"-95.55517",	"-95.55517",	"-98.2628",	"-98.2628",	"-97.9229",	"-97.9312",	"-97.9312",	"-97.9312",	"-97.9312",	"-97.9312",	"-97.9312",	"-97.9312",	"-97.9312",	"-97.9312",	"-97.9312",	"-97.9312",	"-97.9312",	"-97.9312",	"-97.9312",	"-97.8969",	"-97.8969",	"-97.8969",	"-97.8969",	"-97.8969",	"-97.8969",	"-97.8969",	"-97.3902",	"-97.3902",	"-97.3902",	"-97.3902",	"-97.3902",	"-97.3902",	"-97.3902",	"-97.3902",	"-97.3902",	"-97.3902",	"-97.3902",	"-97.3902",	"-96.647",	"-96.647",	"-96.647",	"-96.647",	"-96.647",	"-96.647",	"-96.647",	"-96.647",	"-96.647",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.7431",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.4582",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-97.9612",	"-96.8436",	"-96.8436",	"-96.8436",	"-96.8436",	"-96.8436",	"-96.8436",	"-96.8436",	"-97.2784",	"-97.2784",	"-97.2784",	"-97.2784",	"-96.7007",	"-97.9266",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-99.7",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.585",	"-97.585",	"-97.585",	"-97.585",	"-97.585",	"-97.585",	"-97.585",	"-97.585",	"-97.585",	"-97.585",	"-97.585",	"-97.585",	"-97.585",	"-97.585",	"-97.585",	"-97.585",	"-97.585",	"-97.585",	"-97.9229",	"-97.9229",	"-97.9229",	"-97.9229",	"-97.9229",	"-97.9229",	"-97.9266",	"-97.9266",	"-97.9266",	"-97.9266",	"-97.9266",	"-97.9266",	"-97.9342",	"-97.9342",	"-97.9342",	"-97.9342",	"-97.9342",	"-97.9342",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.8666667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667",	"-97.9166667"))

#zone <- factor(c("Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z1",	"Z1",	"Z1",	"Z2",	"Z2",	"Z2",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z2",	"Z3",	"Z4",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z3",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z4",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1",	"Z1"))

#basin <- factor(c("7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040049270",	"7040049270",	"7040049270",	"7040048420",	"7040048420",	"7040048420",	"7040048310",	"7040048310",	"7040048310",	"7040048310",	"7040048310",	"7040048310",	"7040048310",	"7040048310",	"7040048310",	"7040048310",	"7040048310",	"7040048310",	"7040048310",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040049270",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040048420",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049280",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990",	"7040049990"))

#sub.basin <- factor(c("7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048870",	"7050048870",	"7050048870",	"7050049270",	"7050049270",	"7050049270",	"7050048860",	"7050048860",	"7050048860",	"7050048310",	"7050048310",	"7050048310",	"7050048310",	"7050048310",	"7050048310",	"7050048310",	"7050048310",	"7050048310",	"7050048310",	"7050048310",	"7050048310",	"7050048310",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048490",	"7050048490",	"7050048490",	"7050048490",	"7050048490",	"7050048490",	"7050048490",	"7050048490",	"7050048490",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050048870",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049130",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050049270",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050048630",	"7050824100",	"7050824100",	"7050824100",	"7050824100",	"7050824100",	"7050824100",	"7050824100",	"7050824100",	"7050824100",	"7050824100",	"7050824100",	"7050824100",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050048620",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050049280",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630",	"7050820630"))

#create a dataframe for sailfin by selecting data from your raw dataframe with P. latipinna as the species.


#create a dataframe for amazons by selecting data from your raw dataframe with P. formosa as the species.


#check the head of your new sailfin and amazon dataframes to make sure it selected correctly.



```


## Generalized Procrustes Analysis

A GPA rotates, scales, and aligns data to ensure that comparisons are not influenced by errors in photography (I think?)

For this step: gpagen(), plot()

Will perform the analysis, then plot old and new data to see the difference

```{r}
#perform GPA

gpaData <- gpagen(gmData)

#plot old data and new data
par(mfrow=c(1,2))
(old.data <- plot(gmData))
(new.data <- plot(gpaData))


```


Lastly, I will create a combined dataframe with the GPA data and the classifiers.

For this step: geomorph.data.frame()

```{r}


#gmDF <- geomorph.data.frame(gpaData, ID = classifiers$ID, spp = classifiers$Species, latitude = classifiers$Latitude, longitude = classifiers$Longitude, zone = classifiers$Zone, basin = classifiers$Basin, sub.basin = classifiers$Sub.basin)


##says that the inputs have different numbers of observations. I double checked the tps file and classifiers file to make sure they had the same individuals and updated both, but still gives the error. 

```
***



# Analysis


## PCA

Briefly describe your first step. If you are unfamiliar with the package or stat used, might be a good idea to take notes on them here as well. NEVER DELETE CODE THAT DOESN'T WORK. Always best to either describe what you did and why it didn't work and what you'll try next here, OR comment out (put a # in front of the lines of code) the code in the code chunk. That way you'll never make the same mistake twice.

For this step I will need: [list packages and functions]

The end goal of this step is: [briefly describe the comparisons you are making/the data manipulations you are looking for etc]

 

```{r}
PCA<- gm.prcomp(gpaData$coords)
summary(PCA)

(loadings <- PCA$rotation[, 1:5])

sorted.loadings.1 <- loadings[order(loadings[, 1]), 1]
myTitle <- "Loadings Plot for PC1" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.1, main=myTitle, xlab=myXlab, cex=0.5, col="slateblue4")

sorted.loadings.2 <- loadings[order(loadings[, 2]), 2]
myTitle <- "Loadings Plot for PC2" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.2, main=myTitle, xlab=myXlab, cex=0.5, col="slateblue4")

sorted.loadings.3 <- loadings[order(loadings[, 3]), 3]
myTitle <- "Loadings Plot for PC3" 
myXlab  <- "Variable Loadings"
dotchart(sorted.loadings.3, main=myTitle, xlab=myXlab, cex=0.5, col="slateblue4")


plot(PCA, col=classifiers$Species)


#P2 <- make_ggplot(P)
##this function in geomorph makes the plot useable in ggplot, but for whatever reason, when I assign the plot function above, it just produces a list not a plot???



##tried printing a prettier pca plot with ellipses but just didn't work. The codes from my other analyses doesn't work with gm.prcomp, and gives a lot of errors even when I try regular prcomp (which I'm not sure is correct anyways). Can't find instructions online for how to make ellipses using base R

#test <- prcomp(gpaData$coords)

#library(AMR) 
#library(ggplot2)
#library(ggfortify)
#library(ggbiplot)


#library(devtools)
#install_github("vqv/ggbiplot")
#library(ggbiplot)
#g <- ggbiplot(test,
              #obs.scale = 1,
              #ar.scale = 1,
              #groups = classifiers$Species,
              #ellipse = TRUE,
              #circle = TRUE,
              #ellipse.prob = 0.68)
#g <- g + scale_color_discrete(name = '')
#g <- g + theme(legend.direction = 'horizontal',
               #legend.position = 'top')
#print(g)

#PCA2<- autoplot(test, x=1, y=2, data = gpaData, colour=classifiers$Species, loadings=FALSE, loadings.label=FALSE, frame=TRUE, frame.type='norm')+scale_color_manual(values=c("#E1AF00","#78B7C5", "#9986A5"))+ scale_fill_manual(values=c("#E1AF00","#78B7C5", "#9986A5"))+  ggtitle("PCA Plot of Morphology traits") + theme_minimal() 
#PCA2

```

***

### Step 1 plots

While numbers are great, a picture is often easier to understand. Use this section to display all your plots for easy viewing. There are several options for making plots, either using base R [plot() functions] or ggplot [ggplot2 package, functions within]. Either works fine. 

  
  * Remember to add x and y lables, and a title! 


```{r}



```


If applicable, write your results for step 1 below.

<font color="deeporchid4"> ***STEP 1 RESULTS:*** </font>


***







## Step 2


Briefly describe your next step. If you are unfamiliar with the package or stat used, might be a good idea to take notes on them here as well. NEVER DELETE CODE THAT DOESN'T WORK. Always best to either describe what you did and why it didn't work and what you'll try next here, OR comment out (put a # in front of the lines of code) the code in the code chunk. That way you'll never make the same mistake twice.

For this step I will need: [list packages and functions]

The end goal of this step is: [briefly describe the comparisons you are making/the data manipulations you are looking for etc]

 

```{r}
#type in code below [this line is an example of a code comment]



```

***

### Step 2 plots

While numbers are great, a picture is often easier to understand. Use this section to display all your plots for easy viewing. There are several options for making plots, either using base R [plot() functions] or ggplot [ggplot2 package, functions within]. Either works fine. 

  
  * Remember to add x and y lables, and a title! 


```{r}



```


If applicable, write your results for step 2 below.

<font color="deeporchid4"> ***STEP 2 RESULTS:*** </font>


***







## Step 3


Briefly describe your next step. If you are unfamiliar with the package or stat used, might be a good idea to take notes on them here as well. NEVER DELETE CODE THAT DOESN'T WORK. Always best to either describe what you did and why it didn't work and what you'll try next here, OR comment out (put a # in front of the lines of code) the code in the code chunk. That way you'll never make the same mistake twice.

For this step I will need: [list packages and functions]

The end goal of this step is: [briefly describe the comparisons you are making/the data manipulations you are looking for etc]

 

```{r}
#type in code below [this line is an example of a code comment]



```

***

### Step 3a

This is an example of how to create further indented tabs under the same section. Say for this analysis you have two alternative stats (a and b) that each have two further steps to them (.1 and .2). Here is how you can tab them out in your html.

```{r}
#main code for step 3a


```


#### Step 3a.1

This is where you describe what you are doing in this sub step.

```{r}
#code for step 3a.1

```


#### Step 3a.2

Same for this sub step.

```{r}
#code for step 3a.2

```


#### Step 3b

You can go up to five # for indents, so you can alternatively do the step a or b as a title, then the sub steps as their own indent.

##### Step 3b.1

This sub step is in its own tab under the Step 3b header.

```{r}
#code for step 3b.1

```

### Step 3 plots

While numbers are great, a picture is often easier to understand. Use this section to display all your plots for easy viewing. There are several options for making plots, either using base R [plot() functions] or ggplot [ggplot2 package, functions within]. Either works fine. 

  
  * Remember to add x and y lables, and a title! 


```{r}



```


If applicable, write your results for step 3 below.

<font color="deeporchid4"> ***STEP 3 RESULTS:*** </font>


***






# Pretty Plots 

This section is for the pretty, final plots we'll use for presentations/publications. You can copy your plot code from the previous plot sections and start playing around with the aesthetics here.

It is always best to have visuals for the analyses you perform. This allows readers to visually grasp the comparisons you are making. Whenever possible, it is also a good idea to overlay the original datapoints so that readers can see how individuals change over trials or how the mean compares to the spread of the data (often this is done with faint grey points so as to not be too crowded on the graph). Lastly, it is important to label axes, add legends (if needed), add titles, and add or remove background grids to ensure the graph is easy to read and interpret. I tend to use ggplot to do all of this, but most can also be done in base R plotting. I'll leave it up to you to decide on which one to use and to google for instructions on how to change the aesthetics. What I'll list below is the plots you should have for your presentation (if you've done the plot already in summary stats, you can just copy it here).


```{r}




```


***




# Methods 

In this section you will write out your analysis methods. Write it as if it was going into a paper (no bullet points, no 'step1', etc). You can edit this section as your methods change, but again make sure to keep a record of what you did previously/what you changed in the sections above!


***










