---
title: "morphology_analysis"
author: "Allison Davis"
date: "November 21, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data collection



```{r}
library(curl)

raw <- curl("https://raw.githubusercontent.com/allisondavis/morphology_analysis/master/adavis_morphology_raw-data.csv")

raw <- read.csv(raw, header = TRUE, sep = ",", stringsAsFactors = TRUE)

head(raw)
raw

lat <- raw[raw$SPP == "p.latipinna",]
head(lat)

form <- raw[raw$SPP == "p.formosa",]
head(form)

```

## Preliminary data inspection

I will first examine the data using box plots to visually assess any major differences. Then I will run t-tests and F-tests between the species for each characteristic to see what differences in means and variences exist *prior to any correction for body size.* At the end of this section, I will present a table with all characteristics, the mean for each species, the variance for each species, and the *p* results for the t- and F-tests.


#### <font color = "deepskyblue"> BOX PLOTS</font>

```{r echo=FALSE}

library(ggplot2)

ggplot(raw, aes(x=SPP, y=D, fill=SPP)) + geom_boxplot() + xlab("Species") + ylab("Ray count\n") + ggtitle("Dorsal ray comparison\n") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")


ggplot(raw, aes(x=SPP, y=P1, fill=SPP)) + geom_boxplot() + xlab("Species") + ylab("Ray count\n") + ggtitle("Left pectoral ray comparison\n") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")


ggplot(raw, aes(x=SPP, y=P2.L)) + geom_boxplot() + xlab("Species") + ylab("Ray count") + ggtitle("Pelvic ray comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=P2.L)) + geom_boxplot() + xlab("Species") + ylab("Ray count") + ggtitle("Pelvic ray comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")


ggplot(raw, aes(x=SPP, y=A)) + geom_boxplot() + xlab("Species") + ylab("Ray count") + ggtitle("Anal ray comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=P1.R)) + geom_boxplot() + xlab("Species") + ylab("Ray count") + ggtitle("Right pectorcal ray comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=LLSC)) + geom_boxplot() + xlab("Species") + ylab("Scale count") + ggtitle("Lateral line scale comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=SALL)) + geom_boxplot() + xlab("Species") + ylab("Scale count") + ggtitle("Scales above Lateral line comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=SBLL)) + geom_boxplot() + xlab("Species") + ylab("Scale count") + ggtitle("Scales below Lateral line comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=SBDF)) + geom_boxplot() + xlab("Species") + ylab("Scale count") + ggtitle("Scales before dorsal fin comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=TL)) + geom_boxplot() + xlab("Species") + ylab("Size (mm)") + ggtitle("Total length comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=SL)) + geom_boxplot() + xlab("Species") + ylab("Size (mm)") + ggtitle("Standard length comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=BD)) + geom_boxplot() + xlab("Species") + ylab("Size (mm)") + ggtitle("Body depth comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=CPD)) + geom_boxplot() + xlab("Species") + ylab("Size (mm)") + ggtitle("Caudal peduncle comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=CPL)) + geom_boxplot() + xlab("Species") + ylab("Size (mm)") + ggtitle("Caudal peduncle length comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=PreDL)) + geom_boxplot() + xlab("Species") + ylab("Size (mm)") + ggtitle("Pre-dorsal length comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=DbL)) + geom_boxplot() + xlab("Species") + ylab("Size (mm)") + ggtitle("Dorsal base length comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=HL)) + geom_boxplot() + xlab("Species") + ylab("Size (mm)") + ggtitle("Head length comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=HD)) + geom_boxplot() + xlab("Species") + ylab("Size (mm)") + ggtitle("Head depth comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=HW)) + geom_boxplot() + xlab("Species") + ylab("Size (mm)") + ggtitle("Head width comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=SnL)) + geom_boxplot() + xlab("Species") + ylab("Size (mm)") + ggtitle("Snout length comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")

ggplot(raw, aes(x=SPP, y=OL)) + geom_boxplot() + xlab("Species") + ylab("Size (mm)") + ggtitle("Orbital length comparison") + geom_jitter(shape=16, position=position_jitter(0.2)) + theme_classic() + scale_fill_manual(values = c("#9C964A", "#FAD77B")) + theme(text = element_text(size = 16)) + theme(legend.position = "none")



```



#### <font color ="deepskyblue"> T.TESTS</font>

```{r}
#doral ray counts
t.test(lat$D, form$D, alternative = "two.sided")

#left pectoral ray counts
t.test(lat$P1, form$P1, alternative = "two.sided")

#left pelvic
t.test(lat$P2.L, form$P2.L, alternative = "two.sided")

#right pelvic
t.test(lat$P2.R, form$P2.R, alternative = "two.sided")

#anal ray counts
t.test(lat$A, form$A, alternative = "two.sided")

#right pectoral ray counts
t.test(lat$P1.R, form$P1.R, alternative = "two.sided")

#lateral line scale count
t.test(lat$LLSC, form$LLSC, alternative = "two.sided")

#scales above lateral line
t.test(lat$SALL, form$SALL, alternative = "two.sided")

#scales below lateral line
t.test(lat$SBLL, form$SBLL, alternative = "two.sided")

#scales before dorsal fin
t.test(lat$SBDF, form$SBDF, alternative = "two.sided")

#total length
t.test(lat$TL, form$TL, alternative = "two.sided")

#standard length
t.test(lat$SL, form$SL, alternative = "two.sided")

#body depth
t.test(lat$BD, form$BD, alternative = "two.sided")

#caudal peduncle depth
t.test(lat$CPD, form$CPD, alternative = "two.sided")

#caudal peduncle length
t.test(lat$CPL, form$CPL, alternative = "two.sided")

#pre dorsal length
t.test(lat$PreDL, form$PreDL, alternative = "two.sided")

#dorsal base length
t.test(lat$DbL, form$DbL, alternative = "two.sided")

#head length
t.test(lat$HL, form$HL, alternative = "two.sided")

#head depth
t.test(lat$HD, form$HD, alternative = "two.sided")

#head width
t.test(lat$HW, form$HW, alternative = "two.sided")

#snout length
t.test(lat$SnL, form$SnL, alternative = "two.sided")

#oribital length
t.test(lat$OL, form$OL, alternative = "two.sided")

```

#### <font color="deepskyblue"> VARIANCE </font>

```{r}
lat.var <- apply(lat[8:29], 2, var)
lat.var

form.var <- apply(form[8:29], 2, var)
form.var

```

#### <font color="deepskyblue"> VARIANCE TESTS </font>

```{r}
#doral ray counts
var.test(lat$D, form$D, alternative = "two.sided")

#left pectoral ray counts
var.test(lat$P1, form$P1, alternative = "two.sided")

#left pelvic
var.test(lat$P2.L, form$P2.L, alternative = "two.sided")

#right pelvic
var.test(lat$P2.R, form$P2.R, alternative = "two.sided")

#anal ray counts
var.test(lat$A, form$A, alternative = "two.sided")

#right pectoral ray counts
var.test(lat$P1.R, form$P1.R, alternative = "two.sided")

#lateral line scale count
var.test(lat$LLSC, form$LLSC, alternative = "two.sided")

#scales above lateral line
var.test(lat$SALL, form$SALL, alternative = "two.sided")

#scales below lateral line
var.test(lat$SBLL, form$SBLL, alternative = "two.sided")

#scales before dorsal fin
var.test(lat$SBDF, form$SBDF, alternative = "two.sided")

#total length
var.test(lat$TL, form$TL, alternative = "two.sided")

#standard length
var.test(lat$SL, form$SL, alternative = "two.sided")

#body depth
var.test(lat$BD, form$BD, alternative = "two.sided")

#caudal peduncle depth
var.test(lat$CPD, form$CPD, alternative = "two.sided")

#caudal peduncle length
var.test(lat$CPL, form$CPL, alternative = "two.sided")

#pre dorsal length
var.test(lat$PreDL, form$PreDL, alternative = "two.sided")

#dorsal base length
var.test(lat$DbL, form$DbL, alternative = "two.sided")

#head length
var.test(lat$HL, form$HL, alternative = "two.sided")

#head depth
var.test(lat$HD, form$HD, alternative = "two.sided")

#head width
var.test(lat$HW, form$HW, alternative = "two.sided")

#snout length
var.test(lat$SnL, form$SnL, alternative = "two.sided")

#oribital length
var.test(lat$OL, form$OL, alternative = "two.sided")

```



## Corrections for body size

The amazon mollies are on averages larger than the sailfin mollies, as shown in the t-test above. To ensure body size is not affecting our comparison of these characteristics, I will correct for body size in two main ways:
    1) I will run an ANCOVA for each characteristic using standard length (SL) as a covariate. Each model will resemble: trait ~ species * size. This will tell us if the mean differences of the traits are driven by the species or by the body length.
    2) To correct for body size with variance, I will regress each trait against SL and use the residuals in the F-test. The original residuals should express both negative and positive values. Additionally, I will run F-tests with the absolute value of the residuals and note any differences in levels of significance (i.e. if it was sig in one but not the other). *need to figure out why doing this was logical again.*
    

#### <font color="deepskyblue"> ANCOVA </font>

* first I will visualize the regression of each trait against SL for each species to check linearity assumption.

```{r echo=FALSE}

library(tidyverse)
library(ggpubr)
library(rstatix)
library(broom)

ggscatter(raw, x="SL", y="D", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="P1", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="P2.L", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="P2.R", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="A", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="P1.R", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="LLSC", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="SALL", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="SBLL", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="SBDF", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="TL", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="BD", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="CPD", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="CPL", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="PreDL", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="DbL", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="HL", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="HD", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="HW", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="SnL", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))

ggscatter(raw, x="SL", y="OL", color="SPP", add = "reg.line")+ stat_regline_equation(aes(label=paste(..eq.label.., ..rr.label.., sep = "~~~"), color=SPP))


```

* all the scatter plots were linear, although as predicted, only the size measurements (such as body depth or snout length) were affected by SL. 
  + there was a slight correlation between LLSC and SL but as this metric (along with other scale counts and ray counts) was on a discrete rather than continuous scale, this correlation is difficult to interpret.

* I will now run anova tests to check for homogeneity of regression slopes for the size metrics (assumption 2 of ANCOVA)
  + looking to see that the interaction term is not significant

```{r}

Anova(lm(data = raw, TL~SPP*SL))

Anova(lm(data = raw, BD ~ SPP * SL))

Anova(lm(data = raw, CPD ~ SPP * SL))

Anova(lm(data = raw, CPL ~ SPP * SL))

Anova(lm(data = raw, PreDL ~ SPP * SL))

Anova(lm(data = raw, DbL ~ SPP * SL))

Anova(lm(data = raw, HL ~ SPP * SL))

Anova(lm(data = raw, HD ~ SPP * SL))

Anova(lm(data = raw, HW ~ SPP * SL))

Anova(lm(data = raw, SnL ~ SPP * SL))

Anova(lm(data = raw, OL ~ SPP * SL))


```

* the only interactions that come out significant are PreDL and DbL
  + this makes sense that there's an interaction between SPP and SL, since this is a diagnostic characteristic for the species
  + ***DO I REMOVE THIS FROM FURTHER ANALYSES?***
  
### <font color = "darkorchid4"> ***Unable to perform ANCOVA using the method from https://www.datanovia.com/en/lessons/ancova-in-r/. Not sure what's going on with the data.***</font>

#### <font color = "deepskyblue"> Regression analysis</font>

I will now use residuals to account for body size in each trait, and compare the variance for each species. Again, the first round will take the raw residuals, and the second round with use the absolute value of the raw residuals. 


* STEP ONE: plot trait vs body size

```{r echo=FALSE}
reg.lat.D <- lm(lat$D ~ lat$SL)

reg.lat.P1 <- lm(lat$P1 ~ lat$SL)

reg.lat.P2.L <- lm(lat$P2.L ~ lat$SL)

reg.lat.P2.R <- lm(lat$P2.R ~ lat$SL)

reg.lat.A <- lm(lat$A ~ lat$SL)

reg.lat.P1.R <- lm(lat$P1.R ~ lat$SL)

reg.lat.LLSC <- lm(lat$LLSC ~ lat$SL)

reg.lat.SALL <- lm(lat$SALL ~ lat$SL)

reg.lat.SBLL <- lm(lat$SBLL ~ lat$SL)

reg.lat.SBDF <- lm(lat$SBDF ~ lat$SL)

reg.lat.TL <- lm(lat$TL ~ lat$SL)

reg.lat.BD <- lm(lat$BD ~ lat$SL)

reg.lat.CPD <- lm(lat$CPD ~ lat$SL)

reg.lat.CPL <- lm(lat$CPL ~ lat$SL)

reg.lat.PreDL <- lm(lat$PreDL ~ lat$SL)

reg.lat.DbL <- lm(lat$DbL ~ lat$SL)

reg.lat.HL <- lm(lat$HL ~ lat$SL)

reg.lat.HD <- lm(lat$HD ~ lat$SL)

reg.lat.HW <- lm(lat$HW ~ lat$SL)

reg.lat.SnL <- lm(lat$SnL ~ lat$SL)

reg.lat.OL <- lm(lat$OL ~ lat$SL)



reg.form.D <- lm(form$D ~ form$SL)

reg.form.P1 <- lm(form$P1 ~ form$SL)

reg.form.P2.L <- lm(form$P2.L ~ form$SL)

reg.form.P2.R <- lm(form$P2.R ~ form$SL)

reg.form.A <- lm(form$A ~ form$SL)

reg.form.P1.R <- lm(form$P1.R ~ form$SL)

reg.form.LLSC <- lm(form$LLSC ~ form$SL)

reg.form.SALL <- lm(form$SALL ~ form$SL)

reg.form.SBLL <- lm(form$SBLL ~ form$SL)

reg.form.SBDF <- lm(form$SBDF ~ form$SL)

reg.form.TL <- lm(form$TL ~ form$SL)

reg.form.BD <- lm(form$BD ~ form$SL)

reg.form.CPD <- lm(form$CPD ~ form$SL)

reg.form.CPL <- lm(form$CPL ~ form$SL)

reg.form.PreDL <- lm(form$PreDL ~ form$SL)

reg.form.DbL <- lm(form$DbL ~ form$SL)

reg.form.HL <- lm(form$HL ~ form$SL)

reg.form.HD <- lm(form$HD ~ form$SL)

reg.form.HW <- lm(form$HW ~ form$SL)

reg.form.SnL <- lm(form$SnL ~ form$SL)

reg.form.OL <- lm(form$OL ~ form$SL)


```

* STEP TWO: get residuals for each individual for that trait

```{r echo=FALSE}
res.lat.D <- resid(reg.lat.D)

res.lat.P1 <- resid(reg.lat.P1)

res.lat.P2.L <- resid(reg.lat.P2.L)

res.lat.P2.R <- resid(reg.lat.P2.R)

res.lat.A <- resid(reg.lat.A)

res.lat.P1.R <- resid(reg.lat.P1.R)

res.lat.LLSC <- resid(reg.lat.LLSC)

res.lat.SALL <- resid(reg.lat.SALL)

res.lat.SBLL <- resid(reg.lat.SBLL)

res.lat.SBDF <- resid(reg.lat.SBDF)

res.lat.TL <- resid(reg.lat.TL)

res.lat.BD <- resid(reg.lat.BD)

res.lat.CPD <- resid(reg.lat.CPD)

res.lat.CPL <- resid(reg.lat.CPL)

res.lat.PreDL <- resid(reg.lat.PreDL)

res.lat.DbL <- resid(reg.lat.DbL)

res.lat.HL <- resid(reg.lat.HL)

res.lat.HD <- resid(reg.lat.HD)

res.lat.HW <- resid(reg.lat.HW)

res.lat.SnL <- resid(reg.lat.SnL)

res.lat.OL <- resid(reg.lat.OL)


res.form.D <- resid(reg.form.D)

res.form.P1 <- resid(reg.form.P1)

res.form.P2.L <- resid(reg.form.P2.L)

res.form.P2.R <- resid(reg.form.P2.R)

res.form.A <- resid(reg.form.A)

res.form.P1.R <- resid(reg.form.P1.R)

res.form.LLSC <- resid(reg.form.LLSC)

res.form.SALL <- resid(reg.form.SALL)

res.form.SBLL <- resid(reg.form.SBLL)

res.form.SBDF <- resid(reg.form.SBDF)

res.form.TL <- resid(reg.form.TL)

res.form.BD <- resid(reg.form.BD)

res.form.CPD <- resid(reg.form.CPD)

res.form.CPL <- resid(reg.form.CPL)

res.form.PreDL <- resid(reg.form.PreDL)

res.form.DbL <- resid(reg.form.DbL)

res.form.HL <- resid(reg.form.HL)

res.form.HD <- resid(reg.form.HD)

res.form.HW <- resid(reg.form.HW)

res.form.SnL <- resid(reg.form.SnL)

res.form.OL <- resid(reg.form.OL)

```


* STEP THREE: convert residuals to absolute value

```{r echo=FALSE}
abs.lat.D <- abs(res.lat.D)

abs.lat.P1 <- abs(res.lat.P1)

abs.lat.P2.L <- abs(res.lat.P2.L)

abs.lat.P2.R <- abs(res.lat.P2.R)

abs.lat.A <- abs(res.lat.A)

abs.lat.P1.R <- abs(res.lat.P1.R)

abs.lat.LLSC <- abs(res.lat.LLSC)

abs.lat.SALL <- abs(res.lat.SALL)

abs.lat.SBLL <- abs(res.lat.SBLL)

abs.lat.SBDF <- abs(res.lat.SBDF)

abs.lat.TL <- abs(res.lat.TL)

abs.lat.BD <- abs(res.lat.BD)

abs.lat.CPD <- abs(res.lat.CPD)

abs.lat.CPL <- abs(res.lat.CPL)

abs.lat.PreDL <- abs(res.lat.PreDL)

abs.lat.DbL <- abs(res.lat.DbL)

abs.lat.HL <- abs(res.lat.HL)

abs.lat.HD <- abs(res.lat.HD)

abs.lat.HW <- abs(res.lat.HW)

abs.lat.SnL <- abs(res.lat.SnL)

abs.lat.OL <- abs(res.lat.OL)


abs.form.D <- abs(res.form.D)

abs.form.P1 <- abs(res.form.P1)

abs.form.P2.L <- abs(res.form.P2.L)

abs.form.P2.R <- abs(res.form.P2.R)

abs.form.A <- abs(res.form.A)

abs.form.P1.R <- abs(res.form.P1.R)

abs.form.LLSC <- abs(res.form.LLSC)

abs.form.SALL <- abs(res.form.SALL)

abs.form.SBLL <- abs(res.form.SBLL)

abs.form.SBDF <- abs(res.form.SBDF)

abs.form.TL <- abs(res.form.TL)

abs.form.BD <- abs(res.form.BD)

abs.form.CPD <- abs(res.form.CPD)

abs.form.CPL <- abs(res.form.CPL)

abs.form.PreDL <- abs(res.form.PreDL)

abs.form.DbL <- abs(res.form.DbL)

abs.form.HL <- abs(res.form.HL)

abs.form.HD <- abs(res.form.HD)

abs.form.HW <- abs(res.form.HW)

abs.form.SnL <- abs(res.form.SnL)

abs.form.OL <- abs(res.form.OL)


```

* STEP 4a: perform F tests to compare species (raw)
    + just looked at greater (where lat is more variable than form)

    + first, lat>form, then form>lat
    
```{r echo=FALSE}
#lat>form

var.res.D <- var.test(res.lat.D, res.form.D, alternative = "greater")
var.res.D

var.res.P1 <- var.test(res.lat.P1, res.form.P1, alternative = "greater")
var.res.P1

var.res.P2.L <- var.test(res.lat.P2.L, res.form.P2.L, alternative = "greater")
var.res.P2.L

var.res.P2.R <- var.test(res.lat.P2.R, res.form.P2.R, alternative = "greater")
var.res.P2.R

var.res.A <- var.test(res.lat.A, res.form.A, alternative = "greater")
var.res.A

var.res.P1.R <- var.test(res.lat.P1.R, res.form.P1.R, alternative = "greater")
var.res.P1.R

var.res.LLSC <- var.test(res.lat.LLSC, res.form.LLSC, alternative = "greater")
var.res.LLSC

var.res.SALL <- var.test(res.lat.SALL, res.form.SALL, alternative = "greater")
var.res.SALL

var.res.SBLL <- var.test(res.lat.SBLL, res.form.SBLL, alternative = "greater")
var.res.SBLL

var.res.SBDF <- var.test(res.lat.SBDF, res.form.SBDF, alternative = "greater")
var.res.SBDF

var.res.TL <- var.test(res.lat.TL, res.form.TL, alternative = "greater")
var.res.TL

var.res.BD <- var.test(res.lat.BD, res.form.BD, alternative = "greater")
var.res.BD

var.res.CPD <- var.test(res.lat.CPD, res.form.CPD, alternative = "greater")
var.res.CPD

var.res.CPL <- var.test(res.lat.CPL, res.form.CPL, alternative = "greater")
var.res.CPL

var.res.PreDL <- var.test(res.lat.PreDL, res.form.PreDL, alternative = "greater")
var.res.PreDL

var.res.DbL <- var.test(res.lat.DbL, res.form.DbL, alternative = "greater")
var.res.DbL

var.res.HL <- var.test(res.lat.HL, res.form.HL, alternative = "greater")
var.res.HL

var.res.HD <- var.test(res.lat.HD, res.form.HD, alternative = "greater")
var.res.HD

var.res.HW <- var.test(res.lat.HW, res.form.HW, alternative = "greater")
var.res.HW

var.res.SnL <- var.test(res.lat.SnL, res.form.SnL, alternative = "greater")
var.res.SnL

var.res.OL <- var.test(res.lat.OL, res.form.OL, alternative = "greater")
var.res.OL



#form>lat

var.res.D <- var.test(res.lat.D, res.form.D, alternative = "less")
var.res.D

var.res.P1 <- var.test(res.lat.P1, res.form.P1, alternative = "less")
var.res.P1

var.res.P2.L <- var.test(res.lat.P2.L, res.form.P2.L, alternative = "less")
var.res.P2.L

var.res.P2.R <- var.test(res.lat.P2.R, res.form.P2.R, alternative = "less")
var.res.P2.R

var.res.A <- var.test(res.lat.A, res.form.A, alternative = "less")
var.res.A

var.res.P1.R <- var.test(res.lat.P1.R, res.form.P1.R, alternative = "less")
var.res.P1.R

var.res.LLSC <- var.test(res.lat.LLSC, res.form.LLSC, alternative = "less")
var.res.LLSC

var.res.SALL <- var.test(res.lat.SALL, res.form.SALL, alternative = "less")
var.res.SALL

var.res.SBLL <- var.test(res.lat.SBLL, res.form.SBLL, alternative = "less")
var.res.SBLL

var.res.SBDF <- var.test(res.lat.SBDF, res.form.SBDF, alternative = "less")
var.res.SBDF

var.res.TL <- var.test(res.lat.TL, res.form.TL, alternative = "less")
var.res.TL

var.res.BD <- var.test(res.lat.BD, res.form.BD, alternative = "less")
var.res.BD

var.res.CPD <- var.test(res.lat.CPD, res.form.CPD, alternative = "less")
var.res.CPD

var.res.CPL <- var.test(res.lat.CPL, res.form.CPL, alternative = "less")
var.res.CPL

var.res.PreDL <- var.test(res.lat.PreDL, res.form.PreDL, alternative = "less")
var.res.PreDL

var.res.DbL <- var.test(res.lat.DbL, res.form.DbL, alternative = "less")
var.res.DbL

var.res.HL <- var.test(res.lat.HL, res.form.HL, alternative = "less")
var.res.HL

var.res.HD <- var.test(res.lat.HD, res.form.HD, alternative = "less")
var.res.HD

var.res.HW <- var.test(res.lat.HW, res.form.HW, alternative = "less")
var.res.HW

var.res.SnL <- var.test(res.lat.SnL, res.form.SnL, alternative = "less")
var.res.SnL

var.res.OL <- var.test(res.lat.OL, res.form.OL, alternative = "less")
var.res.OL


```


* STEP 4b: will compare variances with absolute value of residuals and mark which ones produce an alternative p-value
    + first lat>form, then form>lat

```{r echo=FALSE}

#lat>form


var.abs.D <- var.test(abs.lat.D, abs.form.D, alternative = "greater")
var.abs.D

var.abs.P1 <- var.test(abs.lat.P1, abs.form.P1, alternative = "greater")
var.abs.P1

var.abs.P2.L <- var.test(abs.lat.P2.L, abs.form.P2.L, alternative = "greater")
var.abs.P2.L

var.abs.P2.R <- var.test(abs.lat.P2.R, abs.form.P2.R, alternative = "greater")
var.abs.P2.R

var.abs.A <- var.test(abs.lat.A, abs.form.A, alternative = "greater")
var.abs.A

var.abs.P1.R <- var.test(abs.lat.P1.R, abs.form.P1.R, alternative = "greater")
var.abs.P1.R

var.abs.LLSC <- var.test(abs.lat.LLSC, abs.form.LLSC, alternative = "greater")
var.abs.LLSC

var.abs.SALL <- var.test(abs.lat.SALL, abs.form.SALL, alternative = "greater")
var.abs.SALL

var.abs.SBLL <- var.test(abs.lat.SBLL, abs.form.SBLL, alternative = "greater")
var.abs.SBLL

var.abs.SBDF <- var.test(abs.lat.SBDF, abs.form.SBDF, alternative = "greater")
var.abs.SBDF

var.abs.TL <- var.test(abs.lat.TL, abs.form.TL, alternative = "greater")
var.abs.TL

var.abs.BD <- var.test(abs.lat.BD, abs.form.BD, alternative = "greater")
var.abs.BD

var.abs.CPD <- var.test(abs.lat.CPD, abs.form.CPD, alternative = "greater")
var.abs.CPD

var.abs.CPL <- var.test(abs.lat.CPL, abs.form.CPL, alternative = "greater")
var.abs.CPL

var.abs.PreDL <- var.test(abs.lat.PreDL, abs.form.PreDL, alternative = "greater")
var.abs.PreDL

var.abs.DbL <- var.test(abs.lat.DbL, abs.form.DbL, alternative = "greater")
var.abs.DbL

var.abs.HL <- var.test(abs.lat.HL, abs.form.HL, alternative = "greater")
var.abs.HL

var.abs.HD <- var.test(abs.lat.HD, abs.form.HD, alternative = "greater")
var.abs.HD

var.abs.HW <- var.test(abs.lat.HW, abs.form.HW, alternative = "greater")
var.abs.HW

var.abs.SnL <- var.test(abs.lat.SnL, abs.form.SnL, alternative = "greater")
var.abs.SnL

var.abs.OL <- var.test(abs.lat.OL, abs.form.OL, alternative = "greater")
var.abs.OL



#form>lat

var.abs.D <- var.test(abs.lat.D, abs.form.D, alternative = "less")
var.abs.D

var.abs.P1 <- var.test(abs.lat.P1, abs.form.P1, alternative = "less")
var.abs.P1

var.abs.P2.L <- var.test(abs.lat.P2.L, abs.form.P2.L, alternative = "less")
var.abs.P2.L

var.abs.P2.R <- var.test(abs.lat.P2.R, abs.form.P2.R, alternative = "less")
var.abs.P2.R

var.abs.A <- var.test(abs.lat.A, abs.form.A, alternative = "less")
var.abs.A

var.abs.P1.R <- var.test(abs.lat.P1.R, abs.form.P1.R, alternative = "less")
var.abs.P1.R

var.abs.LLSC <- var.test(abs.lat.LLSC, abs.form.LLSC, alternative = "less")
var.abs.LLSC

var.abs.SALL <- var.test(abs.lat.SALL, abs.form.SALL, alternative = "less")
var.abs.SALL

var.abs.SBLL <- var.test(abs.lat.SBLL, abs.form.SBLL, alternative = "less")
var.abs.SBLL

var.abs.SBDF <- var.test(abs.lat.SBDF, abs.form.SBDF, alternative = "less")
var.abs.SBDF

var.abs.TL <- var.test(abs.lat.TL, abs.form.TL, alternative = "less")
var.abs.TL

var.abs.BD <- var.test(abs.lat.BD, abs.form.BD, alternative = "less")
var.abs.BD

var.abs.CPD <- var.test(abs.lat.CPD, abs.form.CPD, alternative = "less")
var.abs.CPD

var.abs.CPL <- var.test(abs.lat.CPL, abs.form.CPL, alternative = "less")
var.abs.CPL

var.abs.PreDL <- var.test(abs.lat.PreDL, abs.form.PreDL, alternative = "less")
var.abs.PreDL

var.abs.DbL <- var.test(abs.lat.DbL, abs.form.DbL, alternative = "less")
var.abs.DbL

var.abs.HL <- var.test(abs.lat.HL, abs.form.HL, alternative = "less")
var.abs.HL

var.abs.HD <- var.test(abs.lat.HD, abs.form.HD, alternative = "less")
var.abs.HD

var.abs.HW <- var.test(abs.lat.HW, abs.form.HW, alternative = "less")
var.abs.HW

var.abs.SnL <- var.test(abs.lat.SnL, abs.form.SnL, alternative = "less")
var.abs.SnL

var.abs.OL <- var.test(abs.lat.OL, abs.form.OL, alternative = "less")
var.abs.OL



```

* <font color="darkorchid4"> The only p-value that changes direction after the absolute value transformation is anal ray count (A)</font>



 





