---
title: "morphology-analysis_final"
author: "Allison Davis Connelly"
date: "9/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## *Morphological variation of amazon and sailfin mollies* {.tabset .tabset-fade .tabset-pills}

In this study, I am looking at various populations of amazon and sailfin mollies across their native/introduced range to assess morphological variation both within and among the species. I am using the Pickle fish collections for my samples.

### Data collection {.tabset .tabset-fade .tabset-pills}

```{r, echo=FALSE}
library(curl)

raw <- curl("https://raw.githubusercontent.com/allisondavis/morphology_analysis/master/adavis_morphology_raw-data.csv")

raw <- read.csv(raw, header = TRUE, sep = ",", stringsAsFactors = TRUE)

head(raw)


lat <- raw[raw$SPP == "p.latipinna",]
head(lat)

form <- raw[raw$SPP == "p.formosa",]
head(form)

```

### PCA analysis {.tabset .tabset-fade .tabset-pills}

In this analysis, I will compare the principle components after centering and scaling the data. A PCA analysis will help us determine what aspects of morphology influence the variation in our data the most without worrying about differences in scales/measurements. Currently, data consists of **30 Sailfin and 71 Amazon**. (31 sailfin and 75 Amazon including NAs)

***Will use the original raw data set rather than any outlier-removed data sets.***

*Variable chart:*

  - D: dorsal ray count
  - P1: left pectoral ray count
  - P2.L: left pelvic rays
  - P2.R: right pelvic rays
  - A: anal rays
  - P1.R: right pectoral ray count
  - LLSC: lateral line scale count
  - SALL: scales above lateral line
  - SBLL: scales below lateral line
  - SBDF: scales before dorsal fin
  - TL: total length
  - SL: standard length
  - BD: body depth
  - CPD: caudal peduncle depth
  - CPL: caudal peduncle length
  - PreDL: pre-dorsal length
  - DbL: dorsal base length
  - HL/HW/HD: head length/width/depth
  - SnL: snout length
  - OL: ocular length

-   summary of PCs

```{r, echo=FALSE}


# error when running PCA on raw data due to the two NA values; will remove these rows.

raw1 <- raw[complete.cases(raw[, 8:29]),]



#excluded ray and scale counts for whatever reason...PCA1 <- prcomp(raw1[, 18:29], center=TRUE, scale. = TRUE)
#summary(PCA1)
#PCA1$rotation

PCA2 <- prcomp(raw1[, 8:29], center=TRUE, scale. = TRUE)
summary(PCA2)
loadings <- PCA2$rotation
loadings[, 1:5]

```

-   visualization of PCA

```{r, echo=FALSE}

library(AMR)
library(ggplot2)
raw.spp <- c(rep("Sailfin", 30), rep("Amazon", 71))


plot1c<- ggplot_pca(PCA2) + ggtitle("Plot 1") + theme_minimal()
plot1c

plot2c <- ggplot_pca(PCA2, ellipse = TRUE, color= raw.spp, groups = raw.spp) + ggtitle("Plot 2") + theme_minimal()
plot2c

library(ggfortify)

plot6<- autoplot(PCA2, data = raw1, colour='SPP', loadings=TRUE, loadings.colour='navyblue', loadings.label=TRUE, loadings.label.colour='navyblue', loadings.label.size=5, loadings.label.vjust= 1, loadings.label.hjust= 1.2, frame=TRUE, frame.type='norm')+ ggtitle("Plot 5") + theme_minimal()
plot6

```

*Variable chart:*

  - D: dorsal ray count
  - P1: left pectoral ray count
  - P2.L: left pelvic rays
  - P2.R: right pelvic rays
  - A: anal rays
  - P1.R: right pectoral ray count
  - LLSC: lateral line scale count
  - SALL: scales above lateral line
  - SBLL: scales below lateral line
  - SBDF: scales before dorsal fin
  - TL: total length
  - SL: standard length
  - BD: body depth
  - CPD: caudal peduncle depth
  - CPL: caudal peduncle length
  - PreDL: pre-dorsal length
  - DbL: dorsal base length
  - HL/HW/HD: head length/width/depth
  - SnL: snout length
  - OL: ocular length
  
  
  
### T and F tests (body size corrected) {.tabset .tabset-fade .tabset-pills}

#### Calculated residuals {.tabset .tabset-fade .tabset-pills}


I will now use residuals to account for body size in each trait, and compare the variance for each species. Again, the first round will take the raw residuals, and the second round with use the absolute value of the raw residuals.

-   STEP ONE: plot trait vs body size

```{r echo=FALSE}
lat.1 <- lat[complete.cases(lat[, 8:29]),]
form.1 <- form[complete.cases(form[, 8:29]),]


reg.lat.D1 <- lm(lat.1$D ~ lat.1$SL)

reg.lat.P11 <- lm(lat.1$P1 ~ lat.1$SL)

reg.lat.P2.L1 <- lm(lat.1$P2.L ~ lat.1$SL)

reg.lat.P2.R1 <- lm(lat.1$P2.R ~ lat.1$SL)

reg.lat.A1 <- lm(lat.1$A ~ lat.1$SL)

reg.lat.P1.R1 <- lm(lat.1$P1.R ~ lat.1$SL)

reg.lat.LLSC1 <- lm(lat.1$LLSC ~ lat.1$SL)

reg.lat.SALL1 <- lm(lat.1$SALL ~ lat.1$SL)

reg.lat.SBLL1 <- lm(lat.1$SBLL ~ lat.1$SL)

reg.lat.SBDF1 <- lm(lat.1$SBDF ~ lat.1$SL)

reg.lat.TL1 <- lm(lat.1$TL ~ lat.1$SL)

reg.lat.BD1 <- lm(lat.1$BD ~ lat.1$SL)

reg.lat.CPD1 <- lm(lat.1$CPD ~ lat.1$SL)

reg.lat.CPL1 <- lm(lat.1$CPL ~ lat.1$SL)

reg.lat.PreDL1 <- lm(lat.1$PreDL ~ lat.1$SL)

reg.lat.DbL1 <- lm(lat.1$DbL ~ lat.1$SL)

reg.lat.HL1 <- lm(lat.1$HL ~ lat.1$SL)

reg.lat.HD1 <- lm(lat.1$HD ~ lat.1$SL)

reg.lat.HW1 <- lm(lat.1$HW ~ lat.1$SL)

reg.lat.SnL1 <- lm(lat.1$SnL ~ lat.1$SL)

reg.lat.OL1 <- lm(lat.1$OL ~ lat.1$SL)



reg.form.D1 <- lm(form.1$D ~ form.1$SL)

reg.form.P11 <- lm(form.1$P1 ~ form.1$SL)

reg.form.P2.L1 <- lm(form.1$P2.L ~ form.1$SL)

reg.form.P2.R1 <- lm(form.1$P2.R ~ form.1$SL)

reg.form.A1 <- lm(form.1$A ~ form.1$SL)

reg.form.P1.R1 <- lm(form.1$P1.R ~ form.1$SL)

reg.form.LLSC1 <- lm(form.1$LLSC ~ form.1$SL)

reg.form.SALL1 <- lm(form.1$SALL ~ form.1$SL)

reg.form.SBLL1 <- lm(form.1$SBLL ~ form.1$SL)

reg.form.SBDF1 <- lm(form.1$SBDF ~ form.1$SL)

reg.form.TL1 <- lm(form.1$TL ~ form.1$SL)

reg.form.BD1 <- lm(form.1$BD ~ form.1$SL)

reg.form.CPD1 <- lm(form.1$CPD ~ form.1$SL)

reg.form.CPL1 <- lm(form.1$CPL ~ form.1$SL)

reg.form.PreDL1 <- lm(form.1$PreDL ~ form.1$SL)

reg.form.DbL1 <- lm(form.1$DbL ~ form.1$SL)

reg.form.HL1 <- lm(form.1$HL ~ form.1$SL)

reg.form.HD1 <- lm(form.1$HD ~ form.1$SL)

reg.form.HW1 <- lm(form.1$HW ~ form.1$SL)

reg.form.SnL1 <- lm(form.1$SnL ~ form.1$SL)

reg.form.OL1 <- lm(form.1$OL ~ form.1$SL)


```

-   STEP TWO: get residuals for each individual for that trait

```{r echo=FALSE}
res.lat.D1 <- resid(reg.lat.D1)

res.lat.P11 <- resid(reg.lat.P11)

res.lat.P2.L1 <- resid(reg.lat.P2.L1)

res.lat.P2.R1 <- resid(reg.lat.P2.R1)

res.lat.A1 <- resid(reg.lat.A1)

res.lat.P1.R1 <- resid(reg.lat.P1.R1)

res.lat.LLSC1 <- resid(reg.lat.LLSC1)

res.lat.SALL1 <- resid(reg.lat.SALL1)

res.lat.SBLL1 <- resid(reg.lat.SBLL1)

res.lat.SBDF1 <- resid(reg.lat.SBDF1)

res.lat.TL1 <- resid(reg.lat.TL1)

res.lat.BD1 <- resid(reg.lat.BD1)

res.lat.CPD1 <- resid(reg.lat.CPD1)

res.lat.CPL1 <- resid(reg.lat.CPL1)

res.lat.PreDL1 <- resid(reg.lat.PreDL1)

res.lat.DbL1 <- resid(reg.lat.DbL1)

res.lat.HL1 <- resid(reg.lat.HL1)

res.lat.HD1 <- resid(reg.lat.HD1)

res.lat.HW1 <- resid(reg.lat.HW1)

res.lat.SnL1 <- resid(reg.lat.SnL1)

res.lat.OL1 <- resid(reg.lat.OL1)


res.form.D1 <- resid(reg.form.D1)

res.form.P11 <- resid(reg.form.P11)

res.form.P2.L1 <- resid(reg.form.P2.L1)

res.form.P2.R1 <- resid(reg.form.P2.R1)

res.form.A1 <- resid(reg.form.A1)

res.form.P1.R1 <- resid(reg.form.P1.R1)

res.form.LLSC1 <- resid(reg.form.LLSC1)

res.form.SALL1 <- resid(reg.form.SALL1)

res.form.SBLL1 <- resid(reg.form.SBLL1)

res.form.SBDF1 <- resid(reg.form.SBDF1)

res.form.TL1 <- resid(reg.form.TL1)

res.form.BD1 <- resid(reg.form.BD1)

res.form.CPD1 <- resid(reg.form.CPD1)

res.form.CPL1 <- resid(reg.form.CPL1)

res.form.PreDL1 <- resid(reg.form.PreDL1)

res.form.DbL1 <- resid(reg.form.DbL1)

res.form.HL1 <- resid(reg.form.HL1)

res.form.HD1 <- resid(reg.form.HD1)

res.form.HW1 <- resid(reg.form.HW1)

res.form.SnL1 <- resid(reg.form.SnL1)

res.form.OL1 <- resid(reg.form.OL1)

```

-   STEP THREE: convert residuals to absolute value

```{r echo=FALSE}
abs.lat.D1 <- abs(res.lat.D1)

abs.lat.P11 <- abs(res.lat.P11)

abs.lat.P2.L1 <- abs(res.lat.P2.L1)

abs.lat.P2.R1 <- abs(res.lat.P2.R1)

abs.lat.A1 <- abs(res.lat.A1)

abs.lat.P1.R1 <- abs(res.lat.P1.R1)

abs.lat.LLSC1 <- abs(res.lat.LLSC1)

abs.lat.SALL1 <- abs(res.lat.SALL1)

abs.lat.SBLL1 <- abs(res.lat.SBLL1)

abs.lat.SBDF1 <- abs(res.lat.SBDF1)

abs.lat.TL1 <- abs(res.lat.TL1)

abs.lat.BD1 <- abs(res.lat.BD1)

abs.lat.CPD1 <- abs(res.lat.CPD1)

abs.lat.CPL1 <- abs(res.lat.CPL1)

abs.lat.PreDL1 <- abs(res.lat.PreDL1)

abs.lat.DbL1 <- abs(res.lat.DbL1)

abs.lat.HL1 <- abs(res.lat.HL1)

abs.lat.HD1 <- abs(res.lat.HD1)

abs.lat.HW1 <- abs(res.lat.HW1)

abs.lat.SnL1 <- abs(res.lat.SnL1)

abs.lat.OL1 <- abs(res.lat.OL1)


abs.form.D1 <- abs(res.form.D1)

abs.form.P11 <- abs(res.form.P11)

abs.form.P2.L1 <- abs(res.form.P2.L1)

abs.form.P2.R1 <- abs(res.form.P2.R1)

abs.form.A1 <- abs(res.form.A1)

abs.form.P1.R1 <- abs(res.form.P1.R1)

abs.form.LLSC1 <- abs(res.form.LLSC1)

abs.form.SALL1 <- abs(res.form.SALL1)

abs.form.SBLL1 <- abs(res.form.SBLL1)

abs.form.SBDF1 <- abs(res.form.SBDF1)

abs.form.TL1 <- abs(res.form.TL1)

abs.form.BD1 <- abs(res.form.BD1)

abs.form.CPD1 <- abs(res.form.CPD1)

abs.form.CPL1 <- abs(res.form.CPL1)

abs.form.PreDL1 <- abs(res.form.PreDL1)

abs.form.DbL1 <- abs(res.form.DbL1)

abs.form.HL1 <- abs(res.form.HL1)

abs.form.HD1 <- abs(res.form.HD1)

abs.form.HW1 <- abs(res.form.HW1)

abs.form.SnL1 <- abs(res.form.SnL1)

abs.form.OL1 <- abs(res.form.OL1)


```

#### F tests to compare species (raw) {.tabset .tabset-fade .tabset-pills}


    -   first, lat\>form, then form\>lat

```{r echo=FALSE}
#lat>form

var.res.D1 <- var.test(res.lat.D1, res.form.D1, alternative = "greater")


var.res.P11 <- var.test(res.lat.P11, res.form.P11, alternative = "greater")


var.res.P2.L1 <- var.test(res.lat.P2.L1, res.form.P2.L1, alternative = "greater")


var.res.P2.R1 <- var.test(res.lat.P2.R1, res.form.P2.R1, alternative = "greater")


var.res.A1 <- var.test(res.lat.A1, res.form.A1, alternative = "greater")


var.res.P1.R1 <- var.test(res.lat.P1.R1, res.form.P1.R1, alternative = "greater")


var.res.LLSC1 <- var.test(res.lat.LLSC1, res.form.LLSC1, alternative = "greater")


var.res.SALL1 <- var.test(res.lat.SALL1, res.form.SALL1, alternative = "greater")


var.res.SBLL1 <- var.test(res.lat.SBLL1, res.form.SBLL1, alternative = "greater")


var.res.SBDF1 <- var.test(res.lat.SBDF1, res.form.SBDF1, alternative = "greater")


var.res.TL1 <- var.test(res.lat.TL1, res.form.TL1, alternative = "greater")


var.res.BD1 <- var.test(res.lat.BD1, res.form.BD1, alternative = "greater")


var.res.CPD1 <- var.test(res.lat.CPD1, res.form.CPD1, alternative = "greater")


var.res.CPL1 <- var.test(res.lat.CPL1, res.form.CPL1, alternative = "greater")


var.res.PreDL1 <- var.test(res.lat.PreDL1, res.form.PreDL1, alternative = "greater")


var.res.DbL1 <- var.test(res.lat.DbL1, res.form.DbL1, alternative = "greater")


var.res.HL1 <- var.test(res.lat.HL1, res.form.HL1, alternative = "greater")


var.res.HD1 <- var.test(res.lat.HD1, res.form.HD1, alternative = "greater")


var.res.HW1 <- var.test(res.lat.HW1, res.form.HW1, alternative = "greater")


var.res.SnL1 <- var.test(res.lat.SnL1, res.form.SnL1, alternative = "greater")


var.res.OL1 <- var.test(res.lat.OL1, res.form.OL1, alternative = "greater")




#form>lat

var.res.D2 <- var.test(res.lat.D1, res.form.D1, alternative = "less")


var.res.P12 <- var.test(res.lat.P11, res.form.P11, alternative = "less")


var.res.P2.L2 <- var.test(res.lat.P2.L1, res.form.P2.L1, alternative = "less")


var.res.P2.R2 <- var.test(res.lat.P2.R1, res.form.P2.R1, alternative = "less")


var.res.A2 <- var.test(res.lat.A1, res.form.A1, alternative = "less")


var.res.P1.R2 <- var.test(res.lat.P1.R1, res.form.P1.R1, alternative = "less")


var.res.LLSC2 <- var.test(res.lat.LLSC1, res.form.LLSC1, alternative = "less")


var.res.SALL2 <- var.test(res.lat.SALL1, res.form.SALL1, alternative = "less")


var.res.SBLL2 <- var.test(res.lat.SBLL1, res.form.SBLL1, alternative = "less")


var.res.SBDF2 <- var.test(res.lat.SBDF1, res.form.SBDF1, alternative = "less")


var.res.TL2 <- var.test(res.lat.TL1, res.form.TL1, alternative = "less")


var.res.BD2 <- var.test(res.lat.BD1, res.form.BD1, alternative = "less")


var.res.CPD2 <- var.test(res.lat.CPD1, res.form.CPD1, alternative = "less")


var.res.CPL2 <- var.test(res.lat.CPL1, res.form.CPL1, alternative = "less")


var.res.PreDL2 <- var.test(res.lat.PreDL1, res.form.PreDL1, alternative = "less")


var.res.DbL2 <- var.test(res.lat.DbL1, res.form.DbL1, alternative = "less")


var.res.HL2 <- var.test(res.lat.HL1, res.form.HL1, alternative = "less")


var.res.HD2 <- var.test(res.lat.HD1, res.form.HD1, alternative = "less")


var.res.HW2 <- var.test(res.lat.HW1, res.form.HW1, alternative = "less")


var.res.SnL2 <- var.test(res.lat.SnL1, res.form.SnL1, alternative = "less")


var.res.OL2 <- var.test(res.lat.OL1, res.form.OL1, alternative = "less")



##ABSOLUTE VALUES

#lat>form


var.abs.D1 <- var.test(abs.lat.D1, abs.form.D1, alternative = "greater")


var.abs.P11 <- var.test(abs.lat.P11, abs.form.P11, alternative = "greater")


var.abs.P2.L1 <- var.test(abs.lat.P2.L1, abs.form.P2.L1, alternative = "greater")


var.abs.P2.R1 <- var.test(abs.lat.P2.R1, abs.form.P2.R1, alternative = "greater")


var.abs.A1 <- var.test(abs.lat.A1, abs.form.A1, alternative = "greater")


var.abs.P1.R1 <- var.test(abs.lat.P1.R1, abs.form.P1.R1, alternative = "greater")


var.abs.LLSC1 <- var.test(abs.lat.LLSC1, abs.form.LLSC1, alternative = "greater")


var.abs.SALL1 <- var.test(abs.lat.SALL1, abs.form.SALL1, alternative = "greater")


var.abs.SBLL1 <- var.test(abs.lat.SBLL1, abs.form.SBLL1, alternative = "greater")


var.abs.SBDF1 <- var.test(abs.lat.SBDF1, abs.form.SBDF1, alternative = "greater")


var.abs.TL1 <- var.test(abs.lat.TL1, abs.form.TL1, alternative = "greater")


var.abs.BD1 <- var.test(abs.lat.BD1, abs.form.BD1, alternative = "greater")


var.abs.CPD1 <- var.test(abs.lat.CPD1, abs.form.CPD1, alternative = "greater")


var.abs.CPL1 <- var.test(abs.lat.CPL1, abs.form.CPL1, alternative = "greater")


var.abs.PreDL1 <- var.test(abs.lat.PreDL1, abs.form.PreDL1, alternative = "greater")


var.abs.DbL1 <- var.test(abs.lat.DbL1, abs.form.DbL1, alternative = "greater")


var.abs.HL1 <- var.test(abs.lat.HL1, abs.form.HL1, alternative = "greater")


var.abs.HD1 <- var.test(abs.lat.HD1, abs.form.HD1, alternative = "greater")


var.abs.HW1 <- var.test(abs.lat.HW1, abs.form.HW1, alternative = "greater")


var.abs.SnL1 <- var.test(abs.lat.SnL1, abs.form.SnL1, alternative = "greater")


var.abs.OL1 <- var.test(abs.lat.OL1, abs.form.OL1, alternative = "greater")




#form>lat

var.abs.D2 <- var.test(abs.lat.D1, abs.form.D1, alternative = "less")


var.abs.P12 <- var.test(abs.lat.P11, abs.form.P11, alternative = "less")


var.abs.P2.L2 <- var.test(abs.lat.P2.L1, abs.form.P2.L1, alternative = "less")


var.abs.P2.R2 <- var.test(abs.lat.P2.R1, abs.form.P2.R1, alternative = "less")


var.abs.A2 <- var.test(abs.lat.A1, abs.form.A1, alternative = "less")


var.abs.P1.R2 <- var.test(abs.lat.P1.R1, abs.form.P1.R1, alternative = "less")


var.abs.LLSC2 <- var.test(abs.lat.LLSC1, abs.form.LLSC1, alternative = "less")


var.abs.SALL2 <- var.test(abs.lat.SALL1, abs.form.SALL1, alternative = "less")


var.abs.SBLL2 <- var.test(abs.lat.SBLL1, abs.form.SBLL1, alternative = "less")


var.abs.SBDF2 <- var.test(abs.lat.SBDF1, abs.form.SBDF1, alternative = "less")


var.abs.TL2 <- var.test(abs.lat.TL1, abs.form.TL1, alternative = "less")


var.abs.BD2 <- var.test(abs.lat.BD1, abs.form.BD1, alternative = "less")


var.abs.CPD2 <- var.test(abs.lat.CPD1, abs.form.CPD1, alternative = "less")


var.abs.CPL2 <- var.test(abs.lat.CPL1, abs.form.CPL1, alternative = "less")


var.abs.PreDL2 <- var.test(abs.lat.PreDL1, abs.form.PreDL1, alternative = "less")


var.abs.DbL2 <- var.test(abs.lat.DbL1, abs.form.DbL1, alternative = "less")


var.abs.HL2 <- var.test(abs.lat.HL1, abs.form.HL1, alternative = "less")


var.abs.HD2 <- var.test(abs.lat.HD1, abs.form.HD1, alternative = "less")


var.abs.HW2 <- var.test(abs.lat.HW1, abs.form.HW1, alternative = "less")


var.abs.SnL2 <- var.test(abs.lat.SnL1, abs.form.SnL1, alternative = "less")


var.abs.OL2 <- var.test(abs.lat.OL1, abs.form.OL1, alternative = "less")


```


```{r, echo=FALSE}

variance.table <- matrix(c(var.res.D1$p.value, var.res.D2$p.value, var.abs.D1$p.value, var.abs.D2$p.value, var.res.P11$p.value, var.res.P12$p.value, var.abs.P11$p.value, var.abs.P12$p.value, var.res.P2.L1$p.value, var.res.P2.L2$p.value, var.abs.P2.L1$p.value, var.abs.P2.L2$p.value, var.res.P2.R1$p.value, var.res.P2.R2$p.value, var.abs.P2.R1$p.value, var.abs.P2.R2$p.value, var.res.A1$p.value, var.res.A2$p.value, var.abs.A1$p.value, var.abs.A2$p.value, var.res.P1.R1$p.value, var.res.P1.R2$p.value, var.abs.P1.R1$p.value, var.abs.P1.R2$p.value, var.res.LLSC1$p.value, var.res.LLSC2$p.value, var.abs.LLSC1$p.value, var.abs.LLSC2$p.value, var.res.SALL1$p.value, var.res.SALL2$p.value, var.abs.SALL1$p.value, var.abs.SALL2$p.value, var.res.SBLL1$p.value, var.res.SBLL2$p.value, var.abs.SBLL1$p.value, var.abs.SBLL2$p.value, var.res.SBDF1$p.value, var.res.SBDF2$p.value, var.abs.SBDF1$p.value, var.abs.SBDF2$p.value, var.res.TL1$p.value, var.res.TL2$p.value, var.abs.TL1$p.value, var.abs.TL2$p.value, var.res.BD1$p.value, var.res.BD2$p.value, var.abs.BD1$p.value, var.abs.BD2$p.value, var.res.CPD1$p.value, var.res.CPD2$p.value, var.abs.CPD1$p.value, var.abs.CPD2$p.value, var.res.CPL1$p.value, var.res.CPL2$p.value, var.abs.CPL1$p.value, var.abs.CPL2$p.value, var.res.PreDL1$p.value, var.res.PreDL2$p.value, var.abs.PreDL1$p.value, var.abs.PreDL2$p.value, var.res.DbL1$p.value, var.res.DbL2$p.value, var.abs.DbL1$p.value, var.abs.DbL2$p.value, var.res.HL1$p.value, var.res.HL2$p.value, var.abs.HL1$p.value, var.abs.HL2$p.value, var.res.HD1$p.value, var.res.HD2$p.value, var.abs.HD1$p.value, var.abs.HD2$p.value, var.res.HW1$p.value, var.res.HW2$p.value, var.abs.HW1$p.value, var.abs.HW2$p.value, var.res.SnL1$p.value, var.res.SnL2$p.value, var.abs.SnL1$p.value, var.abs.SnL2$p.value, var.res.OL1$p.value, var.res.OL2$p.value, var.abs.OL1$p.value, var.abs.OL2$p.value), ncol=4, byrow=TRUE)

colnames(variance.table)<- c('sail>amz', 'amz>sail', '|sail>amz|', '|amz>sail|')

rownames(variance.table) <- c('dorsal rays', 'L pect rays', 'L pelvic rays', 'R pelvic rays', 'Anal rays', 'R pect rays', 'lat line scales', 'scales above LL', 'scales below LL', 'scales before dor', 'total len', 'body dep', 'peduncle dep', 'peduncle leng', 'pre-dorsal', 'dorsal base', 'head lengtt', 'head depth', 'head width', 'snout leng', 'orbital leng')

variance.table


```
