---
title: "morphology_analysis"
author: "Allison Davis"
date: "November 21, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data collection

  * raw data: occurance number and preference score for each individual fish (left, right, NA)
  * pref data: preference scores for tank A, B, combo of A and B, and average of A and B
      - combo score determined as left/(left+right) for total left and right occurances
      - average score determined by averaging tank A pref score and tank B pref score
  * separated all data by species

```{r}
library(curl)

rawRE <- curl("https://raw.githubusercontent.com/allisondavis/mirror-exp_wild-pop/master/adavis_mirror-exp_wild-pop_rescored-raw.csv")

rawRE <- read.csv(rawRE, header = TRUE, sep = ",", stringsAsFactors = TRUE)

prefRE <- curl("https://raw.githubusercontent.com/allisondavis/mirror-exp_wild-pop/master/adavis_mirror-exp_wild-pop_rescored-prefs.csv")

prefRE <- read.csv(prefRE, header = TRUE, sep = ",", stringsAsFactors = TRUE)